### 2019-03-15 0_abstract.tex
Add abstract.

\begin{abstract}
% 150 words
We introduce \emph{minimum effective task granularity} (METG), a new
metric for quantifying the fundamental performance behavior of
parallel and distributed runtime systems. METG is either directly
related to, or correlates with a variety of measures of direct
interest to application developers: namely the smallest problem size
that can be efficiently weak scaled, or the limit of strong
scalability. METG can be used to study the baseline overhead, scaling
behavior, and communication hiding ability of runtime systems, among
other properties.

We present Task Bench, a novel \emph{parameterized} benchmark for
exploring the behavior of parallel and distributed runtime
systems. Task Bench executes a configurable \emph{task graph} with
tasks for each point in an iteration space and dependencies determined
by a \emph{dependence relation}. We use Task Bench to conduct a
comprehensive evaluation of 13 runtime systems, and show that the
baseline overhead of these systems varies by up to 5 orders of
magnitude.
\end{abstract}

### 2019-03-15 1_introduction.tex
Add introduction text.

\section{Introduction}

Challenges in heterogeneous and extreme-scale parallel and distributed
computing have lead to the rise of a wide variety of new programming
models and runtime systems. However, efforts to understand the
performance behavior of these system have been limited. In many cases,
programming models are evaluated with weak and/or strong scaling of a
bespoke set of carefully hand-written applications, which make it
difficult to understand how the results generalize to other
applications. Furthermore, these are not limit studies, and thus do
not help to elucide the boundaries of the performance envelops of the
respective systems.

We introduce \emph{minimum effective task granularity} (METG), a
metric of application performance specifically designed to explore the
fundamental performance behavior of parallel and distributed runtime
systems. METG can be used to identify the baseline overhead, limit of
scalability, and ability to hide and overlap communication of such
systems. A key insight of the METG metric is that large \emph{tasks},
or units of work, can be used to hide almost any amount of
overhead. Thus METG considers the \emph{minimum} task granularity that
can be executed. This is a lower bound, so as long as the average task
granularity of a given application exceeds this value, the runtime
system should be able to execute it efficiently.

A variety of runtime systems are able to make use of parallelism
internally for a variety of purposes, from analysis of tasks, to
scheduling or the orchestration of data movement and
synchronization. In order to mitigate unwanted influence due to
unreasonable allocation of resources to such internal uses, METG
concerns \emph{effective} task execution, where the overall
computational efficiency meets some threshold (say, 50\%). If the
METG(50\%) of an application is, say, 1~ms, then this indicates that
the application can be executed with as small as 1~ms tasks while
maintaining at least 50\% overall efficiency.

We present Task Bench, a benchmark designed specifically to study the
performance behavior of parallel and distributed runtime systems at
the limits of runtime overhead. Task Bench is a \emph{parameterized}
benchmark, characterized by an \emph{iteration space} of tasks (with
one task per point in the iteration space) and a \emph{dependence
  relation} which determines how tasks depend on other tasks. By
formulating the benchmark in this way, we make it possible to rapidly
simulate a wide variety of communication patterns relevant in
scientific computing, such as trivial parallelism, halo exchanges (for
structured and unstructured mesh-based computations), sweeps (such as
for the discrete ordinates method), FFTs, trees (for divide and
conquer algorithms), and so on. Task Bench can be configured to
execute a variety of kernels, such as compute- and memory-bound
kernels, and load imbalanced variants of the above. Task Bench can
also be used to stress communication-bound cases, by configuring the
amount of data that is produced by each task.

In order to explore the impact of task parallelism on runtime
performance, Task Bench can also execute multiple task graphs
simultaneously. Runtime systems which support asynchronous execution
can use this additional scheduling flexibility to hide communication
overheads or to mitigate load imbalance. METG makes it possible to
quantify the degree to which various systems are successful at doing
so, and to measure if (and when) such factors are more, or less,
important than baseline runtime overhead.

We provide a comprehensive evaluation of Task Bench implementations in
13 different runtime systems, including Chapel~\cite{Chapel07},
Charm++~\cite{Charmpp93}, MPI~\cite{MPI}, OmpSs~\cite{OmpSs11},
OpenMP~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13},
Realm~\cite{Realm14}, Regent~\cite{Regent15}, Spark~\cite{Spark10},
StarPU~\cite{StarPU11}, Swift/T~\cite{Wozniak13},
TensorFlow~\cite{TensorFlow15}, and X10~\cite{X1005}. Using the METG
metric, we determine the baseline runtime overhead of each system
along with overheads associated with increasing number and/or
variations in the pattern of dependencies. We also explore each
system's ability to scale, hide communication overhead, and mitigate
load imbalance.

The following four sections each describe a contribution of this
paper:

\begin{itemize}
\item Section~\ref{sec:metg} defines the METG metric.
\item Section~\ref{sec:task-bench} describe the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 13 runtime systems.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench performance on up to 64 nodes of the Cori
  supercomputer.
\end{itemize}

Section~\ref{sec:related-work} relates this work to previous efforst,
and Section~\ref{sec:conclusion} concludes.

### 2019-03-15 1_introduction.tex
Add new section.

@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 \begin{itemize}
-\item Section~\ref{sec:metg} defines the METG metric.
-\item Section~\ref{sec:task-bench} describe the design of Task Bench.
+\item Section~\ref{sec:metg} defines the METG metric and its
+  relationship to quantities of interest to application developers.
+\item Section~\ref{sec:task-bench} describes the design of Task Bench.
 \item Section~\ref{sec:implementation} details the implementation of

### 2019-03-15 2_metg.tex
Add new section.

\section{METG}
\label{sec:metg}

\input{f1_flops}
\input{f2_efficiency}

### 2019-03-15 2_metg.tex
More text in section 2.

@@
 \input{f2_efficiency}
+
+Minimum effective task granularity (METG) is defined intuitively as
+the smallest average task granularity such that the application
+achieves overall efficiency exceeding some threshold. For example, in
+a compute-bound application efficiency might be measure as the
+percentage of the available FLOPS achieved in a given application
+run. On a machine with a peak capability of 1.2 TFLOPS per node,
+METG(50\%) would correspond to smallest task granularity achieved
+while maintaing at least 0.6 TFLOPS. Absolute efficiency (i.e., FLOPS
+or bytes per second) is preferrable, but parallel efficiency (relative
+to a sufficiently large problem size) can also be used in applications
+which are not amenable to being characterized by peak resource
+utilization.
+
+Figure~\ref{fig:flops} shows the way in which METG is calculated. The
+application is run with a large enough problem size to ensure that it
+achieves peak utilization of the machine. This confirms that the
+application is properly configured and that the choice of efficiency
+metric is achievable. Once this is established, the problem size is
+shrunk (in a manner analogous to what happens on a node during a
+strong scaling run). The expectation is as problem size shrinks, the
+efficiency will begin to drop and eventually approach zero. As seen in
+Figure~\ref{fig:flops}, the systems with lower overheads will maintain
+higher efficiency at smaller problem sizes compared to systems with
+higher overheads.
+
+To calculate METG, this data is replotted along the axes of efficiency
+(i.e., as a percentage of the peak FLOPS achieved) and task
+granularity (i.e., $\text{wall time} \times \text{number of
+  cores}/\text{number of tasks}$). Figure~\ref{fig:efficiency} shows
+the replotted data from above. In this chart we can see that
+efficiency starts near 100\% for the vast majority of systems, and
+that initially task granularity drops quickly with minimal change in
+efficiency. However as task granularity shrinks, efficiency begins to
+drop as well, leading most systems to reach a plateau. Intuitively,
+this plateau represents the fundamental baseline overhead of the
+system, the point beyond which the system simply cannot execute tasks
+efficiently.
+
+METG is thus the result of drawing a vertical line through the chart
+in Figure~\ref{fig:efficiency} at the desired level of efficiency. For
+example, at 50\% efficiency, MPI achieves an average task granularity
+of 4.5 us while Charm++ achieves an average task granularity of 12
+us. Thus the METG(50\%) of MPI and Charm++ are 4.5 and 12 us,
+respectively, for this application. The number 50\% is chosen
+arbitrarily, but is a reasonable choice for this application because
+it represents a point at which most systems have plateaued, yet the
+overall efficiency is still reasonable. Lower values would likely not
+be reasonable, because in most cases operators of computing facilities
+prefer to maintain higher levels of utilization of the machines. And
+higher values would misrepresent the performance of more systems.
+
+METG is an appealing metric because it has a well-understood
+relationship with quantities of interest to application developers,
+namely weak and strong scaling. METG corresponds to the smallest
+problem size that can be weak scaled to a given node count with a
+given level of efficiency. Larger problem sizes would correspond with
+larger tasks, and therefore have lower overhead relative to useful
+computation. Similarly, METG corresponds to the point at which strong
+scaling can be expected to stop; as strong scaling runs start with a
+larger problem size and progressively shrink it until overheads begin
+to dwarf the gains due to continued scaling.
+
+The METG metric has one particularly useful property in comparison to
+weak and strong scaling. Because METG is measured ``in place'' (i.e.,
+without changing the number of nodes or cores available to the
+application), it is possible to use METG to isolate effects that are
+due to shrinking problem size from effects that are due to e.g.,
+increased communication latencies as progressively larger portions of
+the machine are being used. Section~\ref{sec:evaluation} contains a
+comprehensive evaluation of METG including how it changes with node
+count.
+
+% talk about TPS / MTG ?

### 2019-03-18 1_introduction.tex
Edits, text, and new figures.

@@
 difficult to understand how the results generalize to other
-applications. Furthermore, these are not limit studies, and thus do
-not help to elucide the boundaries of the performance envelops of the
-respective systems.
+applications, and impossible to compare across systems. Furthermore,
+these are not limit studies, and thus do nothing to elucide the
+boundaries of the performance envelops of the respective systems.
 
@@
 METG(50\%) of an application is, say, 1~ms, then this indicates that
-the application can be executed with as small as 1~ms tasks while
+the application can be executed with tasks as small as 1~ms while
 maintaining at least 50\% overall efficiency.
@@
 \item Section~\ref{sec:evaluation} provides a comprehensive evaluation
-  of Task Bench performance on up to 64 nodes of the Cori
+  of Task Bench performance on up to 256 nodes of the Cori
   supercomputer.
@@
 
-Section~\ref{sec:related-work} relates this work to previous efforst,
+Section~\ref{sec:related-work} relates this work to previous efforts,
 and Section~\ref{sec:conclusion} concludes.

### 2019-03-18 2_metg.tex
Edits, text, and new figures.

--- 2019-03-15 2_metg.tex
@@
 
-\input{f1_flops}
-\input{f2_efficiency}
+\input{f1_flops_mpi}
+\input{f2_efficiency_mpi}
+
+Minimum effective task granularity (METG) is defined intuitively as
+the smallest average task granularity such that the application
+achieves overall efficiency exceeding some threshold. For example, in
+a compute-bound application efficiency might be measured as the
+percentage of the available FLOPS achieved. On a machine with a peak
+capability of 1.2 TFLOPS per node, METG(50\%) would correspond to
+smallest task granularity achieved while maintaing at least 0.6
+TFLOPS. Absolute efficiency (i.e., FLOPS or bytes per second) is
+preferrable, but parallel efficiency (relative to a sufficiently large
+problem size) can also be used in applications which are not amenable
+to being characterized by peak resource utilization.
+
+Figure~\ref{fig:flops-mpi} shows the way in which METG is
+calculated. The application is run with a large enough problem size to
+ensure that it achieves peak utilization of the machine. This confirms
+that the application is properly configured and that the choice of
+efficiency metric is achievable. Once this is established, the problem
+size is shrunk (in a manner analogous to what happens on a node during
+a strong scaling run). The expectation is as problem size shrinks, the
+efficiency will begin to drop and eventually approach zero. Systems
+with lower overheads will maintain higher efficiency at smaller
+problem sizes compared to systems with higher overheads.
+
+To calculate METG, this data is replotted along axes of efficiency
+(i.e., as a percentage of the peak FLOPS achieved) and task
+granularity (i.e., $\text{wall time} \times \text{number of
+  cores}/\text{number of tasks}$). Note that a \emph{task} is defined
+broadly to be any continuously-executing unit of work in the
+application, thus it makes sense to discuss tasks even in systems
+without an explicit notion of tasking, such as in MPI codes written in
+a bulk synchronous style.
+
+Figure~\ref{fig:efficiency-mpi}
+shows the replotted data from above. In this chart we can see that
+efficiency starts near 100\%, and that initially task granularity
+drops quickly with minimal change in efficiency. However as task
+granularity shrinks, efficiency begins to drop as well, leading
+eventually to a plateau in task granularity. Intuitively, this plateau
+represents the fundamental baseline overhead of the system, the point
+beyond which the system simply cannot execute tasks efficiently.
+
+METG is thus the result of drawing a vertical line through
+Figure~\ref{fig:efficiency-mpi} at the desired level of efficiency,
+and taking the intersection of this line with the task granularity
+curve. In the figures, the red, dashed line shows the 50\% efficiency
+level. At 50\% efficiency, MPI achieves an average task granularity of
+4.5 us, thus the METG(50\%) of MPI is 4.5 us for this configuration of
+the application. The number 50\% is chosen arbitrarily, but is a
+reasonable choice for this application because it represents a point
+at which many systems will have plateaued, yet the overall efficiency
+is still reasonable. Lower values would likely not be reasonable,
+because in most cases operators of computing facilities prefer to
+maintain higher levels of utilization of the machines. And higher
+values risks misrepresenting the performance of some systems (not
+shown in the figures above).
+
+METG is an appealing metric in part because it has a well-defined
+relationship with quantities of interest to application developers,
+namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
+\ref{fig:strong-scaling-mpi} show the behavior of MPI weak and strong
+scaling, respectively, at a variety of problem sizes. Intuitively, at
+larger problem sizes MPI is perfectly efficient. This can be seen at
+the top of each figurs, with flat lines when weak scaling and
+ideally-sloped downward lines when strong scaling. Inefficiency begins
+to show up at smaller problem sizes, towards the bottom of the graph,
+where lines become more compressed and begin to trend upward. At the
+very bottom, the lines compress together as it the fundamental
+overhead of MPI dwarfs any gains from running smaller
+problems. Critically, the shape of the curve at the bottom is the same
+for both weak and strong scaling; this is the precisely the shape of
+the METG curve.
+
+\input{f3_weak_scaling_mpi}
+\input{f4_strong_scaling_mpi}
+
+METG therefore has a direct relationship with the smallest problem
+size that can be weak scaled to a given node count with a given level
+of efficiency. Any smaller and the overhead begins to dominate useful
+computation. Similarly, METG corresponds to the point at which strong
+scaling can be expected to stop; as strong scaling runs start with a
+larger problem size and progressively shrink it until overheads begin
+to dwarf the gains due to continued scaling. In
+Figures~\ref{fig:weak-scaling-mpi} and \ref{fig:strong-scaling-mpi},
+the dashed, bright red line show the value of METG(50\%) at each node
+count.
+
+The METG metric has one particularly useful property in comparison to
+weak and strong scaling. Because METG is measured ``in place'' (i.e.,
+without changing the number of nodes or cores available to the
+application), it is possible to use METG to isolate effects that are
+due to shrinking problem size from effects that are due to e.g.,
+increased communication latencies as progressively larger portions of
+the machine are being used. In comparison, with only a single strong
+or weak scaling run, it is impossible to determine if the shape of the
+curve is due to the overhead of the runtime system, or communication
+effects as progressively larger portions of the machine are being
+used.
+
+Section~\ref{sec:evaluation} contains a comprehensive evaluation of
+METG including how it changes with node count for a wide variety of
+programming systems.
+
+% talk about TPS / MTG ?

### 2019-03-18 3_task_bench.tex
Add section on Task Bench.

\section{Task Bench}
\label{sec:task-bench}

In order to measure properties such as METG for a wide variety of
systems and application scenarios, we developed a novel
\emph{parameterized} benchmark called Task Bench. Task Bench
implementations execute a \emph{task graph}, with tasks for each point
in an \emph{iteration space} and dependencies between tasks determined
by a \emph{dependence relation}. Each task can execute any one of a
number of kernels (compute-bound, memory-bound, etc.) and can generate
a configurable amount of result data in order to the amount of
communication that must be performed. Thus Task Bench can be used to
rapidly explore a large space of possible application scenarios in
order to understand the performance behaviors of parallel and
distributed languages and runtime systems.

For simplicity but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time flowing along
the vertical axis and the width of the task graph along the
horizontal. Tasks can depend only on tasks from the immediately
preceding time step. Many problems can be described by a task graph of
constant width, in which case the iteration space is rectangular, but
Task Bench is capable of handling non-rectangular iteration spaces,
such as required by sweeps and tree-based dependence
patterns. Figure~\ref{fig:task-graphs} show a number of sample task
graphs that can be implemented with Task Bench.

% FIXME: Make figure showing example task graphs

Dependencies between tasks are controlled by a configurable dependence
relation. Dependencies are conceived as flowing backwards in time. The
dependence relation enumerates for each task, the set of tasks in the
previous time step that task depends on. This permits a wide variety
of dependence patterns to be implemented that are relevant to real
applications in high-performance scientific computing: stencils,
sweeps, FFTs, trees, and so on. Dependence patterns may also be
parameterized, such as picking the $K$ nearest neighbors, or $K$
distant neighbors. These dependence patterns can be seen in
Figure~\ref{fig:task-graphs}.

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. Most
Task Bench implementations are several hundred lines long, with the
shortest being 88 lines and the longest being 1500 lines of executable
code. Much of this is possible because the core aspects of generating
the task graph and enumerating dependencies are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all cases, eliminating a potential
source of performance disparity which is common in mini-apps
implementations. Finally, the Task Bench core library manages the
parsing of input parameters and the display out output results,
ensuring that all implementations behave in a uniform manner, and can
be scripted consistently.

The task implementation in the Task Bench core library is fully
validating. Because the task graph configuration is explicitly
represented (though unmaterialized) in Task Bench, this representation
can be queried to determine exactly what dependencies a task should
expect. The output of every task in Task Bench is uniquely generated,
and all inputs are verified. An assertion is thrown if validation
fails. This ensures that every execution of Task Bench, if it
completes successfully, is correct. This validation provides high
assurance that all Task Bench implementations are correct, and
safeguards against corner cases that may not have been adequately
tested during development. An evaluation of the performance impact of
this validation showed it to be less than 3\% at the smallest task
granularities in any Task Bench implementation, with a negligible
effect on overall results.

### 2019-03-18 4_implementation.tex
Initial text for implementation section.

\section{Implementations}
\label{sec:implementation}

We have implemented Task Bench in 13 parallel and distributed
programming languages and runtime systems. Below we describe the
systems, and any salient details of their Task Bench implementations.

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Their insight was invaluable
in ensuring that we provide the highest level of quality in the Task
Bench implementations for each system.

\subsection{Chapel}

Chapel~\cite{Chapel07} is an explicitly parallel programming language
that follows a \emph{multi-resolution} approach to parallel
programming, with support for parallelism at a variety of
levels. Chapel's core features are a partitioned global address space
(PGAS) which permits remote access to objects from anywhere in the
machine, a data model for the distribution of data structures over the
global address space via \emph{domain maps}, tasks, and
synchronization primitives. Chapel thus permits a great deal of
flexibility in the choice of implementation strategy for applications.

For Task Bench, we decided to target a relatively low-level
implementation in Chapel, using explicit task instantiation (via
\lstinline[language=Chapel]{coforall}), bulk access to distributed
arrays for efficient data movement, and atomic integers for
synchronization. The advantage of this approach is that it permits
fine-tuning of the implementation for performance, and avoids relying
on advanced optimizations in the Chapel compiler for performance.

Chapel also supports a more high-level approach to parallel
programming in which the granularities of tasks are determined
automatically by the compiler (via
\lstinline[language=Chapel]{forall}). This style was not evaluated in
Task Bench as we were concerned that the dynamic, input-dependent
nature of the task dependencies would defeat optimizations in the
Chapel compiler that might turn out to be critical for performance.

\subsection{Charm++}

Charm++~\cite{Charmpp93} is an explicitly parallel programming system
based on the actor model of distributed computation. Actors, or
\emph{chares}, are objects that live in their own address space and
can be migrated automatically between nodes for better load
balance. Chares are permitted to send messages to other chares via a
form of remote procedure call (RPC). Calling a method of another chare
is the primary form of synchronization and data movement in
Charm++. Additional support is provided for collective operations
between chares.

Our Task Bench implementation was straightforward: we create a chare
array for the task graph, with one chare for each column. Methods are
used to implement dependencies, and a task is executed as soon as its
dependecies are all available. Collectives are used to synchronize the
completion of the task graph execution.

\subsection{MPI}

MPI~\cite{MPI}

\subsection{OmpSs}

OmpSs~\cite{OmpSs11}

\subsection{OpenMP}

OpenMP~\cite{OpenMPSpec40}

\subsection{PaRSEC}

PaRSEC~\cite{PARSEC13}

\subsection{Realm}

Realm~\cite{Realm14}

\subsection{Regent}

Regent~\cite{Regent15}

\subsection{Spark}

Spark~\cite{Spark10}

\subsection{StarPU}

StarPU~\cite{StarPU11}

\subsection{Swift/T}

Swift/T~\cite{Wozniak13}

\subsection{TensorFlow}

TensorFlow~\cite{TensorFlow15}

\subsection{X10}

X10~\cite{X1005}

### 2019-03-19 4_implementation.tex
More text on implementations.

@@
 fine-tuning of the implementation for performance, and avoids relying
-on advanced optimizations in the Chapel compiler for performance.
+on advanced optimizations in the Chapel compiler, which might struggle
+with the dynamic, input-dependent nature of the task graphs in Task
+Bench.
 
@@
 automatically by the compiler (via
-\lstinline[language=Chapel]{forall}). This style was not evaluated in
-Task Bench as we were concerned that the dynamic, input-dependent
-nature of the task dependencies would defeat optimizations in the
-Chapel compiler that might turn out to be critical for performance.
+\lstinline[language=Chapel]{forall}). This is the level that users
+would typically target in their programs. The specific dependence
+patterns supported by Task Bench might also be implementable at this
+level, though a full Task Bench implementation would be challenging.
+
+% FIXME: Does the Chapel compiler handle: bulk data movement, point-to-point synchronization
+
+Internally, Chapel implements custom support for Cray networks, and
+uses the GASNet~\cite{GASNET07} compatibility layer on other networks.
 
@@
 
-Our Task Bench implementation was straightforward: we create a chare
+Our Task Bench implementation is straightforward: we create a chare
 array for the task graph, with one chare for each column. Methods are
@@
 
+% FIXME: Discuss SMP vs non-SMP versions
+
+Internally, Charm++ implements support for a variety of networks
+directly on top of the low-level, vendor-provided APIs. For Cori, we
+use Charm++'s native support for uGNI.
+
 \subsection{MPI}
 
-MPI~\cite{MPI}
+MPI~\cite{MPI} is the industry standard message-passing API for
+programming supercomputers, and is the most widely used programming
+model in high-performance scientific computing today. The goal of our
+MPI implementation of Task Bench is to represent the common (if
+unexciting) case which continues to dominate high-performance
+computing: bulk synchronous programming with distinct computation and
+communication phases. We use \lstinline[language=C++]{MPI_Isend} and
+\lstinline[language=C++]{MPI_Irecv} to implement the communication
+phase, posting receives before sends. Each task dependency maps to one
+send/receive pair in MPI. Despite the relatively straightforward
+design, our MPI implementation is among the best performing.
+
+In our evaluation, we use the default MPI implementation installed on
+the machine, which is Cray MPI on the Cori supercomputer. This
+implementation is vetted by the machine operators and vendor and is
+the highest performing available implementation for the hardware.
+
+For completeness, we also explored the use of additional MPI calls:
+namely buffered sends (\lstinline[language=C++]{MPI_Bsend}),
+broadcasts (\lstinline[language=C++]{MPI_Bcast}), and all-to-all
+communication (\lstinline[language=C++]{MPI_Alltoallv}). For the
+configurations used in our evaluation, we found non-blocking sends and
+receives to be the best performing, often by a wide margin.
 
@@
 
-OmpSs~\cite{OmpSs11}
+OmpSs~\cite{OmpSs11} is a programming model for loop-based parallelism
+that is source-compatible with OpenMP. Our Task Bench implementation
+uses OpenMP 4.0-style tasks with task dependencies. Because tasks in
+OpenMP 4.0 have a fixed number of dependencies, we use a switch
+statement to implement the dynamic dependencies required for Task
+Bench. The implementation is otherwise straightforward.
+
+% FIXME: More to say about OmpSs? Is there anything different vs the OpenMP implementation?
 
@@
 
-OpenMP~\cite{OpenMPSpec40}
+OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
+parallelism on shared-memory systems. Our Task Bench implementation
+for OpenMP is very similar to OmpSs and uses OpenMP 4.0 tasks with
+task dependencies.
+
+In our evaluation, we test with both GNU GOMP and Intel KMP
+implementations of OpenMP. We found KMP to provide better performance
+and thus report KMP numbers in our evaluation.
 
@@
 
-PaRSEC~\cite{PARSEC13}
+PaRSEC~\cite{PARSEC13} is a task-based programming system which
+supports two distinct programming models: \emph{parameterized task
+  graphs} (PTG) and \emph{sequential task flow} (STF). In the PTG
+model, the program consists of an analytical function which describes
+the expansion of the task graph. PTG is therefore a more limited
+programming model, but can potentially provide better performance
+because the graph description is amenable to static analysis and can
+be pruned efficiently for horizontally scalable execution.
+
+In the STF model, tasks are enumerated in program order (by
+dynamically executing the program), and dependencies between tasks are
+identified automatically based on the arguments to tasks. A task
+depends on another task if it reads data written by the other task,
+and the data is copied automatically if the two tasks are executed on
+different nodes. While the basic STF model is implicitly parallel
+because the system is responsible for the discovery of parallelism, in
+order to achieve horizontal scalability, the program is executed in a
+SPMD fashion on all nodes, and the user is responsible for eliding
+tasks that are not directly connected to those that are to be executed
+on the current node.
+
+Data in PaRSEC is stored in \emph{regions}, or dense arrays of
+contiguous memory. Regions may be passed to tasks on other nodes, but
+can only be written on the node where the data is originally
+allocated. In our Task Bench implementation, each task writes to a
+region as output, and these regions are passed as inputs to the
+subsequent set of tasks. As in the OmpSs and OpenMP implementations,
+the task launch code uses a switch statement as the PaRSEC API assumes
+tasks have a fixed number of dependencies.
+
+% FIXME: What does PaRSEC call regions?
+
+Internally, PaRSEC uses MPI for network communication.
 
@@
 
-Realm~\cite{Realm14}
+Realm~\cite{Realm14} is an explicitly-parallel task-based programming
+model that is used internally by Legion~\cite{Legion12} and
+Regent~\cite{Regent15}, which are implicitly parallel layers written
+on top of Realm. Thus the Realm implementation of Task Bench can be
+seen as a limit study of what is possible with Legion or Regent.
+
+Tasks in Realm are executed in a deferred manner, with dependencies
+between tasks explicitly defined by \emph{events} passed from one task
+to another. Realm's data model supports regions which live on a single
+node. Data must be explicitly copied, and the copies synchronized with
+tasks via events.
+
+Realm also supports a subgraph API where sets of tasks and copies are
+issued together. This mode permits additional optimizations to be
+performed. We have implemented Task Bench in both individual and
+subgraph styles, but report numbers for the subgraph version as it
+provides better performance.
+
+Internally, Realm uses GASNet~\cite{GASNET07} for communication.
 
@@
 
-Regent~\cite{Regent15}
+Regent~\cite{Regent15} is an implicitly-parallel task-based language
+which implements the Legion programming model~\cite{Legion12}. We use
+Regent rather than Legion directly because the Regent compiler
+provides a critical optimization, \emph{control replication}, that
+enables superior horizontal scalability of
+applications~\cite{ControlReplication17}.
+
+In Regent tasks are executed sequentially and parallelism is
+discovered automatically based on the data that is read or written by
+tasks. Regent programs are agnostic to the machine architecture and
+are separately \emph{mapped} for execution on a distributed memory
+system. Copies are inserted automatically by the compiler and runtime
+whenever data is read on a different node from where it is written.
+
+The Regent compiler performs control replication to slice the Regent
+program into \emph{shards} of execution which can scalably execute on
+large numbers of nodes. Our Regent implementation also makes use of a
+tracing optimization~\cite{LegionTracing18} which substantially
+reduces the cost of dynamic analysis of tasks in Regent.
 

### 2019-03-19 4_implementation.tex
More systems.

@@
 
-Spark~\cite{Spark10}
+Spark~\cite{Spark10} is an implicitly-parallel programming model for
+data analytics, widely used in industrial data center applications. As
+one of the programming models in this study not originally intended
+for scientific computing, Spark makes very different tradeoffs from
+other systems under consideration, and therefore makes an interesting
+point of comparison.
+
+Spark's programming model is based of \emph{functional parallelism}, a
+generalization of the MapReduce programming
+model~\cite{MapReduce04}. The core computational abstraction in Spark
+are operators such as map, reduce, filter, join, etc. which by virtue
+of being pure functions are trivially known to be
+parallizable. Functions in Spark operate on \emph{resilient
+  distributed datasets} (RDDs), which are globally-visible,
+dynamically single-assignment data structures that cached in memory
+and written to disk for durability. One of Spark's primary
+optimizations therefore is the caching of RDDs to avoid unnecessary
+disk traffic, resulting in substantial speedups over MapReduce.
+
+Though Spark has tasks internally, these are not exposed to the user,
+so one of the challenges in developing a Task Bench implementation is
+mapping the task graph to a set of operators which will result in the
+desired performance behavior. We use a combination of
+\lstinline[language=Scala]{flatMap} and
+\lstinline[language=Scala]{groupByKey} operations to generate task
+dependencies, and then \lstinline[language=Scala]{map} to execute a
+series of tasks. We use an explicit hash partitioner to ensure that
+Spark does not attempt to group multiple Task Bench tasks into a
+single Spark task.
+
+We performed an extensive set of experiments to verify that no
+extraneous factors interfered with our Spark measurements. We
+completely disabled logging, ensured that no RDDs are actually written
+to disk, confirmed that there is no measurable overhead due to JNI
+calls from Java to C, and cross-checked our results with known cases
+that hit optimal task throughput in Spark, among other things. Our
+results are discussed in detail in Section~\ref{sec:evaluation}.
 
@@
 
-StarPU~\cite{StarPU11}
+StarPU~\cite{StarPU11} is a task-based system that supports an STF
+programming model similar to PaRSEC. Our Task Bench implementation in
+StarPU is very similar to PaRSEC, including the use of a switch
+statement to handle dynamic numbers of task dependencies, and manual
+elision of tasks on unrelated nodes to provide horizontal scalability.
+
+Internally, StarPU uses MPI for communication.
 
@@
 
-Swift/T~\cite{Wozniak13}
+Swift/T~\cite{Wozniak13} is a parallel scripting language intended
+primarily for the composition of workflows in high-performance
+scientific computing. Tasks in Swift/T can be written in any
+programming language, and may even be themselves parallel. Swift/T
+programs follow dataflow semantics, where every statement may
+potentially execute in parallel as soon as its dependencies are
+satisfied; statements only execute sequentially when explicitly
+requested. The Swift/T compiler performs a number of optimizations to
+improve performance of highly parallel programs~\cite{Armstrong14}.
+
+Our Task Bench in Swift/T is straightforward, using Swift/T's dataflow
+semantics to capture dependencies on other tasks.
+
+Internally, Swift/T uses MPI for communication.
 
@@
 
-TensorFlow~\cite{TensorFlow15}
+TensorFlow~\cite{TensorFlow15} is a programming system intended for
+the executing deep learning workloads, widely used in industrial data
+center applications. Although TensorFlow's high-level API exposes
+concepts specific to machine learning, internally TensorFlow is
+implemented as a task graph execution engine, making it a good fit for
+Task Bench and one of the shorter implementations at only 88 lines.
+
+Our Task Bench implementation in TensorFlow works by defining a custom
+operator, which represents the execution of one task. We then use
+TensorFlow APIs to instantiate the custom operator for each point in
+the task graph and connect the task dependencies with the TensorFlow
+API.
 
@@
 
-X10~\cite{X1005}
+X10~\cite{X1005} is an explicitly parallel programming language for
+place-based programming. The core features of X10 are \emph{places}
+which represent distributed memories, a PGAS model where references to
+remote objects can be held (but can only be deferenced on the local
+place), asynchronous tasks, and a place-changing construct to move the
+execution of a task to a remote place. X10 also supports a variety of
+synchronization primitives.
+
+Our Task Bench implementation uses
+\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
+between places, along with place-changing and atomic integers for
+synchronization. We use the C++ backend of X10, which uses MPI
+internally for portability to a variety of networks, and which
+conveniently integrates well with the Task Bench API.

### 2019-03-19 1_introduction.tex
read intro

@@
 models and runtime systems. However, efforts to understand the
-performance behavior of these system have been limited. In many cases,
+performance behavior of these systems have been limited. In many cases,
 programming models are evaluated with weak and/or strong scaling of a
@@
 applications, and impossible to compare across systems. Furthermore,
-these are not limit studies, and thus do nothing to elucide the
+these are not limit studies, and thus do nothing to elucidate the
 boundaries of the performance envelops of the respective systems.
@@
 maintaining at least 50\% overall efficiency. 
+\textcolor{red}{TODO:explain what does efficiency mean here?}
 

### 2019-03-19 5_evaluation.tex
Initial experiment figures.

\section{Evaluation}
\label{sec:evaluation}

\input{e1_metg_compute}

### 2019-03-19 2_metg.tex
read metg

@@
 capability of 1.2 TFLOPS per node, METG(50\%) would correspond to
-smallest task granularity achieved while maintaing at least 0.6
+smallest task granularity achieved while maintaining at least 0.6
 TFLOPS. Absolute efficiency (i.e., FLOPS or bytes per second) is
-preferrable, but parallel efficiency (relative to a sufficiently large
+preferable, but parallel efficiency (relative to a sufficiently large
 problem size) can also be used in applications which are not amenable
-to being characterized by peak resource utilization.
+to be characterized by peak resource utilization.
 
@@
 application, thus it makes sense to discuss tasks even in systems
-without an explicit notion of tasking, such as in MPI codes written in
+without an explicit notion of tasking, such as in MPI codes are written in
 a bulk synchronous style.
@@
 Figure~\ref{fig:efficiency-mpi}
-shows the replotted data from above. In this chart we can see that
+shows the replotted data from above. In this chart, we can see that
 efficiency starts near 100\%, and that initially task granularity
-drops quickly with minimal change in efficiency. However as task
+drops quickly with a minimal change in efficiency. However as task
 granularity shrinks, efficiency begins to drop as well, leading
@@
 represents the fundamental baseline overhead of the system, the point
-beyond which the system simply cannot execute tasks efficiently.
+below which the system simply cannot execute tasks efficiently.
 
@@
 larger problem sizes MPI is perfectly efficient. This can be seen at
-the top of each figurs, with flat lines when weak scaling and
+the top of each figure, with flat lines when weak scaling and
 ideally-sloped downward lines when strong scaling. Inefficiency begins
@@
 problems. Critically, the shape of the curve at the bottom is the same
-for both weak and strong scaling; this is the precisely the shape of
+for both weak and strong scaling; this is precisely the shape of
 the METG curve.
@@
 size that can be weak scaled to a given node count with a given level
-of efficiency. Any smaller and the overhead begins to dominate useful
+of efficiency. With any smaller problem size, 
+the overhead begins to dominate useful
 computation. Similarly, METG corresponds to the point at which strong

### 2019-03-19 4_implementation.tex

@@
 implemented as a task graph execution engine, making it a good fit for
-Task Bench and one of the shorter implementations at only 88 lines.
+Task Bench. TensorFlow programs are written by creating nodes in the
+task graph, and then connecting these nodes with edges to implement
+the desired dataflow. Task graphs are composed in a high-level
+language such as Python, but then serialized and run by an efficient
+execution engine written in C++.
 
 Our Task Bench implementation in TensorFlow works by defining a custom
-operator, which represents the execution of one task. We then use
-TensorFlow APIs to instantiate the custom operator for each point in
-the task graph and connect the task dependencies with the TensorFlow
-API.
+operator, which we instantiate for each point in the task graph. We
+then define dataflow edges to correspond to the task dependencies.
 

### 2019-03-19 5_evaluation.tex
Add table with version numbers.
Update graphs for evaluation.

@@
 
-\input{e1_metg_compute}
+\input{e0_flags}
+
+We now present a comprehensive evaluation of Task Bench on up to 256
+Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
+machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
+processors (total of 32 physical cores per node), 128 GB RAM, and a
+Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
+implementations, and (where applicable) the system default MPI
+implementation, Cray MPICH 7.7.3. Additional versions and flags for the
+various systems are shown in Table~\ref{tab:flags}.
+
+\subsection{Peak Performance and Efficiency}
+
+\input{e1_flops}
+\input{e2_efficiency}
+
+\subsection{Overhead and Scalability}
+
+\input{e3_metg_compute}
+
+\subsection{Load Imbalance}
+
+\subsection{Communication Hiding}

### 2019-03-19 e3_metg_compute.tex
Update graphs for evaluation.

\begin{figure*}[t]

\subfloat[Stencil.\label{fig:metg-compute-stencil}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_stencil.pdf}
}
\subfloat[Nearest 5 neighbors.\label{fig:metg-compute-nearest}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_nearest.pdf}
}

\subfloat[Spread 5 neighbors.\label{fig:metg-compute-spread}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_spread.pdf}
}
\subfloat[Nearest 5 neighbors, 4 simultaneous task graphs.\label{fig:metg-compute-4x-nearest}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_ngraphs_4_nearest.pdf}
}

\caption{METG vs node count.\label{fig:metg-compute}}
\end{figure*}

### 2019-03-19 5_evaluation.tex
More text in evaluation.

@@
 
+In theory, any system should be able to hit peak efficiency as long as
+the kernels are well-written and of sufficiently large granularity. In
+practice, achieving peak performance is tricky, subject to a number of
+pitfalls, and unfortunately, frequently skipped in evaluations of
+runtime system overhead. However, verifying that peak performance is
+achieved is important to ensuring that such evaluation of overhead and
+efficiency are well-grounded and avoid common mistakes in
+configuration.
+
+Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
+kernel with varying problem sizes. This is the full version of
+Figure~\ref{fig:flops-mpi}. In the best case, we measure peak FLOPS of
+$1.26 \times 10^{12}$, which compares favorably with the officially
+reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
+of measuring efficiency we use our emperically determined number as
+the baseline for 100\% efficiency.
+
+The vast majority of systems achieve at or near peak FLOPS. A number
+of systems reserve a number of cores (usually 1 or 2) for internal
+runtime usage; these systems take a minor hit in peak FLOPS compared
+to systems which use all cores for application tasks. Some of the
+slower systems struggle to achieve peak FLOPs, though in most cases
+the curves suggest that performance would continue to improve if we
+were to run larger problem sizes. Unfortunately, the excessive
+computational cost of running such tests makes it prohibitively
+expensive to run larger problem sizes. For example, the Spark job in
+this configuration ran for over 5 hours.
+
+Figure~\ref{fig:efficiency} shows the efficiency (as a percentage of
+peak FLOPS achieved) plotted against task granularity. This plot is
+used to calculate METG, as described in Section~\ref{sec:metg}. The
+red, dashed line shows the point where 50\% efficiency is achieved.
+
+In most cases, task granularity has plateaued prior to this point,
+though a number of systems continue to improve as efficiency continues
+to drop. Accounting for this effect is one of the main arguments in
+favor of METG with a reasonable efficiency threshold compared to
+measuring task scheduling throughput in the presence of empty tasks
+(effectively the equivalent of METG(0\%)). Empty tasks permit runtime
+systems to make unrealistic tradeoffs which are not applicable to real
+application workloads where some degree of efficiency is necessarily a
+goal. Reporting numbers with empty tasks therefore gives unfair
+advantage to systems that make such unrealistic tradeoffs.
+
 \subsection{Overhead and Scalability}
@@
 
+\subsection{Number of Dependencies}
+
 \subsection{Load Imbalance}

### 2019-03-19 2_metg.tex
Update METG result for MPI.

@@
 level. At 50\% efficiency, MPI achieves an average task granularity of
-4.5 us, thus the METG(50\%) of MPI is 4.5 us for this configuration of
+4.6 us, thus the METG(50\%) of MPI is 4.6 us for this configuration of
 the application. The number 50\% is chosen arbitrarily, but is a

### 2019-03-19 5_evaluation.tex
More work on evaluation.

@@
 \input{e2_efficiency}
+
+% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
+\input{e3_metg_compute}
 
@@
 
-\input{e3_metg_compute}
+METG is a useful metric in part because it summarizes the basic task
+scheduling overhead of a system in a single number. This makes it
+possible to evaluate METG at different node counts to see how it is
+impacted by changes in communication topology and latency.
+
+Figure~\ref{fig:metg-compute} shows how METG(50\%) varies with node
+count for a variety of dependence patterns supported by Task Bench:
+
+\begin{itemize}
+\item Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each
+  task depends on 3 other tasks (including the same point in the
+  previous timestep).
+\item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
+  task depends on its 5 nearest neighbors.
+\item Figure~\ref{fig:metg-compute-spread} is a pattern where each
+  task depends on 5 distant neighbors, spread as widely as possible
+  throughout the machine.
+\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 simultaneous
+  copies of the nearest dependence pattern.
+\end{itemize}
+
+In general, the least complicated pattern (stencil) is most favorable
+to MPI, as it represents a truly bulk synchronous task graph with no
+opportunity for task parallelism. For the stencil pattern, the
+dominating factor is the basic overhead of executing a task, which is
+minimal for MPI as the Task Bench implementation is bulk synchronous
+and simply executes tasks one after another in alternation with
+communication phases. In contrast the asynchronous execution
+mechanisms of the other systems is pure overhead in this scenario.
+
+The gap between MPI and other systems shrinks as the complexity of the
+communication pattern grows, can even reverse as task parallelism is
+added in the form of multiple task graphs.
+
+% FIXME: update this when we have more data
+
+Most systems for high-performance computing are horizontally scalable,
+but this is not true of all the systems included in this
+evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and a
+flat line is ideal. Systems with non-scalable implementations shows
+lines that go up with node count. Most notably, Spark is typically
+intended for use in industrial data center applications with task
+granularities measured in tens of seconds. Spark uses a centralized
+controller which limits throughput, and this is visible in the figure
+as line for Spark immediately rises with node count. Keep in mind also
+that Spark is being evaluated here with a non-trivial dependence
+pattern which is relatively unrepresentative of Spark's normal use
+cases. Spark is more efficient with trivial parallelism, as described
+in Section~\ref{subsec:number-of-dependencies}.
+
+% FIXME: talk about starpu and parsec?
 
 \subsection{Number of Dependencies}
+\label{subsec:number-of-dependencies}
 
@@
 \subsection{Communication Hiding}
+
+\subsection{Comparison with Published Results}

### 2019-03-20 3_task_bench.tex
read sec 3

@@
 a configurable amount of result data in order to the amount of
-communication that must be performed. Thus Task Bench can be used to
+communications that must be performed. Thus Task Bench can be used to
 rapidly explore a large space of possible application scenarios in
@@
 such as required by sweeps and tree-based dependence
-patterns. Figure~\ref{fig:task-graphs} show a number of sample task
+patterns. Figure~\ref{fig:task-graphs} shows a number of sample task
 graphs that can be implemented with Task Bench.
@@
 code. Much of this is possible because the core aspects of generating
-the task graph and enumerating dependencies are encapsulated in a core
+task graphs and enumerating dependencies are encapsulated in a core
 library that is shared among all the Task Bench implementations. The

### 2019-03-20 e5_imbalance.tex
Efficiency plots.

\begin{figure*}[t]

\subfloat[Imbalance 0.5.\label{fig:efficiency-imbalance-0.5}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_0.5.pdf}
}
\subfloat[Imbalance 1.0.\label{fig:efficiency-imbalance-1.0}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_1.0.pdf}
}

\subfloat[Imbalance 1.5.\label{fig:efficiency-imbalance-1.5}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_1.5.pdf}
}
\subfloat[Imbalance 2.0.\label{fig:efficiency-imbalance-2.0}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_2.0.pdf}
}

\caption{Task granularity vs efficiency for varying load imbalance.\label{fig:metg-imbalance}}
\end{figure*}

### 2019-03-20 e6_communication.tex
Communication graphs.

\begin{figure*}[t]

\subfloat[16 bytes per task dependency.\label{fig:efficiency-communication-16}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_16.pdf}
}
\subfloat[256 bytes per task dependency.\label{fig:efficiency-communication-256}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_256.pdf}
}

\subfloat[4096 bytes per task dependency.\label{fig:efficiency-communication-16}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_4096.pdf}
}
\subfloat[65536 bytes per task dependency.\label{fig:efficiency-communication-256}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_65536.pdf}
}

\caption{Task granularity vs efficiency for varying communication.\label{fig:metg-communication}}
\end{figure*}

### 2019-03-21 1_introduction.tex
Update graphs for evaluation.
Rewrite introduction.

--- 2019-03-18 1_introduction.tex
@@
 computing have lead to the rise of a wide variety of new programming
-models and runtime systems. However, efforts to understand the
-performance behavior of these system have been limited. In many cases,
-programming models are evaluated with weak and/or strong scaling of a
-bespoke set of carefully hand-written applications, which make it
-difficult to understand how the results generalize to other
-applications, and impossible to compare across systems. Furthermore,
-these are not limit studies, and thus do nothing to elucide the
-boundaries of the performance envelops of the respective systems.
+models, languages and runtime systems. These systems often provide
+capabilities such as asynchronous execution of units of work, called
+\emph{tasks}, coordinated in some manner with data movement and
+synchronization. But while the features of these systems are widely
+written about in the literature, a large and often difficult to answer
+question is how different systems will perform for a given
+application.
 
-We introduce \emph{minimum effective task granularity} (METG), a
-metric of application performance specifically designed to explore the
-fundamental performance behavior of parallel and distributed runtime
-systems. METG can be used to identify the baseline overhead, limit of
-scalability, and ability to hide and overlap communication of such
-systems. A key insight of the METG metric is that large \emph{tasks},
-or units of work, can be used to hide almost any amount of
-overhead. Thus METG considers the \emph{minimum} task granularity that
-can be executed. This is a lower bound, so as long as the average task
-granularity of a given application exceeds this value, the runtime
-system should be able to execute it efficiently.
+One approach to answering this question has focused around the
+development of \emph{proxy-} or \emph{mini-apps}. So called because
+they embody the logic of a larger application, but in reduced form,
+mini-apps are popular because they can provide insight without
+incurring the costly expense of developing a full-blown application
+code. However, despite the name, mini-apps still requite significant
+investment to develop, especially to a level of quality need to reach
+a publishable state. In many cases, the effort to tune for performance
+exceeds the effort to develop the mini-app in the first place. This
+can be exacerbated if the system in question is undergoing concurrent
+development. As a result, implementations of mini-apps, though
+well-intentioned, often do not reach a level of maturity required to
+perform a broad comparison between systems, and as a result there are
+very few comparative studies involving multiple programming systems
+that reach publication.
 
-A variety of runtime systems are able to make use of parallelism
-internally for a variety of purposes, from analysis of tasks, to
-scheduling or the orchestration of data movement and
-synchronization. In order to mitigate unwanted influence due to
-unreasonable allocation of resources to such internal uses, METG
-concerns \emph{effective} task execution, where the overall
-computational efficiency meets some threshold (say, 50\%). If the
-METG(50\%) of an application is, say, 1~ms, then this indicates that
-the application can be executed with tasks as small as 1~ms while
-maintaining at least 50\% overall efficiency.
+A compounding issue is one of methodology: how does one predict how
+the performance observed for a mini-app will generalize to a full
+application, even one closely related? Part of the problem here is
+that weak and strong scaling, which are by far the most commonly
+reported metrics for performance of both mini- and full applications,
+do nothing to elucidate the performance envelop of a particular
+system. Weak scaling is notorious for being able to hide arbitrary
+amounts of overhead, as long as the problem size is sufficiently
+large. And strong scaling runs the risk of entangling unrelated
+effects: namely, the baseline overhead of a given system, with the
+growth of communication latencies, and potentially the saturation of
+global resources such as communication and I/O bandwidth, when using
+progressively larger portions of the machine.
 
-We present Task Bench, a benchmark designed specifically to study the
-performance behavior of parallel and distributed runtime systems at
-the limits of runtime overhead. Task Bench is a \emph{parameterized}
-benchmark, characterized by an \emph{iteration space} of tasks (with
-one task per point in the iteration space) and a \emph{dependence
-  relation} which determines how tasks depend on other tasks. By
-formulating the benchmark in this way, we make it possible to rapidly
-simulate a wide variety of communication patterns relevant in
-scientific computing, such as trivial parallelism, halo exchanges (for
-structured and unstructured mesh-based computations), sweeps (such as
-for the discrete ordinates method), FFTs, trees (for divide and
-conquer algorithms), and so on. Task Bench can be configured to
-execute a variety of kernels, such as compute- and memory-bound
-kernels, and load imbalanced variants of the above. Task Bench can
-also be used to stress communication-bound cases, by configuring the
-amount of data that is produced by each task.
+To address the first of these issues, we present Task Bench, a
+\emph{parameterized} benchmark designed for exploring the performance
+behaviors of programming systems under a variety of conditions. A
+critical property of Task Bench is that, while powerful, it is simple
+enough to be implemented in a wide variety of systems with reasonable
+effort, to a level of quality necessary for comparative study.
 
-In order to explore the impact of task parallelism on runtime
-performance, Task Bench can also execute multiple task graphs
-simultaneously. Runtime systems which support asynchronous execution
-can use this additional scheduling flexibility to hide communication
-overheads or to mitigate load imbalance. METG makes it possible to
-quantify the degree to which various systems are successful at doing
-so, and to measure if (and when) such factors are more, or less,
-important than baseline runtime overhead.
+Task Bench consists of a configurable \emph{task graph} with tasks for
+each point in an iteration space. Tasks are connected by dependencies
+as determined by a \emph{dependence relation}. This design permits a
+very concise description of a wide variety of patterns relevant to
+scientific computing: trivial parallelism, halo exchanges (such as
+seen in structured and unstructured mesh codes), sweeps (such as used
+in the discrete ordinates method of radiation simulation), FFTs, trees
+(for divide and conquer algorithms), and so on. The tasks at each
+point in the task graph can be configured to execute kernels with a
+variety of computational properties, including compute- and
+memory-bound operations of varying durations. The dependencies between
+tasks can be configured to carry payloads of varying size, permitting
+the design of communication-bound cases. Finally, multiple
+(potentially heterogeneous) task graphs can be executed concurrently
+to introduce task parallelism into the workload. Together, these
+design elements permit the exploration of a broad space of application
+performance behaviors.
 
-We provide a comprehensive evaluation of Task Bench implementations in
-13 different runtime systems, including Chapel~\cite{Chapel07},
-Charm++~\cite{Charmpp93}, MPI~\cite{MPI}, OmpSs~\cite{OmpSs11},
-OpenMP~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13},
-Realm~\cite{Realm14}, Regent~\cite{Regent15}, Spark~\cite{Spark10},
-StarPU~\cite{StarPU11}, Swift/T~\cite{Wozniak13},
-TensorFlow~\cite{TensorFlow15}, and X10~\cite{X1005}. Using the METG
-metric, we determine the baseline runtime overhead of each system
-along with overheads associated with increasing number and/or
-variations in the pattern of dependencies. We also explore each
-system's ability to scale, hide communication overhead, and mitigate
-load imbalance.
+To resolve the methodological issue, we introduce a new metric called
+\emph{minimum effective task granularity} (METG). Intuitively, METG is
+the smallest task granularity which can be executed while maintaining
+a given level of computational efficiency (e.g., as measured by the
+percentage of peak FLOPS achieved). METG captures the essence of what
+is important in a weak or strong scaling study: that is, the behavior
+at the limit of scalability. For weak scaling, METG corresponds
+directly to the smallest problem size that can be weak scaled with a
+given level of efficiency. For strong scaling, METG can be used to
+compute the point at which strong scaling can be expected to tail
+off. In either case, METG is a useful, intuitive measure as it
+encapsulates the task granularity at which systems can be relied on to
+execute efficiently. As a rule of thumb, as long as an application's
+average task granularity exceeds the METG, the application should
+execute efficiently.
+
+Together, METG and Task Bench help to mitigate issues which are common
+in limit studies on the task scheduling throughput of runtime
+systems. Such studies typically employ a metric called \emph{tasks per
+  second} (TPS), which is almost universally measured with trivial
+dependencies and empty tasks. While TPS is certainly an upper bound on
+task scheduling throughput, it is not a tight bound, as the cost of
+non-trivial dependencies can be significant and sometimes adversely
+impact scalability. This issue is easily fixed by running non-trivial
+configurations of Task Bench.
+
+But there is a deeper issue with the use of empty tasks. In all
+realistic scenarios, runtime systems must share resources with the
+application, either explicitly by e.g., allocating cores for runtime
+use, or implicitly by sharing a time slice on the same cores used for
+application tasks, and other resources such as memory bandwidth and
+processor cache capacity. The use of empty tasks makes the application
+share of these resources appear to be zero, which permits usage by the
+runtime system to be unconstrained. How then is anyone supposed to
+determine what resource usage to consider reasonable on the part of
+the runtime system? Only by requiring the application usage of
+resources to meet a certain minimum level, as in METG, can we ensure
+that runtime system resource usage is constrained to realistic levels.
+
+We conduct a comprehensive comparative study of Task Bench
+implementations in 13 different programming systems, including
+Chapel~\cite{Chapel07}, Charm++~\cite{Charmpp93}, MPI~\cite{MPI},
+OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40},
+PaRSEC~\cite{PARSEC13}, Realm~\cite{Realm14}, Regent~\cite{Regent15},
+Spark~\cite{Spark10}, StarPU~\cite{StarPU11},
+Swift/T~\cite{Wozniak13}, TensorFlow~\cite{TensorFlow15}, and
+X10~\cite{X1005}. Using the METG metric, we determine the baseline
+runtime overhead of each system along with overheads associated with
+increasing number and/or variations in the pattern of dependencies. We
+also explore each system's ability to scale, hide communication
+overhead, and mitigate load imbalance.
 
@@
 \item Section~\ref{sec:implementation} details the implementation of
-  Task Bench in 13 runtime systems.
+  Task Bench in 13 programming systems.
 \item Section~\ref{sec:evaluation} provides a comprehensive evaluation
-  of Task Bench performance on up to 256 nodes of the Cori
+  of Task Bench performance on up to 256 Haswell nodes of the Cori
   supercomputer.

### 2019-03-21 0_abstract.tex
Rewrite abstract.

@@
 \begin{abstract}
-% 150 words
-We introduce \emph{minimum effective task granularity} (METG), a new
-metric for quantifying the fundamental performance behavior of
-parallel and distributed runtime systems. METG is either directly
-related to, or correlates with a variety of measures of direct
-interest to application developers: namely the smallest problem size
-that can be efficiently weak scaled, or the limit of strong
-scalability. METG can be used to study the baseline overhead, scaling
-behavior, and communication hiding ability of runtime systems, among
-other properties.
+% 153 words
+We present Task Bench, a \emph{parameterized} benchmark designed to
+explore the performance envelops of parallel and distributed
+programming systems under a variety of application scenarios. Task
+Bench consists of a configurable \emph{task graph} with one task per
+point in an iteration space, and dependencies between tasks determined
+by a dependence relation. Tasks can execute a variety of kernels,
+including compute- and memory-bound, and task dependencies carry
+payloads of configurable size, enabling communication-bound use
+cases. Multiple task graphs can be executed to enable task
+parallelism.
 
-We present Task Bench, a novel \emph{parameterized} benchmark for
-exploring the behavior of parallel and distributed runtime
-systems. Task Bench executes a configurable \emph{task graph} with
-tasks for each point in an iteration space and dependencies determined
-by a \emph{dependence relation}. We use Task Bench to conduct a
-comprehensive evaluation of 13 runtime systems, and show that the
-baseline overhead of these systems varies by up to 5 orders of
-magnitude.
+We conduct a comprehensive comparative study with implementations of
+Task Bench in 13 programming systems on up to 256 Haswell nodes of the
+Cori supercomputer. We use a novel metric, \emph{minimum effective
+  task granularity} to study the baseline runtime overhead of each
+system and the impact of varying dependence patterns and number of
+dependencies. We also study each system's scalability, and ability to
+hide communication and mitigate load imbalance.
 \end{abstract}

### 2019-03-21 1_introduction.tex
Edit intro.

@@
 code. However, despite the name, mini-apps still requite significant
-investment to develop, especially to a level of quality need to reach
+investment to develop, especially to a level of quality needed to reach
 a publishable state. In many cases, the effort to tune for performance
@@
 perform a broad comparison between systems, and as a result there are
-very few comparative studies involving multiple programming systems
+few comparative studies involving multiple programming systems
 that reach publication.
@@
 given level of efficiency. For strong scaling, METG can be used to
-compute the point at which strong scaling can be expected to tail
-off. In either case, METG is a useful, intuitive measure as it
+compute the scale at which performance can be expected to dip below
+that level of efficiency. In either case, METG is a useful, intuitive
+measure as it
 encapsulates the task granularity at which systems can be relied on to
@@
 share of these resources appear to be zero, which permits usage by the
-runtime system to be unconstrained. How then is anyone supposed to
+runtime system to be unconstrained. How then is one supposed to
 determine what resource usage to consider reasonable on the part of

### 2019-03-21 t1_systems.tex
Add table listing systems.

\begin{table}[t]
\begin{tabular}{l | l | l | l | l}
System & Paradigm & Parallelism & Distrib. & Network \\
\hline
Chapel & multi-resolution & explicit & yes & uGNI \\
Charm++ & actor model & explicit & yes & uGNI \\
MPI & message passing & explicit & yes & uGNI \\
OmpSs & loop-, task-based & expl., impl. & no & \\
OpenMP & loop-, task-based & expl., impl. & no & \\
PaRSEC & task-based & imlicit & yes & MPI \\
Realm & task-based & explicit & yes & GASNet \\
Regent & task-based & implicit & yes & GASNet \\
Spark & functional & implicit & yes & sockets \\
StarPU & task-based & implicit & yes & MPI \\
Swift/T & dataflow & implicit & yes & MPI \\
TensorFlow & dataflow & explicit & yes\footnote{Our evaluation only considers TensorFlow on a single node.} & sockets \\
X10 & place-based & explicit & yes & MPI
\end{tabular}

\caption{Systems used in Task Bench implementations.\label{tab:systems}}
\end{table}

### 2019-03-21 0_abstract.tex

@@
 \begin{abstract}
-% 153 words
+% 149 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
@@
 point in an iteration space, and dependencies between tasks determined
-by a dependence relation. Tasks can execute a variety of kernels,
-including compute- and memory-bound, and task dependencies carry
-payloads of configurable size, enabling communication-bound use
-cases. Multiple task graphs can be executed to enable task
-parallelism.
+by a \emph{dependence relation}. Tasks execute a variety of kernels,
+including compute- and memory-bound, and dependencies carry payloads
+of configurable size, enabling communication-bound scenarios. Multiple
+task graphs can be executed to enable task parallelism.
 
-We conduct a comprehensive comparative study with implementations of
-Task Bench in 13 programming systems on up to 256 Haswell nodes of the
-Cori supercomputer. We use a novel metric, \emph{minimum effective
-  task granularity} to study the baseline runtime overhead of each
-system and the impact of varying dependence patterns and number of
+We conduct a comprehensive study with implementations of Task Bench in
+13 programming systems on up to 256 Haswell nodes of the Cori
+supercomputer. We use a novel metric, \emph{minimum effective task
+  granularity} to study the baseline runtime overhead of each system
+and the impact of varying dependence patterns and numbers of
 dependencies. We also study each system's scalability, and ability to

### 2019-03-21 3_task_bench.tex
Fix typo.

@@
 number of kernels (compute-bound, memory-bound, etc.) and can generate
-a configurable amount of result data in order to the amount of
-communications that must be performed. Thus Task Bench can be used to
+a configurable amount of result data in order to control the amount of
+communication that must be performed with each task dependency. Thus Task Bench can be used to
 rapidly explore a large space of possible application scenarios in

### 2019-03-21 4_implementation.tex
Add table listing systems.
More text in implementation.

@@
 
-We have implemented Task Bench in 13 parallel and distributed
-programming languages and runtime systems. Below we describe the
-systems, and any salient details of their Task Bench implementations.
+\input{t1_systems}
+
+We have implemented Task Bench in the 13 parallel and distributed
+programming systems listed in Table~\ref{tab:systems}. We describe the
+systems, and any salient details of their Task Bench implementations,
+below.
+
+One challenge in implementing Task Bench for such a wide variety of
+systems is that the capabilities of the systems vary considerably. For
+example, some systems are \emph{implicitly parallel}, and provide some
+form of parallelism discovery from sequential programs, whereas others
+are \emph{explicitly parallel} and require users to specify the
+parallelism in the program. As a rule of thumb we aim for the Task
+Bench implementations to be represenative of how actual applications
+would be written in the respective systems. Places where the Task
+Bench implementations may be unidiomatic are noted below.
+
+There are also aspects of some systems that we do not consider in our
+evaluation. For example, a number of systems provide some level of
+support for heterogeneous processors such as GPUs, but the support
+varies so widely (if it exists at all) that it is challenging to come
+up with a representative benchmark which allows these features to be
+compared in an apples-to-apples manner.
 
@@
 corresponding Task Bench implementations. Their insight was invaluable
-in ensuring that we provide the highest level of quality in the Task
+to ensuring that we provide the highest level of quality in the Task
 Bench implementations for each system.

### 2019-03-21 t1_systems.tex
Update footnotes.

@@
 \hline
-Chapel & multi-resolution & explicit & yes & uGNI \\
-Charm++ & actor model & explicit & yes & uGNI \\
-MPI & message passing & explicit & yes & uGNI \\
+Chapel & multi-resolution & explicit & yes & uGNI\footnote{Chapel uses GASNet to support non-Cray networks.} \\
+Charm++ & actor model & explicit & yes & uGNI\footnote{Charm++ provides additional backends for other networks.} \\
+MPI & message passing & explicit & yes & uGNI\footnote{Most MPI implementations provide additional backends for other networks.} \\
 OmpSs & loop-, task-based & expl., impl. & no & \\
@@
 TensorFlow & dataflow & explicit & yes\footnote{Our evaluation only considers TensorFlow on a single node.} & sockets \\
-X10 & place-based & explicit & yes & MPI
+X10 & place-based & explicit & yes & MPI\footnote{X10 also provides support for PAMI on supported networks.}
 \end{tabular}

### 2019-03-21 5_evaluation.tex
Comment.
Add radix figure.
Efficiency plots.
Communication graphs.
More text in evaluation.

@@
 to systems which use all cores for application tasks. Some of the
-slower systems struggle to achieve peak FLOPs, though in most cases
+slower systems struggle to achieve peak FLOPS, though in most cases
 the curves suggest that performance would continue to improve if we
@@
 
-\subsection{Overhead and Scalability}
+\subsection{Baseline Overhead}
 
-METG is a useful metric in part because it summarizes the basic task
-scheduling overhead of a system in a single number. This makes it
-possible to evaluate METG at different node counts to see how it is
-impacted by changes in communication topology and latency.
+One of the basic questions when considering different programming
+systems is how much overhead the system introduces. This can be a
+tricky question to answer directly because some systems introduce
+overhead \emph{inline} (i.e., by running system internal processes on
+the same cores as application tasks), while other systems introduce
+overhead \emph{out-of-line} (i.e., by dedicating one or more cores
+solely to runtime use). Some systems, like Charm++, Realm, and Regent,
+even support both configurations.
 
-Figure~\ref{fig:metg-compute} shows how METG(50\%) varies with node
-count for a variety of dependence patterns supported by Task Bench:
+To answer this question, we use the METG metric to determine the
+smallest task granularity which can be executed at a given level of
+efficiency, as a proxy for overhead. Figure~\ref{fig:metg-compute}
+shows how METG(50\%) varies with node count for a variety of
+dependence patterns supported by Task Bench:
 
@@
 \item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
-  task depends on its 5 nearest neighbors.
+  task depends 5 others, chosen to be as close as possible.
 \item Figure~\ref{fig:metg-compute-spread} is a pattern where each
-  task depends on 5 distant neighbors, spread as widely as possible
-  throughout the machine.
-\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 simultaneous
-  copies of the nearest dependence pattern.
+  task depends on 5 others, spread as widely as possible.
+\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical
+  copies of the nearest dependence pattern executing concurrently.
 \end{itemize}
+
+We observe in the results that the baseline overheads of different
+systems can vary by over 5 orders of magnitude. It is worth
+remembering when considering this metric that this is a \emph{minimum}
+effective task granularity. Therefore applications with an average
+task granularity of \emph{at least} this value can usually be expected
+to execute efficiently. Typical average task granularity will
+generally be determined by the application domain being
+considered. Most notably, for data analytics workloads being run in
+industrial data centers, the higher METG values observed for Spark and
+TensorFlow are sufficient. In contrast, for high-performance
+scientific simulations, task granularities in the millisecond range
+are useful, as such applications communicate (e.g., for halo
+exchanges) much more frequently.
+
+\subsection{Scalability}
+
+METG is a useful metric in part because it summarizes the basic task
+scheduling overhead of a system in a single number. This makes it
+possible to evaluate METG at different node counts (shown in
+Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
+communication topology and latency.
 
@@
 
+% FIXME: talk about overhead spanning 5 orders of magnitude
+
 \subsection{Number of Dependencies}
@@
 
+\input{e4_radix}
+
 \subsection{Load Imbalance}
+
+\input{e5_imbalance}
 
@@
 
+\input{e6_communication}
+
 \subsection{Comparison with Published Results}

### 2019-03-21 e3_metg_compute.tex
More text in evaluation.

@@
 
-\subfloat[Stencil.\label{fig:metg-compute-stencil}]{
+\subfloat[Stencil pattern.\label{fig:metg-compute-stencil}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_stencil.pdf}
 }
-\subfloat[Nearest 5 neighbors.\label{fig:metg-compute-nearest}]{
+\subfloat[Nearest pattern, 5 dependencies per task.\label{fig:metg-compute-nearest}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_nearest.pdf}
@@
 
-\subfloat[Spread 5 neighbors.\label{fig:metg-compute-spread}]{
+\subfloat[Spread pattern, 5 dependencies per task.\label{fig:metg-compute-spread}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_spread.pdf}
 }
-\subfloat[Nearest 5 neighbors, 4 simultaneous task graphs.\label{fig:metg-compute-4x-nearest}]{
+\subfloat[Nearest pattern, 5 dependencies per task, 4 task graphs.\label{fig:metg-compute-4x-nearest}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_ngraphs_4_nearest.pdf}
@@
 
-\caption{METG vs node count.\label{fig:metg-compute}}
+\caption{METG vs node count for different dependence patterns. Lower is better.\label{fig:metg-compute}}
 \end{figure*}

### 2019-03-21 5_evaluation.tex
More evaluation text.

@@
 
+The number of dependencies per task has a strong influence on the
+overhead incurred by each system, as shown in
+Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
+dependence pattern, when varying the number of dependencies per task
+from 0 to 9.
+
+% FIXME: make sure these numbers are up to date for final paper
+
+The ratio in METG between 0 and 3 dependencies per task ranges from
+$1.3\times$ to $59\times$ (median $3.0\times$). From 1 to 3
+dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
+$2.1\times$). The difference is most pronounced in systems which
+perform runtime work inline. For example, MPI achieves an METG of 390
+ns with 0 dependencies, but this rises to 4.6 us with 3 dependencies,
+a factor of $12\times$ increase. This is unsurprising as in the case
+of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
+issued at all, so MPI has nothing to do aside from executing kernels
+as quickly as possible.
+
+For this reason, we feel especially strongly that it is important to
+measure the performance behaviors of programming systems under
+realistic workloads, rather than trivial dependencies.
+
 \subsection{Load Imbalance}

### 2019-03-21 5_evaluation.tex
More text for evaluation section.

@@
 
+One advantage of systems with asynchronous execution capabilities is
+the ability to mitigate load imbalance, especially in the presence of
+task parallelism. To quantify this effect,
+Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
+efficiency curves under varying amounts of load imbalance. The
+imbalanced workloads are generated by randomly varying the task
+durations according by a factor determined by the formula $1+(X-0.5)I$
+where $X$ is a uniform random variable in the range $[0,1]$ and $I$ is
+the degree of imbalance in the range $[0,2]$. The random numbers are
+generated with a deterministic pseudo random number generator with a
+deterministic seed to ensure that the task durations under all systems
+are identical.
+
+The MPI implementation of Task Bench, being bulk synchronous,
+obviously suffers the most under load imbalance. The biggest
+difference is at large task granularities, where the imbalance has the
+effect of putting an upper bound on MPI efficiency. At smaller task
+granularities the effect shrinks and may even reverse as systems hit
+their fundamental limits due to overhead.
+
+The difference between the remaining systems is due primarily to the
+scheduling behavior of the various systems. For example, OpenMP
+executes tasks via a thread pool, resulting in a form of work
+stealing. This on-node work stealing, in combination with the task
+parallelism due to the presence of 4 task graphs, is sufficient to
+completely mitigate load imbalance. Systems which exploit task
+parallelism but not work stealing on a node achieve lower peak
+efficiency, but may achieve smaller task granularities as a result of
+not relying on a shared global queue for tasks.
+
 \subsection{Communication Hiding}

### 2019-03-21 5_evaluation.tex
More text for section 5.

@@
 
-\subsection{Comparison with Published Results}
+Also of interest is the ability of asynchronous systems to be able to
+hide communication latencies in the presence of task
+parallelism. Figure~\ref{fig:efficiency-communication} plots task
+granularity vs efficiency curves for varying amounts of
+communication. The degree of communication is configured by the number
+of bytes produced by each task (and therefore communicated with each
+task dependency).
+
+Asynchronous systems such as Charm++ are able to show two benefits in
+these plots. First, by overlapping communication with computation,
+such systems are able to execute smaller task granularities at higher
+levels of efficiency compared to the bulk synchronous MPI
+implementation. Second, the asynchrony also makes the curves smoother,
+as spikes in latency due to interference from other jobs can be
+mitigated, leading to more predictable performance, especially at
+smaller message sizes.

### 2019-03-21 6_related.tex
Some text for related work.

\section{Related Work}
\label{sec:related-work}

Programming systems for parallel and distributed computing are most
commonly evaluated using proxy- or mini-apps, or
microbenchmarks. Mini-apps are explicitly derived from larger
applications and therefore have the advantage of bearing some
relationship to the original. This advantage typically does not hold
for microbenchmarks.

A common failure mode for mini-apps is to bite off more complexity
than can be reasonably implemented to a level of quality sufficient
for publication. This can lead to publications for mini-apps omitting
details or performance results for systems beyond the reference
implementation. Follow-up papers, if any, often compare only against a
reference implementation and omit any other implementations that might
exist.

\begin{itemize}

\item
The initial paper on PENNANT~\cite{PENNANT} includes an MPI+OpenMP
implementation; follow-up papers present an implementation in
Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.

\item
The initial report on LULESH~\cite{LULESH12} includes no performance
results at all; follow-up papers present performance results for
CnC~\cite{LULESHCNC15} (a self-comparison) and X10~\cite{LULESHX1015}
(comparison against reference only).

\item
The mini-app CoMD received a software release but no initial formal
publication. One follow-up paper describes a Chapel
implementation~\cite{CoMDChapel16} (comparison against reference
only). Additional follow-up papers consider aspects of the reference
implementation only~\cite{CoMDLoadImbalance17,
  CoMDThreadedModels14}.

\item
A report on the Mantevo project~\cite{Mantevo09} describes a number of
mini-apps, but only includes self-comparisons based on reference
implementations.

\item
A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
four implementations of the mini-app, but only includes performance
results for three, and of those three only two can be compared in an
apples-to-apples manner as the last implementation uses a structured
rather than an unstructured mesh. Follow-up papers describe an
additional implementation in Regent~\cite{Regent15,
  ControlReplication17, LegionTracing18} (comparison against reference
only).

\end{itemize}

Microbenchmarks are often easier to implement, but may not be
representative of real-world applications. From the PRK
suite~\cite{PRK14}, the PRK Stencil code has been reasonably widely
implemented, and a follow-up paper~\cite{PRKRuntimes16} compares MPI
variants, SHMEM, UPC, Charm++, and Grappa. However, PRK Stencil is a
simple code involving 2 kernels (which combined fit in under 50 lines
of C++) and a halo exchange on a structured grid; it is unclear to
what extent this is representative of real-world applications. Other
codes in PRK are even less representative.

The NAS benchmark suite~\cite{NAS91, NAS95} are widely studied with
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. However the benchmark consider mostly small
kernels for dense array computations, and do not generalize to larger
applications.

### 2019-03-21 6_related.tex
More related work.

@@
 applications.
+
+System-specific benchmarks have been used to quantify specific aspects
+of system performance, such as MPI communication or collective
+latency~\cite{MPPTest99, MPIBench01}. These measurements typically do
+not generalize beyond the immediate system they attempt to quantify,
+nor in an obvious way to real-world applications.
+
+Limit studies of task scheduling throughput in various runtime systems
+often make additional assumptions. A popular assumption is the use of
+trivially parallel tasks~\cite{Canary16, Armstrong14}, which as shown
+in Section~\ref{subsec:number-of-dependencies} underestimates (often
+substantially) the cost of scheduling a task. Worse, trivial
+parallelism overstates the scalability of the runtime system; our
+results for Swift/T in Section~\ref{subsec:scalability} differ rather
+substantial from those published in~\cite{Armstrong14} for this
+reason.

### 2019-03-21 7_conclusion.tex
Add conclusion.

\section{Conclusion}
\label{sec:conclusion}

We presented Task Bench, a parameterized benchmark powerful enough to
enable the exploration of a broad space of application scenarios, but
also simple enough to permit implementation in a wide array of systems
to a level of quality sufficient for comparative study. Our study of
13 programming systems is novel in its breadth, as well as the depth
of insight it delivers about different systems. We discover, among
other things, that the baseline overhead of systems varies by up to 5
orders of magnitude.

Helping us in this endeavor is a new metric, minimum effective task
granularity (METG), which helps to sidestep a number of potential
issues in the evaluation of task scheduling performance, and helps to
separate such factors from changes in communication topology and
latency at increasing scales of the machine.

### 2019-03-22 4_implementation.tex
Run spell check.

@@
 parallelism in the program. As a rule of thumb we aim for the Task
-Bench implementations to be represenative of how actual applications
+Bench implementations to be representative of how actual applications
 would be written in the respective systems. Places where the Task
@@
 used to implement dependencies, and a task is executed as soon as its
-dependecies are all available. Collectives are used to synchronize the
+dependencies are all available. Collectives are used to synchronize the
 completion of the task graph execution.
@@
   graphs} (PTG) and \emph{sequential task flow} (STF). In the PTG
-model, the program consists of an analytical function which describes
+model, the program consists of an analytic function which describes
 the expansion of the task graph. PTG is therefore a more limited
@@
 of being pure functions are trivially known to be
-parallizable. Functions in Spark operate on \emph{resilient
+parallelizable. Functions in Spark operate on \emph{resilient
   distributed datasets} (RDDs), which are globally-visible,
@@
 which represent distributed memories, a PGAS model where references to
-remote objects can be held (but can only be deferenced on the local
+remote objects can be held (but can only be dereferenced on the local
 place), asynchronous tasks, and a place-changing construct to move the

### 2019-03-22 1_introduction.tex
Edits.

@@
 models, languages and runtime systems. These systems often provide
-capabilities such as asynchronous execution of units of work, called
+capabilities such as the asynchronous execution of units of work, called
 \emph{tasks}, coordinated in some manner with data movement and
@@
 written about in the literature, a large and often difficult to answer
-question is how different systems will perform for a given
+question is how different systems will perform for any given
 application.
@@
   second} (TPS), which is almost universally measured with trivial
-dependencies and empty tasks. While TPS is certainly an upper bound on
+dependencies. While TPS is certainly an upper bound on
 task scheduling throughput, it is not a tight bound, as the cost of
 non-trivial dependencies can be significant and sometimes adversely
-impact scalability. This issue is easily fixed by running non-trivial
+impacts scalability. This issue is easily fixed by running non-trivial
 configurations of Task Bench.
 
-But there is a deeper issue with the use of empty tasks. In all
-realistic scenarios, runtime systems must share resources with the
-application, either explicitly by e.g., allocating cores for runtime
-use, or implicitly by sharing a time slice on the same cores used for
-application tasks, and other resources such as memory bandwidth and
-processor cache capacity. The use of empty tasks makes the application
-share of these resources appear to be zero, which permits usage by the
-runtime system to be unconstrained. How then is one supposed to
-determine what resource usage to consider reasonable on the part of
-the runtime system? Only by requiring the application usage of
-resources to meet a certain minimum level, as in METG, can we ensure
-that runtime system resource usage is constrained to realistic levels.
+But there is a deeper issue with the lack of an efficiency constraint
+in measurements of TPS. TPS may be measured with empty
+tasks~\cite{LegionTracing18}, or with tasks of some
+duration~\cite{Canary16, Armstrong14}. If empty tasks are used to
+measure TPS, this leads to an upper bound on the task scheduling
+throughput, but fails to represent any realistic application, as no
+useful work is being performed. If non-empty tasks are used, \emph{but
+  the efficiency of the overall application is not reported}, then TPS
+ceases to be a measurement of the limit of task scheduling
+throughput. Large tasks may be used to hide any amount of overhead,
+while tasks that are too small may result in a drop in total
+application throughput even as TPS increases. Only by constraining
+efficiency, as in METG, can we ensure a limit study that represents
+reasonable throughput on useful work by the application.
 

### 2019-03-22 4_implementation.tex
Edits.

@@
 
-Internally, Chapel implements custom support for Cray networks, and
-uses the GASNet~\cite{GASNET07} compatibility layer on other networks.
-
 \subsection{Charm++}
@@
 
-Internally, Charm++ implements support for a variety of networks
-directly on top of the low-level, vendor-provided APIs. For Cori, we
-use Charm++'s native support for uGNI.
-
 \subsection{MPI}
@@
 
-In our evaluation, we use the default MPI implementation installed on
-the machine, which is Cray MPI on the Cori supercomputer. This
-implementation is vetted by the machine operators and vendor and is
-the highest performing available implementation for the hardware.
-
 For completeness, we also explored the use of additional MPI calls:
@@
 
-OmpSs~\cite{OmpSs11} is a programming model for loop-based parallelism
+OmpSs~\cite{OmpSs11} is a programming model for loop- and task-based parallelism
 that is source-compatible with OpenMP. Our Task Bench implementation
@@
 OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
-parallelism on shared-memory systems. Our Task Bench implementation
-for OpenMP is very similar to OmpSs and uses OpenMP 4.0 tasks with
+parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
+for OpenMP is very similar to OmpSs and uses tasks with
 task dependencies.
@@
 and thus report KMP numbers in our evaluation.
+
+% FIXME: keep this in sync with experiments
 
@@
 
-Internally, PaRSEC uses MPI for network communication.
-
 \subsection{Realm}
@@
 Realm~\cite{Realm14} is an explicitly-parallel task-based programming
-model that is used internally by Legion~\cite{Legion12} and
-Regent~\cite{Regent15}, which are implicitly parallel layers written
-on top of Realm. Thus the Realm implementation of Task Bench can be
-seen as a limit study of what is possible with Legion or Regent.
+model that used internally by Legion~\cite{Legion12} and
+Regent~\cite{Regent15}. Thus the Realm implementation of Task Bench can be
+seen as a limit study of what can be achieved with Legion or Regent.
 
@@
 
-Internally, Realm uses GASNet~\cite{GASNET07} for communication.
-
 \subsection{Regent}
@@
 provides a critical optimization, \emph{control replication}, that
-enables superior horizontal scalability of
-applications~\cite{ControlReplication17}.
+enables superior horizontal scalability~\cite{ControlReplication17}.
 
@@
 program into \emph{shards} of execution which can scalably execute on
-large numbers of nodes. Our Regent implementation also makes use of a
-tracing optimization~\cite{LegionTracing18} which substantially
-reduces the cost of dynamic analysis of tasks in Regent.
+large numbers of nodes.
 
@@
   distributed datasets} (RDDs), which are globally-visible,
-dynamically single-assignment data structures that cached in memory
+dynamically single-assignment data structures that are cached in memory
 and written to disk for durability. One of Spark's primary
@@
 \lstinline[language=Scala]{groupByKey} operations to generate task
-dependencies, and then \lstinline[language=Scala]{map} to execute a
+dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
 series of tasks. We use an explicit hash partitioner to ensure that
@@
 calls from Java to C, and cross-checked our results with known cases
-that hit optimal task throughput in Spark, among other things. Our
-results are discussed in detail in Section~\ref{sec:evaluation}.
+that hit optimal task throughput in Spark, among other things.
 
@@
 elision of tasks on unrelated nodes to provide horizontal scalability.
-
-Internally, StarPU uses MPI for communication.
 
@@
 semantics to capture dependencies on other tasks.
-
-Internally, Swift/T uses MPI for communication.
 
@@
 between places, along with place-changing and atomic integers for
-synchronization. We use the C++ backend of X10, which uses MPI
-internally for portability to a variety of networks, and which
-conveniently integrates well with the Task Bench API.
+synchronization. We use the native backend of X10, which compiles to
+C++.

### 2019-03-22 4_implementation.tex
read impl

--- 2019-03-21 4_implementation.tex
@@
 are \emph{explicitly parallel} and require users to specify the
-parallelism in the program. As a rule of thumb we aim for the Task
-Bench implementations to be represenative of how actual applications
-would be written in the respective systems. Places where the Task
-Bench implementations may be unidiomatic are noted below.
+parallelism in the program. As a rule of thumb, we aim for the Task
+Bench implementations to be representative of how actual applications
+would be written in the respective systems. Places, where the Task
+Bench implementations may be unidiomatic, are noted below.
 
@@
 corresponding Task Bench implementations. Their insight was invaluable
-to ensuring that we provide the highest level of quality in the Task
+to ensure that we provide the highest level of quality in the Task
 Bench implementations for each system.
@@
 supports two distinct programming models: \emph{parameterized task
-  graphs} (PTG) and \emph{sequential task flow} (STF). In the PTG
-model, the program consists of an analytical function which describes
-the expansion of the task graph. PTG is therefore a more limited
+  graphs} (PTG) and \emph{dynamic task discovery} (DTD)~\cite{PARSEC_DTD}. 
+PTG is a data flow model, where programmers write applications 
+in a concise description of all data flows by describing tasks 
+and how data flows between tasks. This creates a compressed algebraic
+representation of the task graph, which is then transformed into C 
+code by PaRSEC source-to-source compiler. PTG is therefore a more limited
 programming model, but can potentially provide better performance
 because the graph description is amenable to static analysis and can
-be pruned efficiently for horizontally scalable execution.
+be pruned efficiently for better horizontally scalable execution.
 
-In the STF model, tasks are enumerated in program order (by
+In the DTD model, tasks are enumerated in program order (by
 dynamically executing the program), and dependencies between tasks are
-identified automatically based on the arguments to tasks. A task
-depends on another task if it reads data written by the other task,
+identified automatically based on the input and output data of tasks. 
+A task depends on another task if it reads data written by the other task,
 and the data is copied automatically if the two tasks are executed on
-different nodes. While the basic STF model is implicitly parallel
+different nodes. While the basic DTD model is implicitly parallel
 because the system is responsible for the discovery of parallelism, in
@@
 
-Data in PaRSEC is stored in \emph{regions}, or dense arrays of
-contiguous memory. Regions may be passed to tasks on other nodes, but
+The PaRSEC build-in 2D block cyclic matrix is used to store data. 
+A matrix is divided into M by N tiles where N equals to the width of
+task graphs and M varies from 2 to the height of task graphs depends
+on the optimization level of alleviating write-after-read depdenencies. 
+Tiles are distributed across nodes in pre-defined patterns, 
+and can be passed to tasks on other nodes, but
 can only be written on the node where the data is originally
 allocated. In our Task Bench implementation, each task writes to a
-region as output, and these regions are passed as inputs to the
+tile as output, and these tiles are passed as inputs to the
 subsequent set of tasks. As in the OmpSs and OpenMP implementations,
@@
 
-Spark's programming model is based of \emph{functional parallelism}, a
+Spark's programming model is based on \emph{functional parallelism}, a
 generalization of the MapReduce programming
@@
 
-StarPU~\cite{StarPU11} is a task-based system that supports an STF
+StarPU~\cite{StarPU11} is a task-based system that supports an DTD
 programming model similar to PaRSEC. Our Task Bench implementation in
@@
 elision of tasks on unrelated nodes to provide horizontal scalability.
+A simplified version of 2D block cyclic matrix from the Chameleon 
+project~\cite{Chameleon} is used to describe data of tasks in 
+our Task Bench implementation.
 

### 2019-03-22 2_metg.tex

@@
 application, thus it makes sense to discuss tasks even in systems
-without an explicit notion of tasking, such as in MPI codes are written in
+without an explicit notion of tasking, such as in MPI codes that are written in
 a bulk synchronous style.
@@
 efficiency starts near 100\%, and that initially task granularity
-drops quickly with a minimal change in efficiency. However as task
+drops quickly with minimal change in efficiency. However as task
 granularity shrinks, efficiency begins to drop as well, leading

### 2019-03-22 3_task_bench.tex

@@
 Task Bench is constrained to be 2-dimensional, with time flowing along
-the vertical axis and the width of the task graph along the
+the vertical axis and parallel tasks along the
 horizontal. Tasks can depend only on tasks from the immediately
@@
 that the kernels are identical in all cases, eliminating a potential
-source of performance disparity which is common in mini-apps
+source of performance disparity which can be a pitfall for mini-apps
 implementations. Finally, the Task Bench core library manages the
@@
 assurance that all Task Bench implementations are correct, and
-safeguards against corner cases that may not have been adequately
-tested during development. An evaluation of the performance impact of
+safeguards against corner cases that may only be exercised at
+scale. An evaluation of the performance impact of
 this validation showed it to be less than 3\% at the smallest task

### 2019-03-22 4_implementation.tex

@@
 Bench implementations to be representative of how actual applications
-would be written in the respective systems. Places, where the Task
-Bench implementations may be unidiomatic, are noted below.
+would be written in the respective systems. Places where the Task
+Bench implementations may be unidiomatic are noted below.
 
@@
 our Task Bench implementation.
-
-Internally, StarPU uses MPI for communication.
 

### 2019-03-22 5_evaluation.tex
More related work.
Run spell check.

@@
 runtime system overhead. However, verifying that peak performance is
-achieved is important to ensuring that such evaluation of overhead and
-efficiency are well-grounded and avoid common mistakes in
+achieved is important to ensuring that such evaluations of overhead and
+efficiency are well-grounded and avoid common pitfalls in
 configuration.
@@
 reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
-of measuring efficiency we use our emperically determined number as
+of measuring efficiency we use our empirically determined number as
 the baseline for 100\% efficiency.
 
-The vast majority of systems achieve at or near peak FLOPS. A number
-of systems reserve a number of cores (usually 1 or 2) for internal
+The vast majority of systems achieve at or near peak FLOPS. Some
+systems reserve a number of cores (usually 1 or 2) for internal
 runtime usage; these systems take a minor hit in peak FLOPS compared
-to systems which use all cores for application tasks. Some of the
+to systems which share all cores between the application and runtime. Some of the
 slower systems struggle to achieve peak FLOPS, though in most cases
@@
 expensive to run larger problem sizes. For example, the Spark job in
-this configuration ran for over 5 hours.
+this configuration ran for over 6 hours.
 
-Figure~\ref{fig:efficiency} shows the efficiency (as a percentage of
+Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
 peak FLOPS achieved) plotted against task granularity. This plot is
@@
 
-In most cases, task granularity has plateaued prior to this point,
+In most cases, task granularity has plateaued prior to the 50\% efficiency point,
 though a number of systems continue to improve as efficiency continues
@@
 \subsection{Scalability}
+\label{subsec:scalability}
 

### 2019-03-22 6_related.tex

--- 2019-03-21 6_related.tex
@@
 for publication. This can lead to publications for mini-apps omitting
-details or performance results for systems beyond the reference
+implementations or performance results for systems beyond the reference
 implementation. Follow-up papers, if any, often compare only against a
@@
 \item
-The initial paper on PENNANT~\cite{PENNANT} includes an MPI+OpenMP
+The initial paper on PENNANT~\cite{PENNANT} includes an MPI/OpenMP/MPI+OpenMP
 implementation; follow-up papers present an implementation in
@@
 MPI+OpenMP~\cite{NASMPIOpenMP00}, and
-Charm++~\cite{NASCharm96}. However the benchmark consider mostly small
+Charm++~\cite{NASCharm96}. However the benchmarks consist mostly of small
 kernels for dense array computations, and do not generalize to larger
 applications.
+
+System-specific benchmarks have been used to quantify specific aspects
+of system performance, such as MPI communication or collective
+latency~\cite{MPPTest99, MPIBench01}. These measurements typically do
+not generalize beyond the immediate system they attempt to quantify,
+nor in an obvious way to real-world applications.
+
+Limit studies of task scheduling throughput in various runtime systems
+often make additional assumptions. A popular assumption is the use of
+trivially parallel tasks~\cite{Canary16, Armstrong14}, which as shown
+in Section~\ref{subsec:number-of-dependencies} underestimates (often
+substantially) the cost of scheduling a task, and can also impact scalability.

### 2019-03-22 f5_task_graphs.tex
Add figures for task graphs.

\begin{figure}[t]

\subfloat[Trivial.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/trivial.pdf}
}
\hfill
\subfloat[Stencil.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/stencil.pdf}
}
\hfill
\subfloat[FFT.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/fft.pdf}
}

\hspace*{\fill}
\subfloat[Sweep.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/dom.pdf}
}
\hfill
\subfloat[Tree.]{
\includegraphics[width=0.45\columnwidth]{figs/sample-task-graphs/tree.pdf}
}
\hspace*{\fill}

\caption{Sample task graphs.\label{fig:task-graphs}}
\end{figure}

### 2019-03-22 4_implementation.tex
Edits.

@@
 are \emph{explicitly parallel} and require users to specify the
-parallelism in the program. As a rule of thumb, we aim for the Task
+parallelism in the program. As a rule of thumb we aim for the Task
 Bench implementations to be representative of how actual applications
@@
 corresponding Task Bench implementations. Their insight was invaluable
-to ensure that we provide the highest level of quality in the Task
+in helping to ensure that we provide the highest level of quality in the Task
 Bench implementations for each system.
@@
 
-PaRSEC~\cite{PARSEC13} is a task-based programming system which
-supports two distinct programming models: \emph{parameterized task
-  graphs} (PTG) and \emph{dynamic task discovery} (DTD)~\cite{PARSEC_DTD}. 
-PTG is a data flow model, where programmers write applications 
-in a concise description of all data flows by describing tasks 
-and how data flows between tasks. This creates a compressed algebraic
-representation of the task graph, which is then transformed into C 
-code by PaRSEC source-to-source compiler. PTG is therefore a more limited
-programming model, but can potentially provide better performance
-because the graph description is amenable to static analysis and can
-be pruned efficiently for better horizontally scalable execution.
+PaRSEC is a task-based programming system which supports two distinct
+programming models: \emph{parameterized task graphs}
+(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
+(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which programmers
+write a concise, algebraic description of the tasks and dataflows in
+the program. This compressed representation is expanded into a full
+task graph by a source-to-source compiler. PTG is therefore a more
+limited programming model, but can potentially provide better
+performance because the graph description is amenable to static
+analysis and can be pruned efficiently for better horizontal
+scalability.
 
@@
 
-The PaRSEC build-in 2D block cyclic matrix is used to store data. 
-A matrix is divided into M by N tiles where N equals to the width of
-task graphs and M varies from 2 to the height of task graphs depends
-on the optimization level of alleviating write-after-read depdenencies. 
+Data in PaRSEC is stored in a built-in matrix data structure with a 2D block cyclic distribution. 
+The Task Bench implementation divides this matrix into $M$ by $N$ tiles where $N$ equals to the width of
+the task graph and $M$ varies from 2 to the height of the task graph depending
+on optimizations performed to alleviate write-after-read depdenencies. 
 Tiles are distributed across nodes in pre-defined patterns, 
@@
 tile as output, and these tiles are passed as inputs to the
-subsequent set of tasks. As in the OmpSs and OpenMP implementations,
+subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
 the task launch code uses a switch statement as the PaRSEC API assumes
 tasks have a fixed number of dependencies.
-
-% FIXME: What does PaRSEC call regions?
 
@@
 between tasks explicitly defined by \emph{events} passed from one task
-to another. Realm's data model supports regions which live on a single
+to another. Realm's data model supports \emph{regions} of memory which live on a single
 node. Data must be explicitly copied, and the copies synchronized with
@@
 elision of tasks on unrelated nodes to provide horizontal scalability.
-A simplified version of 2D block cyclic matrix from the Chameleon 
-project~\cite{Chameleon} is used to describe data of tasks in 
+A simplified version of the 2D block cyclic matrix from the Chameleon 
+project~\cite{Chameleon} is used to describe the data produced by tasks in 
 our Task Bench implementation.

### 2019-03-26 0_abstract.tex
Apply Alex's comments to abstract and introduction.

@@
 \begin{abstract}
-% 149 words
+% 129 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
-explore the performance envelops of parallel and distributed
+explore the performance envelopes of parallel and distributed
 programming systems under a variety of application scenarios. Task
-Bench consists of a configurable \emph{task graph} with one task per
-point in an iteration space, and dependencies between tasks determined
-by a \emph{dependence relation}. Tasks execute a variety of kernels,
-including compute- and memory-bound, and dependencies carry payloads
-of configurable size, enabling communication-bound scenarios. Multiple
-task graphs can be executed to enable task parallelism.
+Bench seeks to provide the ease of implementation and portability of
+microbenchmarks, which makes comparisons between many systems
+possible, with the ability to easily model a wide range of application
+scenarios, giving it coverage approximating large benchmarks.
 
@@
   granularity} to study the baseline runtime overhead of each system
-and the impact of varying dependence patterns and numbers of
-dependencies. We also study each system's scalability, and ability to
-hide communication and mitigate load imbalance.
+and the impact of scenarious with varying dependence patterns. We also
+study each system's scalability, and ability to hide communication and
+mitigate load imbalance.
 \end{abstract}

### 2019-03-26 1_introduction.tex
Apply Alex's comments to abstract and introduction.

--- 2019-03-21 1_introduction.tex
@@
 
-Challenges in heterogeneous and extreme-scale parallel and distributed
-computing have lead to the rise of a wide variety of new programming
-models, languages and runtime systems. These systems often provide
-capabilities such as asynchronous execution of units of work, called
-\emph{tasks}, coordinated in some manner with data movement and
-synchronization. But while the features of these systems are widely
-written about in the literature, a large and often difficult to answer
-question is how different systems will perform for a given
-application.
+The challenge of parallel and distributed computation has lead to a
+wide variety of proposals for programming models, languages, and
+runtime systems. Some, like MPI~\cite{MPI}, have achieved widespread
+usage while many more are in the research and development stage. Our
+goal in this paper is to develop a meaningful and useful framework for
+comparing the performance of parallel and distributed programming
+systems.
 
-One approach to answering this question has focused around the
-development of \emph{proxy-} or \emph{mini-apps}. So called because
-they embody the logic of a larger application, but in reduced form,
-mini-apps are popular because they can provide insight without
-incurring the costly expense of developing a full-blown application
-code. However, despite the name, mini-apps still requite significant
-investment to develop, especially to a level of quality need to reach
-a publishable state. In many cases, the effort to tune for performance
-exceeds the effort to develop the mini-app in the first place. This
-can be exacerbated if the system in question is undergoing concurrent
-development. As a result, implementations of mini-apps, though
-well-intentioned, often do not reach a level of maturity required to
-perform a broad comparison between systems, and as a result there are
-very few comparative studies involving multiple programming systems
-that reach publication.
+% note: removed definition of task
 
-A compounding issue is one of methodology: how does one predict how
-the performance observed for a mini-app will generalize to a full
+One approach to comparing the performance of different systems has
+focused on the development of \emph{proxy-} or
+\emph{mini-apps}. Because they embody the logic of a larger
+application, but in reduced form, mini-apps can provide insight
+without incurring the costly expense of developing a full-blown
+application code. However, despite the name, our experience is that
+mini-apps still requite significant investment to develop, especially
+to a level of quality needed for useful benchmarking. In many cases,
+the effort to tune for performance exceeds the effort to develop the
+mini-app in the first place. As a result, implementations of mini-apps
+often do not reach a level of maturity required to compare
+systems. Indeed, there are few such published studies, and none to our
+knowledge that involve more than a handful of systems.
+
+There is another important issue: How does one predict how
+the performance of a mini-app generalizes to a full
 application, even one closely related? Part of the problem here is
@@
 reported metrics for performance of both mini- and full applications,
-do nothing to elucidate the performance envelop of a particular
-system. Weak scaling is notorious for being able to hide arbitrary
-amounts of overhead, as long as the problem size is sufficiently
-large. And strong scaling runs the risk of entangling unrelated
-effects: namely, the baseline overhead of a given system, with the
-growth of communication latencies, and potentially the saturation of
-global resources such as communication and I/O bandwidth, when using
-progressively larger portions of the machine.
+do not directly characterize the performance of the underlying
+programming system. Weak scaling can hide arbitrary amounts of runtime
+system overhead with sufficiently large problem sizes. And strong
+scaling does not separate the overhead of the underlying runtime
+system from the growth in communication latencies, and potential
+saturation of resources such as communication and I/O bandwidth, when
+using progressively larger portions of the machine.
 
@@
 \emph{parameterized} benchmark designed for exploring the performance
-behaviors of programming systems under a variety of conditions. A
-critical property of Task Bench is that, while powerful, it is simple
-enough to be implemented in a wide variety of systems with reasonable
-effort, to a level of quality necessary for comparative study.
+behaviors of parallel and distributed programming systems under a
+variety of conditions. A critical property of Task Bench is that,
+while able to express a wide variety of typical parallel programming
+patterns, it is simple enough to be implemented with reasonable effort
+to a level of quality sufficient for comparative studies.
 
-Task Bench consists of a configurable \emph{task graph} with tasks for
+A key insight of Task Bench is that many applications can be modeled
+as a set of \emph{tasks}, or coarse-grain units of work, with
+dependencies between tasks representing the communication required for
+distributed execution. Task Bench consists of a configurable \emph{task graph} with tasks for
 each point in an iteration space. Tasks are connected by dependencies
@@
 very concise description of a wide variety of patterns relevant to
-scientific computing: trivial parallelism, halo exchanges (such as
+scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
 seen in structured and unstructured mesh codes), sweeps (such as used
@@
 to introduce task parallelism into the workload. Together, these
-design elements permit the exploration of a broad space of application
-performance behaviors.
+design elements permit the exploration of a large space of application
+behaviors.
 
-To resolve the methodological issue, we introduce a new metric called
-\emph{minimum effective task granularity} (METG). Intuitively, METG is
-the smallest task granularity which can be executed while maintaining
-a given level of computational efficiency (e.g., as measured by the
-percentage of peak FLOPS achieved). METG captures the essence of what
-is important in a weak or strong scaling study: that is, the behavior
-at the limit of scalability. For weak scaling, METG corresponds
-directly to the smallest problem size that can be weak scaled with a
-given level of efficiency. For strong scaling, METG can be used to
-compute the point at which strong scaling can be expected to tail
-off. In either case, METG is a useful, intuitive measure as it
-encapsulates the task granularity at which systems can be relied on to
-execute efficiently. As a rule of thumb, as long as an application's
-average task granularity exceeds the METG, the application should
-execute efficiently.
+To isolate the contribution of the programming model's runtime system
+to application performance, we introduce a new metric called
+\emph{minimum effective task granularity} (METG). Intuitively,
+METG(50\%) is the smallest application task granularity which can be
+executed while maintaining at least 50\% computational efficiency,
+meaning that the application achieves at least 50\% of peak
+performance. METG captures the essence of what is important in a weak
+or strong scaling study, the behavior at the limit of scalability. For
+weak scaling, METG(50\%) corresponds directly to the smallest problem
+size that can be weak scaled with 50\% efficiency. For strong scaling,
+METG(50\%) can be used to compute the scale at which performance can
+be expected to dip below 50\% efficiency. A good rule of thumb is that
+as long as an application's \emph{average} task granularity exceeds
+METG(50\%), the application should achieve at least 50\% efficiently.
 
-Together, METG and Task Bench help to mitigate issues which are common
-in limit studies on the task scheduling throughput of runtime
-systems. Such studies typically employ a metric called \emph{tasks per
-  second} (TPS), which is almost universally measured with trivial
-dependencies and empty tasks. While TPS is certainly an upper bound on
-task scheduling throughput, it is not a tight bound, as the cost of
-non-trivial dependencies can be significant and sometimes adversely
-impact scalability. This issue is easily fixed by running non-trivial
+Task Bench and METG mitigate issues common in limit studies of runtime
+systems for parallel and distributed programming. Such studies often
+employ the metric \emph{tasks per second} (TPS), which is almost
+universally measured with trivial (i.e., no) dependencies. While
+phrased in terms of tasks, TPS can be measured for any system as long
+as the application in question has identifiable units of work that run
+to completion without interruption. TPS is an upper bound on
+runtime-limited application throughput, but not a tight bound, as the
+cost of non-trivial dependencies can be significant and may impact
+scalability. This issue is easily fixed by running non-trivial
 configurations of Task Bench.
 
-But there is a deeper issue with the use of empty tasks. In all
-realistic scenarios, runtime systems must share resources with the
-application, either explicitly by e.g., allocating cores for runtime
-use, or implicitly by sharing a time slice on the same cores used for
-application tasks, and other resources such as memory bandwidth and
-processor cache capacity. The use of empty tasks makes the application
-share of these resources appear to be zero, which permits usage by the
-runtime system to be unconstrained. How then is anyone supposed to
-determine what resource usage to consider reasonable on the part of
-the runtime system? Only by requiring the application usage of
-resources to meet a certain minimum level, as in METG, can we ensure
-that runtime system resource usage is constrained to realistic levels.
+There is another, deeper issue with TPS. TPS may be measured with
+empty tasks~\cite{LegionTracing18} or with tasks of some
+duration~\cite{Canary16, Armstrong14}. If empty tasks are used, the
+resulting upper bound on task scheduling throughput fails to represent
+any realistic application, as no
+useful work is being performed. If non-empty tasks are used, \emph{but
+  the efficiency of the overall application is not reported}, then TPS
+is not a measurement of runtime-limited performance. Large tasks may
+be used to hide any amount of runtime overhead,
+while tasks that are too small may result in a drop in total
+application throughput even as TPS increases. Only by constraining
+efficiency, as in METG, can we measure runtime-limited application
+throughput while performing useful application work.
 
@@
 Swift/T~\cite{Wozniak13}, TensorFlow~\cite{TensorFlow15}, and
-X10~\cite{X1005}. Using the METG metric, we determine the baseline
+X10~\cite{X1005}. Using METG(50\%), we determine the baseline
 runtime overhead of each system along with overheads associated with
-increasing number and/or variations in the pattern of dependencies. We
+increasing numbers of and/or variations in the pattern of dependencies. We
 also explore each system's ability to scale, hide communication
@@
 \begin{itemize}
-\item Section~\ref{sec:metg} defines the METG metric and its
+\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
   relationship to quantities of interest to application developers.

### 2019-03-26 1_introduction.tex
Tweak introduction line breaking.

@@
 implementations in 13 different programming systems, including
-Chapel~\cite{Chapel07}, Charm++~\cite{Charmpp93}, MPI~\cite{MPI},
-OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40},
-PaRSEC~\cite{PARSEC13}, Realm~\cite{Realm14}, Regent~\cite{Regent15},
-Spark~\cite{Spark10}, StarPU~\cite{StarPU11},
-Swift/T~\cite{Wozniak13}, TensorFlow~\cite{TensorFlow15}, and
-X10~\cite{X1005}. Using METG(50\%), we determine the baseline
+Chapel \cite{Chapel07}, Charm++ \cite{Charmpp93}, MPI \cite{MPI},
+OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
+PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
+Spark \cite{Spark10}, StarPU \cite{StarPU11},
+Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
+X10 \cite{X1005}. Using METG(50\%), we determine the baseline
 runtime overhead of each system along with overheads associated with

### 2019-03-26 2_metg.tex
Apply Alex's comments to section 2.

--- 2019-03-19 2_metg.tex
@@
-\section{METG}
+\section{METG (50\%)}
 \label{sec:metg}
@@
 
-Minimum effective task granularity (METG) is defined intuitively as
-the smallest average task granularity such that the application
-achieves overall efficiency exceeding some threshold. For example, in
+The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
+the smallest average task granularity such that $A$
+achieves overall efficiency exceeding 50\%. For example, in
 a compute-bound application efficiency might be measured as the
 percentage of the available FLOPS achieved. On a machine with a peak
-capability of 1.2 TFLOPS per node, METG(50\%) would correspond to
-smallest task granularity achieved while maintaining at least 0.6
-TFLOPS. Absolute efficiency (i.e., FLOPS or bytes per second) is
-preferable, but parallel efficiency (relative to a sufficiently large
-problem size) can also be used in applications which are not amenable
-to be characterized by peak resource utilization.
+capability of 1.26 TFLOPS per node, METG(50\%) would correspond to
+smallest task granularity achieved while maintaining at least 0.63
+TFLOPS. Absolute efficiency (i.e., FLOPS or B/s) as well
+as parallel efficiency (relative to a sufficiently large problem size)
+can be used.
 
 Figure~\ref{fig:flops-mpi} shows the way in which METG is
-calculated. The application is run with a large enough problem size to
+calculated. The application, in this case an MPI implementation of the
+Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
+run on a single Haswell node of Cori with a large enough problem size to
 ensure that it achieves peak utilization of the machine. This confirms
@@
 efficiency metric is achievable. Once this is established, the problem
-size is shrunk (in a manner analogous to what happens on a node during
-a strong scaling run). The expectation is as problem size shrinks, the
+size is repeatedly reduced while maintaining exactly the same hardware
+configuration (in particular, the same number of nodes). The
+expectation is that as problem size shrinks, the
 efficiency will begin to drop and eventually approach zero. Systems
-with lower overheads will maintain higher efficiency at smaller
+with lower runtime overheads maintain higher efficiency at smaller
 problem sizes compared to systems with higher overheads.
@@
   cores}/\text{number of tasks}$). Note that a \emph{task} is defined
-broadly to be any continuously-executing unit of work in the
-application, thus it makes sense to discuss tasks even in systems
-without an explicit notion of tasking, such as in MPI codes are written in
-a bulk synchronous style.
+broadly to be any continuously-executing unit of application code,
+and thus it makes sense to discuss tasks even in systems
+without an explicit notion of tasking, such as in MPI programs that
+are written in a bulk synchronous style. In this case, the tasks run a
+compute-bound kernel included in the Task Bench implementation,
+described in more detail in Section~\ref{sec:task-bench}.
 
-Figure~\ref{fig:efficiency-mpi}
-shows the replotted data from above. In this chart, we can see that
+Figure~\ref{fig:efficiency-mpi} replots the data in
+Figure~\ref{fig:flops-mpi}. In Figure~\ref{fig:efficiency-mpi}, we see that
 efficiency starts near 100\%, and that initially task granularity
-drops quickly with a minimal change in efficiency. However as task
+drops quickly with minimal change in efficiency. However as task
 granularity shrinks, efficiency begins to drop as well, leading
@@
 
-METG is thus the result of drawing a vertical line through
-Figure~\ref{fig:efficiency-mpi} at the desired level of efficiency,
-and taking the intersection of this line with the task granularity
-curve. In the figures, the red, dashed line shows the 50\% efficiency
+METG(50\%) is the intersection of this curve with a vertical line through
+Figure~\ref{fig:efficiency-mpi} at 50\% efficiency. In the figures, the red, dashed line shows the 50\% efficiency
 level. At 50\% efficiency, MPI achieves an average task granularity of
 4.6 us, thus the METG(50\%) of MPI is 4.6 us for this configuration of
-the application. The number 50\% is chosen arbitrarily, but is a
-reasonable choice for this application because it represents a point
-at which many systems will have plateaued, yet the overall efficiency
-is still reasonable. Lower values would likely not be reasonable,
-because in most cases operators of computing facilities prefer to
-maintain higher levels of utilization of the machines. And higher
-values risks misrepresenting the performance of some systems (not
-shown in the figures above).
+the application. We use 50\% because it is a reasonable level of
+efficiency in practice. Lower values would likely not be reasonable,
+because operators of computing facilities prefer higher levels of
+utilization. Values above 90\% can misrepresent the performance of
+some systems (see
+Section~\ref{subsec:peak-performance-and-efficiency}).
 
@@
 namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
-\ref{fig:strong-scaling-mpi} show the behavior of MPI weak and strong
+\ref{fig:strong-scaling-mpi} show the behavior of the MPI Task Bench stencil's weak and strong
 scaling, respectively, at a variety of problem sizes. Intuitively, at
@@
 where lines become more compressed and begin to trend upward. At the
-very bottom, the lines compress together as it the fundamental
-overhead of MPI dwarfs any gains from running smaller
-problems. Critically, the shape of the curve at the bottom is the same
+very bottom, the lines compress together as the fundamental
+overhead of the MPI implementation dwarfs any gains from running smaller
+problems. The shape of the curve at the bottom is the same
 for both weak and strong scaling; this is precisely the shape of
@@
 of efficiency. With any smaller problem size, 
-the overhead begins to dominate useful
+the runtime overhead begins to dominate useful
 computation. Similarly, METG corresponds to the point at which strong
@@
 application), it is possible to use METG to isolate effects that are
-due to shrinking problem size from effects that are due to e.g.,
-increased communication latencies as progressively larger portions of
-the machine are being used. In comparison, with only a single strong
-or weak scaling run, it is impossible to determine if the shape of the
-curve is due to the overhead of the runtime system, or communication
-effects as progressively larger portions of the machine are being
-used.
+due to shrinking problem size from effects that are due to
+increased communication latencies and other resource issues as
+progressively larger portions of the machine are used.
 
@@
 programming systems.
-
-% talk about TPS / MTG ?

### 2019-03-26 3_task_bench.tex
Add figures for task graphs.
Apply Alex's comments to Section 3.

@@
 
-In order to measure properties such as METG for a wide variety of
+\input{f5_task_graphs}
+
+To measure properties such as METG for a wide variety of
 systems and application scenarios, we developed a novel
@@
 in an \emph{iteration space} and dependencies between tasks determined
-by a \emph{dependence relation}. Each task can execute any one of a
+by a \emph{dependence relation}. A task that depends on another task
+must execute after that task, and receives the output data from that
+task as input. Each task can execute any one of a
 number of kernels (compute-bound, memory-bound, etc.) and can generate
-a configurable amount of result data in order to control the amount of
-communication that must be performed with each task dependency. Thus Task Bench can be used to
-rapidly explore a large space of possible application scenarios in
-order to understand the performance behaviors of parallel and
+a configurable amount of output to control the amount of communication
+performed with each task dependency. Thus Task Bench can be used to
+rapidly explore a large space of possible application scenarios to
+understand the performance of parallel and
 distributed languages and runtime systems.
@@
 For simplicity but without loss of generality, the iteration space in
-Task Bench is constrained to be 2-dimensional, with time flowing along
+Task Bench is constrained to be 2-dimensional, with time along
 the vertical axis and parallel tasks along the
@@
 patterns. Figure~\ref{fig:task-graphs} shows a number of sample task
-graphs that can be implemented with Task Bench.
-
-% FIXME: Make figure showing example task graphs
+graphs that can be implemented with Task Bench. Note that layout is
+significant---in particular, each column $i$ corresponds to all tasks
+with index $i$ in the iteration space over all the time steps.
 
 Dependencies between tasks are controlled by a configurable dependence
-relation. Dependencies are conceived as flowing backwards in time. The
-dependence relation enumerates for each task, the set of tasks in the
+relation. Dependencies point backwards in time. The
+dependence relation enumerates, for each task, the set of tasks in the
 previous time step that task depends on. This permits a wide variety
 of dependence patterns to be implemented that are relevant to real
-applications in high-performance scientific computing: stencils,
+applications in high-performance scientific computing and data analysis: stencils,
 sweeps, FFTs, trees, and so on. Dependence patterns may also be
 parameterized, such as picking the $K$ nearest neighbors, or $K$
-distant neighbors. These dependence patterns can be seen in
-Figure~\ref{fig:task-graphs}.
+distant neighbors. Table~\ref{tab:equations} show equations for the
+dependence relations of the patterns in Figure~\ref{fig:task-graphs}.
+
+\input{t0_equations}
 
 Despite its generality, Task Bench is easy to implement, making it
-tractable to develop a suite of high-quality implementations. Most
-Task Bench implementations are several hundred lines long, with the
-shortest being 88 lines and the longest being 1500 lines of executable
-code. Much of this is possible because the core aspects of generating
-task graphs and enumerating dependencies are encapsulated in a core
+tractable to develop a suite of high-quality implementations. The core aspects of Task Bench, such as generating
+task graphs and enumerating dependencies, are encapsulated in a core
 library that is shared among all the Task Bench implementations. The
@@
 that the kernels are identical in all cases, eliminating a potential
-source of performance disparity which can be a pitfall for mini-apps
-implementations. Finally, the Task Bench core library manages the
+source of performance disparity that can be a pitfall for
+implementations of mini-apps. Finally, the Task Bench core library manages the
 parsing of input parameters and the display out output results,
 ensuring that all implementations behave in a uniform manner, and can
-be scripted consistently.
+be scripted consistently. Because much of the functionality needed for
+a Task Bench implementation is in the core library, implementations of
+Task Bench are small: Our 13 Task Bench implementations range from 88
+to 1500 lines, with several hundred lines being typical.
 
-The task implementation in the Task Bench core library is fully
+The Task Bench core library is fully
 validating. Because the task graph configuration is explicitly
@@
 can be queried to determine exactly what dependencies a task should
-expect. The output of every task in Task Bench is uniquely generated,
+expect. The output of every task in Task Bench is unique,
 and all inputs are verified. An assertion is thrown if validation
-fails. This ensures that every execution of Task Bench, if it
+fails. This level of self-checking ensures that every execution of Task Bench, if it
 completes successfully, is correct. This validation provides high
@@
 effect on overall results.
+
+Task Bench provides two main kernels which can be called from tasks,
+which are compute- and memory-bound, respectively. The compute-bound
+kernel executes a tight loop and is hand-written using AVX2 FMA
+intrinsics. The memory-bound kernel performs sequential reads and
+writes over an array, and is again hand-optimized with AVX2
+intrinsics. The duration of both kernels can be configured by setting
+the number of iterations to execute; we use this ability to simulate
+the effects of varying application problem sizes. The memory-bound
+kernel is carefully written to keep the working set size constant as
+the number of iterations decrease, to avoid unwanted speedups due to
+cache effects.

### 2019-03-26 4_implementation.tex
Apply Alex's comments to section 4.

@@
 are \emph{explicitly parallel} and require users to specify the
-parallelism in the program. As a rule of thumb we aim for the Task
-Bench implementations to be representative of how actual applications
+parallelism in the program. The Task
+Bench implementations are intended to reflect how actual applications
 would be written in the respective systems. Places where the Task
@@
 varies so widely (if it exists at all) that it is challenging to come
-up with a representative benchmark which allows these features to be
+up with a representative benchmark that allows these features to be
 compared in an apples-to-apples manner.
@@
 were consulted during the development and evaluation of the
-corresponding Task Bench implementations. Their insight was invaluable
+corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
 in helping to ensure that we provide the highest level of quality in the Task
@@
 levels. Chapel's core features are a partitioned global address space
-(PGAS) which permits remote access to objects from anywhere in the
-machine, a data model for the distribution of data structures over the
-global address space via \emph{domain maps}, tasks, and
-synchronization primitives. Chapel thus permits a great deal of
+(PGAS), a data model for the distribution of data structures, tasks, and
+synchronization primitives. Certain operations can be lifted to entire arrays, enabling forms of data-parallelism. Chapel thus permits a great deal of
 flexibility in the choice of implementation strategy for applications.
 
-For Task Bench, we decided to target a relatively low-level
-implementation in Chapel, using explicit task instantiation (via
+For Task Bench, we decided to target a low-level implementation in
+Chapel, using explicit task instantiation (via
 \lstinline[language=Chapel]{coforall}), bulk access to distributed
 arrays for efficient data movement, and atomic integers for
-synchronization. The advantage of this approach is that it permits
-fine-tuning of the implementation for performance, and avoids relying
-on advanced optimizations in the Chapel compiler, which might struggle
-with the dynamic, input-dependent nature of the task graphs in Task
-Bench.
-
-Chapel also supports a more high-level approach to parallel
-programming in which the granularities of tasks are determined
-automatically by the compiler (via
-\lstinline[language=Chapel]{forall}). This is the level that users
-would typically target in their programs. The specific dependence
-patterns supported by Task Bench might also be implementable at this
-level, though a full Task Bench implementation would be challenging.
-
-% FIXME: Does the Chapel compiler handle: bulk data movement, point-to-point synchronization
+synchronization. This approach is a good match for Task Bench, as the
+Task Bench core provides implementations of the kernels.
 
@@
 Charm++~\cite{Charmpp93} is an explicitly parallel programming system
-based on the actor model of distributed computation. Actors, or
+based on the actor model. Actors, or
 \emph{chares}, are objects that live in their own address space and
-can be migrated automatically between nodes for better load
-balance. Chares are permitted to send messages to other chares via a
+can be migrated automatically between nodes to
+balance load. Chares send messages to other chares via a
 form of remote procedure call (RPC). Calling a method of another chare
@@
 Our Task Bench implementation is straightforward: we create a chare
-array for the task graph, with one chare for each column. Methods are
+array for the task graph, with one chare for each column of the task graph. Methods are
 used to implement dependencies, and a task is executed as soon as its
@@
 
-MPI~\cite{MPI} is the industry standard message-passing API for
-programming supercomputers, and is the most widely used programming
-model in high-performance scientific computing today. The goal of our
-MPI implementation of Task Bench is to represent the common (if
-unexciting) case which continues to dominate high-performance
-computing: bulk synchronous programming with distinct computation and
+MPI~\cite{MPI} is the standard message-passing API for
+programming supercomputers and remains the most widely used programming
+model in high-performance scientific computing today. Our
+MPI implementation of Task Bench represents the common case of bulk synchronous programs with distinct computation and
 communication phases. We use \lstinline[language=C++]{MPI_Isend} and
@@
 
-PaRSEC is a task-based programming system which supports two distinct
+PaRSEC is a task-based programming system supporting two distinct
 programming models: \emph{parameterized task graphs}
@@
 different nodes. While the basic DTD model is implicitly parallel
-because the system is responsible for the discovery of parallelism, in
-order to achieve horizontal scalability, the program is executed in a
+because the system is responsible for the discovery of parallelism, to
+achieve horizontal scalability, the program is executed in a
 SPMD fashion on all nodes, and the user is responsible for eliding
@@
 
-Data in PaRSEC is stored in a built-in matrix data structure with a 2D block cyclic distribution. 
-The Task Bench implementation divides this matrix into $M$ by $N$ tiles where $N$ equals to the width of
-the task graph and $M$ varies from 2 to the height of the task graph depending
-on optimizations performed to alleviate write-after-read depdenencies. 
+Data in PaRSEC is stored in a built-in matrix data structure with a 2D
+block cyclic distribution.  The Task Bench implementation divides this
+matrix into $M$ by $N$ tiles where $N$ equals to the width of the task
+graph and $M$ varies from 2 to the height of the task graph depending
+on optimizations performed to alleviate write-after-read dependencies.
 Tiles are distributed across nodes in pre-defined patterns,
@@
 Realm~\cite{Realm14} is an explicitly-parallel task-based programming
-model that used internally by Legion~\cite{Legion12} and
+model used internally by Legion~\cite{Legion12} and
 Regent~\cite{Regent15}. Thus the Realm implementation of Task Bench can be
@@
 between tasks explicitly defined by \emph{events} passed from one task
-to another. Realm's data model supports \emph{regions} of memory which live on a single
-node. Data must be explicitly copied, and the copies synchronized with
-tasks via events.
+to another. Realm's data model supports collections that live in a
+specific memory. Data must be explicitly copied, and the copies
+synchronized with tasks via events.
 
@@
 
-StarPU~\cite{StarPU11} is a task-based system that supports an DTD
-programming model similar to PaRSEC. Our Task Bench implementation in
+StarPU~\cite{StarPU11} is a task-based system that supports a \emph{sequential task flow} (STF)
+programming model similar to PaRSEC's DTD. Our Task Bench implementation in
 StarPU is very similar to PaRSEC, including the use of a switch

### 2019-03-26 5_evaluation.tex
Apply Alex's comments to section 2.
Apply Alex's comments to section 5.

@@
 
-We now present a comprehensive evaluation of Task Bench on up to 256
+We present a comprehensive evaluation of Task Bench on up to 256
 Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
 machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
-processors (total of 32 physical cores per node), 128 GB RAM, and a
+processors (a total of 32 physical cores per node), 128 GB RAM, and a
 Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
@@
 \subsection{Peak Performance and Efficiency}
+\label{subsec:peak-performance-and-efficiency}
 
@@
 
-The vast majority of systems achieve at or near peak FLOPS. Some
+The vast majority of systems achieve or nearly achieve peak FLOPS. Some
 systems reserve a number of cores (usually 1 or 2) for internal
@@
 peak FLOPS achieved) plotted against task granularity. This plot is
-used to calculate METG, as described in Section~\ref{sec:metg}. The
+used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
 red, dashed line shows the point where 50\% efficiency is achieved.
@@
 In most cases, task granularity has plateaued prior to the 50\% efficiency point,
-though a number of systems continue to improve as efficiency continues
+though some systems continue to improve as efficiency continues
 to drop. Accounting for this effect is one of the main arguments in
@@
 measuring task scheduling throughput in the presence of empty tasks
-(effectively the equivalent of METG(0\%)). Empty tasks permit runtime
+(effectively METG(0\%)). Empty tasks permit runtime
 systems to make unrealistic tradeoffs which are not applicable to real
@@
 \item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
-  task depends 5 others, chosen to be as close as possible.
+  task depends on 5 others, chosen to be as close as possible.
 \item Figure~\ref{fig:metg-compute-spread} is a pattern where each
@@
 communication phases. In contrast the asynchronous execution
-mechanisms of the other systems is pure overhead in this scenario.
+mechanisms of the other systems are pure overhead in this scenario.
 
 The gap between MPI and other systems shrinks as the complexity of the
-communication pattern grows, can even reverse as task parallelism is
+communication pattern grows, and even reverses as task parallelism is
 added in the form of multiple task graphs.
@@
 evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and a
-flat line is ideal. Systems with non-scalable implementations shows
-lines that go up with node count. Most notably, Spark is typically
+flat line is ideal. Lines that rise with node count indicate less than
+ideal scaling. Most notably, Spark is typically
 intended for use in industrial data center applications with task
 granularities measured in tens of seconds. Spark uses a centralized
-controller which limits throughput, and this is visible in the figure
+controller, which limits throughput, and this is visible in the figure
 as line for Spark immediately rises with node count. Keep in mind also
@@
 imbalanced workloads are generated by randomly varying the task
-durations according by a factor determined by the formula $1+(X-0.5)I$
+durations by $1+(X-0.5)I$
 where $X$ is a uniform random variable in the range $[0,1]$ and $I$ is
@@
 The MPI implementation of Task Bench, being bulk synchronous,
-obviously suffers the most under load imbalance. The biggest
+suffers the most under load imbalance. The biggest
 difference is at large task granularities, where the imbalance has the
@@
 parallelism due to the presence of 4 task graphs, is sufficient to
-completely mitigate load imbalance. Systems which exploit task
+completely mitigate load imbalance. Systems that exploit task
 parallelism but not work stealing on a node achieve lower peak
@@
 
-Also of interest is the ability of asynchronous systems to be able to
-hide communication latencies in the presence of task
+Also of interest is the ability of asynchronous systems to
+hide communication latency in the presence of task
 parallelism. Figure~\ref{fig:efficiency-communication} plots task

### 2019-03-26 5_evaluation.tex
Apply Alex's comments to sections 6 and 7.

@@
 
-For this reason, we feel especially strongly that it is important to
-measure the performance behaviors of programming systems under
-realistic workloads, rather than trivial dependencies.
+For this reason, we recommend evaluating the performance of
+programming systems under realistic workloads, rather than with
+trivial dependencies.
 

### 2019-03-26 6_related.tex
Apply Alex's comments to sections 6 and 7.

@@
 than can be reasonably implemented to a level of quality sufficient
-for publication. This can lead to publications for mini-apps omitting
-implementations or performance results for systems beyond the reference
-implementation. Follow-up papers, if any, often compare only against a
-reference implementation and omit any other implementations that might
-exist.
+for meaningful measurement. New mini-apps often provide only a
+reference implementation, or omit performance evaluations of different
+implementations, if those are presented. Follow-up experiments, if
+any, normally present a single system, and compare against the
+reference implementation (or not at all) and not with any other
+implementations that might be available. As a result, comparative
+studies across many implementations are rare.
 
@@
 \item
-The mini-app CoMD received a software release but no initial formal
-publication. One follow-up paper describes a Chapel
+One follow-up paper for the mini-app CoMD describes a Chapel
 implementation~\cite{CoMDChapel16} (comparison against reference
@@
 variants, SHMEM, UPC, Charm++, and Grappa. However, PRK Stencil is a
-simple code involving 2 kernels (which combined fit in under 50 lines
-of C++) and a halo exchange on a structured grid; it is unclear to
-what extent this is representative of real-world applications. Other
-codes in PRK are even less representative.
-
-The NAS benchmark suite~\cite{NAS91, NAS95} are widely studied with
+single, simple code involving 2 kernels (which combined fit in under 50 lines
+of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} are also relatively widely studied with
 implementations in OpenMP~\cite{NASOpenMP99}, MPI and
 MPI+OpenMP~\cite{NASMPIOpenMP00}, and
-Charm++~\cite{NASCharm96}. However the benchmarks consist mostly of small
-kernels for dense array computations, and do not generalize to larger
-applications.
+Charm++~\cite{NASCharm96}. The benchmarks consist mostly of small
+kernels for dense array computations. Neither the PRK nor the NAS
+benchmark suite achieve the breadth of coverage, flexibility, or ease
+of implementation of Task Bench. We believe the evidence for ease of
+implementation is clear; Task Bench is implemented for more systems
+than all previous mini-apps and microbenchmarks. Task Bench is also
+not a fixed set of benchmarks, and it is even easier to use Task Bench
+to generate a new benchmark to measure some aspect of system
+performance.
 
@@
 latency~\cite{MPPTest99, MPIBench01}. These measurements typically do
-not generalize beyond the immediate system they attempt to quantify,
-nor in an obvious way to real-world applications.
+not generalize beyond the immediate system they measure.
 

### 2019-03-26 7_conclusion.tex
Apply Alex's comments to sections 6 and 7.

@@
 
-We presented Task Bench, a parameterized benchmark powerful enough to
+We presented Task Bench, a parameterized benchmark flexible enough to
 enable the exploration of a broad space of application scenarios, but
@@
 of insight it delivers about different systems. We discover, among
-other things, that the baseline overhead of systems varies by up to 5
-orders of magnitude.
+other things, that the baseline overhead of parallel and distributed
+programming systems varies by up to 5 orders of magnitude.
 
 Helping us in this endeavor is a new metric, minimum effective task
-granularity (METG), which helps to sidestep a number of potential
-issues in the evaluation of task scheduling performance, and helps to
-separate such factors from changes in communication topology and
-latency at increasing scales of the machine.
+granularity (METG), which isolates the inherent overhead of the
+runtime system for a given application from factors such as changes in
+communication topology and latency when using progressively larger
+portions of a machine.

### 2019-03-26 5_evaluation.tex
Cut down on the number of load imbalance plots.

@@
 Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
-efficiency curves under varying amounts of load imbalance. The
-imbalanced workloads are generated by randomly varying the task
-durations by $1+(X-0.5)I$
-where $X$ is a uniform random variable in the range $[0,1]$ and $I$ is
-the degree of imbalance in the range $[0,2]$. The random numbers are
-generated with a deterministic pseudo random number generator with a
-deterministic seed to ensure that the task durations under all systems
-are identical.
+efficiency curves under load imbalance with uniformly distributed task
+durations. The task durations are generated with a deterministic
+pseudo random number generator with a deterministic seed to ensure
+that the task durations under all systems are identical.
 

### 2019-03-26 e5_imbalance.tex
More text in evaluation.
More text for evaluation section.
Cut down on the number of load imbalance plots.

@@
-\begin{figure*}[t]
+\begin{figure}[t]
 
-\subfloat[Imbalance 0.5.\label{fig:efficiency-imbalance-0.5}]{
-\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_0.5.pdf}
-}
-\subfloat[Imbalance 1.0.\label{fig:efficiency-imbalance-1.0}]{
-\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_1.0.pdf}
-}
+\centering
+\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_2.0.pdf}
 
-\subfloat[Imbalance 1.5.\label{fig:efficiency-imbalance-1.5}]{
-\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_1.5.pdf}
-}
-\subfloat[Imbalance 2.0.\label{fig:efficiency-imbalance-2.0}]{
-\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_2.0.pdf}
-}
-
-\caption{Task granularity vs efficiency for varying load imbalance.\label{fig:metg-imbalance}}
-\end{figure*}
+\caption{Task granularity vs efficiency under load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Lower is better.\label{fig:efficiency-imbalance}}
+\end{figure}

### 2019-03-26 3_task_bench.tex
Apply Alex's comments to section 4.
Update section 3 text.

@@
 significant---in particular, each column $i$ corresponds to all tasks
-with index $i$ in the iteration space over all the time steps.
+with index $i$ in the iteration space over all the time
+steps. Generally speaking each column of the task graph will be
+assigned to execute on a different processor core.
 
@@
 distant neighbors. Table~\ref{tab:equations} show equations for the
-dependence relations of the patterns in Figure~\ref{fig:task-graphs}.
+dependence relations of the patterns in Figure~\ref{fig:task-graphs},
+where $t$ is timestep, $i$ is column, and $W$ is the width of the task
+graph.
 

### 2019-03-27 5_evaluation.tex
read evaluation

@@
 runtime system overhead. However, verifying that peak performance is
-achieved is important to ensuring that such evaluations of overhead and
+achieved is important to ensure that such evaluations of overhead and
 efficiency are well-grounded and avoid common pitfalls in
@@
 reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
-of measuring efficiency we use our empirically determined number as
+of measuring efficiency, we use our empirically determined number as
 the baseline for 100\% efficiency.
@@
 shows how METG(50\%) varies with node count for a variety of
-dependence patterns supported by Task Bench:
+dependence patterns supported by Task Bench. The experiment is conducted
+in a weak scale fashion, in which we keep the number of tasks per node
+constant while increasing the number of nodes. The detailed explanations
+of each dependence pattern are as follows:
 
@@
 cases. Spark is more efficient with trivial parallelism, as described
-in Section~\ref{subsec:number-of-dependencies}.
+in Section~\ref{subsec:number-of-dependencies}. Systems like PaRSEC
+and StarPU build the DAG on the fly, so they require to examine all tasks
+of the entire DAG on each process. Even though DAG trimming is 
+implemented on both systems to not launching tasks on a specific node
+if data is not local, the overhead of sweeping all the tasks to 
+check data localities goes up with the increasement of
+node counts in the weak scale experiment, resulting in poor scalability. 
+Even though the PTG implementation relieves the overhead of DAG expansion
+from runtime to compile stage, the overhead of determining local tasks
+based on data locality leads to un-horizontally scalable. 
+The ``PaRSEC DTD Shard'' is an modified implementation of ``PaRSEC DTD''
+by pre-partitioning the DAG when assuming data distribution is fixed 
+(1D block is used in Task Bench). It removes the cost of sweeping the
+entire DAG, and is therefore horizontally scalable.
 
-% FIXME: talk about starpu and parsec?
+% FIXME: remove parsec dtd shard if no space.
 
@@
 scheduling behavior of the various systems. For example, OpenMP
-executes tasks via a thread pool, resulting in a form of work
+execute tasks via a thread pool, resulting in a form of work
 stealing. This on-node work stealing, in combination with the task
@@
 not relying on a shared global queue for tasks.
+
+% FIXME: I am not sure if OpenMP has only 1 thread pool or multiple pools. 
+% Systems like StarPU and PaRSEC have multiple pools, but still allows on-node
+% work stealing. So I also think the last sentence " may achieve smaller 
+% task granularities as a result of not relying on a 
+% shared global queue for tasks." is not accurate.
 
@@
 smaller message sizes.
+
+% FIXME why MPI can be affected by other jobs but asynchronous systems can not?

### 2019-03-29 5_evaluation.tex
Some edits on section 5.

--- 2019-03-26 5_evaluation.tex
@@
 shows how METG(50\%) varies with node count for a variety of
-dependence patterns supported by Task Bench:
+dependence patterns supported by Task Bench. METG(50\%) is calculated
+separately at each node count, to isolate runtime system behavior from
+changes in communication latencies and topology when using
+progressively larger portions of the machine. We consider the
+following configurations of Task Bench:
 
@@
 
-% FIXME: talk about starpu and parsec?
+Implicitly parallel systems, such as PaRSEC, StarPU, and Regent, that
+perform runtime analysis in order to build the DAG can suffer from
+scalability bottlenecks if every node must consider the tasks
+executing on all other nodes. PaRSEC DTD and StarPU mitigate this
+partially be allowing the user to omit tasks that are not direct
+dependencies of those executed on the current node. However, this DAG
+trimming approach still requires some dynamic checks that scale with
+the number of nodes. Approaches based on compile-time analysis can
+partially or fully mitigate this overhead. PaRSEC PTG provides some
+improvement over DTD by performing DAG expansion at compile time, but
+still retains dynamic checks leading to limited scalability. Regent
+uses a compile-time optimization that enables full horizontal
+scalability~\cite{ControlReplication17}. A similar, manual
+optimization is shown in the plots as PaRSEC DTD Shard.
+
+% FIXME: remove parsec dtd shard if no space.
 
@@
 The difference between the remaining systems is due primarily to the
-scheduling behavior of the various systems. For example, OpenMP
-executes tasks via a thread pool, resulting in a form of work
-stealing. This on-node work stealing, in combination with the task
-parallelism due to the presence of 4 task graphs, is sufficient to
-completely mitigate load imbalance. Systems that exploit task
-parallelism but not work stealing on a node achieve lower peak
-efficiency, but may achieve smaller task granularities as a result of
-not relying on a shared global queue for tasks.
+scheduling behavior of the various systems. The execution of 4
+simultaneous task graphs partially, but does not fully, mitigate the
+load imbalance in the tasks. Therefore, systems which provide an
+additional on-node work stealing capability see additional gains in
+efficiency at large task granularities. However, the use of
+work-stealing queues can also impact throughput at very small task
+granularities. This can be seen in the difference between Chapel
+Distrib (which uses a work stealing scheduler) vs Chapel with its
+default (non-work-stealing) scheduler.
 
@@
 levels of efficiency compared to the bulk synchronous MPI
-implementation. Second, the asynchrony also makes the curves smoother,
+implementation. Second, the asynchrony and scheduling flexibility from
+executing multiple graphs also makes the curves smoother,
 as spikes in latency due to interference from other jobs can be

### 2019-03-29 5_evaluation.tex
Tweak text.

@@
 
-Implicitly parallel systems, such as PaRSEC, StarPU, and Regent, that
-perform runtime analysis in order to build the DAG can suffer from
+Implicitly parallel systems such as PaRSEC, StarPU and Regent that
+rely on runtime analysis to build the DAG can suffer from
 scalability bottlenecks if every node must consider the tasks
 executing on all other nodes. PaRSEC DTD and StarPU mitigate this
-partially be allowing the user to omit tasks that are not direct
+partially by allowing the user to omit tasks that are not direct
 dependencies of those executed on the current node. However, this DAG
 trimming approach still requires some dynamic checks that scale with
-the number of nodes. Approaches based on compile-time analysis can
-partially or fully mitigate this overhead. PaRSEC PTG provides some
-improvement over DTD by performing DAG expansion at compile time, but
+the number of nodes and thus limits scalability. Approaches based on compile-time analysis can
+partially or fully mitigate this overhead. PaRSEC PTG improves over DTD by performing DAG expansion at compile time, but
 still retains dynamic checks leading to limited scalability. Regent

### 2019-03-29 0_abstract.tex
Apply Alex's comments to sections 1-4.

@@
 \begin{abstract}
-% 129 words
+% 126 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
@@
 microbenchmarks, which makes comparisons between many systems
-possible, with the ability to easily model a wide range of application
+possible, with the ability to model a wide range of application
 scenarios, giving it coverage approximating large benchmarks.
@@
   granularity} to study the baseline runtime overhead of each system
-and the impact of scenarious with varying dependence patterns. We also
+and the impact of different dependence patterns. We also
 study each system's scalability, and ability to hide communication and

### 2019-03-29 1_introduction.tex
Apply Alex's comments to sections 1-4.

@@
 wide variety of proposals for programming models, languages, and
-runtime systems. Some, like MPI~\cite{MPI}, have achieved widespread
-usage while many more are in the research and development stage. Our
+runtime systems. But while the features of these systems are widely
+written about in the literature, comprehensive performance evaluations
+of these systems remain difficult to find. Our
 goal in this paper is to develop a meaningful and useful framework for
@@
 focused on the development of \emph{proxy-} or
-\emph{mini-apps}. Because they embody the logic of a larger
-application, but in reduced form, mini-apps can provide insight
-without incurring the costly expense of developing a full-blown
-application code. However, despite the name, our experience is that
+\emph{mini-apps}. Because they distill the structure of a larger
+application, mini-apps can provide insight
+without the expense of developing a full-blown
+application. However, despite the name, our experience is that
 mini-apps still requite significant investment to develop, especially
-to a level of quality needed for useful benchmarking. In many cases,
+to the level of quality needed for useful benchmarking. In many cases,
 the effort to tune for performance exceeds the effort to develop the
 mini-app in the first place. As a result, implementations of mini-apps
-often do not reach a level of maturity required to compare
+often do not reach the level of maturity required to compare
 systems. Indeed, there are few such published studies, and none to our
@@
 programming system. Weak scaling can hide arbitrary amounts of runtime
-system overhead with sufficiently large problem sizes. And strong
-scaling does not separate the overhead of the underlying runtime
-system from the growth in communication latencies, and potential
-saturation of resources such as communication and I/O bandwidth, when
-using progressively larger portions of the machine.
+system overhead with sufficiently large problem sizes, and strong
+scaling does not separate runtime system overhead from other costs
+(such as increased communication, among other things) when
+using progressively larger portions of a machine.
 
-To address the first of these issues, we present Task Bench, a
+To address the first issue, we present Task Bench, a
 \emph{parameterized} benchmark designed for exploring the performance
@@
 
-To isolate the contribution of the programming model's runtime system
+To address the second issue and isolate the contribution of the programming model's runtime system
 to application performance, we introduce a new metric called
@@
 meaning that the application achieves at least 50\% of peak
-performance. METG captures the essence of what is important in a weak
+performance on a given machine. METG captures the essence of what is important in a weak
 or strong scaling study, the behavior at the limit of scalability. For
@@
 METG(50\%) can be used to compute the scale at which performance can
-be expected to dip below 50\% efficiency. A good rule of thumb is that
-as long as an application's \emph{average} task granularity exceeds
-METG(50\%), the application should achieve at least 50\% efficiently.
+be expected to dip below 50\% efficiency.
 
-Task Bench and METG mitigate issues common in limit studies of runtime
+A lower METG does not necessarily mean that performance will be
+better. For example, two systems with METG(50\%) of 100~us and 1~ms,
+respectively, running an application with an average task granularity
+of 10~ms, are both likely to perform well. Only when the task
+granularity approaches (or drops below) METG(50\%) is performance
+likely to diverge. Thus METG helps to identify the regime in which a
+given system is likely to perform well.
+
+Task Bench and METG address issues common in limit studies of runtime
 systems for parallel and distributed programming. Such studies often
@@
 
-The following four sections each describe a contribution of this
-paper:
+The paper is organized as follows:
 

### 2019-03-29 2_metg.tex
Apply Alex's comments to sections 1-4.

@@
 The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
-the smallest average task granularity such that $A$
+the smallest average task granularity (i.e., task duration) such that $A$
 achieves overall efficiency exceeding 50\%. For example, in
@@
 capability of 1.26 TFLOPS per node, METG(50\%) would correspond to
-smallest task granularity achieved while maintaining at least 0.63
-TFLOPS. Absolute efficiency (i.e., FLOPS or B/s) as well
-as parallel efficiency (relative to a sufficiently large problem size)
-can be used.
+the smallest task granularity achieved while maintaining at least 0.63
+TFLOPS. Relative parallel efficiency (vs a sufficiently large problem
+size) can also be used in cases where the application isn't easily
+characterized by peak resource usage.
 
-Figure~\ref{fig:flops-mpi} shows the way in which METG is
+Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
 calculated. The application, in this case an MPI implementation of the
 Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
-run on a single Haswell node of Cori with a large enough problem size to
-ensure that it achieves peak utilization of the machine. This confirms
+run on a single Haswell node of Cori with a problem size large enough to
+ensure that it achieves peak FLOPS. This starting point confirms
 that the application is properly configured and that the choice of
-efficiency metric is achievable. Once this is established, the problem
-size is repeatedly reduced while maintaining exactly the same hardware
+efficiency metric is achievable. The problem
+size is then repeatedly reduced while maintaining exactly the same hardware
 configuration (in particular, the same number of nodes). The
-expectation is that as problem size shrinks, the
-efficiency will begin to drop and eventually approach zero. Systems
-with lower runtime overheads maintain higher efficiency at smaller
+expectation is that as problem size shrinks,
+performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
+with lower runtime overheads maintain higher performance at smaller
 problem sizes compared to systems with higher overheads.
 
-To calculate METG, this data is replotted along axes of efficiency
+To calculate METG, the data is replotted along axes of efficiency
 (i.e., as a percentage of the peak FLOPS achieved) and task
 granularity (i.e., $\text{wall time} \times \text{number of
-  cores}/\text{number of tasks}$). Note that a \emph{task} is defined
+  cores}/\text{number of tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
 broadly to be any continuously-executing unit of application code,
@@
 
-Figure~\ref{fig:efficiency-mpi} replots the data in
-Figure~\ref{fig:flops-mpi}. In Figure~\ref{fig:efficiency-mpi}, we see that
+In Figure~\ref{fig:efficiency-mpi}, we see that
 efficiency starts near 100\%, and that initially task granularity
 drops quickly with minimal change in efficiency. However as task
-granularity shrinks, efficiency begins to drop as well, leading
+granularity shrinks further, efficiency begins to drop as well, leading
 eventually to a plateau in task granularity. Intuitively, this plateau
@@
 ideally-sloped downward lines when strong scaling. Inefficiency begins
-to show up at smaller problem sizes, towards the bottom of the graph,
-where lines become more compressed and begin to trend upward. At the
-very bottom, the lines compress together as the fundamental
-overhead of the MPI implementation dwarfs any gains from running smaller
-problems. The shape of the curve at the bottom is the same
-for both weak and strong scaling; this is precisely the shape of
+to show at smaller problem sizes, towards the bottom of the graph,
+where lines become more compressed. At the
+very bottom, the lines compress together as running time becomes dominated by overhead. The shape of the curve at the bottom is the same
+for both weak and strong scaling; this is the shape of
 the METG curve.
@@
 
-The METG metric has one particularly useful property in comparison to
-weak and strong scaling. Because METG is measured ``in place'' (i.e.,
+The METG metric has another useful property. Because METG is measured ``in place'' (i.e.,
 without changing the number of nodes or cores available to the
-application), it is possible to use METG to isolate effects that are
+application), METG isolates effects that are
 due to shrinking problem size from effects that are due to
-increased communication latencies and other resource issues as
+increased communication and other resource issues as
 progressively larger portions of the machine are used.

### 2019-03-29 3_task_bench.tex
Tweak spacing around figure captions.
Apply Alex's comments to sections 1-4.

@@
 
-\input{f5_task_graphs}
-
-To measure properties such as METG for a wide variety of
+To measure a wide variety of
 systems and application scenarios, we developed a novel
@@
 a configurable amount of output to control the amount of communication
-performed with each task dependency. Thus Task Bench can be used to
-rapidly explore a large space of possible application scenarios to
+performed with each task dependency. Thus Task Bench can explore a large space of application scenarios to
 understand the performance of parallel and
-distributed languages and runtime systems.
+distributed programming systems.
 
-For simplicity but without loss of generality, the iteration space in
+For simplicity, but without loss of generality, the iteration space in
 Task Bench is constrained to be 2-dimensional, with time along
@@
 horizontal. Tasks can depend only on tasks from the immediately
-preceding time step. Many problems can be described by a task graph of
-constant width, in which case the iteration space is rectangular, but
-Task Bench is capable of handling non-rectangular iteration spaces,
-such as required by sweeps and tree-based dependence
-patterns. Figure~\ref{fig:task-graphs} shows a number of sample task
+preceding time step. Figure~\ref{fig:task-graphs} shows a number of sample task
 graphs that can be implemented with Task Bench. Note that layout is
@@
 Dependencies between tasks are controlled by a configurable dependence
-relation. Dependencies point backwards in time. The
+relation. The
 dependence relation enumerates, for each task, the set of tasks in the
@@
 
+\input{f5_task_graphs}
 \input{t0_equations}
@@
 implementations of mini-apps. Finally, the Task Bench core library manages the
-parsing of input parameters and the display out output results,
+parsing of input parameters and the display of output results,
 ensuring that all implementations behave in a uniform manner, and can
@@
 fails. This level of self-checking ensures that every execution of Task Bench, if it
-completes successfully, is correct. This validation provides high
-assurance that all Task Bench implementations are correct, and
-safeguards against corner cases that may only be exercised at
-scale. An evaluation of the performance impact of
-this validation showed it to be less than 3\% at the smallest task
+completes successfully, is correct. An evaluation of the impact of
+validation showed it to be less than 3\% at the smallest task
 granularities in any Task Bench implementation, with a negligible
@@
 
-Task Bench provides two main kernels which can be called from tasks,
+Task Bench provides two main kernels that can be called from tasks,
 which are compute- and memory-bound, respectively. The compute-bound

### 2019-03-29 4_implementation.tex
Apply Alex's comments to section 5.
Apply Alex's comments to sections 1-4.

@@
 levels. Chapel's core features are a partitioned global address space
-(PGAS), a data model for the distribution of data structures, tasks, and
-synchronization primitives. Certain operations can be lifted to entire arrays, enabling forms of data-parallelism. Chapel thus permits a great deal of
-flexibility in the choice of implementation strategy for applications.
-
-For Task Bench, we decided to target a low-level implementation in
+(PGAS), data distributions, tasks, and
+synchronization primitives, and array promotion. For Task Bench, we decided to target a low-level implementation in
 Chapel, using explicit task instantiation (via
@@
 arrays for efficient data movement, and atomic integers for
-synchronization. This approach is a good match for Task Bench, as the
-Task Bench core provides implementations of the kernels.
+synchronization.
 
@@
 based on the actor model. Actors, or
-\emph{chares}, are objects that live in their own address space and
-can be migrated automatically between nodes to
-balance load. Chares send messages to other chares via a
-form of remote procedure call (RPC). Calling a method of another chare
-is the primary form of synchronization and data movement in
-Charm++. Additional support is provided for collective operations
-between chares.
-
-Our Task Bench implementation is straightforward: we create a chare
-array for the task graph, with one chare for each column of the task graph. Methods are
+\emph{chares}, are objects that live in their own address space;
+chares communicate data and synchronize via messages, and can be moved
+to balance load. Our Task Bench implementation uses a chare
+array for the task graph, with one chare for each column of the task graph. Messages are
 used to implement dependencies, and a task is executed as soon as its
@@
 
-MPI~\cite{MPI} is the standard message-passing API for
-programming supercomputers and remains the most widely used programming
-model in high-performance scientific computing today. Our
-MPI implementation of Task Bench represents the common case of bulk synchronous programs with distinct computation and
-communication phases. We use \lstinline[language=C++]{MPI_Isend} and
+Our MPI~\cite{MPI} implementation of Task Bench represents the common
+case of bulk synchronous programs with distinct computation and
+communication phases. We experimented with a variety of implementation
+strategies and found the best performing to be using
+\lstinline[language=C++]{MPI_Isend} and
 \lstinline[language=C++]{MPI_Irecv} to implement the communication
 phase, posting receives before sends. Each task dependency maps to one
-send/receive pair in MPI. Despite the relatively straightforward
-design, our MPI implementation is among the best performing.
-
-For completeness, we also explored the use of additional MPI calls:
-namely buffered sends (\lstinline[language=C++]{MPI_Bsend}),
-broadcasts (\lstinline[language=C++]{MPI_Bcast}), and all-to-all
-communication (\lstinline[language=C++]{MPI_Alltoallv}). For the
-configurations used in our evaluation, we found non-blocking sends and
-receives to be the best performing, often by a wide margin.
+send/receive pair in MPI.
 
@@
 the program. This compressed representation is expanded into a full
-task graph by a source-to-source compiler. PTG is therefore a more
-limited programming model, but can potentially provide better
-performance because the graph description is amenable to static
-analysis and can be pruned efficiently for better horizontal
-scalability.
+task graph by a source-to-source compiler.
 
 In the DTD model, tasks are enumerated in program order (by
-dynamically executing the program), and dependencies between tasks are
+executing the program), and dependencies between tasks are
 identified automatically based on the input and output data of tasks. 
@@
 
-In Regent tasks are executed sequentially and parallelism is
-discovered automatically based on the data that is read or written by
-tasks. Regent programs are agnostic to the machine architecture and
-are separately \emph{mapped} for execution on a distributed memory
-system. Copies are inserted automatically by the compiler and runtime
-whenever data is read on a different node from where it is written.
-
-The Regent compiler performs control replication to slice the Regent
-program into \emph{shards} of execution which can scalably execute on
-large numbers of nodes.
-
 \subsection{Spark}
@@
 
-Spark's programming model is based on \emph{functional parallelism}, a
-generalization of the MapReduce programming
-model~\cite{MapReduce04}. The core computational abstraction in Spark
-are operators such as map, reduce, filter, join, etc. which by virtue
-of being pure functions are trivially known to be
-parallelizable. Functions in Spark operate on \emph{resilient
+The core abstractions in Spark are functional operators such as map,
+reduce, filter, join, etc. Functions in Spark operate on
+\emph{resilient
   distributed datasets} (RDDs), which are globally-visible,
@@
 optimizations therefore is the caching of RDDs to avoid unnecessary
-disk traffic, resulting in substantial speedups over MapReduce.
+disk traffic.
 
@@
 the desired dataflow. Task graphs are composed in a high-level
-language such as Python, but then serialized and run by an efficient
+language such as Python, but then serialized and run by an
 execution engine written in C++.

### 2019-03-30 7_conclusion.tex
New version of conclusion.

--- 2019-03-21 7_conclusion.tex
@@
 
-We presented Task Bench, a parameterized benchmark powerful enough to
-enable the exploration of a broad space of application scenarios, but
-also simple enough to permit implementation in a wide array of systems
-to a level of quality sufficient for comparative study. Our study of
-13 programming systems is novel in its breadth, as well as the depth
-of insight it delivers about different systems. We discover, among
-other things, that the baseline overhead of systems varies by up to 5
-orders of magnitude.
+Parameterized benchmarks are a promising new route for evaluating the
+performance of parallel and distribute programming systems. By
+composing a small number of orthogonal primitives, parameterized
+benchmarks make it possible to explore a space of application
+scenarios, while minimizing the marginal cost of developing each
+additional feature. In this case of Task Bench, this has enabled a
+comprehensive comparative study of programming system performance
+which is unique in both the breadth of systems considered and the
+depth of insight achieved about them.
 
-Helping us in this endeavor is a new metric, minimum effective task
-granularity (METG), which helps to sidestep a number of potential
-issues in the evaluation of task scheduling performance, and helps to
-separate such factors from changes in communication topology and
-latency at increasing scales of the machine.
+Our study has both confirmed and challenged our intuitions. We have
+found, unsurprisingly, that systems intended primarily for scientific
+simulation are optimized for smaller task granularities than systems
+for large-scale data analysis. The baseline overhead of these systems
+spans over 5 orders of magnitude. Even among systems intended for HPC
+there is easily 2 orders of magnitude in variation. The gap shrinks
+with node count, as the most efficient systems are the most impacted
+by increasing communication latencies, and also with numbers of
+dependencies and with increasing task parallelism. In certain cases,
+such as when task parallelism is available and compute is balanced
+with communication, asynchronous systems show a substantial benefit.
+
+Not considered in our analysis is the impact of programming system
+features on programmer productivity and performance portability. Most
+applications do not operate at the absolute extreme of runtime-limited
+performance, and thus may choose to trade overhead for better
+usability. Our study helps to quantify exactly how much a tradeoff is
+being made, so that users can be better informed, and developers can
+see the impact of features of the performance of their programming
+systems.

### 2019-03-30 1_introduction.tex
Edits and corrections.

@@
 often do not reach the level of maturity required to compare
-systems. Indeed, there are few such published studies, and none to our
-knowledge that involve more than a handful of systems.
+systems. Indeed, there are few such published studies that involve
+more than a handful of systems.
 
@@
 employ the metric \emph{tasks per second} (TPS), which is almost
-universally measured with trivial (i.e., no) dependencies. While
+universally measured with trivial (i.e., no) dependencies \cite{LegionTracing18, Canary16, Armstrong14}. While
 phrased in terms of tasks, TPS can be measured for any system as long
@@
 implementations in 13 different programming systems, including
-Chapel \cite{Chapel07}, Charm++ \cite{Charmpp93}, MPI \cite{MPI},
+Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, MPI \cite{MPI},
 OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
@@
 \item Section~\ref{sec:evaluation} provides a comprehensive evaluation
-  of Task Bench performance on up to 256 Haswell nodes of the Cori
-  supercomputer.
+  of Task Bench on up to 256 Haswell nodes of the Cori
+  supercomputer~\cite{Cori}.
 \end{itemize}

### 2019-03-30 4_implementation.tex
Edits and corrections.

@@
 
-Chapel~\cite{Chapel07} is an explicitly parallel programming language
+Chapel~\cite{Chapel15} is an explicitly parallel programming language
 that follows a \emph{multi-resolution} approach to parallel

### 2019-03-30 6_related.tex
Tweak related work.
Edits and corrections.

@@
 
-A common failure mode for mini-apps is to bite off more complexity
-than can be reasonably implemented to a level of quality sufficient
-for meaningful measurement. New mini-apps often provide only a
-reference implementation, or omit performance evaluations of different
-implementations, if those are presented. Follow-up experiments, if
-any, normally present a single system, and compare against the
-reference implementation (or not at all) and not with any other
-implementations that might be available. As a result, comparative
-studies across many implementations are rare.
+Though smaller than full applications, mini-apps can be challenging to
+implement to a level of quality sufficient for conducting comparative
+studies between programming systems. The largest study we're aware of,
+for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
+systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI). Other
+studies usually lack a comprehensive evaluation, even if multiple
+implementations are available.
 
@@
 Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.
-
-\item
-The initial report on LULESH~\cite{LULESH12} includes no performance
-results at all; follow-up papers present performance results for
-CnC~\cite{LULESHCNC15} (a self-comparison) and X10~\cite{LULESHX1015}
-(comparison against reference only).
 

### 2019-03-30 t1_systems.tex
Apply Alex's comments to Section 3.
Tweak spacing around figure captions.
Edits and corrections.

@@
 \hline
-Chapel & multi-resolution & explicit & yes & uGNI\footnote{Chapel uses GASNet to support non-Cray networks.} \\
+Chapel & multi-resolution & expl., impl. & yes & uGNI\footnote{Chapel uses GASNet to support non-Cray networks.} \\
 Charm++ & actor model & explicit & yes & uGNI\footnote{Charm++ provides additional backends for other networks.} \\
@@
 
-\caption{Systems used in Task Bench implementations.\label{tab:systems}}
+\caption{Systems for which we implemented Task Bench.\label{tab:systems}}
+\vspace{-0.5cm}
 \end{table}

### 2019-03-30 2_metg.tex
Tweaks to the paper.

@@
 4.6 us, thus the METG(50\%) of MPI is 4.6 us for this configuration of
-the application. We use 50\% because it is a reasonable level of
+Task Bench. We use 50\% because it is a reasonable level of
 efficiency in practice. Lower values would likely not be reasonable,
@@
 ideally-sloped downward lines when strong scaling. Inefficiency begins
-to show at smaller problem sizes, towards the bottom of the graph,
+to appear at smaller problem sizes, towards the bottom of the graph,
 where lines become more compressed. At the
@@
 increased communication and other resource issues as
-progressively larger portions of the machine are used.
-
-Section~\ref{sec:evaluation} contains a comprehensive evaluation of
+progressively larger portions of the machine are used. Section~\ref{sec:evaluation} contains a comprehensive evaluation of
 METG including how it changes with node count for a wide variety of

### 2019-03-30 3_task_bench.tex
Tweaks to the paper.

@@
 core library also includes implementations of the kernels, ensuring
-that the kernels are identical in all cases, eliminating a potential
+that the kernels are identical in all systems, eliminating a potential
 source of performance disparity that can be a pitfall for

### 2019-03-30 4_implementation.tex
Tweaks to the paper.

@@
 
-Chapel~\cite{Chapel15} is an explicitly parallel programming language
-that follows a \emph{multi-resolution} approach to parallel
+Chapel~\cite{Chapel15} is a programming language
+that follows a \emph{multi-res\-o\-lu\-tion} approach to parallel
 programming, with support for parallelism at a variety of
 levels. Chapel's core features are a partitioned global address space
-(PGAS), data distributions, tasks, and
+(PGAS), data distributions, tasks,
 synchronization primitives, and array promotion. For Task Bench, we decided to target a low-level implementation in
@@
 Data in PaRSEC is stored in a built-in matrix data structure with a 2D
-block cyclic distribution.  The Task Bench implementation divides this
-matrix into $M$ by $N$ tiles where $N$ equals to the width of the task
-graph and $M$ varies from 2 to the height of the task graph depending
-on optimizations performed to alleviate write-after-read dependencies.
-Tiles are distributed across nodes in pre-defined patterns,
+block cyclic distribution. Tiles of the matrix are distributed across nodes in pre-defined patterns,
 and can be passed to tasks on other nodes, but

### 2019-03-30 e6_communication.tex
Tweak captions.
More text in evaluation.
More text for evaluation section.
More text for section 5.
Tweak spacing around figure captions.
Tweaks to the paper.

@@
-\begin{figure*}[t]
+\begin{figure*}[t!]
 
@@
 
-\subfloat[4096 bytes per task dependency.\label{fig:efficiency-communication-16}]{
+\subfloat[4096 bytes per task dependency.\label{fig:efficiency-communication-4096}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_4096.pdf}
 }
-\subfloat[65536 bytes per task dependency.\label{fig:efficiency-communication-256}]{
+\subfloat[65536 bytes per task dependency.\label{fig:efficiency-communication-65536}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_65536.pdf}
@@
 
-\caption{Task granularity vs efficiency for varying communication.\label{fig:metg-communication}}
+\vspace{-0.25cm}
+\caption{Task granularity vs efficiency for varying communication (spread pattern, 5 dependencies per task, 4 task graphs, 64 nodes). Lower is better.\label{fig:efficiency-communication}}
+\vspace{-0.05cm}
 \end{figure*}

### 2019-04-01 0_abstract.tex
apply some of George's comments

@@
 13 programming systems on up to 256 Haswell nodes of the Cori
-supercomputer. We use a novel metric, \emph{minimum effective task
+supercomputer. We introduce and use a novel metric, \emph{minimum effective task
   granularity} to study the baseline runtime overhead of each system

### 2019-04-01 1_introduction.tex
apply some of George's comments

@@
 often do not reach the level of maturity required to compare
-systems. Indeed, there are few such published studies that involve
-more than a handful of systems.
+systems. Indeed, there are few such published studies~\cite{LULESH13, PRK14, NAS95}
+that involve more than a handful of systems.
 
@@
 METG(50\%) is the smallest application task granularity which can be
-executed while maintaining at least 50\% computational efficiency,
+executed while maintaining at least 50\% efficiency,
 meaning that the application achieves at least 50\% of peak
-performance on a given machine. METG captures the essence of what is important in a weak
+performance in terms of FLOPS, memory or IO bandwidth 
+on a given machine. METG captures the essence of what is important in a weak
 or strong scaling study, the behavior at the limit of scalability. For

### 2019-04-01 5_evaluation.tex
Tweaks to the paper.
apply some of George's comments

@@
 Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
-peak FLOPS achieved) plotted against task granularity. This plot is
+peak FLOPS) plotted against task granularity. This plot is
 used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
@@
 the number of nodes and thus limits scalability. Approaches based on compile-time analysis can
-partially or fully mitigate this overhead. PaRSEC PTG improves over DTD by performing DAG expansion at compile time, but
+partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
+by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
 still retains dynamic checks leading to limited scalability. Regent
@@
 scheduling behavior of the various systems. The execution of 4
-simultaneous task graphs partially, but does not fully, mitigate the
+simultaneous task graphs partially, mitigate the
 load imbalance in the tasks. Therefore, systems which provide an

### 2019-04-02 1_introduction.tex
Small tweaks and edits.

--- 2019-03-30 1_introduction.tex
@@
 often do not reach the level of maturity required to compare
-systems. Indeed, there are few such published studies that involve
-more than a handful of systems.
+systems. Indeed, there are few such published studies
+that involve more than a handful of systems~\cite{LULESH13}.
 
@@
 METG(50\%) is the smallest application task granularity which can be
-executed while maintaining at least 50\% computational efficiency,
+executed while maintaining at least 50\% efficiency,
 meaning that the application achieves at least 50\% of peak
-performance on a given machine. METG captures the essence of what is important in a weak
+performance (e.g., FLOPS) 
+on a given machine. METG captures the essence of what is important in a weak
 or strong scaling study, the behavior at the limit of scalability. For

### 2019-04-02 4_implementation.tex
apply some of George's comments
Small tweaks and edits.

@@
 
-Data in PaRSEC is stored in a built-in matrix data structure with a 2D
-block cyclic distribution. Tiles of the matrix are distributed across nodes in pre-defined patterns,
+Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
+distributed across nodes in pre-defined patterns,
 and can be passed to tasks on other nodes, but
@@
 elision of tasks on unrelated nodes to provide horizontal scalability.
-A simplified version of the 2D block cyclic matrix from the Chameleon 
-project~\cite{Chameleon} is used to describe the data produced by tasks in 
-our Task Bench implementation.
+We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
+project~\cite{Chameleon}.
 

### 2019-04-02 5_evaluation.tex
Small tweaks and edits.
rename parsec dtd shard to parsec shard

@@
 scalability~\cite{ControlReplication17}. A similar, manual
-optimization is shown in the plots as PaRSEC DTD Shard.
+optimization is shown in the plots as PaRSEC Shard.
 
@@
 scheduling behavior of the various systems. The execution of 4
-simultaneous task graphs partially, mitigate the
-load imbalance in the tasks. Therefore, systems which provide an
+simultaneous task graphs only partially mitigates the
+load imbalance between tasks. Therefore, systems which provide an
 additional on-node work stealing capability see additional gains in

### 2019-04-02 1_introduction.tex
Apply Alex's and George's comments.

--- 2019-03-30 1_introduction.tex
@@
 
-The challenge of parallel and distributed computation has lead to a
+The challenge of parallel and distributed computation has led to a
 wide variety of proposals for programming models, languages, and
-runtime systems. But while the features of these systems are widely
-written about in the literature, comprehensive performance evaluations
-of these systems remain difficult to find. Our
+runtime systems. But while these systems are well-represented in the literature, comprehensive and comparative performance evaluations
+remain difficult to find. Our
 goal in this paper is to develop a meaningful and useful framework for
@@
 
-One approach to comparing the performance of different systems has
-focused on the development of \emph{proxy-} or
+One approach to comparing the performance of different systems is through \emph{proxy-} or
 \emph{mini-apps}. Because they distill the structure of a larger
 application, mini-apps can provide insight
-without the expense of developing a full-blown
-application. However, despite the name, our experience is that
-mini-apps still requite significant investment to develop, especially
+without the expense of developing a production code. However, despite the name, our experience is that
+mini-apps still requite significant investment to develop
 to the level of quality needed for useful benchmarking. In many cases,
@@
 often do not reach the level of maturity required to compare
-systems. Indeed, there are few such published studies that involve
-more than a handful of systems.
+systems. Indeed, there are few published studies
+that involve more than a handful of systems~\cite{LULESH13}.
 
-There is another important issue: How does one predict how
+There is a second important issue: How does one predict how
 the performance of a mini-app generalizes to a full
-application, even one closely related? Part of the problem here is
-that weak and strong scaling, which are by far the most commonly
-reported metrics for performance of both mini- and full applications,
+application, even one closely related? Part of the problem is
+that \emph{weak} and \emph{strong} scaling, which are by far the most commonly
+reported performance metrics,
 do not directly characterize the performance of the underlying
@@
 scaling does not separate runtime system overhead from other costs
-(such as increased communication, among other things) when
+(such as increased communication, among many other things) when
 using progressively larger portions of a machine.
@@
 memory-bound operations of varying durations. The dependencies between
-tasks can be configured to carry payloads of varying size, permitting
+tasks can be configured to carry communication payloads of varying size, permitting
 the design of communication-bound cases. Finally, multiple
@@
 METG(50\%) is the smallest application task granularity which can be
-executed while maintaining at least 50\% computational efficiency,
+executed while maintaining at least 50\% efficiency,
 meaning that the application achieves at least 50\% of peak
-performance on a given machine. METG captures the essence of what is important in a weak
+performance (e.g., FLOPS or memory bandwidth) 
+on a given machine. METG captures the essence of what is important in a weak
 or strong scaling study, the behavior at the limit of scalability. For
@@
 
-A lower METG does not necessarily mean that performance will be
+A lower METG does not necessarily mean that performance for a particular workload will be
 better. For example, two systems with METG(50\%) of 100~us and 1~ms,
 respectively, running an application with an average task granularity
-of 10~ms, are both likely to perform well. Only when the task
-granularity approaches (or drops below) METG(50\%) is performance
-likely to diverge. Thus METG helps to identify the regime in which a
-given system is likely to perform well.
+of 10~ms, are both very likely to perform well. Only when the task
+granularity approaches (or drops below) METG(50\%) will performance
+likely diverge. Thus METG helps to identify the regime in which a
+given system can deliver good performance.
 
@@
 application throughput even as TPS increases. Only by constraining
-efficiency, as in METG, can we measure runtime-limited application
-throughput while performing useful application work.
+efficiency, as in METG, can we meaningfully measure how runtime
+overhead impacts the ability perform useful application work.
 

### 2019-04-02 2_metg.tex
Apply Alex's and George's comments.

@@
-\section{METG (50\%)}
+\section{METG}
 \label{sec:metg}
@@
 the smallest task granularity achieved while maintaining at least 0.63
-TFLOPS. Relative parallel efficiency (vs a sufficiently large problem
-size) can also be used in cases where the application isn't easily
-characterized by peak resource usage.
+TFLOPS. Absolute (rather than relative) efficiency is preferred
+because it makes it easier to conduct fair comparisons with
+programming systems that reserve one or more cores for internal use.
 
@@
 drops quickly with minimal change in efficiency. However as task
-granularity shrinks further, efficiency begins to drop as well, leading
-eventually to a plateau in task granularity. Intuitively, this plateau
+granularity shrinks further, efficiency begins to drop more rapidly, leading
+eventually to asymptote. Intuitively, this asymptote
 represents the fundamental baseline overhead of the system, the point
@@
 
-METG(50\%) is the intersection of this curve with a vertical line through
-Figure~\ref{fig:efficiency-mpi} at 50\% efficiency. In the figures, the red, dashed line shows the 50\% efficiency
-level. At 50\% efficiency, MPI achieves an average task granularity of
+% FIXME: keep this up-to-date with the graphs
+
+METG(50\%) is the intersection of this curve with 50\% efficiency, as
+shown by the red lines in Figure~\ref{fig:flops-mpi} and
+\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
+task granularity of
 4.6 us, thus the METG(50\%) of MPI is 4.6 us for this configuration of
-Task Bench. We use 50\% because it is a reasonable level of
-efficiency in practice. Lower values would likely not be reasonable,
-because operators of computing facilities prefer higher levels of
-utilization. Values above 90\% can misrepresent the performance of
-some systems (see
+Task Bench. We use 50\% because that is generally an acceptable level
+of efficiency in practice, and values above 90\% can misrepresent the
+performance of some systems (see
 Section~\ref{subsec:peak-performance-and-efficiency}).
 
-METG is an appealing metric in part because it has a well-defined
+METG has a well-defined
 relationship with quantities of interest to application developers,
@@
 \ref{fig:strong-scaling-mpi} show the behavior of the MPI Task Bench stencil's weak and strong
-scaling, respectively, at a variety of problem sizes. Intuitively, at
+scaling, respectively, at a variety of problem sizes. In these
+figures, the vertical axis is shown as wall time to emphasize the
+relationship to time-to-solution, but it could equivalently be shown
+as task granularity (as the number of tasks per execution is
+fixed). Intuitively, at
 larger problem sizes MPI is perfectly efficient. This can be seen at
@@
 very bottom, the lines compress together as running time becomes dominated by overhead. The shape of the curve at the bottom is the same
-for both weak and strong scaling; this is the shape of
+for both weak and strong scaling; this is
 the METG curve.
@@
 size that can be weak scaled to a given node count with a given level
-of efficiency. With any smaller problem size, 
-the runtime overhead begins to dominate useful
+of efficiency. With smaller problem sizes runtime overhead increasingly dominates useful
 computation. Similarly, METG corresponds to the point at which strong

### 2019-04-02 3_task_bench.tex
Apply Alex's and George's comments.

@@
 graphs that can be implemented with Task Bench. Note that layout is
-significant---in particular, each column $i$ corresponds to all tasks
+significant---in particular, column $i$ represents all tasks
 with index $i$ in the iteration space over all the time
@@
 Despite its generality, Task Bench is easy to implement, making it
-tractable to develop a suite of high-quality implementations. The core aspects of Task Bench, such as generating
+tractable to develop a suite of high-quality implementations. The central aspects of Task Bench, such as generating
 task graphs and enumerating dependencies, are encapsulated in a core
@@
 parsing of input parameters and the display of output results,
-ensuring that all implementations behave in a uniform manner, and can
+ensuring that all implementations behave in a uniform manner and can
 be scripted consistently. Because much of the functionality needed for
@@
 fails. This level of self-checking ensures that every execution of Task Bench, if it
-completes successfully, is correct. An evaluation of the impact of
+completes successfully, is correct. An evaluation of the performance impact of
 validation showed it to be less than 3\% at the smallest task

### 2019-04-02 4_implementation.tex
Apply Alex's and George's comments.

@@
 support for heterogeneous processors such as GPUs, but the support
-varies so widely (if it exists at all) that it is challenging to come
-up with a representative benchmark that allows these features to be
-compared in an apples-to-apples manner.
+varies so widely (if it exists at all) that it seems difficult to
+compare systems with direct GPU support to those wihtout in an
+apples-to-apples manner.
 
@@
 
-Chapel~\cite{Chapel15} is a programming language
-that follows a \emph{multi-res\-o\-lu\-tion} approach to parallel
-programming, with support for parallelism at a variety of
+Chapel~\cite{Chapel15} is a parallel programming language
+with a \emph{multi-res\-o\-lu\-tion} approach, supporting parallelism at a variety of
 levels. Chapel's core features are a partitioned global address space
 (PGAS), data distributions, tasks,
-synchronization primitives, and array promotion. For Task Bench, we decided to target a low-level implementation in
+synchronization primitives, and array promotion. For Task Bench, we target a low level of
 Chapel, using explicit task instantiation (via
@@
 
-Charm++~\cite{Charmpp93} is an explicitly parallel programming system
-based on the actor model. Actors, or
-\emph{chares}, are objects that live in their own address space;
-chares communicate data and synchronize via messages, and can be moved
+Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors, or
+\emph{chares}, are objects in their own address space.
+Chares communicate data and synchronize via messages and can be moved
 to balance load. Our Task Bench implementation uses a chare
-array for the task graph, with one chare for each column of the task graph. Messages are
-used to implement dependencies, and a task is executed as soon as its
-dependencies are all available. Collectives are used to synchronize the
+array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
+dependencies are all available. Collectives synchronize the
 completion of the task graph execution.
@@
 
+% FIXME: bulk synchronous or point-to-point
+
 Our MPI~\cite{MPI} implementation of Task Bench represents the common
-case of bulk synchronous programs with distinct computation and
+case of point-to-point communication with distinct computation and
 communication phases. We experimented with a variety of implementation
@@
 implementations of OpenMP. We found KMP to provide better performance
-and thus report KMP numbers in our evaluation.
+and report KMP numbers in our evaluation.
 
@@
 and the data is copied automatically if the two tasks are executed on
-different nodes. While the basic DTD model is implicitly parallel
-because the system is responsible for the discovery of parallelism, to
-achieve horizontal scalability, the program is executed in a
+different nodes. To
+improve horizontal scalability, the program is executed in
 SPMD fashion on all nodes, and the user is responsible for eliding
@@
 model used internally by Legion~\cite{Legion12} and
-Regent~\cite{Regent15}. Thus the Realm implementation of Task Bench can be
-seen as a limit study of what can be achieved with Legion or Regent.
+Regent~\cite{Regent15}. A Realm implementation of Task Bench is a limit study of what can be achieved with Legion or Regent.
 
@@
 
-Realm also supports a subgraph API where sets of tasks and copies are
-issued together. This mode permits additional optimizations to be
-performed. We have implemented Task Bench in both individual and
-subgraph styles, but report numbers for the subgraph version as it
-provides better performance.
+Realm also supports a \emph{subgraph} API that enables additional
+optimizations. We report numbers for the subgraph version of Task
+Bench as it provides better performance.
 
@@
   distributed datasets} (RDDs), which are globally-visible,
-dynamically single-assignment data structures that are cached in memory
-and written to disk for durability. One of Spark's primary
-optimizations therefore is the caching of RDDs to avoid unnecessary
+dynamically single-assignment data structures. One of Spark's primary
+optimizations is the caching of RDDs to avoid unnecessary
 disk traffic.
@@
 so one of the challenges in developing a Task Bench implementation is
-mapping the task graph to a set of operators which will result in the
+mapping the task graph to a set of operators that result in the
 desired performance behavior. We use a combination of
@@
 Spark does not attempt to group multiple Task Bench tasks into a
-single Spark task.
+single Spark task, as the tasks in Task Bench already represent coarse-grained units of work.
 
-We performed an extensive set of experiments to verify that no
+We performed extensive experiments to verify that no
 extraneous factors interfered with our Spark measurements. We
-completely disabled logging, ensured that no RDDs are actually written
-to disk, confirmed that there is no measurable overhead due to JNI
-calls from Java to C, and cross-checked our results with known cases
+completely disabled logging, ensured no RDDs are written
+to disk, confirmed there is no measurable overhead due to JNI
+calls from Java to C, and cross-checked results with known cases
 that hit optimal task throughput in Spark, among other things.
@@
 Swift/T~\cite{Wozniak13} is a parallel scripting language intended
-primarily for the composition of workflows in high-performance
-scientific computing. Tasks in Swift/T can be written in any
+primarily for the composition of HPC workflows. Tasks in Swift/T can be written in any
 programming language, and may even be themselves parallel. Swift/T
@@
 
-TensorFlow~\cite{TensorFlow15} is a programming system intended for
-the executing deep learning workloads, widely used in industrial data
-center applications. Although TensorFlow's high-level API exposes
-concepts specific to machine learning, internally TensorFlow is
-implemented as a task graph execution engine, making it a good fit for
+TensorFlow~\cite{TensorFlow15} is a programming system designed for
+deep learning workloads. Although TensorFlow's API exposes
+machine learning concepts, internally TensorFlow is a task graph execution engine, making it a good fit for
 Task Bench. TensorFlow programs are written by creating nodes in the

### 2019-04-02 5_evaluation.tex
Apply Alex's and George's comments.

@@
 
-We present a comprehensive evaluation of Task Bench on up to 256
+We present a comprehensive evaluation of our Task Bench implementations on up to 256
 Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
@@
 \label{subsec:peak-performance-and-efficiency}
-
-\input{e1_flops}
-\input{e2_efficiency}
-
-% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
-\input{e3_metg_compute}
 
@@
 kernel with varying problem sizes. This is the full version of
-Figure~\ref{fig:flops-mpi}. In the best case, we measure peak FLOPS of
+Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOPS of
 $1.26 \times 10^{12}$, which compares favorably with the officially
@@
 the baseline for 100\% efficiency.
+
+\input{e1_flops}
+\input{e2_efficiency}
+
+% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
+\input{e3_metg_compute}
 
@@
 
-In most cases, task granularity has plateaued prior to the 50\% efficiency point,
+In most cases, task granularity asymptotes prior to the 50\% efficiency point,
 though some systems continue to improve as efficiency continues
 to drop. Accounting for this effect is one of the main arguments in
-favor of METG with a reasonable efficiency threshold compared to
-measuring task scheduling throughput in the presence of empty tasks
-(effectively METG(0\%)). Empty tasks permit runtime
-systems to make unrealistic tradeoffs which are not applicable to real
-application workloads where some degree of efficiency is necessarily a
-goal. Reporting numbers with empty tasks therefore gives unfair
-advantage to systems that make such unrealistic tradeoffs.
+favor of METG with a reasonable efficiency threshold instead of
+measuring task scheduling throughput of empty tasks
+(effectively METG(0\%)). Measuring performance using empty tasks can
+reward implementation strategies, such as devoting nearly 100\% of
+system resources to the runtime system, that make no sense for real
+applications.
 
@@
 One of the basic questions when considering different programming
-systems is how much overhead the system introduces. This can be a
-tricky question to answer directly because some systems introduce
+systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
 overhead \emph{inline} (i.e., by running system internal processes on
@@
 To answer this question, we use the METG metric to determine the
-smallest task granularity which can be executed at a given level of
-efficiency, as a proxy for overhead. Figure~\ref{fig:metg-compute}
+smallest task granularity that can be executed at a given level of
+efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
 shows how METG(50\%) varies with node count for a variety of
@@
 We observe in the results that the baseline overheads of different
-systems can vary by over 5 orders of magnitude. It is worth
+systems vary by over 5 orders of magnitude. It is worth
 remembering when considering this metric that this is a \emph{minimum}
@@
 
-\subsection{Scalability}
-\label{subsec:scalability}
-
-METG is a useful metric in part because it summarizes the basic task
-scheduling overhead of a system in a single number. This makes it
-possible to evaluate METG at different node counts (shown in
-Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
-communication topology and latency.
-
 In general, the least complicated pattern (stencil) is most favorable
@@
 and simply executes tasks one after another in alternation with
-communication phases. In contrast the asynchronous execution
+communication phases. In contrast, the asynchronous execution
 mechanisms of the other systems are pure overhead in this scenario.
@@
 communication pattern grows, and even reverses as task parallelism is
-added in the form of multiple task graphs.
+added in the form of multiple task graphs. Some of the slower systems
+are omitted from comparisons with more complicated dependence
+patterns, as they failed to complete in a reasonable time.
+
+\subsection{Scalability}
+\label{subsec:scalability}
+
+METG is a useful metric in part because it summarizes the basic task
+scheduling overhead of each programming system in a single number. This makes it
+possible to evaluate METG at different node counts (shown in
+Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
+communication topology and latency.
 
@@
 scalability~\cite{ControlReplication17}. A similar, manual
-optimization is shown in the plots as PaRSEC Shard.
+optimization is shown in the plots as PaRSEC DTD Shard.
 
@@
 dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
-$2.1\times$). The difference is most pronounced in systems which
+$2.1\times$). The difference is most pronounced in systems that
 perform runtime work inline. For example, MPI achieves an METG of 390
@@
 pseudo random number generator with a deterministic seed to ensure
-that the task durations under all systems are identical.
+identical task durations for all systems.
 
@@
 simultaneous task graphs only partially mitigates the
-load imbalance between tasks. Therefore, systems which provide an
+load imbalance between tasks. Therefore, systems that provide an
 additional on-node work stealing capability see additional gains in
@@
 granularities. This can be seen in the difference between Chapel
-Distrib (which uses a work stealing scheduler) vs Chapel with its
+distrib (which uses a work stealing scheduler) vs Chapel with its
 default (non-work-stealing) scheduler.
@@
 these plots. First, by overlapping communication with computation,
-such systems are able to execute smaller task granularities at higher
+such systems execute smaller task granularities at higher
 levels of efficiency compared to the bulk synchronous MPI

### 2019-04-02 7_conclusion.tex
Apply Alex's and George's comments.

@@
 Parameterized benchmarks are a promising new route for evaluating the
-performance of parallel and distribute programming systems. By
+performance of parallel and distributed programming systems. By
 composing a small number of orthogonal primitives, parameterized
@@
 scenarios, while minimizing the marginal cost of developing each
-additional feature. In this case of Task Bench, this has enabled a
+additional feature. In the case of Task Bench, this has enabled a
 comprehensive comparative study of programming system performance
@@
 performance, and thus may choose to trade overhead for better
-usability. Our study helps to quantify exactly how much a tradeoff is
-being made, so that users can be better informed, and developers can
-see the impact of features of the performance of their programming
+usability. Our study helps to quantify the performance side of that tradeoff so that users can be better informed and developers can
+see the impact of features on the performance of their programming
 systems.

### 2019-04-02 2_metg.tex
Small tweaks to make text fit better.

@@
 Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
-run on a single Haswell node of Cori with a problem size large enough to
-ensure that it achieves peak FLOPS. This starting point confirms
-that the application is properly configured and that the choice of
+run on a single Haswell node of Cori~\cite{Cori} with a problem size large enough to
+ensure that it achieves peak FLOPS. This confirms
+that the application is properly configured and that the
 efficiency metric is achievable. The problem
@@
 and thus it makes sense to discuss tasks even in systems
-without an explicit notion of tasking, such as in MPI programs that
-are written in a bulk synchronous style. In this case, the tasks run a
+without an explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
 compute-bound kernel included in the Task Bench implementation,

### 2019-04-02 3_task_bench.tex
Small tweaks to make text fit better.

@@
 with index $i$ in the iteration space over all the time
-steps. Generally speaking each column of the task graph will be
+steps. Generally speaking each column will be
 assigned to execute on a different processor core.
@@
 relation. The
-dependence relation enumerates, for each task, the set of tasks in the
-previous time step that task depends on. This permits a wide variety
+dependence relation enumerates the set of tasks from the
+previous time step each task depends on. This permits a wide variety
 of dependence patterns to be implemented that are relevant to real
 applications in high-performance scientific computing and data analysis: stencils,
-sweeps, FFTs, trees, and so on. Dependence patterns may also be
+sweeps, FFTs, trees, etc. Dependence patterns may also be
 parameterized, such as picking the $K$ nearest neighbors, or $K$

### 2019-04-02 4_implementation.tex
Small tweaks to make text fit better.

@@
 that is source-compatible with OpenMP. Our Task Bench implementation
-uses OpenMP 4.0-style tasks with task dependencies. Because tasks in
-OpenMP 4.0 have a fixed number of dependencies, we use a switch
+uses OpenMP 4.0-style task dependencies. Because OpenMP tasks have a fixed number of dependencies, we use a switch
 statement to implement the dynamic dependencies required for Task
@@
 for OpenMP is very similar to OmpSs and uses tasks with
-task dependencies.
-
-In our evaluation, we test with both GNU GOMP and Intel KMP
-implementations of OpenMP. We found KMP to provide better performance
-and report KMP numbers in our evaluation.
+task dependencies. We tested both GNU GOMP and Intel KMP
+implementations of OpenMP and found KMP to be better performing.
 

### 2019-04-02 5_evaluation.tex
Fix typo.
Flip sections around.

@@
 
-% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
-\input{e3_metg_compute}
-
 The vast majority of systems achieve or nearly achieve peak FLOPS. Some
@@
 
+\subsection{Number of Dependencies}
+\label{subsec:number-of-dependencies}
+
+\input{e4_radix}
+
+The number of dependencies per task has a strong influence on the
+overhead incurred by each system, as shown in
+Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
+dependence pattern, when varying the number of dependencies per task
+from 0 to 9.
+
+% FIXME: make sure these numbers are up to date for final paper
+
+The ratio in METG between 0 and 3 dependencies per task ranges from
+$1.3\times$ to $59\times$ (median $3.0\times$). From 1 to 3
+dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
+$2.1\times$). The difference is most pronounced in systems that
+perform runtime work inline. For example, MPI achieves an METG of 390
+ns with 0 dependencies, but this rises to 4.6 us with 3 dependencies,
+a factor of $12\times$ increase. This is unsurprising as in the case
+of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
+issued at all, so MPI has nothing to do aside from executing kernels
+as quickly as possible.
+
+For this reason, we recommend evaluating the performance of
+programming systems under realistic workloads, rather than with
+trivial dependencies.
+
 \subsection{Baseline Overhead}
+
+% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
+\input{e3_metg_compute}
 
@@
 trimming approach still requires some dynamic checks that scale with
-the number of nodes and thus limits scalability. Approaches based on compile-time analysis can
+the number of nodes and thus limits scalability~\cite{PARSEC_DTD}. Approaches based on compile-time analysis can
 partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
@@
 scalability~\cite{ControlReplication17}. A similar, manual
-optimization is shown in the plots as PaRSEC DTD Shard.
+optimization is shown in the plots as PaRSEC shard.
 
@@
 % FIXME: talk about overhead spanning 5 orders of magnitude
-
-\subsection{Number of Dependencies}
-\label{subsec:number-of-dependencies}
-
-\input{e4_radix}
-
-The number of dependencies per task has a strong influence on the
-overhead incurred by each system, as shown in
-Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
-dependence pattern, when varying the number of dependencies per task
-from 0 to 9.
-
-% FIXME: make sure these numbers are up to date for final paper
-
-The ratio in METG between 0 and 3 dependencies per task ranges from
-$1.3\times$ to $59\times$ (median $3.0\times$). From 1 to 3
-dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
-$2.1\times$). The difference is most pronounced in systems that
-perform runtime work inline. For example, MPI achieves an METG of 390
-ns with 0 dependencies, but this rises to 4.6 us with 3 dependencies,
-a factor of $12\times$ increase. This is unsurprising as in the case
-of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
-issued at all, so MPI has nothing to do aside from executing kernels
-as quickly as possible.
-
-For this reason, we recommend evaluating the performance of
-programming systems under realistic workloads, rather than with
-trivial dependencies.
 

### 2019-04-02 6_related.tex
Tweaks to the paper.
Add Conceptual to related work.

@@
 commonly evaluated using proxy- or mini-apps, or
-microbenchmarks. Mini-apps are explicitly derived from larger
+micro\-benchmarks. Mini-apps are explicitly derived from larger
 applications and therefore have the advantage of bearing some
@@
 
+\textsc{coNCePTuaL}~\cite{Conceptual07} is a domain-specific language
+for writing network performance tests. \textsc{coNCePTuaL} and Task
+Bench both enable the easy creation of new benchmarks though
+\textsc{coNCePTuaL} does so via scripting whereas Task Bench provides
+a set of configurable parameters. \textsc{coNCePTuaL} also targets a
+lower level of abstraction, optimized more for testing messaging
+layers, whereas Task Bench is closer to application level and
+therefore enables comparisons of a broader set of parallel and
+distributed programming systems.
+
 Limit studies of task scheduling throughput in various runtime systems

### 2019-04-02 1_introduction.tex
Small tweaks.

@@
 wide variety of proposals for programming models, languages, and
-runtime systems. But while these systems are well-represented in the literature, comprehensive and comparative performance evaluations
+runtime systems. While these systems are well-represented in the literature, comprehensive and comparative performance evaluations
 remain difficult to find. Our
@@
 without the expense of developing a production code. However, despite the name, our experience is that
-mini-apps still requite significant investment to develop
+mini-apps still require significant investment to develop
 to the level of quality needed for useful benchmarking. In many cases,
@@
 often do not reach the level of maturity required to compare
-systems. Indeed, there are few published studies
-that involve more than a handful of systems~\cite{LULESH13}.
+systems. Few published studies involve more than a handful of systems~\cite{LULESH13}.
 
@@
 application, even one closely related? Part of the problem is
-that \emph{weak} and \emph{strong} scaling, which are by far the most commonly
+that \emph{weak} and \emph{strong} scaling, by far the most commonly
 reported performance metrics,
@@
 
-To address the second issue and isolate the contribution of the programming model's runtime system
-to application performance, we introduce a new metric called
-\emph{minimum effective task granularity} (METG). Intuitively,
-METG(50\%) is the smallest application task granularity which can be
-executed while maintaining at least 50\% efficiency,
-meaning that the application achieves at least 50\% of peak
-performance (e.g., FLOPS or memory bandwidth) 
-on a given machine. METG captures the essence of what is important in a weak
+To address the second issue and isolate the contribution of the programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG).
+Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
+performance (e.g., FLOPS or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
 or strong scaling study, the behavior at the limit of scalability. For
-weak scaling, METG(50\%) corresponds directly to the smallest problem
-size that can be weak scaled with 50\% efficiency. For strong scaling,
-METG(50\%) can be used to compute the scale at which performance can
-be expected to dip below 50\% efficiency.
+weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency.
+For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.
 
-A lower METG does not necessarily mean that performance for a particular workload will be
+A lower METG does not necessarily mean that performance for a particular workload is
 better. For example, two systems with METG(50\%) of 100~us and 1~ms,
@@
 cost of non-trivial dependencies can be significant and may impact
-scalability. This issue is easily fixed by running non-trivial
+scalability. This issue can be easily fixed by running non-trivial
 configurations of Task Bench.

### 2019-04-02 1_introduction.tex
A few more small edits and rewording to deal with some passive voice.

@@
 empty tasks~\cite{LegionTracing18} or with tasks of some
-duration~\cite{Canary16, Armstrong14}. If empty tasks are used, the
+duration~\cite{Canary16, Armstrong14}. When using empty tasks, the
 resulting upper bound on task scheduling throughput fails to represent
-any realistic application, as no
-useful work is being performed. If non-empty tasks are used, \emph{but
-  the efficiency of the overall application is not reported}, then TPS
-is not a measurement of runtime-limited performance. Large tasks may
-be used to hide any amount of runtime overhead,
-while tasks that are too small may result in a drop in total
+useful work within a realistic application. With non-empty tasks,
+\emph{where the efficiency of the overall application is not
+ reported}, TPS is not a measurement of runtime-limited
+performance. Large tasks may be used to hide any amount of runtime
+overhead, while tasks that are too small may result in a drop in total
 application throughput even as TPS increases. Only by constraining
 efficiency, as in METG, can we meaningfully measure how runtime
-overhead impacts the ability perform useful application work.
+overhead impacts the ability to perform useful application work.
 

### 2019-04-05 5_evaluation.tex
Code for memory graph (currently commented out).
Text for memory-bound kernels.

@@
 
-\subsection{Peak Performance and Efficiency}
+\subsection{Compute-Bound Performance}
 \label{subsec:peak-performance-and-efficiency}
@@
 
+\subsection{Memory-Bound Performance}
+
+\input{e7_bytes}
+
+Figure~\ref{fig:bytes} shows the performance of Task Bench
+implementations with a memory-bound kernel. We measure peak memory
+bandwidth of 108 GB/s, which compares favorably with the
+OpenMP-enabled STREAM benchmarks~\cite{STREAM} which report betweeen
+79 and 90 GB/s.
+
+Not all cores are required to saturate memory bandwidth. As a result,
+systems that reserve one or more cores are able to hit 100\% of peak
+performance, unlike with compute-bound kernels. Otherwise the
+performance curves are similar. We are currently investigating the
+performance of Charm++.
+
+The remaining experiments use compute-bound kernels.
+
 \subsection{Number of Dependencies}
@@
 issued at all, so MPI has nothing to do aside from executing kernels
-as quickly as possible.
-
-For this reason, we recommend evaluating the performance of
+as quickly as possible. For this reason, we recommend evaluating the performance of
 programming systems under realistic workloads, rather than with

### 2019-04-05 5_evaluation.tex
Tweak text.

@@
 implementations with a memory-bound kernel. We measure peak memory
-bandwidth of 108 GB/s, which compares favorably with the
-OpenMP-enabled STREAM benchmarks~\cite{STREAM} which report betweeen
-79 and 90 GB/s.
+bandwidth of 108~GB/s, which compares favorably with the
+OpenMP-enabled STREAM benchmarks~\cite{STREAM} which report up to 98~GB/s.
 

### 2019-04-05 7_conclusion.tex
Edit conclusion.

@@
 
-Parameterized benchmarks are a promising new route for evaluating the
-performance of parallel and distributed programming systems. By
-composing a small number of orthogonal primitives, parameterized
-benchmarks make it possible to explore a space of application
-scenarios, while minimizing the marginal cost of developing each
-additional feature. In the case of Task Bench, this has enabled a
-comprehensive comparative study of programming system performance
-which is unique in both the breadth of systems considered and the
-depth of insight achieved about them.
-
-Our study has both confirmed and challenged our intuitions. We have
-found, unsurprisingly, that systems intended primarily for scientific
-simulation are optimized for smaller task granularities than systems
-for large-scale data analysis. The baseline overhead of these systems
-spans over 5 orders of magnitude. Even among systems intended for HPC
-there is easily 2 orders of magnitude in variation. The gap shrinks
-with node count, as the most efficient systems are the most impacted
-by increasing communication latencies, and also with numbers of
-dependencies and with increasing task parallelism. In certain cases,
-such as when task parallelism is available and compute is balanced
-with communication, asynchronous systems show a substantial benefit.
+Task Bench is a promising new approach for evaluating the performance
+of parallel and distributed programming systems. We have seen that
+systems with support for asynchronous (and/or task-based) programming
+appear to show benefits with more complicated dependency patterns,
+higher node counts (because of higher communication latencies) and
+more communication (because of ability to overlap it with
+computation). However, all these benefits can be eclipsed if baseline
+runtime overheads are too high. The current generation of task-based
+systems (PaRSEC, Realm, Regent, and StarPU) show higher overheads
+than traditional HPC systems, indicating that there continues to be
+work to be done in this area.
 
@@
 performance, and thus may choose to trade overhead for better
-usability. Our study helps to quantify the performance side of that tradeoff so that users can be better informed and developers can
-see the impact of features on the performance of their programming
+usability. Our study helps to quantify the performance side of that
+tradeoff so that users can be better informed and developers can see
+the impact of features on the performance of their programming
 systems.

### 2019-04-05 0_abstract.tex
Small tweaks and edits.
Small tweak.
Tweak abstract.
Fiddle with spacing and text to make things fit again.

@@
 \begin{abstract}
-% 126 words
+% 128 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
@@
 possible, with the ability to model a wide range of application
-scenarios, giving it coverage approximating large benchmarks.
+scenarios, giving it coverage approximating mini-apps.
 
@@
 and the impact of different dependence patterns. We also
-study each system's scalability, and ability to hide communication and
+study each system's scalability, ability to hide communication and
 mitigate load imbalance.

### 2019-04-05 2_metg.tex
Fiddle with spacing and text to make things fit again.

@@
 the smallest average task granularity (i.e., task duration) such that $A$
-achieves overall efficiency exceeding 50\%. For example, in
-a compute-bound application efficiency might be measured as the
-percentage of the available FLOPS achieved. On a machine with a peak
-capability of 1.26 TFLOPS per node, METG(50\%) would correspond to
+achieves overall efficiency of at least 50\%. For example, in
+compute-bound applications efficiency can be measured as the
+percentage of the available FLOPS achieved. On Cori with 1.26 TFLOPS available per Haswell node, METG(50\%) corresponds to
 the smallest task granularity achieved while maintaining at least 0.63
@@
 Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
-run on a single Haswell node of Cori~\cite{Cori} with a problem size large enough to
+run on a single Haswell node of Cori with a problem size large enough to
 ensure that it achieves peak FLOPS. This confirms
@@
 In Figure~\ref{fig:efficiency-mpi}, we see that
-efficiency starts near 100\%, and that initially task granularity
-drops quickly with minimal change in efficiency. However as task
+efficiency starts at 100\%. Initially task granularity
+shrinks with minimal change in efficiency. However as task
 granularity shrinks further, efficiency begins to drop more rapidly, leading
 eventually to asymptote. Intuitively, this asymptote
-represents the fundamental baseline overhead of the system, the point
-below which the system simply cannot execute tasks efficiently.
+represents the fundamental baseline overhead of the system.
 
@@
 Section~\ref{subsec:peak-performance-and-efficiency}).
+
+\input{f3_weak_scaling_mpi}
+\input{f4_strong_scaling_mpi}
 
@@
 the METG curve.
-
-\input{f3_weak_scaling_mpi}
-\input{f4_strong_scaling_mpi}
 

### 2019-04-05 4_implementation.tex
Fiddle with spacing and text to make things fit again.

@@
 varies so widely (if it exists at all) that it seems difficult to
-compare systems with direct GPU support to those wihtout in an
+compare systems with direct GPU support to those without in an
 apples-to-apples manner.
@@
 primarily for the composition of HPC workflows. Tasks in Swift/T can be written in any
-programming language, and may even be themselves parallel. Swift/T
+programming language. Swift/T
 programs follow dataflow semantics, where every statement may

### 2019-04-05 5_evaluation.tex
Fiddle with spacing and text to make things fit again.

@@
 processors (a total of 32 physical cores per node), 128 GB RAM, and a
-Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
+Cray Aries interconnect. We use GCC 7.3.0 for all Task Bench
 implementations, and (where applicable) the system default MPI
-implementation, Cray MPICH 7.7.3. Additional versions and flags for the
+implementation, Cray MPICH 7.7.3. Versions and flags for the
 various systems are shown in Table~\ref{tab:flags}.
@@
 changes in communication latencies and topology when using
-progressively larger portions of the machine. We consider the
-following configurations of Task Bench:
+progressively larger portions of the machine.
 
-\begin{itemize}
-\item Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each
-  task depends on 3 other tasks (including the same point in the
-  previous timestep).
-\item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
-  task depends on 5 others, chosen to be as close as possible.
-\item Figure~\ref{fig:metg-compute-spread} is a pattern where each
-  task depends on 5 others, spread as widely as possible.
-\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical
-  copies of the nearest dependence pattern executing concurrently.
-\end{itemize}
+We consider the following configurations of Task Bench:
+Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each task
+depends on 3 other tasks (including the same point in the previous
+timestep). Figure~\ref{fig:metg-compute-nearest} is a pattern where
+each task depends on 5 others, chosen to be as close as
+possible. Figure~\ref{fig:metg-compute-spread} is a pattern where each
+task depends on 5 others, spread as widely as possible. And
+Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical copies of
+the nearest dependence pattern executing concurrently.
 
@@
 task granularity of \emph{at least} this value can usually be expected
-to execute efficiently. Typical average task granularity will
+to execute efficiently. Typical task granularities will
 generally be determined by the application domain being
-considered. Most notably, for data analytics workloads being run in
-industrial data centers, the higher METG values observed for Spark and
+considered. Most notably, for large-scale data analytics workloads, the higher METG values observed for Spark and
 TensorFlow are sufficient. In contrast, for high-performance
@@
 
-In general, the least complicated pattern (stencil) is most favorable
+The least complicated pattern (stencil) is most favorable
 to MPI, as it represents a truly bulk synchronous task graph with no
@@
 are omitted from comparisons with more complicated dependence
-patterns, as they failed to complete in a reasonable time.
+patterns, as they did not complete in reasonable time.
 
@@
 work-stealing queues can also impact throughput at very small task
-granularities. This can be seen in the difference between Chapel
-distrib (which uses a work stealing scheduler) vs Chapel with its
-default (non-work-stealing) scheduler.
+granularities. This can be seen in the difference between the Chapel
+distrib (work stealing) vs default (non-work-stealing) schedulers.
 
@@
 hide communication latency in the presence of task
-parallelism. Figure~\ref{fig:efficiency-communication} plots task
-granularity vs efficiency curves for varying amounts of
-communication. The degree of communication is configured by the number
+parallelism. Figure~\ref{fig:efficiency-communication} plots efficiency vs task
+granularity curves for varying amounts of
+communication, determined by the number
 of bytes produced by each task (and therefore communicated with each

### 2019-04-05 1_introduction.tex
Respond to Manolis's comments.

@@
 to the level of quality needed for useful benchmarking. In many cases,
-the effort to tune for performance exceeds the effort to develop the
-mini-app in the first place. As a result, implementations of mini-apps
+the effort to tune for performance exceeds the effort to develop a correct implementation. As a result, implementations of mini-apps
 often do not reach the level of maturity required to compare
@@
 A lower METG does not necessarily mean that performance for a particular workload is
-better. For example, two systems with METG(50\%) of 100~us and 1~ms,
+better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
 respectively, running an application with an average task granularity
@@
 to completion without interruption. TPS is an upper bound on
-runtime-limited application throughput, but not a tight bound, as the
-cost of non-trivial dependencies can be significant and may impact
-scalability. This issue can be easily fixed by running non-trivial
+runtime-limited application throughput. But it is not a tight bound, as the
+cost of non-trivial dependencies can be significant. This issue can be easily fixed by running non-trivial
 configurations of Task Bench.
@@
 
-The paper is organized as follows:
+% The paper is organized as follows:
+This paper makes the following contributions:
+
+% FIXME: if you need more space you can collapsed the itemize:
 

### 2019-04-05 2_metg.tex
Respond to Manolis's comments.

@@
 granularity shrinks further, efficiency begins to drop more rapidly, leading
-eventually to asymptote. Intuitively, this asymptote
-represents the fundamental baseline overhead of the system.
+eventually to an asymptote as overhead dominates useful work.
 
@@
 task granularity of
-4.6 us, thus the METG(50\%) of MPI is 4.6 us for this configuration of
+4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s for this configuration of
 Task Bench. We use 50\% because that is generally an acceptable level
@@
 namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
-\ref{fig:strong-scaling-mpi} show the behavior of the MPI Task Bench stencil's weak and strong
-scaling, respectively, at a variety of problem sizes. In these
+\ref{fig:strong-scaling-mpi} show the weak and strong scaling behavior of the MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
 figures, the vertical axis is shown as wall time to emphasize the

### 2019-04-05 5_evaluation.tex
Respond to Manolis's comments.

@@
 perform runtime work inline. For example, MPI achieves an METG of 390
-ns with 0 dependencies, but this rises to 4.6 us with 3 dependencies,
+ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
 a factor of $12\times$ increase. This is unsurprising as in the case
@@
 controller, which limits throughput, and this is visible in the figure
-as line for Spark immediately rises with node count. Keep in mind also
+as the line for Spark immediately rises with node count. Keep in mind also
 that Spark is being evaluated here with a non-trivial dependence
@@
 trimming approach still requires some dynamic checks that scale with
-the number of nodes and thus limits scalability~\cite{PARSEC_DTD}. Approaches based on compile-time analysis can
+the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Approaches based on compile-time analysis can
 partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 

### 2019-04-05 0_abstract.tex
Tweak abstract.

@@
 programming systems under a variety of application scenarios. Task
-Bench seeks to provide the ease of implementation and portability of
+Bench seeks to combine the ease of implementation and portability of
 microbenchmarks, which makes comparisons between many systems
 possible, with the ability to model a wide range of application
-scenarios, giving it coverage approximating mini-apps.
+scenarios that distill the key characteristics of larger applications.
 

### 2019-04-08 3_task_bench.tex
Fiddle with spacing and text to make things fit again.
apply some of Yuankun's comments

@@
 
-\input{f5_task_graphs}
-\input{t0_equations}
-
 Despite its generality, Task Bench is easy to implement, making it
@@
 to 1500 lines, with several hundred lines being typical.
+
+\input{f5_task_graphs}
+\input{t0_equations}
 
@@
 kernel is carefully written to keep the working set size constant as
-the number of iterations decrease, to avoid unwanted speedups due to
+the number of iterations decreases, to avoid unwanted speedups due to
 cache effects.

### 2019-04-08 1_introduction.tex
Apply Sean's comments.

@@
 comparing the performance of parallel and distributed programming
-systems.
-
-% note: removed definition of task
+systems, to help users and developers evaluate the performance tradeoffs of these systems.
 
@@
 system overhead with sufficiently large problem sizes, and strong
-scaling does not separate runtime system overhead from other costs
-(such as increased communication, among many other things) when
+scaling does not separate runtime system overhead from costs
+(such as increased communication) that scale with the number of nodes when
 using progressively larger portions of a machine.
@@
 as a set of \emph{tasks}, or coarse-grain units of work, with
-dependencies between tasks representing the communication required for
+dependencies between tasks representing the communication and synchronization required for
 distributed execution. Task Bench consists of a configurable \emph{task graph} with tasks for
@@
 We conduct a comprehensive comparative study of Task Bench
-implementations in 13 different programming systems, including
+implementations in 13 different programming systems:
 Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, MPI \cite{MPI},

### 2019-04-08 2_metg.tex
Apply Sean's comments.

@@
 TFLOPS. Absolute (rather than relative) efficiency is preferred
-because it makes it easier to conduct fair comparisons with
-programming systems that reserve one or more cores for internal use.
+because it better quantifies the performance tradeoff made by
+programming systems that reserve processor cores to implement asynchronous operations (such as data movement).
 
@@
 Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
-run on a single Haswell node of Cori with a problem size large enough to
-ensure that it achieves peak FLOPS. This confirms
+run on a single Haswell node of Cori with a problem size large enough
+that runtime is dominated by the application's inner loops. This confirms
 that the application is properly configured and that the
 efficiency metric is achievable. The problem
-size is then repeatedly reduced while maintaining exactly the same hardware
-configuration (in particular, the same number of nodes). The
+size is then repeatedly reduced while maintaining exactly the same hardware and software
+configuration (in particular, the same number of nodes and tasks). The
 expectation is that as problem size shrinks,
@@
 shrinks with minimal change in efficiency. However as task
-granularity shrinks further, efficiency begins to drop more rapidly, leading
-eventually to an asymptote as overhead dominates useful work.
+granularity shrinks further, efficiency drops more rapidly, leading
+to an asymptote as overhead dominates useful work.
 

### 2019-04-08 4_implementation.tex
Apply Sean's comments.

@@
 are \emph{explicitly parallel} and require users to specify the
-parallelism in the program. The Task
-Bench implementations are intended to reflect how actual applications
-would be written in the respective systems. Places where the Task
-Bench implementations may be unidiomatic are noted below.
+parallelism in the program. For systems that provide both implicit and explicit parallelism, the form of parallelism used in Task Bench is emphasized in Table~\ref{tab:systems}.
 
-There are also aspects of some systems that we do not consider in our
-evaluation. For example, a number of systems provide some level of
-support for heterogeneous processors such as GPUs, but the support
-varies so widely (if it exists at all) that it seems difficult to
-compare systems with direct GPU support to those without in an
-apples-to-apples manner.
+%% The Task
+%% Bench implementations are intended to reflect how actual applications
+%% would be written in the respective systems. Places where the Task
+%% Bench implementations may be unidiomatic are noted below.
+
+There are also aspects of some systems that we do not consider. For
+example, tasks may run on CPUs, GPUs, or any execution resource
+supported by a programming system. However, the level of support for
+heterogeneous processors varies widely across programming systems (if it exists at all), so
+in order to include as many systems as possible in our comparisons, we
+evaluate Task Bench only on CPUs in this paper.
 
@@
 
+% FIXME: Discuss load balancer
+
 \subsection{MPI}
@@
 task dependencies. We tested both GNU GOMP and Intel KMP
-implementations of OpenMP and found KMP to be better performing.
+implementations of OpenMP and found KMP to be better performing on Cori.
 
@@
 
-Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
-distributed across nodes in pre-defined patterns,
-and can be passed to tasks on other nodes, but
-can only be written on the node where the data is originally
-allocated. In our Task Bench implementation, each task writes to a
-tile as output, and these tiles are passed as inputs to the
-subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
-the task launch code uses a switch statement as the PaRSEC API assumes
-tasks have a fixed number of dependencies.
+In our Task Bench implementation, we evaluate both PTG and DTD models,
+as well as a third option based on DTD that has additional manual
+optimizations applied to further reduce dynamic checks for what tasks
+should be executed on a node, resulting in further improvements to
+scalability.
+
+%% Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
+%% distributed across nodes in pre-defined patterns,
+%% and can be passed to tasks on other nodes, but
+%% can only be written on the node where the data is originally
+%% allocated. In our Task Bench implementation, each task writes to a
+%% tile as output, and these tiles are passed as inputs to the
+%% subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
+%% the task launch code uses a switch statement as the PaRSEC API assumes
+%% tasks have a fixed number of dependencies.
 
@@
 to disk, confirmed there is no measurable overhead due to JNI
-calls from Java to C, and cross-checked results with known cases
-that hit optimal task throughput in Spark, among other things.
+calls from Java to C or due to the serialization of data structures, and cross-checked results with known cases
+that hit optimal task throughput in Spark.
 
@@
 StarPU is very similar to PaRSEC, including the use of a switch
-statement to handle dynamic numbers of task dependencies, and manual
+statement to handle dynamic numbers of task dependencies and manual
 elision of tasks on unrelated nodes to provide horizontal scalability.

### 2019-04-08 5_evaluation.tex
Apply Sean's comments.

@@
 
-In theory, any system should be able to hit peak efficiency as long as
+In theory, any system should be able to hit peak performance as long as
 the kernels are well-written and of sufficiently large granularity. In
-practice, achieving peak performance is tricky, subject to a number of
-pitfalls, and unfortunately, frequently skipped in evaluations of
+practice, achieving peak performance is tricky, and many subtle pitfalls of implementation or configuration can easily lead to poor performance.
+Unfortunately, peak performance is also frequently skipped in evaluations of
 runtime system overhead. However, verifying that peak performance is
-achieved is important to ensuring that such evaluations of overhead and
-efficiency are well-grounded and avoid common pitfalls in
-configuration.
+achieved is important to ensuring that evaluations of overhead and
+efficiency are well-grounded.
 
@@
 
-Most systems for high-performance computing are horizontally scalable,
+Most systems for HPC are highly scalable,
 but this is not true of all the systems included in this
@@
 by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
-still retains dynamic checks leading to limited scalability. Regent
-uses a compile-time optimization that enables full horizontal
-scalability~\cite{ControlReplication17}. A similar, manual
-optimization is shown in the plots as PaRSEC shard.
+still retains dynamic checks leading to limited scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
+uses a compile-time optimization to improve
+scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.
 
@@
 simultaneous task graphs only partially mitigates the
-load imbalance between tasks. Therefore, systems that provide an
-additional on-node work stealing capability see additional gains in
+load imbalance between tasks. Systems that provide an
+additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
 efficiency at large task granularities. However, the use of
 work-stealing queues can also impact throughput at very small task
-granularities. This can be seen in the difference between the Chapel
-distrib (work stealing) vs default (non-work-stealing) schedulers.
+granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
 

### 2019-04-08 6_related.tex
Apply Sean's comments.

@@
 \item
-The initial paper on PENNANT~\cite{PENNANT} includes an MPI/OpenMP/MPI+OpenMP
-implementation; follow-up papers present an implementation in
+The initial paper on PENNANT~\cite{PENNANT} includes an
+implementation that supports MPI/OpenMP/MPI+OpenMP; follow-up papers present an implementation in
 Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.

### 2019-04-08 t1_systems.tex
Apply Sean's comments.

--- 2019-03-21 t1_systems.tex
@@
 \hline
-Chapel & multi-resolution & explicit & yes & uGNI \\
-Charm++ & actor model & explicit & yes & uGNI \\
-MPI & message passing & explicit & yes & uGNI \\
-OmpSs & loop-, task-based & expl., impl. & no & \\
-OpenMP & loop-, task-based & expl., impl. & no & \\
-PaRSEC & task-based & imlicit & yes & MPI \\
+Chapel & multi-resolution & \emph{expl.}, impl. & yes & uGNI\footnote{Chapel uses GASNet to support non-Cray networks.} \\
+Charm++ & actor model & explicit & yes & uGNI\footnote{Charm++ provides additional backends for other networks.} \\
+MPI & message passing & explicit & yes & uGNI\footnote{Most MPI implementations provide additional backends for other networks.} \\
+OmpSs & loop-, task-based & expl., \emph{impl.} & no & \\
+OpenMP & loop-, task-based & expl., \emph{impl.} & no & \\
+PaRSEC & task-based & implicit & yes & MPI \\
 Realm & task-based & explicit & yes & GASNet \\
@@
 TensorFlow & dataflow & explicit & yes\footnote{Our evaluation only considers TensorFlow on a single node.} & sockets \\
-X10 & place-based & explicit & yes & MPI
+X10 & place-based & explicit & yes & MPI\footnote{X10 also provides support for PAMI on supported networks.}
 \end{tabular}
 
-\caption{Systems used in Task Bench implementations.\label{tab:systems}}
+\caption{Systems for which we implemented Task Bench.\label{tab:systems}}
+\vspace{-0.5cm}
 \end{table}

### 2019-04-08 7_conclusion.tex
Taking Alex's suggestion for the conclusion.

@@
 of parallel and distributed programming systems. We have seen that
-systems with support for asynchronous (and/or task-based) programming
-appear to show benefits with more complicated dependency patterns,
-higher node counts (because of higher communication latencies) and
-more communication (because of ability to overlap it with
-computation). However, all these benefits can be eclipsed if baseline
-runtime overheads are too high. The current generation of task-based
-systems (PaRSEC, Realm, Regent, and StarPU) show higher overheads
-than traditional HPC systems, indicating that there continues to be
-work to be done in this area.
+systems vary by orders of magnitude in runtime overhead, as measured
+by METG. Widely used systems in industry have METG measured in
+seconds, reflecting the very coarse tasks and lack of need for strong
+scaling performance in current workloads. At the other extreme,
+traditional HPC systems have METG in the range of 10-100 \textmu{}s,
+necessary for strong scaling to small task granularities. The newer
+task-based systems, with METG in the range of 100 \textmu{}s to 1 ms
+have performance sufficient for weak scaling many HPC workloads, but
+more work is needed to strong scale the most demanding applications.
 

### 2019-04-08 4_implementation.tex
Trimming some text to make things fit again.

@@
 different nodes. To
-improve horizontal scalability, the program is executed in
+improve scalability, the program is executed in
 SPMD fashion on all nodes, and the user is responsible for eliding
@@
 provides a critical optimization, \emph{control replication}, that
-enables superior horizontal scalability~\cite{ControlReplication17}.
+enables superior scalability~\cite{ControlReplication17}.
 
@@
 Spark~\cite{Spark10} is an implicitly-parallel programming model for
-data analytics, widely used in industrial data center applications. As
+large scale data analytics. As
 one of the programming models in this study not originally intended
@@
 statement to handle dynamic numbers of task dependencies and manual
-elision of tasks on unrelated nodes to provide horizontal scalability.
+elision of tasks on unrelated nodes to improve scalability.
 We use a simplified version of the 2D block cyclic data distribution from the Chameleon 

### 2019-04-08 5_evaluation.tex
Trimming some text to make things fit again.
Get the figures reorganized again.
add one more explaination why some systems are missing in figures

@@
 were to run larger problem sizes. Unfortunately, the excessive
-computational cost of running such tests makes it prohibitively
-expensive to run larger problem sizes. For example, the Spark job in
+computational cost of running such tests makes this prohibitively
+expensive. For example, the Spark job in
 this configuration ran for over 6 hours.
@@
 red, dashed line shows the point where 50\% efficiency is achieved.
-
 In most cases, task granularity asymptotes prior to the 50\% efficiency point,
@@
 
-\input{e7_bytes}
-
 Figure~\ref{fig:bytes} shows the performance of Task Bench
@@
 
+% FIXME: why is Charm++ slow?
+
 The remaining experiments use compute-bound kernels.
+
+% FIXME: why compute-bound kernels?
 
@@
 
+\input{e7_bytes}
 \input{e4_radix}
@@
 are omitted from comparisons with more complicated dependence
-patterns, as they did not complete in reasonable time.
+patterns, as they did not complete in reasonable time 
+or reach METG(50\%) even with the largest problem size.
 

### 2019-04-09 5_evaluation.tex
Update evaluation text.

@@
 communication pattern grows, and even reverses as task parallelism is
-added in the form of multiple task graphs. Some of the slower systems
+added in the form of multiple task graphs. Some slower systems
 are omitted from comparisons with more complicated dependence
-patterns, as they did not complete in reasonable time 
-or reach METG(50\%) even with the largest problem size.
+patterns, as the higher overheads of these systems would have required
+unreasonably large problem sizes (beyond what can be completed in 6 hours) to reach 50\% efficiency.
 

### 2019-04-09 2_metg.tex
Edits from Alex.

@@
 
-In Figure~\ref{fig:efficiency-mpi}, we see that
+In Figure~\ref{fig:efficiency-mpi},
 efficiency starts at 100\%. Initially task granularity
@@
 granularity shrinks further, efficiency drops more rapidly, leading
-to an asymptote as overhead dominates useful work.
+to an asymptote as runtime overhead dominates useful work.
 

### 2019-04-09 5_evaluation.tex
Edits from Alex.

@@
 The least complicated pattern (stencil) is most favorable
-to MPI, as it represents a truly bulk synchronous task graph with no
+to MPI, as it provides no
 opportunity for task parallelism. For the stencil pattern, the
 dominating factor is the basic overhead of executing a task, which is
-minimal for MPI as the Task Bench implementation is bulk synchronous
-and simply executes tasks one after another in alternation with
+minimal for MPI as the Task Bench implementation simply executes tasks one after another in alternation with
 communication phases. In contrast, the asynchronous execution
-mechanisms of the other systems are pure overhead in this scenario.
+mechanisms of other systems are pure overhead in this scenario.
 
@@
 
-The MPI implementation of Task Bench, being bulk synchronous,
+The MPI implementation of Task Bench, with its distinct computation and communication phases,
 suffers the most under load imbalance. The biggest
@@
 
-The difference between the remaining systems is due primarily to the
-scheduling behavior of the various systems. The execution of 4
+The difference between the remaining systems is due primarily to
+different scheduling behaviors. The execution of 4
 simultaneous task graphs only partially mitigates the

### 2019-04-09 6_related.tex
Edits from Alex.

@@
 implement to a level of quality sufficient for conducting comparative
-studies between programming systems. The largest study we're aware of,
+studies between programming systems. The largest study we know of,
 for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
@@
 studies usually lack a comprehensive evaluation, even if multiple
-implementations are available.
+implementations are available:
 

### 2019-04-10 5_evaluation.tex
Add note about Regent and X10 results.

@@
 communication pattern grows, and even reverses as task parallelism is
-added in the form of multiple task graphs. Some slower systems
+added in the form of multiple task graphs.
+
+Spark, Swift/T and TensorFlow
 are omitted from comparisons with more complicated dependence
 patterns, as the higher overheads of these systems would have required
-unreasonably large problem sizes (beyond what can be completed in 6 hours) to reach 50\% efficiency.
+unreasonably large problem sizes (beyond what can be completed in 6
+hours) to reach 50\% efficiency. Regent and X10 are missing from some
+graphs as Cori was highly impacted before the deadline.
 

### 2019-04-10 5_evaluation.tex
Update final notes based on performance.

@@
 Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
-efficiency curves under load imbalance with uniformly distributed task
-durations. The task durations are generated with a deterministic
+efficiency curves under load imbalance where each task's duration is multiplied by a uniform random variable between [0,~1). The task durations are generated with a deterministic
 pseudo random number generator with a deterministic seed to ensure
@@
 suffers the most under load imbalance. The biggest
-difference is at large task granularities, where the imbalance has the
-effect of putting an upper bound on MPI efficiency. At smaller task
+difference is at large task granularities, where the imbalance
+effectively puts an upper bound on MPI efficiency. At smaller task
 granularities the effect shrinks and may even reverse as systems hit
@@
 work-stealing queues can also impact throughput at very small task
-granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
+granularities.
+For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
+We do not consider Charm++ load balancers here because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).
+We leave the analysis of persisitent load imbalance to future work.
 

### 2019-04-10 5_evaluation.tex
Remove resolved FIXMEs.
Update evaluation numbers to match chart.

@@
 The ratio in METG between 0 and 3 dependencies per task ranges from
-$1.3\times$ to $59\times$ (median $3.0\times$). From 1 to 3
-dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
-$2.1\times$). The difference is most pronounced in systems that
+$1.01\times$ to $16\times$ (median $2.9\times$). From 1 to 3
+dependencies, the ratio varies from $0.54\times$ to $7.3\times$ (median
+$1.8\times$). The difference is most pronounced in systems that
 perform runtime work inline. For example, MPI achieves an METG of 390
@@
 
-% FIXME: update this when we have more data
-
 Most systems for HPC are highly scalable,
@@
 scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.
-
-% FIXME: remove parsec dtd shard if no space.
-
-% FIXME: talk about overhead spanning 5 orders of magnitude
 

### 2019-04-11 5_evaluation.tex
Apply Alex's comments.

@@
 Spark, Swift/T and TensorFlow
-are omitted from comparisons with more complicated dependence
-patterns, as the higher overheads of these systems would have required
+are omitted from comparisons with more complicated dependencies, as the higher overheads of these systems would have required
 unreasonably large problem sizes (beyond what can be completed in 6
-hours) to reach 50\% efficiency. Regent and X10 are missing from some
-graphs as Cori was highly impacted before the deadline.
+hours) to reach 50\% efficiency. Regent and X10 are incomplete as Cori was highly impacted before the deadline; complete results will be in the final paper.
 
@@
 One advantage of systems with asynchronous execution capabilities is
-the ability to mitigate load imbalance, especially in the presence of
+the ability to mitigate load imbalance with little or no additional programmer effort, especially in the presence of
 task parallelism. To quantify this effect,

### 2019-04-11 1_introduction.tex
Apply Pat's comments.

@@
 One approach to comparing the performance of different systems is through \emph{proxy-} or
-\emph{mini-apps}. Because they distill the structure of a larger
+\emph{mini-apps}. Because they distill the key computational characteristics of a larger
 application, mini-apps can provide insight

### 2019-04-11 5_evaluation.tex
Apply Pat's comments.

@@
 controller, which limits throughput, and this is visible in the figure
-as the line for Spark immediately rises with node count. Keep in mind also
+as the line for Spark immediately rises with node count. Keep in mind
 that Spark is being evaluated here with a non-trivial dependence

### 2019-04-11 5_evaluation.tex
Apply Pat's feedback.

@@
 that Spark is being evaluated here with a non-trivial dependence
-pattern which is relatively unrepresentative of Spark's normal use
+pattern that is relatively unrepresentative of Spark's normal use
 cases. Spark is more efficient with trivial parallelism, as described

### 2019-04-11 5_evaluation.tex
Final updates for paper submission.

@@
 performance curves are similar. We are currently investigating the
-performance of Charm++.
+performance of Charm++ and Regent.
 
-% FIXME: why is Charm++ slow?
+% FIXME: why are Charm++ and Regent slow?
 
@@
 
+%% From 1 to 3
+%% dependencies, the ratio varies from $0.54\times$ to $7.3\times$ (median
+%% $1.8\times$).
+
 The ratio in METG between 0 and 3 dependencies per task ranges from
-$1.01\times$ to $16\times$ (median $2.9\times$). From 1 to 3
-dependencies, the ratio varies from $0.54\times$ to $7.3\times$ (median
-$1.8\times$). The difference is most pronounced in systems that
+$1.01\times$ to $16\times$ (median $2.9\times$). The difference is most pronounced in systems that
 perform runtime work inline. For example, MPI achieves an METG of 390
@@
 efficiency curves under load imbalance where each task's duration is multiplied by a uniform random variable between [0,~1). The task durations are generated with a deterministic
-pseudo random number generator with a deterministic seed to ensure
+pseudo random number generator with a consistent seed to ensure
 identical task durations for all systems.

### 2019-05-20 1_introduction.tex
Update list of systems.

@@
 We conduct a comprehensive comparative study of Task Bench
-implementations in 13 different programming systems:
-Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, MPI \cite{MPI},
+implementations in 16 different programming systems:
+Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
 OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
@@
 \item Section~\ref{sec:implementation} details the implementation of
-  Task Bench in 13 programming systems.
+  Task Bench in 16 programming systems.
 \item Section~\ref{sec:evaluation} provides a comprehensive evaluation

### 2019-05-20 3_task_bench.tex
Apply Pat's comments.
Update list of systems.

@@
 systems and application scenarios, we developed a novel
-\emph{parameterized} benchmark called Task Bench. Task Bench
+parameterized benchmark called Task Bench. Task Bench
 implementations execute a \emph{task graph}, with tasks for each point
@@
 a Task Bench implementation is in the core library, implementations of
-Task Bench are small: Our 13 Task Bench implementations range from 88
+Task Bench are small: Our 16 Task Bench implementations range from 88
 to 1500 lines, with several hundred lines being typical.

### 2019-05-20 t1_systems.tex
Update list of systems.

@@
 Charm++ & actor model & explicit & yes & uGNI\footnote{Charm++ provides additional backends for other networks.} \\
+Dask & task-based & implicit & yes & sockets \\
 MPI & message passing & explicit & yes & uGNI\footnote{Most MPI implementations provide additional backends for other networks.} \\
+MPI+X & hybrid & explicit & yes & MPI \\
 OmpSs & loop-, task-based & expl., \emph{impl.} & no & \\

### 2019-05-21 reviews/response.md
Some work on response text.

We thank the reviewers for their detailed and insightful feedback.

We have added three new implementations to Task Bench: Dask,
MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
section.

One of the salient features of Task Bench is that it permits the
implementation of a benchmark with M configurations and N systems with
an amount of effort that is O(M+N) rather than O(MN). We have updated
the introduction to reflect this.

In response to specific reviewer questions:

### 2019-05-22 1_introduction.tex
Only claim 15 systems so we don't have to quibble about MPI+CUDA not running on Cori.

--- 2019-04-11 1_introduction.tex
@@
 We conduct a comprehensive comparative study of Task Bench
-implementations in 13 different programming systems:
-Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, MPI \cite{MPI},
+implementations in 15 different programming systems:
+Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
 OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
@@
 \item Section~\ref{sec:implementation} details the implementation of
-  Task Bench in 13 programming systems.
+  Task Bench in 15 programming systems.
 \item Section~\ref{sec:evaluation} provides a comprehensive evaluation

### 2019-05-22 3_task_bench.tex
Add a code sample from Dask implementation.
Only claim 15 systems so we don't have to quibble about MPI+CUDA not running on Cori.

@@
 a Task Bench implementation is in the core library, implementations of
-Task Bench are small: Our 16 Task Bench implementations range from 88
+Task Bench are small: Our 15 Task Bench implementations range from 88
 to 1500 lines, with several hundred lines being typical.
@@
 \input{t0_equations}
+\input{l1_code_sample}
 

### 2019-05-22 4_implementation.tex
Remove resolved FIXMEs.
Update list of systems.
Fiddle with figure placement, labels, listing font size.
Only claim 15 systems so we don't have to quibble about MPI+CUDA not running on Cori.
Add Dask text.

@@
 
-\input{t1_systems}
-
-We have implemented Task Bench in the 13 parallel and distributed
+We have implemented Task Bench in the 15 parallel and distributed
 programming systems listed in Table~\ref{tab:systems}. We describe the
@@
 
+\input{t1_systems}
+
 \subsection{Chapel}
@@
 
-% FIXME: Discuss load balancer
+\subsection{Dask}
+
+Dask is an implicitly parallel task-based programming system for large
+scale data analytics with Python. Dask provides abstractions over
+distributed NumPy arrays and Pandas dataframes, and also a lower-level
+interface for launching tasks. We use the lower-level interface to
+implement Task Bench.
 
 \subsection{MPI}
-
-% FIXME: bulk synchronous or point-to-point
 
@@
 send/receive pair in MPI.
+
+\subsection{MPI+X}
+
+% FIXME: TODO
 
@@
 implementations of OpenMP and found KMP to be better performing on Cori.
-
-% FIXME: keep this in sync with experiments
 

### 2019-05-22 reviews/response.md
More work on response.

@@
 
+Any measure of efficiency can be used with METG. For example, a
+mesh-based application could use mesh cells processed per second (as a
+percentage of the highest throughput achieved with any problem
+size). We use peak performance with Task Bench as it gives us a
+straightforward measure that provides assurance that all
+implementations are configured correctly.
+
 In response to specific reviewer questions:

### 2019-05-22 4_implementation.tex
Add more information on implementations.

@@
 phase, posting receives before sends. Each task dependency maps to one
-send/receive pair in MPI.
+send/receive pair in MPI. We also provide a bulk synchronous
+implementation which enforces the boundary between communication and
+computation with \lstinline[language=C++]{MPI_Barrier}.
 
@@
 
-% FIXME: TODO
+We provide two MPI+X implementations of Task Bench to evaluate the
+performance of hybrid programming models. Our MPI+OpenMP
+implementation uses forall-style parallel loops to execute tasks,
+interleaved with MPI point-to-point communication as above. Our
+MPI+CUDA implementation follows an offload model where data is copied
+to and from the GPU on every timestep.
 

### 2019-05-22 1_introduction.tex
Tweak text to make things fit better.

@@
 performance. Large tasks may be used to hide any amount of runtime
-overhead, while tasks that are too small may result in a drop in total
+overhead, while small tasks may result in a drop in total
 application throughput even as TPS increases. Only by constraining

### 2019-05-22 4_implementation.tex
Tweak text.

@@
 
-There are also aspects of some systems that we do not consider. For
-example, tasks may run on CPUs, GPUs, or any execution resource
-supported by a programming system. However, the level of support for
-heterogeneous processors varies widely across programming systems (if it exists at all), so
-in order to include as many systems as possible in our comparisons, we
-evaluate Task Bench only on CPUs in this paper.
+%% There are also aspects of some systems that we do not consider. For
+%% example, tasks may run on CPUs, GPUs, or any execution resource
+%% supported by a programming system. However, the level of support for
+%% heterogeneous processors varies widely across programming systems (if it exists at all), so
+%% in order to include as many systems as possible in our comparisons, we
+%% evaluate Task Bench only on CPUs in this paper.
 
@@
 We provide two MPI+X implementations of Task Bench to evaluate the
-performance of hybrid programming models. Our MPI+OpenMP
+performance of hierarchical programming models. Our MPI+OpenMP
 implementation uses forall-style parallel loops to execute tasks,

### 2019-05-22 4_implementation.tex
A lot of edits to shrink section 4.

@@
 Spark~\cite{Spark10} is an implicitly-parallel programming model for
-large scale data analytics. As
-one of the programming models in this study not originally intended
-for scientific computing, Spark makes very different tradeoffs from
-other systems under consideration, and therefore makes an interesting
-point of comparison.
-
+large scale data analytics.
 The core abstractions in Spark are functional operators such as map,
-reduce, filter, join, etc. Functions in Spark operate on
+reduce, and join. Data in Spark are stored in
 \emph{resilient
-  distributed datasets} (RDDs), which are globally-visible,
-dynamically single-assignment data structures. One of Spark's primary
-optimizations is the caching of RDDs to avoid unnecessary
+  distributed datasets} (RDDs): globally-visible,
+dynamically single-assignment data structures. Spark caches RDDs in memory to avoid unnecessary
 disk traffic.
 
-Though Spark has tasks internally, these are not exposed to the user,
-so one of the challenges in developing a Task Bench implementation is
-mapping the task graph to a set of operators that result in the
-desired performance behavior. We use a combination of
+Spark has tasks internally but they are not exposed to the user,
+so a Task Bench implementation must map the task graph to a set of operators that result in the
+desired execution. We use
 \lstinline[language=Scala]{flatMap} and
-\lstinline[language=Scala]{groupByKey} operations to generate task
+\lstinline[language=Scala]{groupByKey} to generate task
 dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
-series of tasks. We use an explicit hash partitioner to ensure that
+series of tasks. An explicit hash partitioner ensures that
 Spark does not attempt to group multiple Task Bench tasks into a
-single Spark task, as the tasks in Task Bench already represent coarse-grained units of work.
+single Spark task, as tasks in Task Bench already represent coarse-grained units of work.
 
@@
 extraneous factors interfered with our Spark measurements. We
-completely disabled logging, ensured no RDDs are written
+disabled logging, ensured no RDDs are written
 to disk, confirmed there is no measurable overhead due to JNI
@@
 programming model similar to PaRSEC's DTD. Our Task Bench implementation in
-StarPU is very similar to PaRSEC, including the use of a switch
-statement to handle dynamic numbers of task dependencies and manual
-elision of tasks on unrelated nodes to improve scalability.
-We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
-project~\cite{Chameleon}.
+StarPU is very similar to PaRSEC.
+%% We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
+%% project~\cite{Chameleon}.
 
@@
 machine learning concepts, internally TensorFlow is a task graph execution engine, making it a good fit for
-Task Bench. TensorFlow programs are written by creating nodes in the
-task graph, and then connecting these nodes with edges to implement
-the desired dataflow. Task graphs are composed in a high-level
-language such as Python, but then serialized and run by an
-execution engine written in C++.
-
-Our Task Bench implementation in TensorFlow works by defining a custom
-operator, which we instantiate for each point in the task graph. We
-then define dataflow edges to correspond to the task dependencies.
+Task Bench. TensorFlow programs operate via explicit graph construction. Task graphs are composed in Python and run by an
+execution engine written in C++. Our Task Bench implementation is straightforward.
 

### 2019-05-22 4_implementation.tex
More tweaks.

@@
 Regent rather than Legion directly because the Regent compiler
-provides a critical optimization, \emph{control replication}, that
-enables superior scalability~\cite{ControlReplication17}.
+provides a critical optimization that
+improves scalability~\cite{ControlReplication17}.
 
@@
 Swift/T~\cite{Wozniak13} is a parallel scripting language intended
-primarily for the composition of HPC workflows. Tasks in Swift/T can be written in any
-programming language. Swift/T
+primarily for the composition of HPC workflows. Swift/T
 programs follow dataflow semantics, where every statement may
@@
 requested. The Swift/T compiler performs a number of optimizations to
-improve performance of highly parallel programs~\cite{Armstrong14}.
-
-Our Task Bench in Swift/T is straightforward, using Swift/T's dataflow
+improve performance of highly parallel programs~\cite{Armstrong14}. Our Task Bench in Swift/T is straightforward, using Swift/T's dataflow
 semantics to capture dependencies on other tasks.

### 2019-05-22 t1_systems.tex
More tweaks.

--- 2019-04-08 t1_systems.tex
@@
 TensorFlow & dataflow & explicit & yes\footnote{Our evaluation only considers TensorFlow on a single node.} & sockets \\
-X10 & place-based & explicit & yes & MPI\footnote{X10 also provides support for PAMI on supported networks.}
+X10 & place-based & explicit & yes & MPI\footnote{X10 also provides a PAMI backend on supported networks.}
 \end{tabular}

### 2019-05-22 5_evaluation.tex
Add weak/strong scaling figures.
Add GPU efficiency figure.
Fiddle with figure placement, labels, listing font size.
Fiddle with figure placement.
New section to describe weak and strong scaling curves.

@@
 
+For GPU experiments we use the Piz Daint supercomputer~\cite{PizDaint}, a Cray XC50 with
+one Intel Xeon E5-2690 v3 (12 physical cores) and one NVIDIA Tesla
+P100 per node. We use GCC \textbf{X.Y}, Cray MPICH \textbf{Z.W}, and
+CUDA \textbf{A.B}.
+
+% FIXME: get Piz Daint software versions.
+
 \subsection{Compute-Bound Performance}
@@
 
+\input{e1_flops}
+\input{e2_efficiency}
+\input{e7_bytes}
+\input{e4_radix}
+
 Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
@@
 the baseline for 100\% efficiency.
-
-\input{e1_flops}
-\input{e2_efficiency}
 
@@
 
+% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
+\input{e3_metg_compute}
+
 Figure~\ref{fig:bytes} shows the performance of Task Bench
@@
 \label{subsec:number-of-dependencies}
-
-\input{e7_bytes}
-\input{e4_radix}
 
@@
 \subsection{Baseline Overhead}
-
-% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
-\input{e3_metg_compute}
 
@@
 unreasonably large problem sizes (beyond what can be completed in 6
-hours) to reach 50\% efficiency. Regent and X10 are incomplete as Cori was highly impacted before the deadline; complete results will be in the final paper.
+hours) to reach 50\% efficiency.
 
@@
 \label{subsec:scalability}
+
+\input{e8_weak}
+\input{e9_strong}
 
@@
 
+\subsection{Comparison with Weak and Strong Scaling}
+
+Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
+conventional strong and weak scaling of the stencil pattern for
+comparison with the METG results in
+Figure~\ref{fig:metg-compute-stencil}. At small node counts, where the
+average task granularity is larger than the METG for most systems, the
+performance is ideal. For several systems METG rises with node count,
+and this becomes visible in the weak and strong scaling graphs as
+non-ideal performance. Thus the shape of the weak and strong scaling
+graphs effectively interpolates between ideal performance and the METG
+curve depending on how close each system is to being limited by
+overhead.
+
 \subsection{Load Imbalance}
@@
 \input{e5_imbalance}
+\input{e6_communication}
 
@@
 
-\input{e6_communication}
-
 Also of interest is the ability of asynchronous systems to
@@
 smaller message sizes.
+
+\subsection{Heterogeneous Processors}
+
+\input{e10_cuda_efficiency}
+
+% FIXME: TODO

### 2019-05-22 1_introduction.tex
New text for GPU comparison and other tweaks.

@@
 Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
-performance (e.g., FLOPS or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
+performance (e.g., FLOP/s or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
 or strong scaling study, the behavior at the limit of scalability. For

### 2019-05-22 2_metg.tex
Apply Pat's comments.
New text for GPU comparison and other tweaks.

@@
 compute-bound applications efficiency can be measured as the
-percentage of the available FLOPS achieved. On Cori with 1.26 TFLOPS available per Haswell node, METG(50\%) corresponds to
+percentage of the available FLOP/s achieved. On Cori with 1.26 TFLOP/s available per Haswell node, METG(50\%) corresponds to
 the smallest task granularity achieved while maintaining at least 0.63
-TFLOPS. Absolute (rather than relative) efficiency is preferred
+TFLOP/s. Absolute (rather than relative) efficiency is preferred
 because it better quantifies the performance tradeoff made by
@@
 To calculate METG, the data is replotted along axes of efficiency
-(i.e., as a percentage of the peak FLOPS achieved) and task
+(i.e., as a percentage of the peak FLOP/s achieved) and task
 granularity (i.e., $\text{wall time} \times \text{number of
@@
 granularity shrinks further, efficiency drops more rapidly, leading
-to an asymptote as runtime overhead dominates useful work.
+to a vertical asymptote as runtime overhead dominates useful work.
 

### 2019-05-22 5_evaluation.tex
New text for GPU comparison and other tweaks.

@@
 
-Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
+Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
 kernel with varying problem sizes. This is the full version of
-Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOPS of
+Figure~\ref{fig:flops-mpi}.
+Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern.
+In the best case, we measure peak FLOP/s of
 $1.26 \times 10^{12}$, which compares favorably with the officially
@@
 
-The vast majority of systems achieve or nearly achieve peak FLOPS. Some
+The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
 systems reserve a number of cores (usually 1 or 2) for internal
-runtime usage; these systems take a minor hit in peak FLOPS compared
+runtime usage; these systems take a minor hit in peak FLOP/s compared
 to systems which share all cores between the application and runtime. Some of the
-slower systems struggle to achieve peak FLOPS, though in most cases
+slower systems struggle to achieve peak FLOP/s, though in most cases
 the curves suggest that performance would continue to improve if we
@@
 Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
-peak FLOPS) plotted against task granularity. This plot is
+peak FLOP/s) plotted against task granularity. This plot is
 used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
@@
 \input{e9_strong}
+\input{e5_imbalance}
 
@@
 
+\input{e6_communication}
+
 Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
@@
 \subsection{Load Imbalance}
-
-\input{e5_imbalance}
-\input{e6_communication}
 
@@
 
-% FIXME: TODO
+Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
+and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
+compute kernel achieves a peak performance of $4.759 \times 10^{12}$
+FLOP/s, which compares favorably with the officially reported number
+$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
+FLOP/s. Note that the x-axis is measured in the number of FLOP
+performed during the run, as the GPU kernels must perform more work to
+achieve peak performance.
+
+Our MPI+CUDA implementation uses an offload model in which input and
+output data is copied to and from the GPU on each time step. We tested
+two configurations: \lstinline{w1} uses 1 task per GPU, whereas
+\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
+ranks per GPU to push work to the GPU in parallel and thus achieve
+some overlap between communication and computation. We see that the
+overdecomposed \lstinline{w4} configuration achieves higher FLOP/s but
+drops more rapidly at smaller problem sizes, due to the overhead of
+scheduling $4\times$ as many CUDA kernel launches. In either case, the
+GPU requires more work to achieve high performance, and the overhead
+of copying data to and from the GPU dominates at small task
+granularities. Thus at smaller problem sizes the CPU achieves higher
+performance.

### 2019-05-22 e5_imbalance.tex
Tweak spacing around figure captions.
Apply Alex's and George's comments.
New text for GPU comparison and other tweaks.

@@
 
-\caption{Task granularity vs efficiency under load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Lower is better.\label{fig:efficiency-imbalance}}
+\vspace{-0.15cm}
+\caption{Efficiency vs task granularity under load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Higher is better.\label{fig:efficiency-imbalance}}
+\vspace{-0.1cm}
 \end{figure}

### 2019-05-22 e6_communication.tex
Apply Alex's and George's comments.
New text for GPU comparison and other tweaks.

@@
 \vspace{-0.25cm}
-\caption{Task granularity vs efficiency for varying communication (spread pattern, 5 dependencies per task, 4 task graphs, 64 nodes). Lower is better.\label{fig:efficiency-communication}}
+\caption{Efficiency vs task granularity for varying communication (spread pattern, 5 dependencies per task, 4 task graphs, 64 nodes). Higher is better.\label{fig:efficiency-communication}}
 \vspace{-0.05cm}

### 2019-05-23 reviews/response.md
Some work on response.

@@
 
+Major changes in the revised paper:
+
+  * Three new implementations (Dask, MPI+OpenMP, MPI+CUDA)
+  * New sections
+      - Implementation: sections 4.3 (Dask) and 4.5 (MPI+X)
+      - Evaluation: sections 5.6 (weak/strong scaling) and 5.9 (GPUs)
+  * New figures
+      - Listings 1 and 2 (code samples)
+      - Figures 11, 12, 15 (experiments)
+
+In response to specific reviewer questions:
+
+Reviewer 1:
+
+The memory bound kernel has been written to perform a varying number
+of passes (or partial passes) over a fixed-sized buffer, in order to
+keep working set size constant and avoid interference due to cache
+effects, which in our opinion makes the results difficult to
+interpret.
+
+See the code samples in Listing 1 and 2.
+
+Reviewer 2:
+
+Figures 3 and 4 have been updated and new text added.
+
+Reviewer 3:
+
+Weak/strong scaling and TPS are all application and system dependent
+metrics. METG differs from these metrics in that can simultaneously
+(a) capture the behavior of the sytem at the limit of its scalability
+(which is not true of weak/strong scaling) and (b) while achieving a
+specified amount of useful work (which is not true of TPS). METG also
+isolates the impact of changes to the communication topology at scale,
+whereas in a strong scaling study this effect would be entangled with
+issues of system overhead. Examples of how to compute METG from
+application wall time have been added to Section 2.
+
+METG captures the average task granularity that can be supported. It
+is acceptable for some tasks to be smaller and others larger, as
+overheads will even out across tasks.
+
+A comparison with weak/strong scaling has been added in Section 5.6.
+
+Reviewer 4:
+
+The tasks in Task Bench are idealized compute or memory-bound loops; problem size is simulated by controlling the number of iterations in these loops.
+
+Shared memory is permitted and is used in our MPI+OpenMP implementation (among others).
+
+The code for determining the dependence pattern and implementations of
+kernels are contained in the core. A code sample from an
+implementation is included in Listing 2. As the reviewer can see,
+implementations can be quite simple.
+
+Reviewer 5:
+
+As described in Section 6, citations [27] and [37] are the only papers
+we're aware of which provide an empirical performance evaluation of a
+substantial number of systems. Many mini-app and benchmark papers only
+include a small number of systems, or describe systems but do not
+include an empirical evalution of performance. When there are
+follow-up papers which describe additional implementations, they
+rarely make comparisons beyond the reference implementation of the
+code.
+
+The output of a task is an array of bytes overwritten with a repeating
+pattern (timestep, point). If the input does not match an assertion is
+thrown.
+
 Any measure of efficiency can be used with METG. For example, a
@@
 implementations are configured correctly.
-
-In response to specific reviewer questions:

### 2019-05-23 0_abstract.tex
Update word count.
Update list of systems.
Only claim 15 systems so we don't have to quibble about MPI+CUDA not running on Cori.
Apply Alex's comments.

--- 2019-04-05 0_abstract.tex
@@
 \begin{abstract}
-% 128 words
+% 147 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
@@
 programming systems under a variety of application scenarios. Task
-Bench seeks to provide the ease of implementation and portability of
-microbenchmarks, which makes comparisons between many systems
-possible, with the ability to model a wide range of application
-scenarios, giving it coverage approximating mini-apps.
+Bench lowers the barrier to benchmarking multiple systems by making
+the implementation for a given system orthogonal to the benchmarks
+themselves: all benchmarks constructed with Task Bench automatically
+run on all Task Bench implementations, and vice versa. Furthermore,
+Task Bench's parameterization allows a wide variety of benchmark
+scenarios that distill the key characteristics of larger applications.
 
 We conduct a comprehensive study with implementations of Task Bench in
-13 programming systems on up to 256 Haswell nodes of the Cori
+15 programming systems on up to 256 Haswell nodes of the Cori
 supercomputer. We introduce and use a novel metric, \emph{minimum effective task

### 2019-05-23 1_introduction.tex
Apply Alex's comments.

@@
 often do not reach the level of maturity required to compare
-systems. Few published studies involve more than a handful of systems~\cite{LULESH13}.
+systems. Few published studies compare more than a handful of systems~\cite{LULESH13}.
 
@@
 patterns, it is simple enough to be implemented with reasonable effort
-to a level of quality sufficient for comparative studies.
+to a level of quality sufficient for comparative
+studies. Additionally, Task Bench is separated into a core API shared
+by all the system implementations. New benchmarks immediately run on
+all systems, and new systems immediately run all benchmarks included
+in Task Bench. Thus, Task Bench is much more productive for
+benchmarking than the standard approach in which each benchmark must
+be implemented separately for each system.
 

### 2019-05-23 3_task_bench.tex
Add kernel code sample.
Tweak text.
Apply Alex's comments.

@@
 in an \emph{iteration space} and dependencies between tasks determined
-by a \emph{dependence relation}. A task that depends on another task
-must execute after that task, and receives the output data from that
-task as input. Each task can execute any one of a
+by a \emph{dependence relation}. A task $A$ that depends on another task $B$
+must execute after task $B$, and receives the output data from
+task $B$ as input. Each task can execute any one of a
 number of kernels (compute-bound, memory-bound, etc.) and can generate
@@
 
+\input{f5_task_graphs}
+\input{t0_equations}
+
 Despite its generality, Task Bench is easy to implement, making it
@@
 Task Bench are small: Our 15 Task Bench implementations range from 88
-to 1500 lines, with several hundred lines being typical.
-
-\input{f5_task_graphs}
-\input{t0_equations}
-\input{l1_code_sample}
+to 1500 lines, with several hundred lines being
+typical. Listings~\ref{lst:compute-kernel} and \ref{lst:code-sample}
+show excerpts from the Task Bench core, and the Dask implementation,
+respectively. Only the code in \ref{lst:code-sample} need be
+implemented for each system, minimizing the work required for each additional system.
 
@@
 
+\input{l0_compute_kernel}
+\input{l1_code_sample}
+
 Task Bench provides two main kernels that can be called from tasks,

### 2019-05-23 5_evaluation.tex
Apply Alex's comments.

@@
 one Intel Xeon E5-2690 v3 (12 physical cores) and one NVIDIA Tesla
-P100 per node. We use GCC \textbf{X.Y}, Cray MPICH \textbf{Z.W}, and
-CUDA \textbf{A.B}.
+P100 per node. We use GCC 6.2.0, Cray MPICH 7.7.2, and
+CUDA 9.1.85.
 
@@
 
-\subsection{Compute-Bound Performance}
+\subsection{Compute Kernel Performance}
 \label{subsec:peak-performance-and-efficiency}
@@
 
-\subsection{Memory-Bound Performance}
+\subsection{Memory Kernel Performance}
 
@@
 performance curves are similar. We are currently investigating the
-performance of Charm++ and Regent.
+performance of Charm++, Regent and X10.
 
-% FIXME: why are Charm++ and Regent slow?
+% FIXME: why are Charm++, Regent and X10 slow?
 
@@
 Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
-communication topology and latency.
+communication topology and latency. An interesting observation is that
+the systems with the smallest METG on one node have roughly an order
+of magnitude higher METG at 256 nodes. The obvious conclusion is that
+increased communication latency requires significantly larger tasks to
+achieve the same level of efficiency.
 
@@
 $4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
-FLOP/s. Note that the x-axis is measured in the number of FLOP
-performed during the run, as the GPU kernels must perform more work to
+FLOP/s. Note that the x-axis is measured in the number of FLOPs
+performed during the run, as the GPU kernels must be given more work to
 achieve peak performance.
@@
 granularities. Thus at smaller problem sizes the CPU achieves higher
-performance.
+performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
+terms of MTEG (as the peak performance of the GPU and CPU are very
+different), the conclusion here is similar to
+Section~\ref{subsec:scalability}: The cost of communicating data and
+tasks to and from the GPU imposes a higher task granularity to achieve
+the same efficiency than running on the CPU, reducing the advantage of
+MPI's very lightweight mechanisms.

### 2019-05-23 6_related.tex
Apply Alex's comments.

@@
 
-Programming systems for parallel and distributed computing are most
+Parallel and distributed programming systems are most
 commonly evaluated using proxy- or mini-apps, or
@@
 for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
-systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI). Other
+systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI), each of
+which require a wholly separate, tuned implementation (in contrast to
+Task Bench). Other
 studies usually lack a comprehensive evaluation, even if multiple
@@
 single, simple code involving 2 kernels (which combined fit in under 50 lines
-of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} are also relatively widely studied with
+of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} is also relatively widely studied with
 implementations in OpenMP~\cite{NASOpenMP99}, MPI and

### 2019-05-23 7_conclusion.tex
Edits from Alex.
Apply Alex's comments.

@@
 
-Task Bench is a promising new approach for evaluating the performance
+Task Bench is a new approach for evaluating the performance
 of parallel and distributed programming systems. We have seen that
@@
 traditional HPC systems have METG in the range of 10-100 \textmu{}s,
-necessary for strong scaling to small task granularities. The newer
+necessary for strong scaling to small task granularities, though we
+have shown that METG also increases significantly for these systems
+with the size of the cluster and the use offload accelerators. The newer
 task-based systems, with METG in the range of 100 \textmu{}s to 1 ms

### 2019-05-23 reviews/response.md
Update response.

@@
 MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
-section.
-
-One of the salient features of Task Bench is that it permits the
-implementation of a benchmark with M configurations and N systems with
-an amount of effort that is O(M+N) rather than O(MN). We have updated
-the introduction to reflect this.
+section, including preliminary results from experiments with GPUs on
+the Piz Daint supercomputer.
 
@@
 
-  * Three new implementations (Dask, MPI+OpenMP, MPI+CUDA)
+  * New implementations: Dask, MPI+OpenMP and MPI+CUDA
   * New sections
@@
 
-The tasks in Task Bench are idealized compute or memory-bound loops; problem size is simulated by controlling the number of iterations in these loops.
+One of the salient features of Task Bench is that it permits the
+implementation of a benchmark with M configurations and N systems with
+an amount of effort that is O(M+N) rather than O(MN). We have updated
+the introduction to reflect this.
 
-Shared memory is permitted and is used in our MPI+OpenMP implementation (among others).
+The tasks in Task Bench are idealized compute or memory-bound loops;
+problem size is simulated by controlling the number of iterations in
+these loops.
+
+Shared memory is permitted and is used in our MPI+OpenMP
+implementation (among others).
 

### 2019-05-23 1_introduction.tex
Tweak text.

--- 2019-05-22 1_introduction.tex
@@
 patterns, it is simple enough to be implemented with reasonable effort
-to a level of quality sufficient for comparative studies.
+to a level of quality sufficient for comparative
+studies. Task Bench is provides a core API that is shared across the
+various system implementations. New benchmarks created with Task Bench
+immediately run on all systems, and new systems immediately run all
+benchmarks. Thus, Task Bench is much more productive for benchmarking
+a wide range of phenomena than the standard approach in which each
+benchmark must be implemented separately for each system.
 

### 2019-05-23 2_metg.tex
Tweak text.

@@
 the smallest task granularity achieved while maintaining at least 0.63
-TFLOP/s. Absolute (rather than relative) efficiency is preferred
-because it better quantifies the performance tradeoff made by
-programming systems that reserve processor cores to implement asynchronous operations (such as data movement).
+TFLOP/s. For applications not amenable to being characterized by peak
+performance, an application-specific measure of performance can be
+used instead (e.g., mesh cells processed per second).
 

### 2019-05-23 2_metg.tex
More text.

@@
 METG(50\%) is the intersection of this curve with 50\% efficiency, as
-shown by the red lines in Figure~\ref{fig:flops-mpi} and
+shown by the red, dashed lines in Figure~\ref{fig:flops-mpi} and
 \ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
@@
 where lines become more compressed. At the
-very bottom, the lines compress together as running time becomes dominated by overhead. The shape of the curve at the bottom is the same
-for both weak and strong scaling; this is
-the METG curve.
+very bottom, the lines compress together as running time becomes
+dominated by overhead. The red, dashed line in each graph marks
+METG(50\%), the point where efficiency reaches 50\%. Note that the
+shapes at the bottom of the strong and weak scaling curves are identical,
+and conform to the shape of the METG curve.
 
@@
 size that can be weak scaled to a given node count with a given level
-of efficiency. With smaller problem sizes runtime overhead increasingly dominates useful
-computation. Similarly, METG corresponds to the point at which strong
-scaling can be expected to stop; as strong scaling runs start with a
-larger problem size and progressively shrink it until overheads begin
-to dwarf the gains due to continued scaling. In
-Figures~\ref{fig:weak-scaling-mpi} and \ref{fig:strong-scaling-mpi},
-the dashed, bright red line show the value of METG(50\%) at each node
-count.
+of efficiency. Using the formula for task granularity above, each run
+is 32 tasks wide and 1000 timesteps long, so task granularity is wall
+time divided by 1000 (since Cori has 32 cores per node). The $2^{12}$
+problem size in Figure~\ref{fig:weak-scaling-mpi} scales well
+initially because the task granularity of 20~\textmu{}s is greater
+than the METG(50\%) of MPI at small node counts (which is about 4.6-12
+\textmu{}s from 1-64 nodes) but not at higher node counts (which rises
+to 28 \textmu{}s at 128 nodes and 61 \textmu{}s at 256). Similarly,
+METG corresponds to the point at which strong scaling can be expected
+to stop. In Figure~\ref{fig:strong-scaling-mpi} the problem size
+$2^{18}$ begins to peter out at 64 nodes, the point at which the
+scaling curve intersects METG(50\%).
 

### 2019-05-24 reviews/response.md
revise response.md

@@
 
-We have added three new implementations to Task Bench: Dask,
+We have added four new implementations to Task Bench: Dask, MPI bulk synchronous,
 MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
@@
 
-  * New implementations: Dask, MPI+OpenMP and MPI+CUDA
+  * New implementations: Dask, MPI-bulk sync, MPI+OpenMP and MPI+CUDA
   * New sections
@@
 
+We have implemented a MPI-bulk sync style by adding a MPI_Barrier after 
+computation to make sure allcomputations are done before communication. 
+Besides, we have also implemented a MPI+OpenMP,which also can be
+considerred as BSP mode as all OpenMP threads are synchronized before
+entering communication stage.
+
 One of the salient features of Task Bench is that it permits the
@@
 problem size is simulated by controlling the number of iterations in
-these loops.
+these loops. Listing 1 is an example.
 
@@
 implementations can be quite simple.
+
+TODO: explain MPI+CUDA?
 
@@
 percentage of the highest throughput achieved with any problem
-size). We use peak performance with Task Bench as it gives us a
+size). We measured flops and memory bandwidth as 
+peak performance with Task Bench as it gives us a
 straightforward measure that provides assurance that all

### 2019-05-24 reviews/response.md
Update response.

--- 2019-05-23 reviews/response.md
@@
 
-We have added three new implementations to Task Bench: Dask,
+We have added four new implementations to Task Bench: Dask, bulk synchronous MPI,
 MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
@@
 
-  * New implementations: Dask, MPI+OpenMP and MPI+CUDA
+  * New implementations: Dask, bulk synchronous MPI, MPI+OpenMP and MPI+CUDA
   * New sections
@@
 metrics. METG differs from these metrics in that can simultaneously
-(a) capture the behavior of the sytem at the limit of its scalability
+(a) capture the behavior of the system at the limit of its scalability
 (which is not true of weak/strong scaling) and (b) while achieving a
@@
 
+We have added a bulk synchronous implementation in MPI which uses
+MPI_Barrier to enforce the boundary between communication and
+computation phases. Also, the new MPI+OpenMP implementation also
+shares some characteristics with BSP as the OpenMP parallel for
+construct acts as an implicit barrier for the cores within a node.
+
 One of the salient features of Task Bench is that it permits the
@@
 
+We have added an initial evaluation of MPI+CUDA. Our implementation
+uses an offload model in which data is copied to and from the GPU on
+each timestep. We find that even when using multiple ranks per GPU to
+achieve concurrency in data movement and task scheduling, GPUs
+increase the overhead of scheduling small tasks quite substantially.
+
 Reviewer 5:
@@
 include a small number of systems, or describe systems but do not
-include an empirical evalution of performance. When there are
+include an empirical evaluation of performance. When there are
 follow-up papers which describe additional implementations, they
@@
 percentage of the highest throughput achieved with any problem
-size). We use peak performance with Task Bench as it gives us a
-straightforward measure that provides assurance that all
+size). We use peak performance as the measure of efficiency in Task
+Bench because it makes it straightforward to verify that all
 implementations are configured correctly.

### 2019-05-24 2_metg.tex
Apply Alex's comment.

@@
 to stop. In Figure~\ref{fig:strong-scaling-mpi} the problem size
-$2^{18}$ begins to peter out at 64 nodes, the point at which the
+$2^{18}$ strong scales to 64 nodes, the point at which the
 scaling curve intersects METG(50\%).

### 2019-05-24 7_conclusion.tex
Update conclusion.

@@
 
-Task Bench is a new approach for evaluating the performance
-of parallel and distributed programming systems. We have seen that
-systems vary by orders of magnitude in runtime overhead, as measured
-by METG. Widely used systems in industry have METG measured in
-seconds, reflecting the very coarse tasks and lack of need for strong
-scaling performance in current workloads. At the other extreme,
-traditional HPC systems have METG in the range of 10-100 \textmu{}s,
-necessary for strong scaling to small task granularities, though we
-have shown that METG also increases significantly for these systems
-with the size of the cluster and the use offload accelerators. The newer
-task-based systems, with METG in the range of 100 \textmu{}s to 1 ms
-have performance sufficient for weak scaling many HPC workloads, but
-more work is needed to strong scale the most demanding applications.
+Task Bench is a new approach for evaluating the performance of
+parallel and distributed programming systems. By separating the
+specification of a benchmark from implementations in various
+programming systems, it is possible to explore a broad space of
+application scenarios and to do so with a large number of programming
+systems. Our experiments have enabled the following
+insights:
+
+\begin{itemize}
+
+\item METG for current distributed programming systems varies by over
+  5 orders of magnitude.  Clearly understanding the needed task
+  granularity is an important consideration in choosing a programming
+  system for a new application.
+
+\item While some systems support task granularities as small as a few
+  microseconds, this applies only to small CPU-based clusters. Adding
+  accelerators or moving to clusters of hundreds of nodes raises the
+  METG that any system can achieve by over an order of magnitude.
+
+\item Systems that support asynchronous execution show benefits under
+  load imbalanced workloads, or workloads with balanced computation
+  and communication. However, these gains can be easily nullified by
+  high baseline overheads.
+
+\item Systems for large scale data analysis require very large tasks
+  (in the tens of seconds) to scale beyond small numbers of nodes,
+  reflecting the very coarse tasks and lack of need for strong scaling
+  performance in current workloads.
+
+\item The newer task-based systems have performance sufficient for
+  weak scaling many HPC workloads, but more work is needed to strong
+  scale the most demanding applications.
+
+\end{itemize}
 

### 2019-05-24 reviews/response.md
Apply Alex's comments to response.
Update response.

@@
   * New implementations: Dask, bulk synchronous MPI, MPI+OpenMP and MPI+CUDA
-  * New sections
-      - Implementation: sections 4.3 (Dask) and 4.5 (MPI+X)
-      - Evaluation: sections 5.6 (weak/strong scaling) and 5.9 (GPUs)
-  * New figures
+  * New sections:
+      - Implementation: Sections 4.3 (Dask) and 4.5 (MPI+X)
+      - Evaluation: Section 5.9 (GPU experiment)
+  * New figures:
       - Listings 1 and 2 (code samples)
-      - Figures 11, 12, 15 (experiments)
+      - Figure 15 (GPU experiment)
+  * Updated Figures 3 and 4 (see also lines 303-315)
+  * Updated Section 7 with insights derived from Task Bench
 
@@
 
-Weak/strong scaling and TPS are all application and system dependent
+Weak/strong scaling and TPS are also application and system dependent
 metrics. METG differs from these metrics in that can simultaneously

### 2019-05-24 4_implementation.tex
Hide weak/strong scaling figures and fiddle with figure placement.
More changes to shrink text.

@@
 array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
-dependencies are all available. Collectives synchronize the
-completion of the task graph execution.
+dependencies are all available.
+
+%% Collectives synchronize the
+%% completion of the task graph execution.
 
@@
 parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
-for OpenMP is very similar to OmpSs and uses tasks with
-task dependencies. We tested both GNU GOMP and Intel KMP
-implementations of OpenMP and found KMP to be better performing on Cori.
+for OpenMP is similar to OmpSs and uses
+task dependencies. We tested with GNU GOMP and Intel KMP
+and found KMP to be better performing.
 
@@
 
-Tasks in Realm are executed in a deferred manner, with dependencies
-between tasks explicitly defined by \emph{events} passed from one task
+Tasks in Realm are asynchronous, with dependencies
+defined by \emph{events} passed from one task
 to another. Realm's data model supports collections that live in a
@@
 synchronized with tasks via events.
-
 Realm also supports a \emph{subgraph} API that enables additional
-optimizations. We report numbers for the subgraph version of Task
-Bench as it provides better performance.
+optimizations. We use subgraphs in the Task
+Bench implementation.
 

### 2019-05-24 5_evaluation.tex
Hide weak/strong scaling figures and fiddle with figure placement.
More changes to shrink text.

@@
 CUDA 9.1.85.
-
-% FIXME: get Piz Daint software versions.
 
@@
 
+\input{e3_metg_compute}
+
 Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
@@
 \subsection{Memory Kernel Performance}
-
-% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
-\input{e3_metg_compute}
 
@@
 Spark, Swift/T and TensorFlow
-are omitted from comparisons with more complicated dependencies, as the higher overheads of these systems would have required
+are omitted from comparisons with more complicated dependencies, as the higher overheads of these systems require
 unreasonably large problem sizes (beyond what can be completed in 6
@@
 
-\input{e8_weak}
-\input{e9_strong}
 \input{e5_imbalance}
@@
 
+\zap{
+\input{e8_weak}
+\input{e9_strong}
 \subsection{Comparison with Weak and Strong Scaling}
-
-\input{e6_communication}
 
@@
 overhead.
+} % zap
+
+\input{e6_communication}
 
@@
 
-The difference between the remaining systems is due primarily to
+The remaining differences are due primarily to
 different scheduling behaviors. The execution of 4
@@
 efficiency at large task granularities. However, the use of
-work-stealing queues can also impact throughput at very small task
+work-stealing queues can also impact throughput at small task
 granularities.
 For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
-We do not consider Charm++ load balancers here because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).
-We leave the analysis of persisitent load imbalance to future work.
+We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).
+We leave analysis of persisitent load imbalance to future work.
 

### 2019-05-24 reviews/response.md
Apply comments by Sean.

@@
       - Figure 15 (GPU experiment)
-  * Updated Figures 3 and 4 (see also lines 303-315)
+  * Updated Figures 3 and 4 (see also lines 300-312)
   * Updated Section 7 with insights derived from Task Bench

### 2019-05-24 0_abstract.tex
Apply comments by Sean.

@@
 \begin{abstract}
-% 147 words
+% 148 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
@@
 programming systems under a variety of application scenarios. Task
-Bench lowers the barrier to benchmarking multiple systems by making
+Bench lowers the barrier to benchmarking multiple programming systems by making
 the implementation for a given system orthogonal to the benchmarks

### 2019-05-24 1_introduction.tex
Fix Dask citation.

@@
 implementations in 15 different programming systems:
-Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
+Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
 OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},

### 2019-05-24 5_evaluation.tex
Fiddle with figure placement.
Switch to new GPU figure.

@@
 
-\input{e10_cuda_efficiency}
-
 Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
@@
 $4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
-FLOP/s. Note that the x-axis is measured in the number of FLOPs
-performed during the run, as the GPU kernels must be given more work to
-achieve peak performance.
+FLOP/s. Note that the CPU and GPU kernels perform different numbers of
+operations as the GPU must be given more work to achieve peak
+performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
+normalized so that the number of FLOPs is constant for a given problem
+size.
 
@@
 MPI's very lightweight mechanisms.
+
+\input{e10_cuda_efficiency}

### 2019-05-24 7_conclusion.tex
Add other systems to conclusion.

@@
 
+We have implemented Task Bench for a broad set of programming systems
+spanning large scale data analytics and HPC. However, there are
+systems not represented in our evaluation that would be interesting to
+consider in the future. These include GASNet~\cite{GASNET07},
+Habanero~\cite{Habanero11}, Hadoop~\cite{Hadoop},
+HPX~\cite{Kaiser2014}, Nimbus~\cite{Nimbus17}, OCR~\cite{OCR14},
+OpenSHMEM~\cite{OpenSHMEM10}, Ray~\cite{Ray18}, and UPC~\cite{UPC99}.
+
 Not considered in our analysis is the impact of programming system

### 2019-05-24 5_evaluation.tex
Trim some text.

@@
 \lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
-ranks per GPU to push work to the GPU in parallel and thus achieve
-some overlap between communication and computation. We see that the
-overdecomposed \lstinline{w4} configuration achieves higher FLOP/s but
+ranks per GPU to push work to the GPU in parallel and thus overlaps communication and computation. We see that the
+\lstinline{w4} configuration achieves higher FLOP/s but
 drops more rapidly at smaller problem sizes, due to the overhead of

### 2019-05-24 reviews/response.md
Update response.
Fix GPU figure number.

@@
       - Listings 1 and 2 (code samples)
-      - Figure 15 (GPU experiment)
+      - Figure 13 (GPU experiment)
   * Updated Figures 3 and 4 (see also lines 300-312)
@@
 Reviewer 2:
+
+The conclusion has been updated with references to additional systems.
 

### 2019-05-25 reviews/response.md
Fix section references in response.

@@
       - Implementation: Sections 4.3 (Dask) and 4.5 (MPI+X)
-      - Evaluation: Section 5.9 (GPU experiment)
+      - Evaluation: Section 5.8 (GPU experiment)
   * New figures:
@@
 overheads will even out across tasks.
-
-A comparison with weak/strong scaling has been added in Section 5.6.
 

### 2019-05-25 1_introduction.tex
Apply Alex's comments.

@@
 to a level of quality sufficient for comparative
-studies. Task Bench is provides a core API that is shared across the
+studies. Task Bench provides a core API that is shared across the
 various system implementations. New benchmarks created with Task Bench
@@
 
-To address the second issue and isolate the contribution of the programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG).
+To address the second issue and characterize the contribution of a programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG).
 Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak

### 2019-05-25 4_implementation.tex
Fix Dask citation.
Apply Alex's comments.

@@
 
-Dask is an implicitly parallel task-based programming system for large
+Dask~\cite{Dask15} is an implicitly parallel task-based programming system for large
 scale data analytics with Python. Dask provides abstractions over
@@
 optimizations applied to further reduce dynamic checks for what tasks
-should be executed on a node, resulting in further improvements to
+should be executed on a node, which further improves
 scalability.

### 2019-05-25 5_evaluation.tex
Update radix numbers.
Apply Alex's comments.

@@
 to systems which share all cores between the application and runtime. Some of the
-slower systems struggle to achieve peak FLOP/s, though in most cases
+higher overhead systems struggle to achieve peak FLOP/s, though in most cases
 the curves suggest that performance would continue to improve if we
@@
 %% From 1 to 3
-%% dependencies, the ratio varies from $0.54\times$ to $7.3\times$ (median
-%% $1.8\times$).
+%% dependencies, the ratio varies from $1.0\times$ to $150\times$ (median
+%% $2.3\times$).
 
 The ratio in METG between 0 and 3 dependencies per task ranges from
-$1.01\times$ to $16\times$ (median $2.9\times$). The difference is most pronounced in systems that
+$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
 perform runtime work inline. For example, MPI achieves an METG of 390
@@
 solely to runtime use). Some systems, like Charm++, Realm, and Regent,
-even support both configurations.
+support both configurations.
 
@@
 dependence patterns supported by Task Bench. METG(50\%) is calculated
-separately at each node count, to isolate runtime system behavior from
+separately at each node count, to distinguish runtime system behavior from
 changes in communication latencies and topology when using
@@
 the systems with the smallest METG on one node have roughly an order
-of magnitude higher METG at 256 nodes. The obvious conclusion is that
-increased communication latency requires significantly larger tasks to
+of magnitude higher METG at 256 nodes---increased communication latencies require significantly larger tasks to
 achieve the same level of efficiency.
@@
 We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).
-We leave analysis of persisitent load imbalance to future work.
+We leave analysis of persistent load imbalance to future work.
 
@@
 compute kernel achieves a peak performance of $4.759 \times 10^{12}$
-FLOP/s, which compares favorably with the officially reported number
+FLOP/s, which is very close to the officially reported number
 $4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
@@
 of copying data to and from the GPU dominates at small task
-granularities. Thus at smaller problem sizes the CPU achieves higher
+granularities, where the CPU achieves higher
 performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
@@
 the same efficiency than running on the CPU, reducing the advantage of
-MPI's very lightweight mechanisms.
+very lightweight mechanisms such as those in MPI.
 

### 2019-05-25 6_related.tex
Apply Alex's comments.

@@
 for writing network performance tests. \textsc{coNCePTuaL} and Task
-Bench both enable the easy creation of new benchmarks though
+Bench both enable the easy creation of new benchmarks, though
 \textsc{coNCePTuaL} does so via scripting whereas Task Bench provides
@@
 in Section~\ref{subsec:number-of-dependencies} underestimates (often
-substantially) the cost of scheduling a task, and can also impact scalability.
+substantially) the cost of scheduling a task and can also impact scalability.

### 2019-07-20 1_introduction.tex
revised intro

@@
 
-There is a second important issue: How does one predict how
-the performance of a mini-app generalizes to a full
-application, even one closely related? Part of the problem is
-that \emph{weak} and \emph{strong} scaling, by far the most commonly
-reported performance metrics,
-do not directly characterize the performance of the underlying
-programming system. Weak scaling can hide arbitrary amounts of runtime
-system overhead with sufficiently large problem sizes, and strong
-scaling does not separate runtime system overhead from costs
-(such as increased communication) that scale with the number of nodes when
-using progressively larger portions of a machine.
+We present Task Bench, a benchmark designed for exploring the performance
+behaviors of parallel and distributed programming systems under a
+variety of conditions.  The key property of TaskBench is that it completely separates
+the implementation of TaskBench for a new system to be measured from the implementation
+of a new benchmark application to be run on the systems supported by TaskBench.
+In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
+systems is ${\cal O}(mn)$.  TaskBench's design reduce the work to ${\cal O}(m) + {\cal O}(n)$,
+enbaling dramatically more systems and benchmarks to be explored for the same amount of programming
+effort.  New benchmarks created with Task Bench
+immediately run on all systems, and new systems immediately run all
+benchmarks.
 
-To address the first issue, we present Task Bench, a
-\emph{parameterized} benchmark designed for exploring the performance
-behaviors of parallel and distributed programming systems under a
-variety of conditions. A critical property of Task Bench is that,
-while able to express a wide variety of typical parallel programming
-patterns, it is simple enough to be implemented with reasonable effort
-to a level of quality sufficient for comparative
-studies. Task Bench provides a core API that is shared across the
-various system implementations. New benchmarks created with Task Bench
-immediately run on all systems, and new systems immediately run all
-benchmarks. Thus, Task Bench is much more productive for benchmarking
-a wide range of phenomena than the standard approach in which each
-benchmark must be implemented separately for each system.
-
-A key insight of Task Bench is that many applications can be modeled
+Benchmarks in Task Bench are based on the observation that many applications can be modeled
 as a set of \emph{tasks}, or coarse-grain units of work, with
 dependencies between tasks representing the communication and synchronization required for
-distributed execution. Task Bench consists of a configurable \emph{task graph} with tasks for
+distributed execution. A benchmark consists of a parameterized \emph{task graph} with tasks for
 each point in an iteration space. Tasks are connected by dependencies
@@
 
-To address the second issue and characterize the contribution of a programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG).
-Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
-performance (e.g., FLOP/s or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
-or strong scaling study, the behavior at the limit of scalability. For
-weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency.
-For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.
+Adding a new programming system to Task Bench involves  implementing a set of standard
+services, such as executing a task or a data transfer.  This interface is simple enough to 
+be implemented with reasonable effort to a level of quality sufficient for comparative
+studies. Task Bench also provides a core library that encapsulates the functionality
+shared by all system implementations, which not only reduces the implementation effot
+required but also makes it much easier to achieve truly apples-to-apples comparisons
+between systems.
 
-A lower METG does not necessarily mean that performance for a particular workload is
-better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
+Using Task Bench we were able to quickly incorporate 13 very different parallel and distributed
+runtime systems.  By running all systems on common benchmarks we were able to quantify phenomena
+that have never before been measured.
+Most strikingly, the performance of the various systems we examine varies by more than five orders
+of magnitude, with very popular, widely used systems at both ends of the spectrum!  Clearly
+the slower systems have ``good enough'' performance for some class of applications, while presumably
+providing advantages in programmer productivity.
+
+These results point to a natural question:   How does one predict whether the performance of a runtime
+system will in fact be good enough for a particular application?  In trying to answer this question,
+we realized that \emph{weak} and \emph{strong} scaling, by far the most commonly
+reported performance metrics, do not directly characterize the performance of the underlying
+programming system. Weak scaling can hide arbitrary amounts of runtime
+system overhead by using sufficiently large problem sizes, and strong
+scaling does not separate runtime system overhead from costs
+(such as increased communication) that scale with the number of nodes when
+using progressively larger portions of a machine.
+
+To characterize the contribution of a programming model's runtime
+system to application performance, and as an example of the novel studies that can be done
+with Task Bench, we introduce a new metric called
+\emph{minimum effective task granularity} (METG). Intuitively, for a given
+workload, METG(50\%) is the smallest task granularity that maintains
+at least 50\% efficiency, meaning that the application achieves at
+least 50\% of peak performance (e.g., FLOP/s or memory bandwidth) on a
+given machine. METG captures the essence of what is important in a
+weak or strong scaling study, the behavior at the limit of
+scalability. For weak scaling, METG(50\%) corresponds directly to the
+smallest problem size that can be weak-scaled with 50\%
+efficiency. For strong scaling, METG(50\%) can be used to compute the
+scale at which performance can be expected to dip below 50\%
+efficiency.  We note that METG(50\%) for a given runtime system will
+vary with the particular application and the underlying hardware---i.e., METG(50%)
+is not a constant for a given system, but we will see that individual systems have
+a characteristic ranges of METG(50\%) values and that there is additional insight
+in the reasons that the mininum effective task granularity can vary.
+
+Another important characteristic is that a lower METG does not necessarily mean that 
+performance for a particular workload is better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
 respectively, running an application with an average task granularity
@@
 likely diverge. Thus METG helps to identify the regime in which a
-given system can deliver good performance.
+given system can deliver good performance, and helps to explain how
+different systems coexist with METG(50\%) values that vary by orders of magnitude.
 
+%
+% FIXME:  Move these two paragraphs out of the introduction and into one of the METG sections.
+%
 Task Bench and METG address issues common in limit studies of runtime
@@
 \begin{itemize}
-\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
-  relationship to quantities of interest to application developers.
 \item Section~\ref{sec:task-bench} describes the design of Task Bench.
@@
   Task Bench in 15 programming systems.
+\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
+  relationship to quantities of interest to application developers.
 \item Section~\ref{sec:evaluation} provides a comprehensive evaluation

### 2019-07-22 sc19_reviews/response.md
Move SC19 reviews to the side.

--- 2019-05-25 reviews/response.md
@@
 METG captures the average task granularity that can be supported. It
-is acceptable for some tasks to be smaller and others larger, as
-overheads will even out across tasks.
+is acceptable for some tasks to be smaller and others larger. Overhead
+is a function of number of tasks times overhead per task and METG
+captures this well.
+
+The conclusion has been updated with specific insights derived from
+Task Bench.
 

### 2019-07-22 1_introduction.tex
Fix latex build.

@@
 In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
-systems is ${\cal O}(mn)$.  TaskBench's design reduce the work to ${\cal O}(m) + {\cal O}(n)$,
+systems is $\mathcal{O}(mn)$.  TaskBench's design reduce the work to $\mathcal{O}(m) + \mathcal{O}(n)$,
 enbaling dramatically more systems and benchmarks to be explored for the same amount of programming

### 2019-07-22 1_introduction.tex
Run spell check.

@@
 behaviors of parallel and distributed programming systems under a
-variety of conditions.  The key property of TaskBench is that it completely separates
-the implementation of TaskBench for a new system to be measured from the implementation
-of a new benchmark application to be run on the systems supported by TaskBench.
+variety of conditions.  The key property of Task Bench is that it completely separates
+the implementation of Task Bench for a new system to be measured from the implementation
+of a new benchmark application to be run on the systems supported by Task Bench.
 In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
-systems is $\mathcal{O}(mn)$.  TaskBench's design reduce the work to $\mathcal{O}(m) + \mathcal{O}(n)$,
-enbaling dramatically more systems and benchmarks to be explored for the same amount of programming
+systems is $\mathcal{O}(mn)$.  Task Bench's design reduce the work to $\mathcal{O}(m + n)$,
+enabling dramatically more systems and benchmarks to be explored for the same amount of programming
 effort.  New benchmarks created with Task Bench
@@
 studies. Task Bench also provides a core library that encapsulates the functionality
-shared by all system implementations, which not only reduces the implementation effot
+shared by all system implementations, which not only reduces the implementation effort
 required but also makes it much easier to achieve truly apples-to-apples comparisons
@@
 a characteristic ranges of METG(50\%) values and that there is additional insight
-in the reasons that the mininum effective task granularity can vary.
+in the reasons that the minimum effective task granularity can vary.
 

### 2019-07-22 1_introduction.tex
Edits to introduction.

@@
 
-We present Task Bench, a benchmark designed for exploring the performance
+We present Task Bench, a parameterized benchmark designed for exploring the performance
 behaviors of parallel and distributed programming systems under a
 variety of conditions.  The key property of Task Bench is that it completely separates
-the implementation of Task Bench for a new system to be measured from the implementation
-of a new benchmark application to be run on the systems supported by Task Bench.
+the system-specific implementation from the core implementation
+of the benchmarks themselves.
 In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
-systems is $\mathcal{O}(mn)$.  Task Bench's design reduce the work to $\mathcal{O}(m + n)$,
+systems is $\mathcal{O}(mn)$.  Task Bench's design reduces this work to $\mathcal{O}(m + n)$,
 enabling dramatically more systems and benchmarks to be explored for the same amount of programming
@@
 dependencies between tasks representing the communication and synchronization required for
-distributed execution. A benchmark consists of a parameterized \emph{task graph} with tasks for
+distributed execution. A benchmark consists of a \emph{task graph} with tasks for
 each point in an iteration space. Tasks are connected by dependencies
@@
 be implemented with reasonable effort to a level of quality sufficient for comparative
-studies. Task Bench also provides a core library that encapsulates the functionality
+studies. Task Bench provides a core library that encapsulates the functionality
 shared by all system implementations, which not only reduces the implementation effort
@@
 
-Using Task Bench we were able to quickly incorporate 13 very different parallel and distributed
+Using Task Bench we were able to quickly incorporate 15 very different parallel and distributed
 runtime systems.  By running all systems on common benchmarks we were able to quantify phenomena
@@
 at least 50\% efficiency, meaning that the application achieves at
-least 50\% of peak performance (e.g., FLOP/s or memory bandwidth) on a
-given machine. METG captures the essence of what is important in a
+least 50\% of the higest performance (in FLOP/s, B/s, or another
+application-specific measure) that can be achieved on a given
+machine. The efficiency bound in METG is a key innovation over
+previous approaches, such as \emph{tasks per second} (TPS), that fail
+to quantify the amount of useful work peformed (if tasks are
+non-empty~\cite{Canary16, Armstrong14}) or to perform any useful work at all (if tasks are empty~\cite{LegionTracing18}).
+
+METG captures the essence of what is important in a
 weak or strong scaling study, the behavior at the limit of
@@
 efficiency.  We note that METG(50\%) for a given runtime system will
-vary with the particular application and the underlying hardware---i.e., METG(50%)
+vary with the particular application and the underlying hardware---i.e., METG(50\%)
 is not a constant for a given system, but we will see that individual systems have
-a characteristic ranges of METG(50\%) values and that there is additional insight
-in the reasons that the minimum effective task granularity can vary.
+a characteristic range of METG(50\%) values and that there is additional insight
+in the reasons that METG can vary.
 
-Another important characteristic is that a lower METG does not necessarily mean that 
+Note that a lower METG does not necessarily mean that 
 performance for a particular workload is better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
@@
 likely diverge. Thus METG helps to identify the regime in which a
-given system can deliver good performance, and helps to explain how
-different systems coexist with METG(50\%) values that vary by orders of magnitude.
+given system can deliver good performance, and explains how
+different systems coexist with runtime overheads that vary by orders of magnitude.
 
@@
 %
+\zap{
 Task Bench and METG address issues common in limit studies of runtime
@@
 overhead impacts the ability to perform useful application work.
+} % zap
 

### 2019-07-22 4_metg.tex
Flip paper sections around.

--- 2019-05-24 2_metg.tex
@@
 the smallest task granularity achieved while maintaining at least 0.63
-TFLOP/s. For applications not amenable to being characterized by peak
-performance, an application-specific measure of performance can be
+TFLOP/s per node. For applications not amenable to being characterized by raw FLOP/s or B/s, an application-specific measure of performance can be
 used instead (e.g., mesh cells processed per second).
@@
 very bottom, the lines compress together as running time becomes
-dominated by overhead. The red, dashed line in each graph marks
-METG(50\%), the point where efficiency reaches 50\%. Note that the
+dominated by overhead. Note that the
 shapes at the bottom of the strong and weak scaling curves are identical,
-and conform to the shape of the METG curve.
+and conform to the shape of the METG curve (marked by the red, dashed line).
 

### 2019-07-22 1_introduction.tex
Collapse itemize.

@@
 
-% The paper is organized as follows:
-This paper makes the following contributions:
-
-% FIXME: if you need more space you can collapsed the itemize:
-
-\begin{itemize}
-\item Section~\ref{sec:task-bench} describes the design of Task Bench.
-\item Section~\ref{sec:implementation} details the implementation of
-  Task Bench in 15 programming systems.
-\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
-  relationship to quantities of interest to application developers.
-\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
-  of Task Bench on up to 256 Haswell nodes of the Cori
-  supercomputer~\cite{Cori}.
-\end{itemize}
-
-Section~\ref{sec:related-work} relates this work to previous efforts,
-and Section~\ref{sec:conclusion} concludes.
+The paper is organized as follows: Section~\ref{sec:task-bench}
+describes the design of Task Bench, Section~\ref{sec:implementation}
+details the implementation of Task Bench in 15 programming systems,
+Section~\ref{sec:metg} defines the METG(50\%) metric and its
+relationship to quantities of interest to application developers,
+Section~\ref{sec:evaluation} provides a comprehensive evaluation of
+Task Bench on up to 256 Haswell nodes of the Cori
+supercomputer~\cite{Cori}, Section~\ref{sec:related-work} relates this
+work to previous efforts, and Section~\ref{sec:conclusion} concludes.

### 2019-07-22 2_task_bench.tex
Tweak text and font sizes.

--- 2019-05-23 3_task_bench.tex
@@
 
+% FIXME: This seems a bit abrupt but the old intro doesn't really make sense any more.
+\zap{
 To measure a wide variety of
 systems and application scenarios, we developed a novel
-parameterized benchmark called Task Bench. Task Bench
+parameterized benchmark called Task Bench.
+} % zap
+
+Task Bench
 implementations execute a \emph{task graph}, with tasks for each point
@@
 show excerpts from the Task Bench core, and the Dask implementation,
-respectively. Only the code in \ref{lst:code-sample} need be
+respectively. Only the code in Listing~\ref{lst:code-sample} is
 implemented for each system, minimizing the work required for each additional system.

### 2019-07-22 4_metg.tex
Make equation fit on line.

@@
 (i.e., as a percentage of the peak FLOP/s achieved) and task
-granularity (i.e., $\text{wall time} \times \text{number of
-  cores}/\text{number of tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
+granularity (i.e., $\text{wall time} \times \text{num.~cores}/\text{num.~tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
 broadly to be any continuously-executing unit of application code,

### 2019-07-23 t2_parameters.tex
Add table of task bench parameters.

\begin{table}[t]
\small

General options:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Values & Purpose \\
\hline

height & height of graph & number of timesteps \\
width & width of graph & degree of parallelism \\
dependence & trivial, stencil, etc. & dependence relation \\
kernel & compute, memory, etc. & type of kernel \\
output & bytes per dependency & degree of comm.
\end{tabular}

\vspace{1em}

Options specific to dependence relations:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Valid For & Purpose \\
\hline
radix & nearest & dependencies per task
\end{tabular}

\vspace{1em}

Options specific to kernels:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Valid For & Purpose \\
\hline
iterations & compute, memory & task duration/problem size \\
span & memory & bytes used per task per iter. \\
scratch & memory & total working set size \\
imbalance & load imbalanced & degree of imbalance
\end{tabular}

\caption{Task Bench parameters.\label{tab:parameters}}
\vspace{-0.5cm}
\end{table}

### 2019-07-23 2_task_bench.tex
Shuffle figures around.
Add table of task bench parameters.
Edits for top of section 2.

@@
 
-% FIXME: This seems a bit abrupt but the old intro doesn't really make sense any more.
-\zap{
-To measure a wide variety of
-systems and application scenarios, we developed a novel
-parameterized benchmark called Task Bench.
-} % zap
+\input{t2_parameters}
 
-Task Bench
-implementations execute a \emph{task graph}, with tasks for each point
-in an \emph{iteration space} and dependencies between tasks determined
-by a \emph{dependence relation}. A task $A$ that depends on another task $B$
-must execute after task $B$, and receives the output data from
-task $B$ as input. Each task can execute any one of a
-number of kernels (compute-bound, memory-bound, etc.) and can generate
-a configurable amount of output to control the amount of communication
-performed with each task dependency. Thus Task Bench can explore a large space of application scenarios to
-understand the performance of parallel and
-distributed programming systems.
+To explore as broad a space of application scenarios as possible, Task
+Bench provides a large number of configuration parameters. The most
+important parameters are described in
+Table~\ref{tab:parameters}. These parameters control the size and
+structure of the task graph, the type and duration of the kernels
+associated with each task, and the amount of data associated with
+each dependence edge in the graph.
 
+Task graphs are a combination of an iteration space (with a task for
+each point in the space) with a dependence relation.
 For simplicity, but without loss of generality, the iteration space in
@@
 
+\input{l0_compute_kernel}
+\input{l1_code_sample}
+
 The Task Bench core library is fully
@@
 
-\input{l0_compute_kernel}
-\input{l1_code_sample}
-
 Task Bench provides two main kernels that can be called from tasks,

### 2019-07-23 3_implementation.tex
Split body of implementation section into the suplemental materials.

--- 2019-05-25 4_implementation.tex
@@
 \input{t1_systems}
-
-\subsection{Chapel}
-
-Chapel~\cite{Chapel15} is a parallel programming language
-with a \emph{multi-res\-o\-lu\-tion} approach, supporting parallelism at a variety of
-levels. Chapel's core features are a partitioned global address space
-(PGAS), data distributions, tasks,
-synchronization primitives, and array promotion. For Task Bench, we target a low level of
-Chapel, using explicit task instantiation (via
-\lstinline[language=Chapel]{coforall}), bulk access to distributed
-arrays for efficient data movement, and atomic integers for
-synchronization.
-
-\subsection{Charm++}
-
-Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors, or
-\emph{chares}, are objects in their own address space.
-Chares communicate data and synchronize via messages and can be moved
-to balance load. Our Task Bench implementation uses a chare
-array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
-dependencies are all available.
-
-%% Collectives synchronize the
-%% completion of the task graph execution.
-
-% FIXME: Discuss SMP vs non-SMP versions
-
-\subsection{Dask}
-
-Dask~\cite{Dask15} is an implicitly parallel task-based programming system for large
-scale data analytics with Python. Dask provides abstractions over
-distributed NumPy arrays and Pandas dataframes, and also a lower-level
-interface for launching tasks. We use the lower-level interface to
-implement Task Bench.
-
-\subsection{MPI}
-
-Our MPI~\cite{MPI} implementation of Task Bench represents the common
-case of point-to-point communication with distinct computation and
-communication phases. We experimented with a variety of implementation
-strategies and found the best performing to be using
-\lstinline[language=C++]{MPI_Isend} and
-\lstinline[language=C++]{MPI_Irecv} to implement the communication
-phase, posting receives before sends. Each task dependency maps to one
-send/receive pair in MPI. We also provide a bulk synchronous
-implementation which enforces the boundary between communication and
-computation with \lstinline[language=C++]{MPI_Barrier}.
-
-\subsection{MPI+X}
-
-We provide two MPI+X implementations of Task Bench to evaluate the
-performance of hierarchical programming models. Our MPI+OpenMP
-implementation uses forall-style parallel loops to execute tasks,
-interleaved with MPI point-to-point communication as above. Our
-MPI+CUDA implementation follows an offload model where data is copied
-to and from the GPU on every timestep.
-
-\subsection{OmpSs}
-
-OmpSs~\cite{OmpSs11} is a programming model for loop- and task-based parallelism
-that is source-compatible with OpenMP. Our Task Bench implementation
-uses OpenMP 4.0-style task dependencies. Because OpenMP tasks have a fixed number of dependencies, we use a switch
-statement to implement the dynamic dependencies required for Task
-Bench. The implementation is otherwise straightforward.
-
-% FIXME: More to say about OmpSs? Is there anything different vs the OpenMP implementation?
-
-\subsection{OpenMP}
-
-OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
-parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
-for OpenMP is similar to OmpSs and uses
-task dependencies. We tested with GNU GOMP and Intel KMP
-and found KMP to be better performing.
-
-\subsection{PaRSEC}
-
-PaRSEC is a task-based programming system supporting two distinct
-programming models: \emph{parameterized task graphs}
-(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
-(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which programmers
-write a concise, algebraic description of the tasks and dataflows in
-the program. This compressed representation is expanded into a full
-task graph by a source-to-source compiler.
-
-In the DTD model, tasks are enumerated in program order (by
-executing the program), and dependencies between tasks are
-identified automatically based on the input and output data of tasks. 
-A task depends on another task if it reads data written by the other task,
-and the data is copied automatically if the two tasks are executed on
-different nodes. To
-improve scalability, the program is executed in
-SPMD fashion on all nodes, and the user is responsible for eliding
-tasks that are not directly connected to those that are to be executed
-on the current node.
-
-In our Task Bench implementation, we evaluate both PTG and DTD models,
-as well as a third option based on DTD that has additional manual
-optimizations applied to further reduce dynamic checks for what tasks
-should be executed on a node, which further improves
-scalability.
-
-%% Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
-%% distributed across nodes in pre-defined patterns,
-%% and can be passed to tasks on other nodes, but
-%% can only be written on the node where the data is originally
-%% allocated. In our Task Bench implementation, each task writes to a
-%% tile as output, and these tiles are passed as inputs to the
-%% subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
-%% the task launch code uses a switch statement as the PaRSEC API assumes
-%% tasks have a fixed number of dependencies.
-
-\subsection{Realm}
-
-Realm~\cite{Realm14} is an explicitly-parallel task-based programming
-model used internally by Legion~\cite{Legion12} and
-Regent~\cite{Regent15}. A Realm implementation of Task Bench is a limit study of what can be achieved with Legion or Regent.
-
-Tasks in Realm are asynchronous, with dependencies
-defined by \emph{events} passed from one task
-to another. Realm's data model supports collections that live in a
-specific memory. Data must be explicitly copied, and the copies
-synchronized with tasks via events.
-Realm also supports a \emph{subgraph} API that enables additional
-optimizations. We use subgraphs in the Task
-Bench implementation.
-
-\subsection{Regent}
-
-Regent~\cite{Regent15} is an implicitly-parallel task-based language
-which implements the Legion programming model~\cite{Legion12}. We use
-Regent rather than Legion directly because the Regent compiler
-provides a critical optimization that
-improves scalability~\cite{ControlReplication17}.
-
-\subsection{Spark}
-
-Spark~\cite{Spark10} is an implicitly-parallel programming model for
-large scale data analytics.
-The core abstractions in Spark are functional operators such as map,
-reduce, and join. Data in Spark are stored in
-\emph{resilient
-  distributed datasets} (RDDs): globally-visible,
-dynamically single-assignment data structures. Spark caches RDDs in memory to avoid unnecessary
-disk traffic.
-
-Spark has tasks internally but they are not exposed to the user,
-so a Task Bench implementation must map the task graph to a set of operators that result in the
-desired execution. We use
-\lstinline[language=Scala]{flatMap} and
-\lstinline[language=Scala]{groupByKey} to generate task
-dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
-series of tasks. An explicit hash partitioner ensures that
-Spark does not attempt to group multiple Task Bench tasks into a
-single Spark task, as tasks in Task Bench already represent coarse-grained units of work.
-
-We performed extensive experiments to verify that no
-extraneous factors interfered with our Spark measurements. We
-disabled logging, ensured no RDDs are written
-to disk, confirmed there is no measurable overhead due to JNI
-calls from Java to C or due to the serialization of data structures, and cross-checked results with known cases
-that hit optimal task throughput in Spark.
-
-\subsection{StarPU}
-
-StarPU~\cite{StarPU11} is a task-based system that supports a \emph{sequential task flow} (STF)
-programming model similar to PaRSEC's DTD. Our Task Bench implementation in
-StarPU is very similar to PaRSEC.
-%% We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
-%% project~\cite{Chameleon}.
-
-\subsection{Swift/T}
-
-Swift/T~\cite{Wozniak13} is a parallel scripting language intended
-primarily for the composition of HPC workflows. Swift/T
-programs follow dataflow semantics, where every statement may
-potentially execute in parallel as soon as its dependencies are
-satisfied; statements only execute sequentially when explicitly
-requested. The Swift/T compiler performs a number of optimizations to
-improve performance of highly parallel programs~\cite{Armstrong14}. Our Task Bench in Swift/T is straightforward, using Swift/T's dataflow
-semantics to capture dependencies on other tasks.
-
-\subsection{TensorFlow}
-
-TensorFlow~\cite{TensorFlow15} is a programming system designed for
-deep learning workloads. Although TensorFlow's API exposes
-machine learning concepts, internally TensorFlow is a task graph execution engine, making it a good fit for
-Task Bench. TensorFlow programs operate via explicit graph construction. Task graphs are composed in Python and run by an
-execution engine written in C++. Our Task Bench implementation is straightforward.
-
-\subsection{X10}
-
-X10~\cite{X1005} is an explicitly parallel programming language for
-place-based programming. The core features of X10 are \emph{places}
-which represent distributed memories, a PGAS model where references to
-remote objects can be held (but can only be dereferenced on the local
-place), asynchronous tasks, and a place-changing construct to move the
-execution of a task to a remote place. X10 also supports a variety of
-synchronization primitives.
-
-Our Task Bench implementation uses
-\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
-between places, along with place-changing and atomic integers for
-synchronization. We use the native backend of X10, which compiles to
-C++.

### 2019-07-23 a1_implementation.tex
Split body of implementation section into the suplemental materials.

--- 2019-05-25 4_implementation.tex
@@
 \label{sec:implementation}
-
-We have implemented Task Bench in the 15 parallel and distributed
-programming systems listed in Table~\ref{tab:systems}. We describe the
-systems, and any salient details of their Task Bench implementations,
-below.
-
-One challenge in implementing Task Bench for such a wide variety of
-systems is that the capabilities of the systems vary considerably. For
-example, some systems are \emph{implicitly parallel}, and provide some
-form of parallelism discovery from sequential programs, whereas others
-are \emph{explicitly parallel} and require users to specify the
-parallelism in the program. For systems that provide both implicit and explicit parallelism, the form of parallelism used in Task Bench is emphasized in Table~\ref{tab:systems}.
-
-%% The Task
-%% Bench implementations are intended to reflect how actual applications
-%% would be written in the respective systems. Places where the Task
-%% Bench implementations may be unidiomatic are noted below.
-
-%% There are also aspects of some systems that we do not consider. For
-%% example, tasks may run on CPUs, GPUs, or any execution resource
-%% supported by a programming system. However, the level of support for
-%% heterogeneous processors varies widely across programming systems (if it exists at all), so
-%% in order to include as many systems as possible in our comparisons, we
-%% evaluate Task Bench only on CPUs in this paper.
-
-In all cases, members of the respective programming systems' teams
-were consulted during the development and evaluation of the
-corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
-in helping to ensure that we provide the highest level of quality in the Task
-Bench implementations for each system.
-
-\input{t1_systems}
 

### 2019-07-23 3_implementation.tex
Edit implementation section.

@@
 We have implemented Task Bench in the 15 parallel and distributed
-programming systems listed in Table~\ref{tab:systems}. We describe the
-systems, and any salient details of their Task Bench implementations,
-below.
+programming systems listed in Table~\ref{tab:systems}. These include
+traditional HPC programming models: MPI~\cite{MPI} (with bulk
+synchronous and point-to-point communication) and MPI+X (with OpenMP
+and CUDA); PGAS and actor models: Chapel~\cite{Chapel15},
+Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and newer task-based
+systems: OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40} (with task
+dependencies), PaRSEC (with \emph{parameterized task graphs}
+(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
+(DTD)~\cite{PARSEC_DTD}), Realm~\cite{Realm14},
+Regent~\cite{Regent15}, and StarPU~\cite{StarPU11}. Also, we consider
+several systems used in large scale data analytics and workflows:
+Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
+and TensorFlow~\cite{TensorFlow15}. We describe the systems, and any
+salient details of their Task Bench implementations, in our
+supplemental material.
 

### 2019-07-23 4_metg.tex
More text and fiddle with figure placement.

@@
 
-\input{f1_flops_mpi}
-\input{f2_efficiency_mpi}
+Given that Task Bench permits the rapid exploration of a large space
+of application scenarios, one question is how to characterize the
+overall performance of the programming systems under study. As noted
+earlier, the overheads of the systems we consider varies by more than
+five orders of magnitude. This makes it challenging to extract useful
+information from weak and strong scaling runs.
 
-The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
+Existing studies which consider programming system efficiency
+typically use the metric \emph{tasks per second} (TPS). TPS results
+are difficult to interpret and apply, because the efficiency (and thus
+the amount of useful work) is not constrained. When using empty
+tasks~\cite{LegionTracing18}, the resulting upper bound on task
+scheduling throughput fails to represent useful work within a
+realistic application. With non-empty tasks, since the efficiency of
+the overall application is typically not reported~\cite{Canary16,
+  Armstrong14}, TPS is not a measurement of runtime-limited
+performance. Large tasks may be used to hide any amount of runtime
+overhead, while small tasks may result in a drop in total application
+throughput even as TPS increases.
+
+We introduce \emph{minimum effective task granularity}, or METG, an
+efficiency-constrained metric for runtime-limited
+performance. METG(50\%) for an application $A$ is
 the smallest average task granularity (i.e., task duration) such that $A$
@@
 used instead (e.g., mesh cells processed per second).
+
+\input{f1_flops_mpi}
+\input{f2_efficiency_mpi}
 

### 2019-07-25 0_abstract.tex
Changes to trim word count.

@@
 We present Task Bench, a \emph{parameterized} benchmark designed to
-explore the performance envelopes of parallel and distributed
+explore the performance of parallel and distributed
 programming systems under a variety of application scenarios. Task
@@
 the implementation for a given system orthogonal to the benchmarks
-themselves: all benchmarks constructed with Task Bench automatically
-run on all Task Bench implementations, and vice versa. Furthermore,
+themselves: every benchmark constructed with Task Bench
+runs on every Task Bench implementation. Furthermore,
 Task Bench's parameterization allows a wide variety of benchmark

### 2019-07-25 1_introduction.tex
Changes to trim word count.

@@
 We present Task Bench, a parameterized benchmark designed for exploring the performance
-behaviors of parallel and distributed programming systems under a
+of parallel and distributed programming systems under a
 variety of conditions.  The key property of Task Bench is that it completely separates
@@
 effort.  New benchmarks created with Task Bench
-immediately run on all systems, and new systems immediately run all
+immediately run on all systems, and new systems that implement the Task Bench interface immediately run all
 benchmarks.
@@
 studies. Task Bench provides a core library that encapsulates the functionality
-shared by all system implementations, which not only reduces the implementation effort
+shared among systems, which not only reduces the implementation effort
 required but also makes it much easier to achieve truly apples-to-apples comparisons
@@
 
-These results point to a natural question:   How does one predict whether the performance of a runtime
+This result points to a natural question:   How does one predict whether the performance of a runtime
 system will in fact be good enough for a particular application?  In trying to answer this question,
@@
 previous approaches, such as \emph{tasks per second} (TPS), that fail
-to quantify the amount of useful work peformed (if tasks are
+to consider the amount of useful work peformed (if tasks are
 non-empty~\cite{Canary16, Armstrong14}) or to perform any useful work at all (if tasks are empty~\cite{LegionTracing18}).
@@
 weak or strong scaling study, the behavior at the limit of
-scalability. For weak scaling, METG(50\%) corresponds directly to the
+scalability. For weak scaling, METG(50\%) corresponds to the
 smallest problem size that can be weak-scaled with 50\%
@@
 describes the design of Task Bench, Section~\ref{sec:implementation}
-details the implementation of Task Bench in 15 programming systems,
+discusses the implementation of Task Bench in 15 programming systems,
 Section~\ref{sec:metg} defines the METG(50\%) metric and its

### 2019-07-25 3_implementation.tex
More text and fiddle with figure placement.
Changes to trim word count.

@@
 \label{sec:implementation}
+
+\input{t1_systems}
 
@@
 corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
-in helping to ensure that we provide the highest level of quality in the Task
-Bench implementations for each system.
-
-\input{t1_systems}
+in helping to ensure that we provide the highest quality implementations for each system.

### 2019-07-25 4_metg.tex
Changes to trim word count.

@@
 task granularity of
-4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s for this configuration of
-Task Bench. We use 50\% because that is generally an acceptable level
+4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
+configuration. We use 50\% because that is generally an acceptable level
 of efficiency in practice, and values above 90\% can misrepresent the

### 2019-07-25 5_evaluation.tex
More text and fiddle with figure placement.
Changes to trim word count.

@@
 \input{e2_efficiency}
-\input{e7_bytes}
-\input{e4_radix}
 
@@
 expensive. For example, the Spark job in
-this configuration ran for over 6 hours.
+this case ran for over 6 hours.
 
-\input{e3_metg_compute}
+\input{e7_bytes}
+\input{e4_radix}
 
-Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
-peak FLOP/s) plotted against task granularity. This plot is
-used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
+Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
+peak FLOP/s) against task granularity. As described in Section~\ref{sec:metg}, this plot is
+used to calculate METG(50\%). The
 red, dashed line shows the point where 50\% efficiency is achieved.
@@
 systems that reserve one or more cores are able to hit 100\% of peak
-performance, unlike with compute-bound kernels. Otherwise the
-performance curves are similar. We are currently investigating the
+performance, unlike with compute-bound kernels. We are currently investigating the
 performance of Charm++, Regent and X10.
@@
 \subsection{Baseline Overhead}
+
+\input{e3_metg_compute}
 
@@
 
-\input{e5_imbalance}
-
 METG is a useful metric in part because it summarizes the basic task
@@
 
+\input{e5_imbalance}
 \input{e6_communication}

### 2019-07-25 2_task_bench.tex
Changes to trim word count.
Apply Alex's updates.

@@
 parameterized, such as picking the $K$ nearest neighbors, or $K$
-distant neighbors. Table~\ref{tab:equations} show equations for the
+distant neighbors. Table~\ref{tab:equations} shows equations for the
 dependence relations of the patterns in Figure~\ref{fig:task-graphs},
@@
 and all inputs are verified. An assertion is thrown if validation
-fails. This level of self-checking ensures that every execution of Task Bench, if it
+fails. These checks ensure that every execution of Task Bench, if it
 completes successfully, is correct. An evaluation of the performance impact of

### 2019-07-25 4_metg.tex
Apply Alex's updates.

@@
 
-Existing studies which consider programming system efficiency
+Existing studies of programming system efficiency
 typically use the metric \emph{tasks per second} (TPS). TPS results
@@
 METG(50\%) is the intersection of this curve with 50\% efficiency, as
-shown by the red, dashed lines in Figure~\ref{fig:flops-mpi} and
+shown by the red, dashed lines in Figures~\ref{fig:flops-mpi} and
 \ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average

### 2019-07-25 5_evaluation.tex
Apply Alex's updates.

@@
 \input{e2_efficiency}
+\input{e7_bytes}
+
+\input{e3_metg_compute}
 
@@
 this case ran for over 6 hours.
-
-\input{e7_bytes}
-\input{e4_radix}
 
@@
 
-\subsection{Number of Dependencies}
-\label{subsec:number-of-dependencies}
-
-The number of dependencies per task has a strong influence on the
-overhead incurred by each system, as shown in
-Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
-dependence pattern, when varying the number of dependencies per task
-from 0 to 9.
-
-% FIXME: make sure these numbers are up to date for final paper
-
-%% From 1 to 3
-%% dependencies, the ratio varies from $1.0\times$ to $150\times$ (median
-%% $2.3\times$).
-
-The ratio in METG between 0 and 3 dependencies per task ranges from
-$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
-perform runtime work inline. For example, MPI achieves an METG of 390
-ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
-a factor of $12\times$ increase. This is unsurprising as in the case
-of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
-issued at all, so MPI has nothing to do aside from executing kernels
-as quickly as possible. For this reason, we recommend evaluating the performance of
-programming systems under realistic workloads, rather than with
-trivial dependencies.
-
 \subsection{Baseline Overhead}
-
-\input{e3_metg_compute}
 
@@
 Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
-communication topology and latency. An interesting observation is that
+communication topology and latency. A key finding is that
 the systems with the smallest METG on one node have roughly an order
 of magnitude higher METG at 256 nodes---increased communication latencies require significantly larger tasks to
-achieve the same level of efficiency.
+achieve the same level of efficiency, so apparent differences in
+runtime overhead at small node counts can matter much less or not at
+all at larger node counts.
 
@@
 scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.
+
+\subsection{Number of Dependencies}
+\label{subsec:number-of-dependencies}
+
+\input{e4_radix}
+
+The number of dependencies per task has a strong influence on the
+overhead incurred by each system, as shown in
+Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
+dependence pattern, when varying the number of dependencies per task
+from 0 to 9.
+
+% FIXME: make sure these numbers are up to date for final paper
+
+%% From 1 to 3
+%% dependencies, the ratio varies from $1.0\times$ to $150\times$ (median
+%% $2.3\times$).
+
+The ratio in METG between 0 and 3 dependencies per task ranges from
+$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
+perform runtime work inline. For example, MPI achieves an METG of 390
+ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
+a factor of $12\times$ increase. This is unsurprising as in the case
+of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
+issued at all, so MPI has nothing to do aside from executing kernels
+as quickly as possible. For this reason, we recommend evaluating the performance of
+programming systems under realistic workloads, rather than with
+trivial dependencies.
 
@@
 tasks to and from the GPU imposes a higher task granularity to achieve
-the same efficiency than running on the CPU, reducing the advantage of
+the same efficiency than running on the CPU, reducing the advantage at small task granularities of
 very lightweight mechanisms such as those in MPI.

### 2019-07-25 6_related.tex
Apply Alex's updates.

@@
 systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI), each of
-which require a wholly separate, tuned implementation (in contrast to
+which require a separate, tuned implementation (in contrast to
 Task Bench). Other

### 2019-07-25 5_evaluation.tex
Move load balancing results to supplemental material and fiddle with figure placement and spacing.

@@
 
+Load balancing results are in the supplemental material.
+
 \subsection{Compute Kernel Performance}
@@
 \input{e2_efficiency}
-\input{e7_bytes}
-
-\input{e3_metg_compute}
 
@@
 
+\input{e7_bytes}
+
 Figure~\ref{fig:bytes} shows the performance of Task Bench
@@
 
-\input{e4_radix}
-
 The number of dependencies per task has a strong influence on the
@@
 
+\input{e4_radix}
+
 \zap{
@@
 
-\input{e5_imbalance}
 \input{e6_communication}
-
-\subsection{Load Imbalance}
-
-One advantage of systems with asynchronous execution capabilities is
-the ability to mitigate load imbalance with little or no additional programmer effort, especially in the presence of
-task parallelism. To quantify this effect,
-Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
-efficiency curves under load imbalance where each task's duration is multiplied by a uniform random variable between [0,~1). The task durations are generated with a deterministic
-pseudo random number generator with a consistent seed to ensure
-identical task durations for all systems.
-
-The MPI implementation of Task Bench, with its distinct computation and communication phases,
-suffers the most under load imbalance. The biggest
-difference is at large task granularities, where the imbalance
-effectively puts an upper bound on MPI efficiency. At smaller task
-granularities the effect shrinks and may even reverse as systems hit
-their fundamental limits due to overhead.
-
-The remaining differences are due primarily to
-different scheduling behaviors. The execution of 4
-simultaneous task graphs only partially mitigates the
-load imbalance between tasks. Systems that provide an
-additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
-efficiency at large task granularities. However, the use of
-work-stealing queues can also impact throughput at small task
-granularities.
-For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
-We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).
-We leave analysis of persistent load imbalance to future work.
 

### 2019-07-25 a2_evaluation.tex
Move load balancing results to supplemental material and fiddle with figure placement and spacing.

\section{Evaluation}
\label{sec:evaluation}

\input{e5_imbalance}

\subsection{Load Imbalance}

One advantage of systems with asynchronous execution capabilities is
the ability to mitigate load imbalance with little or no additional programmer effort, especially in the presence of
task parallelism. To quantify this effect,
Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
efficiency curves under load imbalance where each task's duration is multiplied by a uniform random variable between [0,~1). The task durations are generated with a deterministic
pseudo random number generator with a consistent seed to ensure
identical task durations for all systems.

The MPI implementation of Task Bench, with its distinct computation and communication phases,
suffers the most under load imbalance. The biggest
difference is at large task granularities, where the imbalance
effectively puts an upper bound on MPI efficiency. At smaller task
granularities the effect shrinks and may even reverse as systems hit
their fundamental limits due to overhead.

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persistent load imbalance to future work.

### 2019-07-25 5_evaluation.tex
More work shrinking space.

@@
 
-Figure~\ref{fig:bytes} shows the performance of Task Bench
-implementations with a memory-bound kernel. We measure peak memory
+Figure~\ref{fig:bytes} shows performance with a memory-bound kernel. We measure peak memory
 bandwidth of 108~GB/s, which compares favorably with the
@@
 
-One of the basic questions when considering different programming
+One question when considering different programming
 systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
@@
 mechanisms of other systems are pure overhead in this scenario.
-
 The gap between MPI and other systems shrinks as the complexity of the
@@
 Spark, Swift/T and TensorFlow
-are omitted from comparisons with more complicated dependencies, as the higher overheads of these systems require
-unreasonably large problem sizes (beyond what can be completed in 6
+are omitted from comparisons with more complicated dependencies, as the overheads of these systems require
+excessive problem sizes (beyond what can be completed in 6
 hours) to reach 50\% efficiency.
@@
 
-METG is a useful metric in part because it summarizes the basic task
-scheduling overhead of each programming system in a single number. This makes it
+METG is useful in part because it summarizes the overhead of each programming system in a single number. This makes it
 possible to evaluate METG at different node counts (shown in
@@
 
-Asynchronous systems such as Charm++ are able to show two benefits in
+Asynchronous systems such as Charm++ demonstrate two benefits in
 these plots. First, by overlapping communication with computation,
 such systems execute smaller task granularities at higher
-levels of efficiency compared to the bulk synchronous MPI
-implementation. Second, the asynchrony and scheduling flexibility from
+levels of efficiency compared to the MPI
+implementations. Second, the asynchrony and scheduling flexibility from
 executing multiple graphs also makes the curves smoother,
@@
 
-Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
-and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
+Figure~\ref{fig:cuda-efficiency} compares MPI
+and MPI+CUDA on the Piz Daint supercomputer. The CUDA
 compute kernel achieves a peak performance of $4.759 \times 10^{12}$

### 2019-07-25 7_conclusion.tex
Apply Alex's comments.
More work shrinking space.

@@
   load imbalanced workloads, or workloads with balanced computation
-  and communication. However, these gains can be easily nullified by
+  and communication. However, these gains can be nullified by
   high baseline overheads.
@@
 
+\zap{
 We have implemented Task Bench for a broad set of programming systems
@@
 OpenSHMEM~\cite{OpenSHMEM10}, Ray~\cite{Ray18}, and UPC~\cite{UPC99}.
+} % zap
 

### 2019-07-25 1_introduction.tex
More work on shrinking text.

@@
 distributed execution. A benchmark consists of a \emph{task graph} with tasks for
-each point in an iteration space. Tasks are connected by dependencies
-as determined by a \emph{dependence relation}. This design permits a
-very concise description of a wide variety of patterns relevant to
+each point in an \emph{iteration space}, and dependencies
+determined by a \emph{dependence relation}. This design permits a
+concise description of a wide variety of patterns relevant to
 scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
@@
 that have never before been measured.
-Most strikingly, the performance of the various systems we examine varies by more than five orders
+Most strikingly, the performance of the systems we examine varies by more than five orders
 of magnitude, with very popular, widely used systems at both ends of the spectrum!  Clearly
@@
 Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
-X10 \cite{X1005}. Using METG(50\%), we determine the baseline
+X10 \cite{X1005}. Using METG, we determine the baseline
 runtime overhead of each system along with overheads associated with
@@
 discusses the implementation of Task Bench in 15 programming systems,
-Section~\ref{sec:metg} defines the METG(50\%) metric and its
+Section~\ref{sec:metg} defines the METG metric and its
 relationship to quantities of interest to application developers,

### 2019-07-25 2_task_bench.tex
More work on shrinking text.

@@
 the vertical axis and parallel tasks along the
-horizontal. Tasks can depend only on tasks from the immediately
+horizontal. Tasks may depend only on tasks from the immediately
 preceding time step. Figure~\ref{fig:task-graphs} shows a number of sample task
@@
 source of performance disparity that can be a pitfall for
-implementations of mini-apps. Finally, the Task Bench core library manages the
-parsing of input parameters and the display of output results,
-ensuring that all implementations behave in a uniform manner and can
+implementations of mini-apps. Finally, the core library manages
+parsing input parameters and displaying results,
+ensuring that all implementations behave uniformly and can
 be scripted consistently. Because much of the functionality needed for
@@
 
-Task Bench provides two main kernels that can be called from tasks,
-which are compute- and memory-bound, respectively. The compute-bound
+Task Bench provides two main kernels that can be called from tasks:
+compute- and memory-bound. The compute-bound
 kernel executes a tight loop and is hand-written using AVX2 FMA
 intrinsics. The memory-bound kernel performs sequential reads and
-writes over an array, and is again hand-optimized with AVX2
+writes over an array, again with AVX2
 intrinsics. The duration of both kernels can be configured by setting

### 2019-07-25 4_metg.tex
More work on shrinking text.

@@
 efficiency starts at 100\%. Initially task granularity
-shrinks with minimal change in efficiency. However as task
-granularity shrinks further, efficiency drops more rapidly, leading
-to a vertical asymptote as runtime overhead dominates useful work.
+shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, leading
+to a vertical asymptote as overhead dominates useful work.
 
@@
 without changing the number of nodes or cores available to the
-application), METG isolates effects that are
-due to shrinking problem size from effects that are due to
+application), METG isolates effects
+due to shrinking problem size from effects due to
 increased communication and other resource issues as
-progressively larger portions of the machine are used. Section~\ref{sec:evaluation} contains a comprehensive evaluation of
-METG including how it changes with node count for a wide variety of
-programming systems.
+progressively larger portions of the machine are used.
+% Elliott: Do I need this sentence at all? The evaluation section is literally immediately after this.
+%% Section~\ref{sec:evaluation} contains a comprehensive evaluation of
+%% METG including how it changes with node count.

### 2019-07-25 5_evaluation.tex
More space tweaking.

@@
 
-In theory, any system should be able to hit peak performance as long as
-the kernels are well-written and of sufficiently large granularity. In
-practice, achieving peak performance is tricky, and many subtle pitfalls of implementation or configuration can easily lead to poor performance.
-Unfortunately, peak performance is also frequently skipped in evaluations of
-runtime system overhead. However, verifying that peak performance is
-achieved is important to ensuring that evaluations of overhead and
+In theory, any system should achieve peak performance if
+the kernels are well-tuned and of sufficiently large granularity. In
+practice, many subtle pitfalls of implementation or configuration can easily lead to poor performance. Verifying that peak performance is
+achieved helps to ensure that evaluations of overhead and
 efficiency are well-grounded.
@@
 
-The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
+Most systems achieve or nearly achieve peak FLOP/s. Some
 systems reserve a number of cores (usually 1 or 2) for internal
@@
 
-\input{e7_bytes}
-
 Figure~\ref{fig:bytes} shows performance with a memory-bound kernel. We measure peak memory
@@
 \subsection{Baseline Overhead}
-
-\input{e3_metg_compute}
 
@@
 
-We observe in the results that the baseline overheads of different
+We observe that the baseline overheads of different
 systems vary by over 5 orders of magnitude. It is worth

### 2019-07-25 5_evaluation.tex
More space saving.

@@
 but this is not true of all the systems included in this
-evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and a
-flat line is ideal. Lines that rise with node count indicate less than
-ideal scaling. Most notably, Spark is typically
-intended for use in industrial data center applications with task
+evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and
+flat is ideal. Lines that rise with node count indicate less than
+ideal scaling. Most notably, Spark is primarily
+intended industrial data center applications with task
 granularities measured in tens of seconds. Spark uses a centralized
@@
 executing on all other nodes. PaRSEC DTD and StarPU mitigate this
-partially by allowing the user to omit tasks that are not direct
-dependencies of those executed on the current node. However, this DAG
-trimming approach still requires some dynamic checks that scale with
-the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Approaches based on compile-time analysis can
+partially by allowing the user to omit tasks not directly
+dependent on those executed by a given node. However, this DAG
+trimming approach requires dynamic checks that scale with
+the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
 partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
 by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
-still retains dynamic checks leading to limited scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
+retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
 uses a compile-time optimization to improve
@@
 
-The number of dependencies per task has a strong influence on the
-overhead incurred by each system, as shown in
+The number of dependencies per task has a strong influence on
+overhead, as shown in
 Figure~\ref{fig:radix}. This plot shows METG(50\%) for the nearest
@@
 ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
-a factor of $12\times$ increase. This is unsurprising as in the case
+a $12\times$ increase. This is unsurprising as in the case
 of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
@@
 
+\input{e10_cuda_efficiency}
+
 \subsection{Communication Hiding}
 
-Also of interest is the ability of asynchronous systems to
+Also of interest is the ability to
 hide communication latency in the presence of task
-parallelism. Figure~\ref{fig:efficiency-communication} plots efficiency vs task
-granularity curves for varying amounts of
+parallelism. Figure~\ref{fig:efficiency-communication} plots efficiency with varying amounts of
 communication, determined by the number
@@
 FLOP/s, which is very close to the officially reported number
-$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
-FLOP/s. Note that the CPU and GPU kernels perform different numbers of
-operations as the GPU must be given more work to achieve peak
+$4.761 \times 10^{12}$. The CPU achieves $5.726 \times 10^{11}$
+FLOP/s. Note that the kernels perform different numbers of
+operations as the GPU requires more work to reach peak
 performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
-normalized so that the number of FLOPs is constant for a given problem
+normalized to keep FLOPs constant for a given problem
 size.
 
-Our MPI+CUDA implementation uses an offload model in which input and
-output data is copied to and from the GPU on each time step. We tested
+Our MPI+CUDA implementation uses an offload model in which data is copied to and from the GPU on each step. We test
 two configurations: \lstinline{w1} uses 1 task per GPU, whereas
-\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
-ranks per GPU to push work to the GPU in parallel and thus overlaps communication and computation. We see that the
-\lstinline{w4} configuration achieves higher FLOP/s but
+\lstinline{w4} overdecomposes by $4\times$, using 4 MPI
+ranks per GPU to push work to the GPU in parallel.
+\lstinline{w4} achieves higher FLOP/s but
 drops more rapidly at smaller problem sizes, due to the overhead of
@@
 GPU requires more work to achieve high performance, and the overhead
-of copying data to and from the GPU dominates at small task
+of copying data dominates at small task
 granularities, where the CPU achieves higher
 performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
-terms of MTEG (as the peak performance of the GPU and CPU are very
+terms of METG (as peak performance on CPU and GPU are very
 different), the conclusion here is similar to
-Section~\ref{subsec:scalability}: The cost of communicating data and
-tasks to and from the GPU imposes a higher task granularity to achieve
+Section~\ref{subsec:scalability}: The cost of sending data and
+tasks to the GPU imposes a higher task granularity to achieve
 the same efficiency than running on the CPU, reducing the advantage at small task granularities of
 very lightweight mechanisms such as those in MPI.
-
-\input{e10_cuda_efficiency}

### 2019-07-25 6_related.tex
More space saving.

@@
 
+% Force this paragraph to page break at a hyphen because otherwise
+% it's going to be nearly impossible to cut enough space out of this
+% paper.
+\let\oldbrokenpenalty\brokenpenalty
+\brokenpenalty=0
+
 Parallel and distributed programming systems are most
 commonly evaluated using proxy- or mini-apps, or
-micro\-benchmarks. Mini-apps are explicitly derived from larger
+microbenchmarks. Mini-apps are explicitly derived from larger
 applications and therefore have the advantage of bearing some
@@
 for microbenchmarks.
+
+\brokenpenalty=\oldbrokenpenalty
 
@@
 
-Microbenchmarks are often easier to implement, but may not be
-representative of real-world applications. From the PRK
-suite~\cite{PRK14}, the PRK Stencil code has been reasonably widely
-implemented, and a follow-up paper~\cite{PRKRuntimes16} compares MPI
-variants, SHMEM, UPC, Charm++, and Grappa. However, PRK Stencil is a
-single, simple code involving 2 kernels (which combined fit in under 50 lines
-of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} is also relatively widely studied with
+Microbenchmarks can be easier to implement, but risk being
+unrepresentative of real applications. PRK Stencil~\cite{PRK14}
+contains 2 kernels (which combined fit in under 50 lines of C++)
+and a halo exchange on a structured grid; an
+evaluation~\cite{PRKRuntimes16} compares MPI variants, SHMEM, UPC,
+Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
+consist mostly of small kernels for dense array computations and has
 implementations in OpenMP~\cite{NASOpenMP99}, MPI and
 MPI+OpenMP~\cite{NASMPIOpenMP00}, and
-Charm++~\cite{NASCharm96}. The benchmarks consist mostly of small
-kernels for dense array computations. Neither the PRK nor the NAS
-benchmark suite achieve the breadth of coverage, flexibility, or ease
-of implementation of Task Bench. We believe the evidence for ease of
-implementation is clear; Task Bench is implemented for more systems
-than all previous mini-apps and microbenchmarks. Task Bench is also
-not a fixed set of benchmarks, and it is even easier to use Task Bench
-to generate a new benchmark to measure some aspect of system
-performance.
+Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieve the breadth of
+coverage, flexibility, or ease of implementation of Task Bench. We
+believe the evidence for ease of implementation is clear; Task Bench
+is implemented for more systems than all previous mini-apps and
+microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
+it is even easier to use Task Bench to generate new benchmarks.
 
-System-specific benchmarks have been used to quantify specific aspects
+System-specific benchmarks quantify specific aspects
 of system performance, such as MPI communication or collective

### 2019-07-29 1_introduction.tex
Tweak introduction.

@@
 
-To characterize the contribution of a programming model's runtime
-system to application performance, and as an example of the novel studies that can be done
+To characterize the contribution of runtime overheads
+to application performance, and as an example of the novel studies that can be done
 with Task Bench, we introduce a new metric called
@@
 at least 50\% efficiency, meaning that the application achieves at
-least 50\% of the higest performance (in FLOP/s, B/s, or another
-application-specific measure) that can be achieved on a given
+least 50\% of the higest performance (in FLOP/s, B/s, or other
+application-specific measure) achieved on a given
 machine. The efficiency bound in METG is a key innovation over
@@
 to consider the amount of useful work peformed (if tasks are
-non-empty~\cite{Canary16, Armstrong14}) or to perform any useful work at all (if tasks are empty~\cite{LegionTracing18}).
+non-empty~\cite{Canary16, Armstrong14}) or to perform useful work at all (if tasks are empty~\cite{LegionTracing18}).
 
@@
 We conduct a comprehensive comparative study of Task Bench
-implementations in 15 different programming systems:
+implementations in 15 programming systems:
 Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
@@
 Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
-X10 \cite{X1005}. Using METG, we determine the baseline
+X10 \cite{X1005}. Using METG, we find a number of factors---node
+count, accelerators, and complex dependencies, among
+others---individually or in combination contribute to an order of
+magnitude or greater increase in METG, even in the most efficient
+systems. While in the most simplified scenarios, some systems can
+achieve a METG(50\%) of 390~ns, we show that a more realistic
+bound for running nearly any application at scale is 100~\textmu{}s with
+current technologies.
+
+%% This implies that simplifying the application space too much
+%% (e.g., by using trivial dependencies) risks overemphasizing the
+%% importance of the lightweight mechanisms in such systems---a risk that
+%% can be mitigated by using Task Bench's more complete coverage of the
+%% space of application scenarios.
+
+\zap{
+Using METG, we determine the baseline
 runtime overhead of each system along with overheads associated with
@@
 overhead, and mitigate load imbalance.
+} % zap
 

### 2019-07-30 0_abstract.tex
Tweak abstract to make performance claims.

@@
 \begin{abstract}
-% 148 words
 We present Task Bench, a \emph{parameterized} benchmark designed to
@@
 runs on every Task Bench implementation. Furthermore,
-Task Bench's parameterization allows a wide variety of benchmark
+Task Bench's parameterization enables a wide variety of benchmark
 scenarios that distill the key characteristics of larger applications.
@@
 15 programming systems on up to 256 Haswell nodes of the Cori
-supercomputer. We introduce and use a novel metric, \emph{minimum effective task
-  granularity} to study the baseline runtime overhead of each system
-and the impact of different dependence patterns. We also
-study each system's scalability, ability to hide communication and
-mitigate load imbalance.
+supercomputer. We introduce a novel metric, \emph{minimum effective task
+  granularity} to study the baseline runtime overhead of each
+system. We show that when running at scale, 100~\textmu{}s is the
+smallest granularity that even the most efficient systems can reliably
+support with current technologies. We also study each system's
+scalability, ability to hide communication and mitigate load
+imbalance.
 \end{abstract}

### 2019-07-30 1_introduction.tex
Tweak abstract to make performance claims.

@@
 
-We present Task Bench, a parameterized benchmark designed for exploring the performance
+We present Task Bench, a parameterized benchmark for exploring the performance
 of parallel and distributed programming systems under a
 variety of conditions.  The key property of Task Bench is that it completely separates
-the system-specific implementation from the core implementation
+the system-specific implementation from the implementation
 of the benchmarks themselves.
-In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
+In all previous benchmarks we know of, the effort to implement $m$ benchmarks on $n$
 systems is $\mathcal{O}(mn)$.  Task Bench's design reduces this work to $\mathcal{O}(m + n)$,
@@
 in the discrete ordinates method of radiation simulation), FFTs, trees
-(for divide and conquer algorithms), and so on. The tasks at each
-point in the task graph can be configured to execute kernels with a
+(for divide and conquer algorithms), and so on. Tasks execute kernels with a
 variety of computational properties, including compute- and
-memory-bound operations of varying durations. The dependencies between
+memory-bound operations of varying durations. Dependencies between
 tasks can be configured to carry communication payloads of varying size, permitting
@@
 to introduce task parallelism into the workload. Together, these
-design elements permit the exploration of a large space of application
+design elements enable the exploration of a large space of application
 behaviors.

### 2019-07-30 7_conclusion.tex
Update conclusion to match intro.

@@
 
-\item While some systems support task granularities as small as a few
-  microseconds, this applies only to small CPU-based clusters. Adding
-  accelerators or moving to clusters of hundreds of nodes raises the
-  METG that any system can achieve by over an order of magnitude.
+\item While some systems support task granularities as small as 390~ns, this applies only to trivial dependencies and small CPU-based clusters.
+A number of factors (non-trivial dependencies, accelerators and cluster sizes in the hundreds of nodes) raise the
+  METG that any system can reliably achieve by over an order of magnitude: 100~\textmu{}s is a reasonable bound for most applications running at scale with current technologies.
 
 \item Systems that support asynchronous execution show benefits under
-  load imbalanced workloads, or workloads with balanced computation
-  and communication. However, these gains can be nullified by
+  balanced computation
+  and communication, and load imbalance. However, these gains can be nullified by
   high baseline overheads.
@@
 \item Systems for large scale data analysis require very large tasks
-  (in the tens of seconds) to scale beyond small numbers of nodes,
+  (tens of seconds) to scale beyond small node counts,
   reflecting the very coarse tasks and lack of need for strong scaling
@@
 
-\item The newer task-based systems have performance sufficient for
+\item Newer task-based systems have performance sufficient for
   weak scaling many HPC workloads, but more work is needed to strong
-  scale the most demanding applications.
+  scale the most demanding codes.
 

### 2019-07-30 7_conclusion.tex
Small wording tweak.

@@
 tradeoff so that users can be better informed and developers can see
-the impact of features on the performance of their programming
+the impact that features have on the performance of their programming
 systems.

### 2019-08-02 3_implementation.tex
More work on shrinking text.
Update text and fiddle with wording and figures to make everything fit again.

@@
 
-One challenge in implementing Task Bench for such a wide variety of
+One challenge in targeting such a wide variety of
 systems is that the capabilities of the systems vary considerably. For
@@
 
-In all cases, members of the respective programming systems' teams
-were consulted during the development and evaluation of the
-corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
-in helping to ensure that we provide the highest quality implementations for each system.
+In all cases, members of the programming systems' teams
+were consulted in the development and evaluation of the
+corresponding Task Bench implementations. Where assistance was provided, the insights helped ensure that we provide the highest quality implementations for each system.

### 2019-08-02 4_metg.tex
Update text and fiddle with wording and figures to make everything fit again.

@@
 \label{sec:metg}
+
+\input{f1_flops_mpi}
+\input{f2_efficiency_mpi}
 
@@
 
-Existing studies of programming system efficiency
-typically use the metric \emph{tasks per second} (TPS). TPS results
-are difficult to interpret and apply, because the efficiency (and thus
-the amount of useful work) is not constrained. When using empty
+Existing studies of system efficiency
+typically report \emph{tasks per second} (TPS). TPS results
+are difficult to interpret and apply, because efficiency (and thus
+the amount of useful work) is not constrained. With empty
 tasks~\cite{LegionTracing18}, the resulting upper bound on task
@@
 
-\input{f1_flops_mpi}
-\input{f2_efficiency_mpi}
+\input{f3_weak_scaling_mpi}
+\input{f4_strong_scaling_mpi}
 
@@
 and thus it makes sense to discuss tasks even in systems
-without an explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
+with no explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
 compute-bound kernel included in the Task Bench implementation,
@@
 
-\input{f3_weak_scaling_mpi}
-\input{f4_strong_scaling_mpi}
-
 METG has a well-defined
@@
 and conform to the shape of the METG curve (marked by the red, dashed line).
+
+\input{e0_flags}
 

### 2019-08-02 5_evaluation.tex
Update conclusion to match intro.
Update text and fiddle with wording and figures to make everything fit again.

@@
 \label{sec:evaluation}
-
-\input{e0_flags}
 
@@
 
-\input{e1_flops}
-\input{e2_efficiency}
-\input{e7_bytes}
-
-\input{e3_metg_compute}
-
 Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
-kernel with varying problem sizes. This is the full version of
+kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations). This is the full version of
 Figure~\ref{fig:flops-mpi}.
@@
 the baseline for 100\% efficiency.
+
+\input{e1_flops}
+\input{e2_efficiency}
+\input{e7_bytes}
+
+\input{e3_metg_compute}
 
@@
 In most cases, task granularity asymptotes prior to the 50\% efficiency point,
-though some systems continue to improve as efficiency continues
-to drop. Accounting for this effect is one of the main arguments in
+though some systems continue to improve as efficiency drops further. Accounting for this effect is one of the main arguments in
 favor of METG with a reasonable efficiency threshold instead of
@@
 
-Figure~\ref{fig:bytes} shows performance with a memory-bound kernel. We measure peak memory
-bandwidth of 108~GB/s, which compares favorably with the
-OpenMP-enabled STREAM benchmarks~\cite{STREAM} which report up to 98~GB/s.
+Figure~\ref{fig:bytes} shows performance with a memory-bound kernel. We measure a peak memory
+bandwidth of 79~GB/s, using a working set size of 0.5 GB.
+As discussed in Section~\ref{sec:task-bench}, the kernels are designed to keep the working set constant as the number of iterations decrease to avoid noisy, superlinear effects in the results.
+For comparison, the
+OpenMP-enabled STREAM benchmarks~\cite{STREAM} report up to 98~GB/s.
 
-Not all cores are required to saturate memory bandwidth. As a result,
-systems that reserve one or more cores are able to hit 100\% of peak
-performance, unlike with compute-bound kernels. We are currently investigating the
-performance of Charm++, Regent and X10.
-
-% FIXME: why are Charm++, Regent and X10 slow?
+Not all cores are required to saturate memory bandwidth, reducing the
+impact of reserving cores for internal use. Most systems hit 100\% of
+peak, unlike the compute-bound case.
 
@@
 issued at all, so MPI has nothing to do aside from executing kernels
-as quickly as possible. For this reason, we recommend evaluating the performance of
-programming systems under realistic workloads, rather than with
-trivial dependencies.
+as quickly as possible. Clearly, choosing a representative dependence
+pattern is important when estimating the performance of a workload or
+class of workloads.
 

### 2019-08-06 1_introduction.tex
Small wording tweak from Alex.
Tweak some wording around GPU text.

@@
 
-Note that a lower METG does not necessarily mean that 
+A lower METG does not necessarily mean that
 performance for a particular workload is better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
@@
 The paper is organized as follows: Section~\ref{sec:task-bench}
-describes the design of Task Bench, Section~\ref{sec:implementation}
-discusses the implementation of Task Bench in 15 programming systems,
+describes the design of Task Bench. Section~\ref{sec:implementation}
+discusses implementations in 15 programming systems.
 Section~\ref{sec:metg} defines the METG metric and its
-relationship to quantities of interest to application developers,
+relationship to quantities of interest to application developers.
 Section~\ref{sec:evaluation} provides a comprehensive evaluation of
 Task Bench on up to 256 Haswell nodes of the Cori
-supercomputer~\cite{Cori}, Section~\ref{sec:related-work} relates this
-work to previous efforts, and Section~\ref{sec:conclusion} concludes.
+supercomputer~\cite{Cori} (GPU results are reported on Piz Daint~\cite{PizDaint}). Section~\ref{sec:related-work} relates to previous efforts; Section~\ref{sec:conclusion} concludes.

### 2019-08-06 5_evaluation.tex
Tweak some wording around GPU text.

@@
 
-Figure~\ref{fig:cuda-efficiency} compares MPI
+In order to determine the cost of scheduling tasks on the GPU, Figure~\ref{fig:cuda-efficiency} compares MPI
 and MPI+CUDA on the Piz Daint supercomputer. The CUDA
@@
 operations as the GPU requires more work to reach peak
-performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
+performance. The x-axis in Figure~\ref{fig:cuda-efficiency} is
 normalized to keep FLOPs constant for a given problem

### 2019-08-06 1_introduction.tex
Add a takeaway to intro, and cuts elsewhere to make space.

@@
 
-Adding a new programming system to Task Bench involves  implementing a set of standard
-services, such as executing a task or a data transfer.  This interface is simple enough to 
+Adding a system to Task Bench involves  implementing a set of standard
+services, such as executing a task or data transfer, simple enough to 
 be implemented with reasonable effort to a level of quality sufficient for comparative
-studies. Task Bench provides a core library that encapsulates the functionality
-shared among systems, which not only reduces the implementation effort
-required but also makes it much easier to achieve truly apples-to-apples comparisons
+studies. Task Bench provides a core library that encapsulates functionality
+shared among systems, which not only reduces implementation effort
+but also makes it much easier to achieve truly apples-to-apples comparisons
 between systems.
@@
 Using Task Bench we were able to quickly incorporate 15 very different parallel and distributed
-runtime systems.  By running all systems on common benchmarks we were able to quantify phenomena
+programming systems.  By running all systems on common benchmarks we were able to quantify phenomena
 that have never before been measured.
-Most strikingly, the performance of the systems we examine varies by more than five orders
-of magnitude, with very popular, widely used systems at both ends of the spectrum!  Clearly
+Most strikingly, the overheads of the systems we examine vary by more than five orders
+of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly
 the slower systems have ``good enough'' performance for some class of applications, while presumably
@@
 
-This result points to a natural question:   How does one predict whether the performance of a runtime
-system will in fact be good enough for a particular application?  In trying to answer this question,
+How does one predict whether the performance of a given
+system will be good enough for a particular application?  In trying to answer this question,
 we realized that \emph{weak} and \emph{strong} scaling, by far the most commonly
@@
 A lower METG does not necessarily mean that
-performance for a particular workload is better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
-respectively, running an application with an average task granularity
-of 10~ms, are both very likely to perform well. Only when the task
-granularity approaches (or drops below) METG(50\%) will performance
-likely diverge. Thus METG helps to identify the regime in which a
+performance for a particular workload is better. Two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
+respectively, running an application with 10~ms average task granularity, are both likely to perform well. Only when task
+granularity approaches (or drops below) METG(50\%) will they
+likely diverge. METG identifies the regime in which a
 given system can deliver good performance, and explains how
@@
 bound for running nearly any application at scale is 100~\textmu{}s with
-current technologies.
+current technologies. We find asynchronous systems provide benefits
+(e.g., overlapping compute and communication), but only if system overheads do not exceed roughly
+100~\textmu{}s.
 

### 2019-08-06 1_introduction.tex
Tweak.

@@
 bound for running nearly any application at scale is 100~\textmu{}s with
-current technologies. We find asynchronous systems provide benefits
+current technologies. We find asynchronous systems can provide benefits
 (e.g., overlapping compute and communication), but only if system overheads do not exceed roughly

### 2019-08-06 1_introduction.tex
Edits from Sean.

--- 2019-08-06 1_introduction.tex
@@
 How does one predict whether the performance of a given
-system will be good enough for a particular application?  In trying to answer this question,
-we realized that \emph{weak} and \emph{strong} scaling, by far the most commonly
-reported performance metrics, do not directly characterize the performance of the underlying
+system will be good enough for a particular application?
+Surprisingly, the most commonly reported performance metrics,
+\emph{weak} and \emph{strong} scaling, do not directly characterize
+the performance of the underlying
 programming system. Weak scaling can hide arbitrary amounts of runtime
 system overhead by using sufficiently large problem sizes, and strong
-scaling does not separate runtime system overhead from costs
+scaling does not separate runtime system overhead from application costs
 (such as increased communication) that scale with the number of nodes when
@@
 at least 50\% efficiency, meaning that the application achieves at
-least 50\% of the higest performance (in FLOP/s, B/s, or other
+least 50\% of the highest performance (in FLOP/s, B/s, or other
 application-specific measure) achieved on a given
@@
 previous approaches, such as \emph{tasks per second} (TPS), that fail
-to consider the amount of useful work peformed (if tasks are
+to consider the amount of useful work performed (if tasks are
 non-empty~\cite{Canary16, Armstrong14}) or to perform useful work at all (if tasks are empty~\cite{LegionTracing18}).
@@
 
+% Seems to be nearly impossible to get rid of the hyphen, so just give
+% up and allow this paragraph to break across it.
+\let\oldbrokenpenalty\brokenpenalty
+\brokenpenalty=0
+
 We conduct a comprehensive comparative study of Task Bench
 implementations in 15 programming systems:
-Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
+Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP, CUDA),
 OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
@@
 Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
-X10 \cite{X1005}. Using METG, we find a number of factors---node
+X10 \cite{X1005}. Using METG, we find that a number of factors---node
 count, accelerators, and complex dependencies, among
@@
 magnitude or greater increase in METG, even in the most efficient
-systems. While in the most simplified scenarios, some systems can
-achieve a METG(50\%) of 390~ns, we show that a more realistic
+systems. While some systems can achieve sub-microsecond METG(50\%) in
+best-case scenarios, we show that a more realistic
 bound for running nearly any application at scale is 100~\textmu{}s with
-current technologies. We find asynchronous systems provide benefits
-(e.g., overlapping compute and communication), but only if system overheads do not exceed roughly
-100~\textmu{}s.
+current technologies. Our study includes several asynchronous systems
+which are designed to provide benefits such as overlapped computation
+and communication. While small-scale benchmarks of these systems
+suffer from increased overhead, we find that the benefits of these systems become
+tangible at scale (provided the runtime overhead doesn't increase
+beyond about 100~\textmu{}s).
+
+\brokenpenalty=\oldbrokenpenalty
+
+%% We find asynchronous systems can provide benefits
+%% (e.g., overlapping compute and communication), but only if system overheads do not exceed roughly
+%% 100~\textmu{}s.
 

### 2019-08-14 3_implementation.tex
Splice everything back into the main paper.

@@
 We have implemented Task Bench in the 15 parallel and distributed
-programming systems listed in Table~\ref{tab:systems}. These include
-traditional HPC programming models: MPI~\cite{MPI} (with bulk
-synchronous and point-to-point communication) and MPI+X (with OpenMP
-and CUDA); PGAS and actor models: Chapel~\cite{Chapel15},
-Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and newer task-based
-systems: OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40} (with task
-dependencies), PaRSEC (with \emph{parameterized task graphs}
-(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
-(DTD)~\cite{PARSEC_DTD}), Realm~\cite{Realm14},
-Regent~\cite{Regent15}, and StarPU~\cite{StarPU11}. Also, we consider
-several systems used in large scale data analytics and workflows:
-Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
-and TensorFlow~\cite{TensorFlow15}. We describe the systems, and any
-salient details of their Task Bench implementations, in our
-supplemental material.
+programming systems listed in Table~\ref{tab:systems}. We describe the systems, and any
+salient details of their Task Bench implementations, below.
 
@@
 corresponding Task Bench implementations. Where assistance was provided, the insights helped ensure that we provide the highest quality implementations for each system.
+
+\subsection{Chapel}
+
+Chapel~\cite{Chapel15} is a parallel programming language
+with a \emph{multi-res\-o\-lu\-tion} approach, supporting parallelism at a variety of
+levels. Chapel's core features are a partitioned global address space
+(PGAS), data distributions, tasks,
+synchronization primitives, and array promotion. For Task Bench, we target a low level of
+Chapel, using explicit task instantiation (via
+\lstinline[language=Chapel]{coforall}), bulk access to distributed
+arrays for efficient data movement, and atomic integers for
+synchronization.
+
+\subsection{Charm++}
+
+Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors, or
+\emph{chares}, are objects in their own address space.
+Chares communicate data and synchronize via messages and can be moved
+to balance load. Our Task Bench implementation uses a chare
+array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
+dependencies are all available.
+
+%% Collectives synchronize the
+%% completion of the task graph execution.
+
+% FIXME: Discuss SMP vs non-SMP versions
+
+\subsection{Dask}
+
+Dask~\cite{Dask15} is an implicitly parallel task-based programming system for large
+scale data analytics with Python. Dask provides abstractions over
+distributed NumPy arrays and Pandas dataframes, and also a lower-level
+interface for launching tasks. We use the lower-level interface to
+implement Task Bench.
+
+\subsection{MPI}
+
+Our MPI~\cite{MPI} implementation of Task Bench represents the common
+case of point-to-point communication with distinct computation and
+communication phases. We experimented with a variety of implementation
+strategies and found the best performing to be using
+\lstinline[language=C++]{MPI_Isend} and
+\lstinline[language=C++]{MPI_Irecv} to implement the communication
+phase, posting receives before sends. Each task dependency maps to one
+send/receive pair in MPI. We also provide a bulk synchronous
+implementation which enforces the boundary between communication and
+computation with \lstinline[language=C++]{MPI_Barrier}.
+
+\subsection{MPI+X}
+
+We provide two MPI+X implementations of Task Bench to evaluate the
+performance of hierarchical programming models. Our MPI+OpenMP
+implementation uses forall-style parallel loops to execute tasks,
+interleaved with MPI point-to-point communication as above. Our
+MPI+CUDA implementation follows an offload model where data is copied
+to and from the GPU on every timestep.
+
+\subsection{OmpSs}
+
+OmpSs~\cite{OmpSs11} is a programming model for loop- and task-based parallelism
+that is source-compatible with OpenMP. Our Task Bench implementation
+uses OpenMP 4.0-style task dependencies. Because OpenMP tasks have a fixed number of dependencies, we use a switch
+statement to implement the dynamic dependencies required for Task
+Bench. The implementation is otherwise straightforward.
+
+% FIXME: More to say about OmpSs? Is there anything different vs the OpenMP implementation?
+
+\subsection{OpenMP}
+
+OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
+parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
+for OpenMP is similar to OmpSs and uses
+task dependencies. We tested with GNU GOMP and Intel KMP
+and found KMP to be better performing.
+
+\subsection{PaRSEC}
+
+PaRSEC is a task-based programming system supporting two distinct
+programming models: \emph{parameterized task graphs}
+(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
+(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which programmers
+write a concise, algebraic description of the tasks and dataflows in
+the program. This compressed representation is expanded into a full
+task graph by a source-to-source compiler.
+
+In the DTD model, tasks are enumerated in program order (by
+executing the program), and dependencies between tasks are
+identified automatically based on the input and output data of tasks. 
+A task depends on another task if it reads data written by the other task,
+and the data is copied automatically if the two tasks are executed on
+different nodes. To
+improve scalability, the program is executed in
+SPMD fashion on all nodes, and the user is responsible for eliding
+tasks that are not directly connected to those that are to be executed
+on the current node.
+
+In our Task Bench implementation, we evaluate both PTG and DTD models,
+as well as a third option based on DTD that has additional manual
+optimizations applied to further reduce dynamic checks for what tasks
+should be executed on a node, which further improves
+scalability.
+
+%% Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
+%% distributed across nodes in pre-defined patterns,
+%% and can be passed to tasks on other nodes, but
+%% can only be written on the node where the data is originally
+%% allocated. In our Task Bench implementation, each task writes to a
+%% tile as output, and these tiles are passed as inputs to the
+%% subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
+%% the task launch code uses a switch statement as the PaRSEC API assumes
+%% tasks have a fixed number of dependencies.
+
+\subsection{Realm}
+
+Realm~\cite{Realm14} is an explicitly-parallel task-based programming
+model used internally by Legion~\cite{Legion12} and
+Regent~\cite{Regent15}. A Realm implementation of Task Bench is a limit study of what can be achieved with Legion or Regent.
+
+Tasks in Realm are asynchronous, with dependencies
+defined by \emph{events} passed from one task
+to another. Realm's data model supports collections that live in a
+specific memory. Data must be explicitly copied, and the copies
+synchronized with tasks via events.
+Realm also supports a \emph{subgraph} API that enables additional
+optimizations. We use subgraphs in the Task
+Bench implementation.
+
+\subsection{Regent}
+
+Regent~\cite{Regent15} is an implicitly-parallel task-based language
+which implements the Legion programming model~\cite{Legion12}. We use
+Regent rather than Legion directly because the Regent compiler
+provides a critical optimization that
+improves scalability~\cite{ControlReplication17}.
+
+\subsection{Spark}
+
+Spark~\cite{Spark10} is an implicitly-parallel programming model for
+large scale data analytics.
+The core abstractions in Spark are functional operators such as map,
+reduce, and join. Data in Spark are stored in
+\emph{resilient
+  distributed datasets} (RDDs): globally-visible,
+dynamically single-assignment data structures. Spark caches RDDs in memory to avoid unnecessary
+disk traffic.
+
+Spark has tasks internally but they are not exposed to the user,
+so a Task Bench implementation must map the task graph to a set of operators that result in the
+desired execution. We use
+\lstinline[language=Scala]{flatMap} and
+\lstinline[language=Scala]{groupByKey} to generate task
+dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
+series of tasks. An explicit hash partitioner ensures that
+Spark does not attempt to group multiple Task Bench tasks into a
+single Spark task, as tasks in Task Bench already represent coarse-grained units of work.
+
+We performed extensive experiments to verify that no
+extraneous factors interfered with our Spark measurements. We
+disabled logging, ensured no RDDs are written
+to disk, confirmed there is no measurable overhead due to JNI
+calls from Java to C or due to the serialization of data structures, and cross-checked results with known cases
+that hit optimal task throughput in Spark.
+
+\subsection{StarPU}
+
+StarPU~\cite{StarPU11} is a task-based system that supports a \emph{sequential task flow} (STF)
+programming model similar to PaRSEC's DTD. Our Task Bench implementation in
+StarPU is very similar to PaRSEC.
+%% We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
+%% project~\cite{Chameleon}.
+
+\subsection{Swift/T}
+
+Swift/T~\cite{Wozniak13} is a parallel scripting language intended
+primarily for the composition of HPC workflows. Swift/T
+programs follow dataflow semantics, where every statement may
+potentially execute in parallel as soon as its dependencies are
+satisfied; statements only execute sequentially when explicitly
+requested. The Swift/T compiler performs a number of optimizations to
+improve performance of highly parallel programs~\cite{Armstrong14}. Our Task Bench in Swift/T is straightforward, using Swift/T's dataflow
+semantics to capture dependencies on other tasks.
+
+\subsection{TensorFlow}
+
+TensorFlow~\cite{TensorFlow15} is a programming system designed for
+deep learning workloads. Although TensorFlow's API exposes
+machine learning concepts, internally TensorFlow is a task graph execution engine, making it a good fit for
+Task Bench. TensorFlow programs operate via explicit graph construction. Task graphs are composed in Python and run by an
+execution engine written in C++. Our Task Bench implementation is straightforward.
+
+\subsection{X10}
+
+X10~\cite{X1005} is an explicitly parallel programming language for
+place-based programming. The core features of X10 are \emph{places}
+which represent distributed memories, a PGAS model where references to
+remote objects can be held (but can only be dereferenced on the local
+place), asynchronous tasks, and a place-changing construct to move the
+execution of a task to a remote place. X10 also supports a variety of
+synchronization primitives.
+
+Our Task Bench implementation uses
+\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
+between places, along with place-changing and atomic integers for
+synchronization. We use the native backend of X10, which compiles to
+C++.

### 2019-08-14 5_evaluation.tex
Splice everything back into the main paper.

@@
 CUDA 9.1.85.
-
-Load balancing results are in the supplemental material.
 
@@
 
+\input{e5_imbalance}
+
 \input{e10_cuda_efficiency}
@@
 smaller message sizes.
+
+\subsection{Load Imbalance}
+
+One advantage of systems with asynchronous execution capabilities is
+the ability to mitigate load imbalance with little or no additional programmer effort, especially in the presence of
+task parallelism. To quantify this effect,
+Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
+efficiency curves under load imbalance where each task's duration is multiplied by a uniform random variable between [0,~1). The task durations are generated with a deterministic
+pseudo random number generator with a consistent seed to ensure
+identical task durations for all systems.
+
+The MPI implementation of Task Bench, with its distinct computation and communication phases,
+suffers the most under load imbalance. The biggest
+difference is at large task granularities, where the imbalance
+effectively puts an upper bound on MPI efficiency. At smaller task
+granularities the effect shrinks and may even reverse as systems hit
+their fundamental limits due to overhead.
+
+The remaining differences are due primarily to
+different scheduling behaviors. The execution of 4
+simultaneous task graphs only partially mitigates the
+load imbalance between tasks. Systems that provide an
+additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
+efficiency at large task granularities. However, the use of
+work-stealing queues can also impact throughput at small task
+granularities.
+For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
+We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).
+We leave analysis of persistent load imbalance to future work.
 

### 2019-08-15 7_conclusion.tex
Add acknowledgment.

@@
 systems.
+
+\section*{Acknowledgment}
+
+This material is based upon work supported by the U.S. Department of
+Energy, Office of Science, Office of ASCR, under the contract number
+DE-AC02-76SF00515, by National Science Foundation under Grant
+No. ACI-1450300, and the Exascale Computing Project (17-SC-20-SC), a
+collaborative effort of the U.S. Department of Energy Office of
+Science and the National Nuclear Security Administration, under prime
+contract DE- AC05-00OR22725, and UT Battelle subawards 4000151974 and
+89233218CNA000001. Experiments on the Cori supercomputer were
+supported by the National Energy Research Scientific Computing Center,
+a DOE Office of Science User Facility supported by the Office of
+Science of the U.S. Department of Energy under Contract
+No. DE-AC02-05CH11231, and experiments on Piz Daint were supported by
+the Swiss National Supercomputing Centre (CSCS) under project ID d80.

### 2019-10-28 ppopp20_reviews/rebuttal.md
Rebuttal draft.

We thank the reviewers for their detailed and insightful feedback.

Task Bench is a new class of benchmark that aims to make it easy to
execute a wide variety of application patterns on a large number of
distributed and parallel programming systems. Despite the name, Task
Bench is not specific to task-based programming models. A "task" in
Task Bench is simply any unit of code that runs to completion without
requiring any network communication. Any application that can be
separated into units of work interleaved with communication can be
adapted to Task Bench.

As a case in point, our MPI implementation is written in an extremely
straightforward style, with interleaved phases of communication and
computation. Each MPI rank is responsible for a set of columns in the
task graph. During communication phases we use non-blocking sends and
receives to communicate task dependencies, and then after calling
`MPI_Waitall()` we execute the corresponding tasks (whose inputs were
just received). We conducted an independent code review of our MPI
implementation with developers of a major MPI implementation to
confirm that the code is written in a manner that is representative of
how real-world MPI users write MPI code, and that it achieves good
performance.

In general, we are confident in the performance of our Task Bench
implementations because each one has been submitted to the developers
of the corresponding systems for code review and suggestions.

Task Bench covers a space of possible application patterns. This space
is easy to extend, e.g. by adding new kernels, dependence patterns, or
by choosing new values for the existing parameters (height, width,
etc.). As can be seen in Listing 2 in the paper, the Task Bench core
API is very general, and includes the ability to include or exclude
arbitrary points within the space (line 17), and to specify arbitrary
dependencies for each point in the task graph (line 19). Note that the
dependence pattern can also change over time; this can be seen in the
FFT and Tree patterns (Figure 1c and e). Thus, while the specific
patterns evaluated in the paper are necessarily limited (as the
reviewers have noted, our evaluation is already extremely
comprehensive), it is entirely possible (and even easy, relatively
speaking) to design patterns and kernels that do reflect actual mini-
or full-size applications in Task Bench.

We believe Task Bench compares very favorably with other benchmarks in
terms of the number of programming system features and behaviors that
it is capable of exploring. Certainly, compared to any benchmark which
is widely implemented, Task Bench provides massively more
expressiveness and configurability. Task Bench provides even more
implementations than the most widely implemented benchmark that we are
aware of. And in terms of the depth of analysis that can be performed,
the only benchmarks that provides substantially more visibility into
system-specific performance behaviors are themselves system-specific.

Task Bench is open source and we intend for it to be an active project
which is open to contribution. In fact, since the initial submission
of the paper we added three new functionally correct implementations
to Task Bench. If these implementations are tuned to our satisfaction
in advance of the deadline they will be included in the final paper.

METG is not fundamentally tied to peak performance (FLOP/s or B/s). If
an application is not amenable to being characterized via peak
performance, another measure of absolute performance can be used. For
example, an aerodynamics application might choose mesh cells processed
per second. The value of 100% efficiency would be set to the highest
throughput that can be achieved on a single node. Other values of
efficiency can be computed as the ratio of that highest throughput
achieved.

Though our evaluation uses METG(50%), this is not a limitation of
METG. We chose 50% because it avoids certain pathologies associated
with low efficiency thresholds (see lines 711-750 in the paper), and
also because it matches what we see implemented in practice. For
example, in a CSCS application for new projects, users are instructed
to "select the most parallel efficient job size (ratio of benchmark
speed-up vs. linear speed-up above 50%)" \[1]. This corresponds to 50\%
parallel efficiency.

\[1]: https://user.cscs.ch/access/report/

In the interest of answering the reviewers' questions as completely as
possible, we include responses to specific reviewer questions below.

## Reviewer A

  * See above for our description of the MPI implementation. A "task"
    in MPI includes the time to execute a set of send/receive calls
    needed to gather the inputs for a given task invocation, the call
    the core API to execute the task.

## Reviewer B

  * In our MPI+OpenMP configuration we use `OMP_NUM_THREADS` equal to
    the number of physical cores on the node (32 on Cori Haswell). In
    general this is set on a per-system basis depending on what is
    most representative of how the system is used in practice.

  * Overlap occurs when there is task parallelism available in the
    task graph. That is, when two tasks are independent (no mutual
    dependencies) their execution can be overlapped with the other's
    data transfers.

## Reviewer C

  * Note that there is an evaluation of trivial dependencies in Figure
    10. Trivial corresponds to 0 dependencies per task in this chart.

  * Note that in Figure 10 Task Bench achieves running times per task
    as small as 390 ns (while still maintaining 50% efficiency). In
    general Task Bench is carefully written to be as optimal as
    possible (e.g. avoiding memory allocations), both in the core API
    and in the implementations themselves. Lines 343-347 describe our
    evaluation of the validation overhead of Task Bench.

# Reviewer D

  * Please see lines 1158-1187 for insights enabled by Task Bench.

  * Task Bench covers a wide variety of features, and does so for
    every programming system that it implements. We believe this
    compares favorably against traditional benchmarks which
    necessarily cover only a subset of features, and typically are
    implemented for only a limited number of programming systems.

  * As described in lines 711-750, measuring overhead alone is prone
    to certain pathologies in evaluation which can make the results
    unrepresentative. For this reason we prefer to use METG(50%) as a
    more representative measure. However, an interested reader can see
    Figure 7: the value at 0% efficiency is by definition the overhead
    of the system (as no useful work is being performed).

  * We do not have hardware metrics at this time but consider this a
    good direction for future work.

# Reviewer E

  * We provide separate kernel implementations for CPU and GPU.

  * See above for a description of the MPI implementation. The Dask
    implementation (which is representative of implicitly parallel
    systems) is shown in Listing 2.

  * Compare Figures 7 and 8. In Figure 7, the spread between systems
    (at the highest efficiency achieved by each system) is larger,
    because some systems reserve cores for internal use. With the
    exception of OmpSs (which we are investigating), this gap is not
    present in Figure 8.

  * A number of slower systems are challenging to run within the time
    limit; TensorFlow is one of these.

  * Regarding task granularity, \[2] describes data analytics
    platforms as executing tasks that "complete in hundreds of
    milliseconds". Note this is after optimizations to the scheduler
    to enable more fine-grained tasks, and task granularity is
    specifically cited as a challenge in scaling such systems. In
    contrast, \[3] (related work section) describes typical HPC
    systems that execute millisecond granularity tasks or smaller.

    \[3] Kay Ousterhout, The Case for Tiny Tasks in Compute Cluster,
    HotOS13.

    \[2]: Reazul Hoque, et al., Dynamic Task Discovery in PaRSEC - A
    Data-flow Task-based Runtime, ScalA17.

### 2019-10-28 ppopp20_reviews/rebuttal.md
Edit.

@@
     milliseconds". Note this is after optimizations to the scheduler
-    to enable more fine-grained tasks, and task granularity is
-    specifically cited as a challenge in scaling such systems. In
+    to enable more fine-grained tasks, and even then task granularity
+    is specifically cited as a challenge in scaling such systems. In
     contrast, \[3] (related work section) describes typical HPC

### 2019-10-29 ppopp20_reviews/rebuttal.md
Edits to the rebuttal.

--- 2019-10-28 ppopp20_reviews/rebuttal.md
@@
 receives to communicate task dependencies, and then after calling
-`MPI_Waitall()` we execute the corresponding tasks (whose inputs were
+`MPI_Waitall()` we execute the corresponding task (whose inputs were
 just received). We conducted an independent code review of our MPI
@@
 speaking) to design patterns and kernels that do reflect actual mini-
-or full-size applications in Task Bench.
+or full-size applications in Task Bench (and e.g. switch between
+multiple patterns over time as the application moves between different
+phases of computation).
 
@@
 is widely implemented, Task Bench provides massively more
-expressiveness and configurability. Task Bench provides even more
-implementations than the most widely implemented benchmark that we are
-aware of. And in terms of the depth of analysis that can be performed,
+expressiveness and configurability. Compared to those most widely implemented benchmarks, Task Bench provides even more
+implementations thanks to its ease of porting to new systems. And in terms of the depth of analysis that can be performed,
 the only benchmarks that provides substantially more visibility into
@@
 Though our evaluation uses METG(50%), this is not a limitation of
-METG. We chose 50% because it avoids certain pathologies associated
+our technique. We chose 50% because it avoids certain pathologies associated
 with low efficiency thresholds (see lines 711-750 in the paper), and
-also because it matches what we see implemented in practice. For
+also because it matches what we see in practice. For
 example, in a CSCS application for new projects, users are instructed
@@
 speed-up vs. linear speed-up above 50%)" \[1]. This corresponds to 50\%
-parallel efficiency.
+efficiency.
 
@@
 
-In the interest of answering the reviewers' questions as completely as
+In the interest of answering reviewers' questions as completely as
 possible, we include responses to specific reviewer questions below.
@@
   * See above for our description of the MPI implementation. A "task"
-    in MPI includes the time to execute a set of send/receive calls
-    needed to gather the inputs for a given task invocation, the call
-    the core API to execute the task.
+    in MPI includes the time to execute one point in the task graph (i.e. one set of send/receive calls
+    to gather inputs and one call to
+    the core API to execute the task body).
 
@@
 
-  * In our MPI+OpenMP configuration we use `OMP_NUM_THREADS` equal to
-    the number of physical cores on the node (32 on Cori Haswell). In
-    general this is set on a per-system basis depending on what is
-    most representative of how the system is used in practice.
+  * The Task Bench core API describes what to do but not how to do
+    it. Implementations are permitted to select the strategy that is
+    most efficient for a given machine. For example, our MPI+OpenMP
+    implementation uses shared memory within a node and uses MPI only
+    for inter-node communication. A similar strategy could be applied
+    on fat-node architectures such as Summit.
 
-  * Overlap occurs when there is task parallelism available in the
-    task graph. That is, when two tasks are independent (no mutual
-    dependencies) their execution can be overlapped with the other's
-    data transfers.
+  * Our MPI+OpenMP experiments set `OMP_NUM_THREADS` to the number of
+    physical cores on the node (32 on Cori Haswell). In general these
+    system-specific settings are chosen to optimize performance while
+    remaining faithful to what is representative of real-world usage.
+
+  * The opportunity for overlap occurs when two tasks are independent
+    (have no mutual dependencies). This occurs whenever multiple task
+    graphs are used (Figures 9d, 11). Task Bench's design also permits
+    this to be included in the dependence pattern itself, though we do
+    not explore this approach in the paper.
 
@@
     Figure 7: the value at 0% efficiency is by definition the overhead
-    of the system (as no useful work is being performed).
+    of the system (as no useful work is being performed, any remaining time is pure overhead).
 
@@
 
-  * We provide separate kernel implementations for CPU and GPU.
+  * We provide separate, optimized kernel implementations for CPU and GPU.
 
   * See above for a description of the MPI implementation. The Dask
-    implementation (which is representative of implicitly parallel
-    systems) is shown in Listing 2.
+    implementation in Listing 2 is implicitly parallel.
 
@@
     because some systems reserve cores for internal use. With the
-    exception of OmpSs (which we are investigating), this gap is not
-    present in Figure 8.
+    exception of OmpSs (which we are investigating), this gap disappears in Figure 8.
 
@@
   * Regarding task granularity, \[2] describes data analytics
-    platforms as executing tasks that "complete in hundreds of
-    milliseconds". Note this is after optimizations to the scheduler
-    to enable more fine-grained tasks, and task granularity is
-    specifically cited as a challenge in scaling such systems. In
-    contrast, \[3] (related work section) describes typical HPC
-    systems that execute millisecond granularity tasks or smaller.
+    platforms as executing "tiny" tasks that "complete in hundreds of
+    milliseconds", implying that traditional task granularities in this domain are larger. Note this is after scheduler optimizations 
+    that enable more fine-grained tasks. In
+    contrast, \[3] (related work) describes typical HPC
+    granularities in the millisecond range.
 
-    \[3] Kay Ousterhout, The Case for Tiny Tasks in Compute Cluster,
+    \[2] Kay Ousterhout, The Case for Tiny Tasks in Compute Cluster,
     HotOS13.
 
-    \[2]: Reazul Hoque, et al., Dynamic Task Discovery in PaRSEC - A
+    \[3]: Reazul Hoque, et al., Dynamic Task Discovery in PaRSEC - A
     Data-flow Task-based Runtime, ScalA17.

### 2019-10-29 ppopp20_reviews/rebuttal.md
Add an additional response to reviewer B.
minor fix on grammar and format

@@
 implementations thanks to its ease of porting to new systems. And in terms of the depth of analysis that can be performed,
-the only benchmarks that provides substantially more visibility into
+the only benchmarks that provide substantially more visibility into
 system-specific performance behaviors are themselves system-specific.
@@
 which is open to contribution. In fact, since the initial submission
-of the paper we added three new functionally correct implementations
+of the paper, we added three new functionally correct implementations
 to Task Bench. If these implementations are tuned to our satisfaction
@@
     for inter-node communication. A similar strategy could be applied
-    on fat-node architectures such as Summit.
+    to fat-node architectures such as Summit.
 
@@
     remaining faithful to what is representative of real-world usage.
+
+  * Note that METG is not a per-link statistic but characterizes the
+    programming system's overall ability to make effective use of a
+    machine. The "two weights" are already taken into account with
+    e.g. MPI+OpenMP and shared memory.
 
@@
 
-  * Note that there is an evaluation of trivial dependencies in Figure
-    10. Trivial corresponds to 0 dependencies per task in this chart.
+  * Note that there is an evaluation of trivial dependencies in Figure 10. 
+    Trivial corresponds to 0 dependencies per task in this chart.
 

### 2019-10-29 ppopp20_reviews/rebuttal.md
Edits.

@@
 
-METG is not fundamentally tied to peak performance (FLOP/s or B/s). If
+METG is not fundamentally tied to peak performance (FLOP/s or B/s), nor is it specific to Task Bench. If
 an application is not amenable to being characterized via peak
@@
 to "select the most parallel efficient job size (ratio of benchmark
-speed-up vs. linear speed-up above 50%)" \[1]. This corresponds to 50\%
+speed-up vs. linear speed-up above 50%)" \[1]. This corresponds to 50%
 efficiency.
@@
 
-# Reviewer D
+## Reviewer D
 
@@
 
-  * Task Bench covers a wide variety of features, and does so for
-    every programming system that it implements. We believe this
+  * Task Bench covers a wide variety of features (compute/memory
+    bound, varying task duration, task graph width and height,
+    dependence pattern, communication per task, etc.). Task Bench is
+    designed so that every feature is supported on every programming
+    system. We believe this
     compares favorably against traditional benchmarks which
@@
 
-# Reviewer E
+## Reviewer E
 

### 2019-10-30 ppopp20_reviews/code_sample.cpp
Check in MPI code sample.

@@ -1,0 +1,49 @@
+void execute_task_graph(Graph g) {
+  char *output = (char *)malloc(g.output_bytes_per_task);
+  char *scratch = (char *)malloc(g.scratch_bytes_per_task);
+  char **inputs = (char **)malloc(g.max_dependencies_per_task * sizeof(char *));
+  char *input_bytes = (size_t *)malloc(g.max_dependencies_per_task * sizeof(size_t));
+  long num_inputs;
+
+  long rank; // each rank executes one column in the task graph
+  std::vector<long> rank_by_point; // maps points to ranks
+  std::vector<MPI_Request> requests;
+
+  // initialize rank and rank_by_point...
+
+  for (long t = 0; t < g.height; ++t) {
+    if (g.contains_point(t, rank)) {
+      num_inputs = 0;
+      requests.clear();
+
+      for (long dep : g.dependencies(t, rank)) {
+        if (g.contains_point(t-1, dep)) {
+          MPI_Request req;
+          MPI_Irecv(
+            inputs[num_inputs], input_bytes[num_inputs], MPI_BYTE,
+            rank_by_point[dep], 0, MPI_COMM_WORLD, req);
+          requests.push_back(req);
+          num_inputs++;
+        }
+      }
+
+      for (long dep : g.reverse_dependencies(t, rank)) {
+        if (g.contains_point(t, dep)) {
+          MPI_Request req;
+          MPI_Isend(
+            output, g.output_bytes_per_task, MPI_BYTE,
+            rank_by_point[dep], 0, MPI_COMM_WORLD, req);
+          requests.push_back(req);
+        }
+      }
+
+      MPI_Waitall(requests.size(), requests.data(), MPI_STATUSES_IGNORE);
+
+      g.execute_point(
+        t, rank,
+        output, g.output_bytes_per_task,
+        inputs, input_bytes, num_inputs,
+        scratch, g.scratch_bytes_per_task);
+    }
+  }
+}

### 2019-10-30 ppopp20_reviews/rebuttal.md
Update rebuttal, include code sample.

--- 2019-10-29 ppopp20_reviews/rebuttal.md
@@
-We thank the reviewers for their detailed and insightful feedback.
+We thank the reviewers for their feedback.
 
-Task Bench is a new class of benchmark that aims to make it easy to
+TaskBench is a new class of benchmark that makes it easy to
 execute a wide variety of application patterns on a large number of
-distributed and parallel programming systems. Despite the name, Task
-Bench is not specific to task-based programming models. A "task" in
-Task Bench is simply any unit of code that runs to completion without
-requiring any network communication. Any application that can be
+distributed and parallel programming systems. Despite the name, TaskBench is not specific to task-based programming models. A "task" in
+TaskBench is simply any unit of code that runs to completion without
+requiring network communication. Any application that can be
 separated into units of work interleaved with communication can be
-adapted to Task Bench.
+adapted to TaskBench.
 
-As a case in point, our MPI implementation is written in an extremely
+As a case in point, our MPI implementation is written in a
 straightforward style, with interleaved phases of communication and
-computation. Each MPI rank is responsible for a set of columns in the
-task graph. During communication phases we use non-blocking sends and
-receives to communicate task dependencies, and then after calling
-`MPI_Waitall()` we execute the corresponding task (whose inputs were
-just received). We conducted an independent code review of our MPI
+computation. A code excerpt is included below and will appear in the
+final paper. We conducted an independent code review of our MPI
 implementation with developers of a major MPI implementation to
@@
 
-In general, we are confident in the performance of our Task Bench
+In general, we are confident in the performance of our TaskBench
 implementations because each one has been submitted to the developers
@@
 
-Task Bench covers a space of possible application patterns. This space
-is easy to extend, e.g. by adding new kernels, dependence patterns, or
-by choosing new values for the existing parameters (height, width,
-etc.). As can be seen in Listing 2 in the paper, the Task Bench core
-API is very general, and includes the ability to include or exclude
-arbitrary points within the space (line 17), and to specify arbitrary
-dependencies for each point in the task graph (line 19). Note that the
-dependence pattern can also change over time; this can be seen in the
-FFT and Tree patterns (Figure 1c and e). Thus, while the specific
-patterns evaluated in the paper are necessarily limited (as the
-reviewers have noted, our evaluation is already extremely
-comprehensive), it is entirely possible (and even easy, relatively
-speaking) to design patterns and kernels that do reflect actual mini-
-or full-size applications in Task Bench (and e.g. switch between
-multiple patterns over time as the application moves between different
-phases of computation).
+TaskBench covers a space of application patterns. This space is easy
+to extend, e.g. by adding kernels, dependence patterns, or by choosing
+new values for existing parameters (height, width, etc.). In the Dask
+implementation in Listing 2 in the paper (and MPI below), methods and
+fields of graph `g` are provided by the core API. This API is very
+general, and includes the ability to include or exclude arbitrary
+points (`contains_point()`), and to specify arbitrary dependencies for
+each task (`dependencies()`). Note that dependencies can also change
+over time (see Figure 1c and 1e). Thus, while the specific patterns
+evaluated in the paper are necessarily limited (as the reviewers have
+noted, our evaluation is very comprehensive), it is entirely possible
+(and even easy, relatively speaking) to design patterns and kernels
+that do reflect actual mini- or full-size applications in TaskBench
+(and switch between patterns over time as the application moves
+between different phases of computation).
 
-We believe Task Bench compares very favorably with other benchmarks in
-terms of the number of programming system features and behaviors that
-it is capable of exploring. Certainly, compared to any benchmark which
-is widely implemented, Task Bench provides massively more
-expressiveness and configurability. Compared to those most widely implemented benchmarks, Task Bench provides even more
-implementations thanks to its ease of porting to new systems. And in terms of the depth of analysis that can be performed,
-the only benchmarks that provide substantially more visibility into
-system-specific performance behaviors are themselves system-specific.
+We believe TaskBench compares very favorably with other benchmarks in
+terms of the programming system features and behaviors it's capable of
+exploring. Compared to any widely implemented benchmark, TaskBench
+provides substantially more expressiveness and configurability, while
+offering even more implementations thanks to its ease of porting to
+new systems. And in terms of the depth of analysis that can be
+performed, while we certainly do not mean to claim that TaskBench can
+do everything, the only benchmarks that provide substantially more
+visibility into system-specific performance behaviors are themselves
+system-specific.
 
-Task Bench is open source and we intend for it to be an active project
+TaskBench is open source and we intend for it to be an active project
 which is open to contribution. In fact, since the initial submission
 of the paper, we added three new functionally correct implementations
-to Task Bench. If these implementations are tuned to our satisfaction
+to TaskBench. If these implementations are tuned to our satisfaction
 in advance of the deadline they will be included in the final paper.
 
-METG is not fundamentally tied to peak performance (FLOP/s or B/s). If
+METG is not fundamentally tied to peak performance, nor is it specific to TaskBench. If
 an application is not amenable to being characterized via peak
@@
 \[1]: https://user.cscs.ch/access/report/
+
+A simplified excerpt from our MPI implementation follows.
+
+```
+void execute_task_graph(Graph g) {
+  char *output = (char *)malloc(g.output_bytes_per_task);
+  char *scratch = (char *)malloc(g.scratch_bytes_per_task);
+  char **inputs = (char **)malloc(g.max_dependencies_per_task * sizeof(char *));
+  char *input_bytes = (size_t *)malloc(g.max_dependencies_per_task * sizeof(size_t));
+  long num_inputs;
+
+  long rank;
+  std::vector<long> rank_by_point; // maps points to ranks
+  std::vector<MPI_Request> requests;
+
+  // initialize data structures...
+
+  for (long t = 0; t < g.height; ++t) {
+    if (g.contains_point(t, rank)) {
+      num_inputs = 0;
+      requests.clear();
+
+      for (long dep : g.dependencies(t, rank)) {
+        MPI_Request req;
+        MPI_Irecv(
+          inputs[num_inputs], input_bytes[num_inputs], MPI_BYTE,
+          rank_by_point[dep], 0, MPI_COMM_WORLD, req);
+        requests.push_back(req);
+        num_inputs++;
+      }
+
+      for (long dep : g.reverse_dependencies(t, rank)) {
+        MPI_Request req;
+        MPI_Isend(
+          output, g.output_bytes_per_task, MPI_BYTE,
+          rank_by_point[dep], 0, MPI_COMM_WORLD, req);
+        requests.push_back(req);
+      }
+
+      MPI_Waitall(requests.size(), requests.data(), MPI_STATUSES_IGNORE);
+
+      g.execute_point(
+        t, rank,
+        output, g.output_bytes_per_task,
+        inputs, input_bytes, num_inputs,
+        scratch, g.scratch_bytes_per_task);
+    }
+  }
+}
+```
 
@@
 
-  * See above for our description of the MPI implementation. A "task"
-    in MPI includes the time to execute one point in the task graph (i.e. one set of send/receive calls
-    to gather inputs and one call to
-    the core API to execute the task body).
+  * See our MPI excerpt above. A "task" in MPI includes the time to
+    execute one point in the task graph (i.e. one iteration of the `t`
+    loop).
 
@@
 
-  * The Task Bench core API describes what to do but not how to do
+  * The core API describes what to do but not how to do
     it. Implementations are permitted to select the strategy that is
@@
     (have no mutual dependencies). This occurs whenever multiple task
-    graphs are used (Figures 9d, 11). Task Bench's design also permits
+    graphs are used (Figures 9d, 11). Our design also permits
     this to be included in the dependence pattern itself, though we do
@@
 
-  * Note that there is an evaluation of trivial dependencies in Figure 10. 
+  * Please see the MPI excerpt above, and also note Listing 2 in the
+    paper shows a Dask implementation.
+
+  * There is an evaluation of trivial dependencies in Figure 10.
     Trivial corresponds to 0 dependencies per task in this chart.
 
-  * Note that in Figure 10 Task Bench achieves running times per task
-    as small as 390 ns (while still maintaining 50% efficiency). In
-    general Task Bench is carefully written to be as optimal as
-    possible (e.g. avoiding memory allocations), both in the core API
-    and in the implementations themselves. Lines 343-347 describe our
-    evaluation of the validation overhead of Task Bench.
+  * Figure 10 TaskBench achieves running times per task as small as
+    390 ns (while still maintaining 50% efficiency). In general
+    TaskBench is carefully written to be as optimal as possible
+    (e.g. avoiding memory allocations), both in the core API and in
+    the implementations themselves. Lines 343-347 describe our
+    evaluation of the validation overhead of TaskBench.
 
-# Reviewer D
+## Reviewer D
 
-  * Please see lines 1158-1187 for insights enabled by Task Bench.
+  * Please see lines 1158-1187 for insights enabled by TaskBench.
 
-  * Task Bench covers a wide variety of features, and does so for
-    every programming system that it implements. We believe this
+  * TaskBench covers a wide variety of features (compute/memory
+    bound, varying task duration, task graph width and height,
+    dependence pattern, communication per task, etc.). TaskBench is
+    designed so that every feature is supported on every programming
+    system. We believe this
     compares favorably against traditional benchmarks which
@@
 
-# Reviewer E
+## Reviewer E
 
-  * We provide separate, optimized kernel implementations for CPU and GPU.
+  * We provide separate, optimized kernel implementations for CPU and
+    GPU. Note the precise computation performed by the kernel is not
+    relevant; what is important is the computational behavior (compute
+    or memory bound), duration, and the amount of data touched, all of
+    which are fully configurable.
 
-  * See above for a description of the MPI implementation. The Dask
-    implementation in Listing 2 is implicitly parallel.
+  * See the MPI excerpt above (which is explicitly parallel), and the
+    Dask implementation in Listing 2 (which is implicitly parallel).
 

### 2019-10-30 ppopp20_reviews/code_sample.cpp

@@ -5,11 +5,11 @@ void execute_task_graph(Graph g) {
   char *input_bytes = (size_t *)malloc(g.max_dependencies_per_task * sizeof(size_t));
   long num_inputs;
 
-  long rank; // each rank executes one column in the task graph
+  long rank;
   std::vector<long> rank_by_point; // maps points to ranks
   std::vector<MPI_Request> requests;
 
-  // initialize rank and rank_by_point...
+  // initialize data structures...
 
   for (long t = 0; t < g.height; ++t) {
     if (g.contains_point(t, rank)) {
@@ -17,24 +17,20 @@ void execute_task_graph(Graph g) {
       requests.clear();
 
       for (long dep : g.dependencies(t, rank)) {
-        if (g.contains_point(t-1, dep)) {
-          MPI_Request req;
-          MPI_Irecv(
-            inputs[num_inputs], input_bytes[num_inputs], MPI_BYTE,
-            rank_by_point[dep], 0, MPI_COMM_WORLD, req);
-          requests.push_back(req);
-          num_inputs++;
-        }
+        MPI_Request req;
+        MPI_Irecv(
+          inputs[num_inputs], input_bytes[num_inputs], MPI_BYTE,
+          rank_by_point[dep], 0, MPI_COMM_WORLD, req);
+        requests.push_back(req);
+        num_inputs++;
       }
 
       for (long dep : g.reverse_dependencies(t, rank)) {
-        if (g.contains_point(t, dep)) {
-          MPI_Request req;
-          MPI_Isend(
-            output, g.output_bytes_per_task, MPI_BYTE,
-            rank_by_point[dep], 0, MPI_COMM_WORLD, req);
-          requests.push_back(req);
-        }
+        MPI_Request req;
+        MPI_Isend(
+          output, g.output_bytes_per_task, MPI_BYTE,
+          rank_by_point[dep], 0, MPI_COMM_WORLD, req);
+        requests.push_back(req);
       }
 
       MPI_Waitall(requests.size(), requests.data(), MPI_STATUSES_IGNORE);

### 2019-10-30 ppopp20_reviews/rebuttal.md
Swap things around.

@@
 
-A simplified excerpt from our MPI implementation follows.
+In the interest of answering reviewers' questions as completely as
+possible, we include responses to specific reviewer questions below.
+
+## Reviewer A
+
+  * See our MPI excerpt below. A "task" in MPI includes the time to
+    execute one point in the task graph (i.e. one iteration of the `t`
+    loop).
+
+## Reviewer B
+
+  * The core API describes what to do but not how to do
+    it. Implementations are permitted to select the strategy that is
+    most efficient for a given machine. For example, our MPI+OpenMP
+    implementation uses shared memory within a node and uses MPI only
+    for inter-node communication. A similar strategy could be applied
+    to fat-node architectures such as Summit.
+
+  * Our MPI+OpenMP experiments set `OMP_NUM_THREADS` to the number of
+    physical cores on the node (32 on Cori Haswell). In general these
+    system-specific settings are chosen to optimize performance while
+    remaining faithful to what is representative of real-world usage.
+
+  * Note that METG is not a per-link statistic but characterizes the
+    programming system's overall ability to make effective use of a
+    machine. The "two weights" are already taken into account with
+    e.g. MPI+OpenMP and shared memory.
+
+  * The opportunity for overlap occurs when two tasks are independent
+    (have no mutual dependencies). This occurs whenever multiple task
+    graphs are used (Figures 9d, 11). Our design also permits
+    this to be included in the dependence pattern itself, though we do
+    not explore this approach in the paper.
+
+## Reviewer C
+
+  * Please see the MPI excerpt below, and also note Listing 2 in the
+    paper shows a Dask implementation.
+
+  * There is an evaluation of trivial dependencies in Figure 10.
+    Trivial corresponds to 0 dependencies per task in this chart.
+
+  * Figure 10 TaskBench achieves running times per task as small as
+    390 ns (while still maintaining 50% efficiency). In general
+    TaskBench is carefully written to be as optimal as possible
+    (e.g. avoiding memory allocations), both in the core API and in
+    the implementations themselves. Lines 343-347 describe our
+    evaluation of the validation overhead of TaskBench.
+
+## Reviewer D
+
+  * Please see lines 1158-1187 for insights enabled by TaskBench.
+
+  * TaskBench covers a wide variety of features (compute/memory
+    bound, varying task duration, task graph width and height,
+    dependence pattern, communication per task, etc.). TaskBench is
+    designed so that every feature is supported on every programming
+    system. We believe this
+    compares favorably against traditional benchmarks which
+    necessarily cover only a subset of features, and typically are
+    implemented for only a limited number of programming systems.
+
+  * As described in lines 711-750, measuring overhead alone is prone
+    to certain pathologies in evaluation which can make the results
+    unrepresentative. For this reason we prefer to use METG(50%) as a
+    more representative measure. However, an interested reader can see
+    Figure 7: the value at 0% efficiency is by definition the overhead
+    of the system (as no useful work is being performed, any remaining time is pure overhead).
+
+  * We do not have hardware metrics at this time but consider this a
+    good direction for future work.
+
+## Reviewer E
+
+  * We provide separate, optimized kernel implementations for CPU and
+    GPU. Note the precise computation performed by the kernel is not
+    relevant; what is important is the computational behavior (compute
+    or memory bound), duration, and the amount of data touched, all of
+    which are fully configurable.
+
+  * See the MPI excerpt below (which is explicitly parallel), and the
+    Dask implementation in Listing 2 (which is implicitly parallel).
+
+  * Compare Figures 7 and 8. In Figure 7, the spread between systems
+    (at the highest efficiency achieved by each system) is larger,
+    because some systems reserve cores for internal use. With the
+    exception of OmpSs (which we are investigating), this gap disappears in Figure 8.
+
+  * A number of slower systems are challenging to run within the time
+    limit; TensorFlow is one of these.
+
+  * Regarding task granularity, \[2] describes data analytics
+    platforms as executing "tiny" tasks that "complete in hundreds of
+    milliseconds", implying that traditional task granularities in this domain are larger. Note this is after scheduler optimizations 
+    that enable more fine-grained tasks. In
+    contrast, \[3] (related work) describes typical HPC
+    granularities in the millisecond range.
+
+    \[2] Kay Ousterhout, The Case for Tiny Tasks in Compute Cluster,
+    HotOS13.
+
+    \[3]: Reazul Hoque, et al., Dynamic Task Discovery in PaRSEC - A
+    Data-flow Task-based Runtime, ScalA17.
+
+## MPI Implementation Excerpt
+
+Note the full implementation is both more general and has additional
+optimizations.
 
@@
 ```
-
-In the interest of answering reviewers' questions as completely as
-possible, we include responses to specific reviewer questions below.
-
-## Reviewer A
-
-  * See our MPI excerpt above. A "task" in MPI includes the time to
-    execute one point in the task graph (i.e. one iteration of the `t`
-    loop).
-
-## Reviewer B
-
-  * The core API describes what to do but not how to do
-    it. Implementations are permitted to select the strategy that is
-    most efficient for a given machine. For example, our MPI+OpenMP
-    implementation uses shared memory within a node and uses MPI only
-    for inter-node communication. A similar strategy could be applied
-    to fat-node architectures such as Summit.
-
-  * Our MPI+OpenMP experiments set `OMP_NUM_THREADS` to the number of
-    physical cores on the node (32 on Cori Haswell). In general these
-    system-specific settings are chosen to optimize performance while
-    remaining faithful to what is representative of real-world usage.
-
-  * Note that METG is not a per-link statistic but characterizes the
-    programming system's overall ability to make effective use of a
-    machine. The "two weights" are already taken into account with
-    e.g. MPI+OpenMP and shared memory.
-
-  * The opportunity for overlap occurs when two tasks are independent
-    (have no mutual dependencies). This occurs whenever multiple task
-    graphs are used (Figures 9d, 11). Our design also permits
-    this to be included in the dependence pattern itself, though we do
-    not explore this approach in the paper.
-
-## Reviewer C
-
-  * Please see the MPI excerpt above, and also note Listing 2 in the
-    paper shows a Dask implementation.
-
-  * There is an evaluation of trivial dependencies in Figure 10.
-    Trivial corresponds to 0 dependencies per task in this chart.
-
-  * Figure 10 TaskBench achieves running times per task as small as
-    390 ns (while still maintaining 50% efficiency). In general
-    TaskBench is carefully written to be as optimal as possible
-    (e.g. avoiding memory allocations), both in the core API and in
-    the implementations themselves. Lines 343-347 describe our
-    evaluation of the validation overhead of TaskBench.
-
-## Reviewer D
-
-  * Please see lines 1158-1187 for insights enabled by TaskBench.
-
-  * TaskBench covers a wide variety of features (compute/memory
-    bound, varying task duration, task graph width and height,
-    dependence pattern, communication per task, etc.). TaskBench is
-    designed so that every feature is supported on every programming
-    system. We believe this
-    compares favorably against traditional benchmarks which
-    necessarily cover only a subset of features, and typically are
-    implemented for only a limited number of programming systems.
-
-  * As described in lines 711-750, measuring overhead alone is prone
-    to certain pathologies in evaluation which can make the results
-    unrepresentative. For this reason we prefer to use METG(50%) as a
-    more representative measure. However, an interested reader can see
-    Figure 7: the value at 0% efficiency is by definition the overhead
-    of the system (as no useful work is being performed, any remaining time is pure overhead).
-
-  * We do not have hardware metrics at this time but consider this a
-    good direction for future work.
-
-## Reviewer E
-
-  * We provide separate, optimized kernel implementations for CPU and
-    GPU. Note the precise computation performed by the kernel is not
-    relevant; what is important is the computational behavior (compute
-    or memory bound), duration, and the amount of data touched, all of
-    which are fully configurable.
-
-  * See the MPI excerpt above (which is explicitly parallel), and the
-    Dask implementation in Listing 2 (which is implicitly parallel).
-
-  * Compare Figures 7 and 8. In Figure 7, the spread between systems
-    (at the highest efficiency achieved by each system) is larger,
-    because some systems reserve cores for internal use. With the
-    exception of OmpSs (which we are investigating), this gap disappears in Figure 8.
-
-  * A number of slower systems are challenging to run within the time
-    limit; TensorFlow is one of these.
-
-  * Regarding task granularity, \[2] describes data analytics
-    platforms as executing "tiny" tasks that "complete in hundreds of
-    milliseconds", implying that traditional task granularities in this domain are larger. Note this is after scheduler optimizations 
-    that enable more fine-grained tasks. In
-    contrast, \[3] (related work) describes typical HPC
-    granularities in the millisecond range.
-
-    \[2] Kay Ousterhout, The Case for Tiny Tasks in Compute Cluster,
-    HotOS13.
-
-    \[3]: Reazul Hoque, et al., Dynamic Task Discovery in PaRSEC - A
-    Data-flow Task-based Runtime, ScalA17.

### 2019-10-30 ppopp20_reviews/rebuttal.md
Tweak wording.

@@
 
-As a case in point, our MPI implementation is written in a
-straightforward style, with interleaved phases of communication and
+As a case in point, our MPI implementation is written in the style of communicating sequential processes, with interleaved phases of communication and
 computation. A code excerpt is included below and will appear in the

### 2019-10-31 ppopp20_reviews/rebuttal.md
Reformat.
Update.
Edits.

--- 2019-10-30 ppopp20_reviews/rebuttal.md
@@
 
-TaskBench is a new class of benchmark that makes it easy to
-execute a wide variety of application patterns on a large number of
-distributed and parallel programming systems. Despite the name, TaskBench is not specific to task-based programming models. A "task" in
-TaskBench is simply any unit of code that runs to completion without
-requiring network communication. Any application that can be
-separated into units of work interleaved with communication can be
-adapted to TaskBench.
+TaskBench is a new class of benchmark that makes it easy to execute a
+wide variety of application patterns on a large number of distributed
+and parallel programming systems. Despite the name, TaskBench is not
+specific to task-based programming models. A "task" in TaskBench is
+simply any unit of code that runs to completion without requiring
+network communication. Any application that can be separated into
+units of work interleaved with communication can be adapted to
+TaskBench.
 
-As a case in point, our MPI implementation is written in a
-straightforward style, with interleaved phases of communication and
-computation. A code excerpt is included below and will appear in the
-final paper. We conducted an independent code review of our MPI
-implementation with developers of a major MPI implementation to
-confirm that the code is written in a manner that is representative of
-how real-world MPI users write MPI code, and that it achieves good
-performance.
+As a case in point, our MPI implementation is written in the style of
+communicating sequential processes. A code excerpt is included below
+and will appear in the final paper. Our MPI implementation was
+independently reviewed by developers of a major MPI implementation to
+confirm that the code is representative of real-world MPI usage and
+follows best practices for efficiency and good performance
+(e.g. posting receives before sends, and avoiding memory allocations
+in the steady state).
 
@@
 
-METG is not fundamentally tied to peak performance, nor is it specific to TaskBench. If
-an application is not amenable to being characterized via peak
-performance, another measure of absolute performance can be used. For
-example, an aerodynamics application might choose mesh cells processed
-per second. The value of 100% efficiency would be set to the highest
-throughput that can be achieved on a single node. Other values of
-efficiency can be computed as the ratio of that highest throughput
+METG is not fundamentally tied to peak performance, nor is it specific
+to TaskBench. If an application is not amenable to being characterized
+via peak performance, another measure of absolute performance can be
+used. For example, an aerodynamics application might choose mesh cells
+processed per second. The value of 100% efficiency would be set to the
+highest throughput that can be achieved on a single node. Other values
+of efficiency can be computed as the ratio of that highest throughput
 achieved.
 
-Though our evaluation uses METG(50%), this is not a limitation of
-our technique. We chose 50% because it avoids certain pathologies associated
-with low efficiency thresholds (see lines 711-750 in the paper), and
-also because it matches what we see in practice. For
+Though our evaluation uses METG(50%), this is not a limitation of our
+technique. We chose 50% because it avoids certain pathologies
+associated with low efficiency thresholds (see lines 711-750 in the
+paper), and also because it matches what we see in practice. For
 example, in a CSCS application for new projects, users are instructed
@@
     (have no mutual dependencies). This occurs whenever multiple task
-    graphs are used (Figures 9d, 11). Our design also permits
-    this to be included in the dependence pattern itself, though we do
-    not explore this approach in the paper.
+    graphs are used (Figures 9d, 11). Our design also permits this to
+    be included in the dependence pattern itself, though we do not
+    explore this approach in the paper.
 
@@
 
-  * TaskBench covers a wide variety of features (compute/memory
-    bound, varying task duration, task graph width and height,
-    dependence pattern, communication per task, etc.). TaskBench is
-    designed so that every feature is supported on every programming
-    system. We believe this
-    compares favorably against traditional benchmarks which
-    necessarily cover only a subset of features, and typically are
-    implemented for only a limited number of programming systems.
+  * TaskBench covers a wide variety of features (compute/memory bound,
+    varying task duration, task graph width and height, dependence
+    pattern, communication per task, etc.). TaskBench is designed so
+    that every feature is supported on every programming system. We
+    believe this compares favorably against traditional benchmarks
+    which necessarily cover only a subset of features, and typically
+    are implemented for only a limited number of programming systems.
 
@@
     Figure 7: the value at 0% efficiency is by definition the overhead
-    of the system (as no useful work is being performed, any remaining time is pure overhead).
+    of the system (as no useful work is being performed, any remaining
+    time is pure overhead).
 
@@
     because some systems reserve cores for internal use. With the
-    exception of OmpSs (which we are investigating), this gap disappears in Figure 8.
+    exception of OmpSs (which we are investigating), this gap
+    disappears in Figure 8.
 
@@
     platforms as executing "tiny" tasks that "complete in hundreds of
-    milliseconds", implying that traditional task granularities in this domain are larger. Note this is after scheduler optimizations 
-    that enable more fine-grained tasks. In
-    contrast, \[3] (related work) describes typical HPC
-    granularities in the millisecond range.
+    milliseconds", implying that traditional task granularities in
+    this domain are larger. Note this is after scheduler optimizations
+    that enable more fine-grained tasks. In contrast, \[3] (related
+    work) describes HPC granularities as sub-millisecond.
 
@@
 
-Note the full implementation is both more general and has additional
-optimizations.
+Note the full implementation is both more general (e.g. it does not
+assume the number of ranks equals the width of the graph) and has
+additional optimizations.
 

### 2019-11-20 3_implementation.tex
Fiddle with figure placement.
Put the implementations back.

@@
 We have implemented Task Bench in the 15 parallel and distributed
-programming systems listed in Table~\ref{tab:systems}. We describe the systems, and any
-salient details of their Task Bench implementations, below.
+programming systems listed in Table~\ref{tab:systems}. These include
+traditional HPC programming models: MPI~\cite{MPI} (with bulk
+synchronous and point-to-point communication) and MPI+X (with OpenMP
+and CUDA); PGAS and actor models: Chapel~\cite{Chapel15},
+Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and newer task-based
+systems: OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40} (with task
+dependencies), PaRSEC (with \emph{parameterized task graphs}
+(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
+(DTD)~\cite{PARSEC_DTD}), Realm~\cite{Realm14},
+Regent~\cite{Regent15}, and StarPU~\cite{StarPU11}. Also, we consider
+several systems used in large scale data analytics and workflows:
+Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
+and TensorFlow~\cite{TensorFlow15}. We describe the systems, and any
+salient details of their Task Bench implementations, in our
+supplemental material.
 

### 2019-11-20 l2_mpi_sample.tex
MPI code sample.

\begin{lstlisting}[language=C++,caption={Excerpt from Task Bench implementation in MPI.},label={lst:code-sample},style=codeblock,float]
void execute_task_graph(Graph g) {
  char *output = (char *)malloc(g.output_bytes);
  char *scratch = (char *)malloc(g.scratch_bytes);
  char **inputs = (char **)malloc(/*...*/);
  long rank;
  // initialize data structures...

  std::vector<MPI_Request> requests;
  for (long row = 0; row < g.height; ++row) {
    if (g.contains_point(row, rank)) {
      long idx = 0;
      requests.clear();

      for (long dep : g.deps(row, rank)) {
        MPI_Request req;
        MPI_Irecv(
          inputs[idx], g.output_bytes, MPI_BYTE,
          dep, 0, MPI_COMM_WORLD, &req);
        requests.push_back(req);
        idx++;
      }

      for (long dep : g.reverse_deps(row, rank)) {
        MPI_Request req;
        MPI_Isend(
          output, g.output_bytes, MPI_BYTE,
          dep, 0, MPI_COMM_WORLD, &req);
        requests.push_back(req);
      }

      MPI_Waitall(
        requests.size(), requests.data(),
        MPI_STATUSES_IGNORE);

      g.execute_point(
        row, rank,
        output, g.output_bytes_per_task,
        inputs, input_bytes, idx,
        scratch, g.scratch_bytes_per_task);
    }
  }
}
\end{lstlisting}

### 2019-11-21 l2_mpi_sample.tex
short the pseudo code

@@
         MPI_Request req;
-        MPI_Irecv(
-          inputs[idx], g.output_bytes, MPI_BYTE,
+        MPI_Irecv(inputs[idx], g.output_bytes, MPI_BYTE,
           dep, 0, MPI_COMM_WORLD, &req);
-        requests.push_back(req);
-        idx++;
+        requests.push_back(req); idx++;
       }
@@
         MPI_Request req;
-        MPI_Isend(
-          output, g.output_bytes, MPI_BYTE,
+        MPI_Isend(output, g.output_bytes, MPI_BYTE,
           dep, 0, MPI_COMM_WORLD, &req);
@@
 
-      MPI_Waitall(
-        requests.size(), requests.data(),
+      MPI_Waitall(requests.size(), requests.data(),
         MPI_STATUSES_IGNORE);
 
-      g.execute_point(
-        row, rank,
-        output, g.output_bytes_per_task,
+      g.execute_point(row, rank, output, g.output_bytes_per_task,
         inputs, input_bytes, idx,

### 2019-11-21 3_implementation.tex
Update reference to supplemental material.
combine ompss with openmp

@@
 and TensorFlow~\cite{TensorFlow15}. We describe the systems, and any
-salient details of their Task Bench implementations, in our
-supplemental material.
+salient details of their Task Bench implementations, below.
 
@@
 
-\subsection{OmpSs}
+\subsection{OmpSs/OpenMP}
 
@@
 that is source-compatible with OpenMP. Our Task Bench implementation
-uses OpenMP 4.0-style task dependencies. Because OpenMP tasks have a fixed number of dependencies, we use a switch
+uses OpenMP 4.0-style~\cite{OpenMPSpec40} task dependencies. 
+Because OpenMP tasks have a fixed number of dependencies, we use a switch
 statement to implement the dynamic dependencies required for Task
 Bench. The implementation is otherwise straightforward. 
+The Task Bench implementation for OpenMP is similar to OmpSs.
+We tested with GNU GOMP and Intel KMP and found KMP to be better performing.
 
@@
 
-\subsection{OpenMP}
+%\subsection{OpenMP}
 
-OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
-parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
-for OpenMP is similar to OmpSs and uses
-task dependencies. We tested with GNU GOMP and Intel KMP
-and found KMP to be better performing.
+%OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
+%parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
+%for OpenMP is similar to OmpSs and uses
+%task dependencies. We tested with GNU GOMP and Intel KMP
+%and found KMP to be better performing.
 

### 2019-11-21 l2_mpi_sample.tex
Fiddle with code, tables.

@@
   std::vector<MPI_Request> requests;
-  for (long row = 0; row < g.height; ++row) {
-    if (g.contains_point(row, rank)) {
+  for (long t = 0; t < g.height; ++t) {
+    if (g.contains_point(t, rank)) {
       long idx = 0;
@@
 
-      for (long dep : g.deps(row, rank)) {
+      for (long dep : g.deps(t, rank)) {
         MPI_Request req;
@@
           dep, 0, MPI_COMM_WORLD, &req);
-        requests.push_back(req); idx++;
+        requests.push_back(req);
+        idx++;
       }
 
-      for (long dep : g.reverse_deps(row, rank)) {
+      for (long dep : g.reverse_deps(t, rank)) {
         MPI_Request req;
@@
 
-      g.execute_point(row, rank, output, g.output_bytes_per_task,
+      g.execute_point(t, rank, output, g.output_bytes_per_task,
         inputs, input_bytes, idx,

### 2019-11-21 1_introduction.tex
Fiddle with figure placement.

@@
 Adding a system to Task Bench involves implementing a set of standard
-services, such as executing a task or data transfer, simple enough to 
-be implemented with reasonable effort to a level of quality sufficient for comparative
-studies. Task Bench provides a core library that encapsulates functionality
-shared among systems, which not only reduces implementation effort
-but also makes it much easier to achieve truly apples-to-apples comparisons
-between systems.
+services, such as executing a task or data transfer. Note that, though
+the benchmarks are described in terms of task graphs, this is simply a
+convenient representation of the computation, and the underlying
+system need not provide any native support for tasks. Tasks are simply
+coarse units of work; dependencies induce communication. We provide
+Task Bench implementations in systems as diverse as MPI and
+Spark. Task Bench provides a core API that encapsulates functionality
+shared among systems, which reduces implementation effort and makes it
+much easier to achieve truly apples-to-apples comparisons between
+systems.
 
@@
 
-How does one predict whether the performance of a given
-system will be good enough for a particular application?
-Surprisingly, the most commonly reported performance metrics,
+How does one predict whether performance will be good enough for a given application?
+Surprisingly, the most commonly reported metrics,
 \emph{weak} and \emph{strong} scaling, do not directly characterize
@@
 \let\oldbrokenpenalty\brokenpenalty
-\brokenpenalty=0
+%% \brokenpenalty=0
 
 We conduct a comprehensive comparative study of Task Bench
-implementations in 15 programming systems:
-Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP, CUDA),
-OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
-PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
-Spark \cite{Spark10}, StarPU \cite{StarPU11},
-Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
-X10 \cite{X1005}. Using METG, we find that a number of factors---node
+implementations in 15 programming systems
+listed in Section~\ref{sec:implementation}.
+%% :
+%% Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP, CUDA),
+%% OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
+%% PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
+%% Spark \cite{Spark10}, StarPU \cite{StarPU11},
+%% Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
+%% X10 \cite{X1005}.
+Using METG, we find that a number of factors---node
 count, accelerators, and complex dependencies, among
@@
 
-\brokenpenalty=\oldbrokenpenalty
+%% \brokenpenalty=\oldbrokenpenalty
 
@@
 
+% Performance issues:
+% Chapel (communication performance, core pinning)
+% Dask (asymptotic complexity of dask.delayed)
+% PaRSEC (task pruning with non-trivial task graphs)
+% Realm (DMA system overhead)
+% TensorFlow (constant folding not parallel)
+
+During the development of Task Bench, we have identified performance
+bugs in five systems: Chapel, Dask, PaRSEC, Realm and
+TensorFlow. Three have been (or are being) fixed and the others have
+been acknowledged by the developers of the respective systems. In some
+cases these correspond to order of magnitude (or even asymptotic)
+improvements in the performance of the underlying systems---benefits
+which apply well beyond Task Bench to all classes of applications.
+
 The paper is organized as follows: Section~\ref{sec:task-bench}
-describes the design of Task Bench. Section~\ref{sec:implementation}
-discusses implementations in 15 programming systems.
-Section~\ref{sec:metg} defines the METG metric and its
-relationship to quantities of interest to application developers.
-Section~\ref{sec:evaluation} provides a comprehensive evaluation of
-Task Bench on up to 256 Haswell nodes of the Cori
-supercomputer~\cite{Cori} (GPU results are reported on Piz Daint~\cite{PizDaint}). Section~\ref{sec:related-work} relates to previous efforts; Section~\ref{sec:conclusion} concludes.
+describes the Task Bench design. Section~\ref{sec:implementation}
+discusses implementations in 15 systems.  Section~\ref{sec:metg}
+defines METG and its relationship to quantities of interest to
+application developers.  Section~\ref{sec:evaluation} provides a
+comprehensive evaluation on the Cori
+supercomputer~\cite{Cori}. Section~\ref{sec:related-work} relates to
+previous efforts; Section~\ref{sec:conclusion} concludes.

### 2019-11-21 t2_parameters.tex
Edits from Sean.
Edit the introduction to make some points more clear.

@@
 width & width of graph & degree of parallelism \\
-dependence & trivial, stencil, etc. & dependence relation \\
+dependence & trivial, stencil, etc. & communication pattern \\
+\quad \raisebox{0.35ex}{$\llcorner$} radix & \quad (for nearest pattern) & dependencies per task \\
 kernel & compute, memory, etc. & type of kernel \\
+\quad \raisebox{0.35ex}{$\llcorner$} iter. & \quad (for all kernels) & task duration \\
+\quad \raisebox{0.35ex}{$\llcorner$} span & \quad (for memory kernel) & bytes used per task per iter. \\
+\quad \raisebox{0.35ex}{$\llcorner$} scratch & \quad (for memory kernel) & total working set size \\
+\quad \raisebox{0.35ex}{$\llcorner$} imbal. & \quad (for load imbalance) & degree of imbalance \\
 output & bytes per dependency & degree of comm.
@@
 
-\vspace{1em}
-
-Options specific to dependence relations:
-
-\vspace{0.25em}
-
-\begin{tabular}{l | l | l}
-Parameter & Valid For & Purpose \\
-\hline
-radix & nearest & dependencies per task
-\end{tabular}
-
-\vspace{1em}
-
-Options specific to kernels:
-
-\vspace{0.25em}
-
-\begin{tabular}{l | l | l}
-Parameter & Valid For & Purpose \\
-\hline
-iterations & compute, memory & task duration/problem size \\
-span & memory & bytes used per task per iter. \\
-scratch & memory & total working set size \\
-imbalance & load imbalanced & degree of imbalance
-\end{tabular}
-
 \caption{Task Bench parameters.\label{tab:parameters}}
-\vspace{-0.5cm}
+\vspace{-0.75cm}
 \end{table}

### 2019-11-21 3_implementation.tex
minior edits on ompss
Edits.

@@
 
-\subsection{OmpSs/OpenMP}
+\subsection{OmpSs and OpenMP}
 
-OmpSs~\cite{OmpSs11} is a programming model for loop- and task-based parallelism
-that is source-compatible with OpenMP. Our Task Bench implementation
-uses OpenMP 4.0-style~\cite{OpenMPSpec40} task dependencies. 
-Because OpenMP tasks have a fixed number of dependencies, we use a switch
-statement to implement the dynamic dependencies required for Task
-Bench. The implementation is otherwise straightforward. 
-The Task Bench implementation for OpenMP is similar to OmpSs.
-We tested with GNU GOMP and Intel KMP and found KMP to be better performing.
+OmpSs~\cite{OmpSs11} and OpenMP~\cite{OpenMPSpec40} are programming
+models for loop- and task-based parallelism. (OmpSs is
+source-compatible with OpenMP.) The implementations use OpenMP 4.0
+task dependencies.  Because tasks have a fixed number of dependencies,
+we use a switch statement to implement dynamic dependencies for Task
+Bench. The implementations are otherwise straightforward.  For OpenMP,
+we tested GNU GOMP and Intel KMP and found KMP to be better
+performing.
 

### 2019-11-21 2_task_bench.tex
Fiddle with figure placement.
MPI code sample.
add a table to describe APIs of the core lib
Edit the introduction to make some points more clear.
Fiddle with figures.
Edits to section 2.

@@
 \input{t2_parameters}
+
+\brokenpenalty=0
 
@@
 each dependence edge in the graph.
+
+\brokenpenalty=\oldbrokenpenalty
 
@@
 
+Listing~\ref{lst:code-sample} shows an excerpt from the Task Bench
+implementation in MPI. Methods of the \lstinline{Graph} object
+\lstinline{g} are provided by Task Bench's core API and are shared
+among all implementations. These methods are summarized in
+Table~\ref{tab:api}. The MPI implementation follows the style of
+communicating sequential processes, and simply executes a set of send
+and receive calls (lines 24 and 16, respectively) followed by
+executing the task body (line 32). For simplicity, the excerpt
+assumes the number of MPI ranks is equal to the width of the graph;
+the full implementation is more general and provides additional
+optimizations. Despite its simplicity, this MPI implementation is
+among our best performing, and shows that Task Bench can be mapped to
+idiomatic code even in programming systems with no native concept of
+task.
+
+In addition specifying the shape and structure of the task graph, the
+Task Bench core API also provides implementations of the kernels
+executed by each task as well as other utility routines (to parse
+inputs and display results). A sample of the core API compute kernel
+is shown in Listing~\ref{lst:compute-kernel}. In addition to reducing
+the effort required to implement Task Bench, providing central
+implementations of these services ensures that all Task Bench
+implementations can be scripted uniformly and eliminates a potential
+source of performance disparity that can be a pitfall for mini-apps.
+
+%% (task(t,i) refers to the task at column i of timestep t)
+
+\input{t3_api}
+\input{l0_compute_kernel}
+%% \input{l1_code_sample}
+\input{l2_mpi_sample}
+
+\zap{
 Despite its generality, Task Bench is easy to implement, making it
@@
 implemented for each system, minimizing the work required for each additional system.
-
-\input{l0_compute_kernel}
-\input{l1_code_sample}
+} % zap
 

### 2019-11-22 4_metg.tex
Fiddle with figure placement.
Update to METG section.

@@
 of application scenarios, one question is how to characterize the
-overall performance of the programming systems under study. As noted
+performance and efficiency of the systems under study. As noted
 earlier, the overheads of the systems we consider varies by more than
@@
 the smallest average task granularity (i.e., task duration) such that $A$
-achieves overall efficiency of at least 50\%. For example, in
+achieves overall efficiency of at least 50\%. Note that METG is parameterized by the efficiency metric. For example, in
 compute-bound applications efficiency can be measured as the
@@
 the smallest task granularity achieved while maintaining at least 0.63
-TFLOP/s per node. For applications not amenable to being characterized by raw FLOP/s or B/s, an application-specific measure of performance can be
-used instead (e.g., mesh cells processed per second).
+TFLOP/s per node. However, METG is not tied to peak performance, and
+in applications not amenable to being characterized in this way,
+another application-specific measure of performance can be used. For
+example, a simulation on a mesh might use the number of mesh cells
+processed per second (i.e., total number of cells divided by wall
+clock time per iteration of the main simulation loop).
+
+The choice of 50\% is also a parameter and is not fundamental to
+METG. We use 50\% in our studies in order to avoid certain pathologies
+associated with lower efficiency thresholds (see
+Section~\ref{subsec:peak-performance-and-efficiency}), and also
+because it aligns with what we observe in practice. For example, in a
+CSCS application for new projects, users are instructed to "select the
+most parallel efficient job size (ratio of benchmark speed-up
+vs. linear speed-up above 50\%)" \cite{CSCSReport}. This corresponds
+to METG(50\%).
 
@@
 and thus it makes sense to discuss tasks even in systems
-with no explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
-compute-bound kernel included in the Task Bench implementation,
-described in more detail in Section~\ref{sec:task-bench}.
+with no explicit notion of tasking, such as MPI. In this case, the
+tasks run the compute-bound kernel shown in
+Section~\ref{sec:task-bench}.
 

### 2019-11-22 3_implementation.tex
More minor edits.
edits on implemntations

@@
 Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
-and TensorFlow~\cite{TensorFlow15}. We describe the systems, and any
-salient details of their Task Bench implementations, below.
+and TensorFlow~\cite{TensorFlow15}. We describe the systems and their implementations below.
 
@@
 
-Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors, or
-\emph{chares}, are objects in their own address space.
+Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors(
+\emph{chares}) are objects in their own address space.
 Chares communicate data and synchronize via messages and can be moved
@@
 distributed NumPy arrays and Pandas dataframes, and also a lower-level
-interface for launching tasks. We use the lower-level interface to
-implement Task Bench.
+interface for launching tasks. We implement Task Bench in the lower-level interface.
 
@@
 OmpSs~\cite{OmpSs11} and OpenMP~\cite{OpenMPSpec40} are programming
-models for loop- and task-based parallelism. (OmpSs is
-source-compatible with OpenMP.) The implementations use OpenMP 4.0
-task dependencies.  Because tasks have a fixed number of dependencies,
+models for loop- and task-based parallelism. (OpenMP is
+source-compatible with OmpSs.) The implementations use OpenMP 4.0
+tasking.  Because tasks have a fixed number of dependencies,
 we use a switch statement to implement dynamic dependencies for Task
-Bench. The implementations are otherwise straightforward.  For OpenMP,
-we tested GNU GOMP and Intel KMP and found KMP to be better
+Bench. For OpenMP, we tested GNU GOMP and Intel KMP and found KMP to be better
 performing.
@@
 which implements the Legion programming model~\cite{Legion12}. We use
-Regent rather than Legion directly because the Regent compiler
-provides a critical optimization that
-improves scalability~\cite{ControlReplication17}.
+Regent rather than Legion since the Regent compiler
+provides a critical optimization to
+improve scalability~\cite{ControlReplication17}.
 
@@
 
-StarPU~\cite{StarPU11} is a task-based system that supports a \emph{sequential task flow} (STF)
+StarPU~\cite{StarPU11} is a task-based system utilizing a \emph{sequential task flow}
 programming model similar to PaRSEC's DTD. Our Task Bench implementation in
@@
 requested. The Swift/T compiler performs a number of optimizations to
-improve performance of highly parallel programs~\cite{Armstrong14}. Our Task Bench in Swift/T is straightforward, using Swift/T's dataflow
+improve performance of highly parallel programs~\cite{Armstrong14}. 
+Our Task Bench in Swift/T uses it's dataflow
 semantics to capture dependencies on other tasks.
@@
 Task Bench. TensorFlow programs operate via explicit graph construction. Task graphs are composed in Python and run by an
-execution engine written in C++. Our Task Bench implementation is straightforward.
+execution engine written in C++.
 

### 2019-11-22 3_implementation.tex
Edits to the implementation section.

@@
 
-Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors(
-\emph{chares}) are objects in their own address space.
-Chares communicate data and synchronize via messages and can be moved
-to balance load. Our Task Bench implementation uses a chare
-array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
-dependencies are all available.
+Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors
+(\emph{chares}) do not share memory and communicate via messages. Our Task Bench implementation uses a chare
+array for the task graph, with one chare for each column, and uses messages to implement dependencies.
 
@@
 
-Our MPI~\cite{MPI} implementation of Task Bench represents the common
-case of point-to-point communication with distinct computation and
-communication phases. We experimented with a variety of implementation
-strategies and found the best performing to be using
-\lstinline[language=C++]{MPI_Isend} and
-\lstinline[language=C++]{MPI_Irecv} to implement the communication
-phase, posting receives before sends. Each task dependency maps to one
-send/receive pair in MPI. We also provide a bulk synchronous
-implementation which enforces the boundary between communication and
-computation with \lstinline[language=C++]{MPI_Barrier}.
+We provide two MPI~\cite{MPI} implementations, written in the style of
+communicating sequential processes (CSP) and bulk synchronous
+programming (BSP), respectively. An excerpt from the CSP
+implementation is shown in Listing~\ref{lst:code-sample}. The BSP
+implementation uses \lstinline[language=C++]{MPI_Barrier} to enforce
+boundaries between phases.
 
@@
 performance of hierarchical programming models. Our MPI+OpenMP
-implementation uses forall-style parallel loops to execute tasks,
-interleaved with MPI point-to-point communication as above. Our
-MPI+CUDA implementation follows an offload model where data is copied
-to and from the GPU on every timestep.
+implementation uses forall-style parallel loops to execute tasks, but
+otherwise follows the CSP implementation above. The code uses shared
+memory for data movement within a rank. Our MPI+CUDA implementation
+follows an offload model where data is copied to and from the GPU on
+every timestep.
 
-\subsection{OmpSs and OpenMP}
+\subsection{OpenMP and OmpSs}
 
-OmpSs~\cite{OmpSs11} and OpenMP~\cite{OpenMPSpec40} are programming
-models for loop- and task-based parallelism. (OpenMP is
-source-compatible with OmpSs.) The implementations use OpenMP 4.0
+OpenMP~\cite{OpenMPSpec40} and OmpSs~\cite{OmpSs11} are programming
+models for loop- and task-based parallelism. (OmpSs is
+source-compatible with OpenMP.) The implementations use OpenMP 4.0
 tasking.  Because tasks have a fixed number of dependencies,
@@
 
-In the DTD model, tasks are enumerated in program order (by
-executing the program), and dependencies between tasks are
-identified automatically based on the input and output data of tasks. 
-A task depends on another task if it reads data written by the other task,
-and the data is copied automatically if the two tasks are executed on
-different nodes. To
-improve scalability, the program is executed in
-SPMD fashion on all nodes, and the user is responsible for eliding
-tasks that are not directly connected to those that are to be executed
-on the current node.
+In the DTD model, tasks are enumerated in program order (by executing
+the program), and dependencies are identified automatically based on
+the arguments to tasks. To improve scalability, the program is
+executed in SPMD fashion on all nodes, and the user is responsible for
+eliding tasks that are not direct predecessors or successors to those
+executed on the current node.
 
@@
 improve performance of highly parallel programs~\cite{Armstrong14}. 
-Our Task Bench in Swift/T uses it's dataflow
+Our Task Bench in Swift/T uses its dataflow
 semantics to capture dependencies on other tasks.

### 2019-11-22 1_introduction.tex
Edits.

@@
 
-Using Task Bench we were able to quickly incorporate 15 very different parallel and distributed
-programming systems.  By running all systems on common benchmarks we were able to quantify phenomena
+This approach has allowed us to incorporate 15 very different parallel
+and distributed programming systems (listed in
+Table~\ref{tab:systems}).  By running all systems on common benchmarks
+we were able to quantify phenomena
 that have never before been measured.
@@
 
-We conduct a comprehensive comparative study of Task Bench
-implementations in 15 programming systems
-listed in Section~\ref{sec:implementation}.
+We conduct a comprehensive study of all 15 Task Bench implementations on
+up to 256 Haswell nodes of the Cori supercomputer~\cite{Cori}.
 %% :
@@
 
-%% We find asynchronous systems can provide benefits
-%% (e.g., overlapping compute and communication), but only if system overheads do not exceed roughly
-%% 100~\textmu{}s.
-
-%% This implies that simplifying the application space too much
-%% (e.g., by using trivial dependencies) risks overemphasizing the
-%% importance of the lightweight mechanisms in such systems---a risk that
-%% can be mitigated by using Task Bench's more complete coverage of the
-%% space of application scenarios.
-
-\zap{
-Using METG, we determine the baseline
-runtime overhead of each system along with overheads associated with
-increasing numbers of and/or variations in the pattern of dependencies. We
-also explore each system's ability to scale, hide communication
-overhead, and mitigate load imbalance.
-} % zap
-
 % Performance issues:
@@
 
-During the development of Task Bench, we have identified performance
-bugs in five systems: Chapel, Dask, PaRSEC, Realm and
-TensorFlow. Three have been (or are being) fixed and the others have
-been acknowledged by the developers of the respective systems. In some
-cases these correspond to order of magnitude (or even asymptotic)
-improvements in the performance of the underlying systems---benefits
-which apply well beyond Task Bench to all classes of applications.
+Beyond comparative study, the ability to explore a large configuration
+space also enables the discovery of bugs in the underlying systems. We
+found five performance issues, ranging from communication efficiency
+(Chapel, Realm), to the efficiency of task pruning, analysis and
+constant folding (PaRSEC, Dask and TensorFlow). Three have been fixed
+and all have been acknowledged by the developers of the respective
+systems. (All were either fixed or worked around in our experiments.)
+In some cases these correspond to order of magnitude (or even
+asymptotic) improvements in the performance of the underlying
+systems---benefits which apply well beyond Task Bench to all classes
+of applications.
 
@@
 application developers.  Section~\ref{sec:evaluation} provides a
-comprehensive evaluation on the Cori
-supercomputer~\cite{Cori}. Section~\ref{sec:related-work} relates to
+comprehensive evaluation on Cori. Section~\ref{sec:related-work} relates to
 previous efforts; Section~\ref{sec:conclusion} concludes.

### 2019-11-22 3_implementation.tex
More edits.
Edits.

@@
 
-\subsection{MPI}
+\subsection{MPI and MPI+X}
 
@@
 
-\subsection{MPI+X}
-
-We provide two MPI+X implementations of Task Bench to evaluate the
-performance of hierarchical programming models. Our MPI+OpenMP
-implementation uses forall-style parallel loops to execute tasks, but
-otherwise follows the CSP implementation above. The code uses shared
-memory for data movement within a rank. Our MPI+CUDA implementation
-follows an offload model where data is copied to and from the GPU on
-every timestep.
+We provide two MPI+X implementations to evaluate hierarchical
+programming models. Our MPI+OpenMP implementation uses forall-style
+parallel loops to execute tasks, but otherwise follows the CSP
+implementation above. The code uses shared memory for data movement
+within a rank. Our MPI+CUDA implementation follows an offload model
+where data is copied to and from the GPU on every timestep.
 
@@
 models for loop- and task-based parallelism. (OmpSs is
-source-compatible with OpenMP.) The implementations use OpenMP 4.0
+source-compatible with OpenMP.) Our implementations use OpenMP 4.0
 tasking.  Because tasks have a fixed number of dependencies,
@@
 
-PaRSEC is a task-based programming system supporting two distinct
+PaRSEC supports two distinct
 programming models: \emph{parameterized task graphs}
 (PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
-(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which programmers
+(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which users
 write a concise, algebraic description of the tasks and dataflows in
@@
 
-In the DTD model, tasks are enumerated in program order (by executing
-the program), and dependencies are identified automatically based on
-the arguments to tasks. To improve scalability, the program is
-executed in SPMD fashion on all nodes, and the user is responsible for
-eliding tasks that are not direct predecessors or successors to those
-executed on the current node.
+In DTD, tasks are enumerated in program order and dependencies are
+identified automatically based on task arguments. To improve
+scalability, the program is executed in SPMD fashion on all nodes, and
+the user is responsible for eliding tasks that are not direct
+predecessors or successors to those executed on the current node.
 
-In our Task Bench implementation, we evaluate both PTG and DTD models,
-as well as a third option based on DTD that has additional manual
-optimizations applied to further reduce dynamic checks for what tasks
-should be executed on a node, which further improves
-scalability.
+We implement Task Bench with both PTG and DTD, as well as a third
+option based on DTD that has additional manual optimizations applied
+to further reduce the cost of dynamic checks for task elision.
 

### 2019-11-22 3_implementation.tex
More shrinking.

@@
 Spark~\cite{Spark10} is an implicitly-parallel programming model for
-large scale data analytics.
+data analytics.
 The core abstractions in Spark are functional operators such as map,
@@
 Spark has tasks internally but they are not exposed to the user,
-so a Task Bench implementation must map the task graph to a set of operators that result in the
+so a Task Bench implementation must map the task graph to operators that result in the
 desired execution. We use
 \lstinline[language=Scala]{flatMap} and
-\lstinline[language=Scala]{groupByKey} to generate task
+\lstinline[language=Scala]{groupByKey} to generate
 dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
@@
 X10~\cite{X1005} is an explicitly parallel programming language for
-place-based programming. The core features of X10 are \emph{places}
+place-based programming. Features include \emph{places}
 which represent distributed memories, a PGAS model where references to

### 2019-11-22 3_implementation.tex
Keep squeezing.

@@
 programming systems listed in Table~\ref{tab:systems}. These include
-traditional HPC programming models: MPI~\cite{MPI} (with bulk
-synchronous and point-to-point communication) and MPI+X (with OpenMP
+traditional HPC programming models: MPI~\cite{MPI} and MPI+X (OpenMP
 and CUDA); PGAS and actor models: Chapel~\cite{Chapel15},
-Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and newer task-based
-systems: OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40} (with task
-dependencies), PaRSEC (with \emph{parameterized task graphs}
-(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
-(DTD)~\cite{PARSEC_DTD}), Realm~\cite{Realm14},
+Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and task-based
+systems: OmpSs~\cite{OmpSs11}, OpenMP 4.0~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13, PARSEC_DTD}, Realm~\cite{Realm14},
 Regent~\cite{Regent15}, and StarPU~\cite{StarPU11}. Also, we consider
@@
 Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
-and TensorFlow~\cite{TensorFlow15}. We describe the systems and their implementations below.
+and TensorFlow~\cite{TensorFlow15}. We describe the systems and implementations below.
 
@@
 Realm~\cite{Realm14} is an explicitly-parallel task-based programming
-model used internally by Legion~\cite{Legion12} and
-Regent~\cite{Regent15}. A Realm implementation of Task Bench is a limit study of what can be achieved with Legion or Regent.
+model used internally by Regent~\cite{Regent15}. Realm is thus a limit
+study of what can be achieved with Regent.
 
@@
 
-TensorFlow~\cite{TensorFlow15} is a programming system designed for
-deep learning workloads. Although TensorFlow's API exposes
-machine learning concepts, internally TensorFlow is a task graph execution engine, making it a good fit for
-Task Bench. TensorFlow programs operate via explicit graph construction. Task graphs are composed in Python and run by an
-execution engine written in C++.
+TensorFlow~\cite{TensorFlow15} is designed for deep learning
+workloads. Although TensorFlow's API exposes machine learning
+concepts, internally TensorFlow is a task graph execution engine,
+making it a good fit for Task Bench. TensorFlow programs operate via
+explicit graph construction. Task graphs are composed in Python and
+run by a C++ execution engine.
 

### 2019-11-22 3_implementation.tex
Now under the limit.

@@
 communicating sequential processes (CSP) and bulk synchronous
-programming (BSP), respectively. An excerpt from the CSP
+programming (BSP). An excerpt from the CSP
 implementation is shown in Listing~\ref{lst:code-sample}. The BSP
@@
 to disk, confirmed there is no measurable overhead due to JNI
-calls from Java to C or due to the serialization of data structures, and cross-checked results with known cases
-that hit optimal task throughput in Spark.
+calls from Java to C or due to serializing data, and cross-checked results with known cases
+with optimal scheduler throughput.
 

### 2019-11-22 0_abstract.tex
Apply Alex's comments.

@@
 programming systems under a variety of application scenarios. Task
-Bench lowers the barrier to benchmarking multiple programming systems by making
+Bench lowers the barrier to benchmarking and comparing multiple programming systems by making
 the implementation for a given system orthogonal to the benchmarks

### 2019-11-22 1_introduction.tex
Apply Alex's comments.

@@
 
-Benchmarks in Task Bench are based on the observation that many applications can be modeled
-as a set of \emph{tasks}, or coarse-grain units of work, with
-dependencies between tasks representing the communication and synchronization required for
-distributed execution. A benchmark consists of a \emph{task graph} with tasks for
-each point in an \emph{iteration space}, and dependencies
-determined by a \emph{dependence relation}. This design permits a
-concise description of a wide variety of patterns relevant to
-scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
-seen in structured and unstructured mesh codes), sweeps (such as used
-in the discrete ordinates method of radiation simulation), FFTs, trees
-(for divide and conquer algorithms), and so on. Tasks execute kernels with a
+Benchmarks in Task Bench are based on the observation that regardless
+of the programming system in which an application is written, many
+applications can be modeled as coarse-grain units of work, called
+\emph{tasks}, with dependencies between tasks representing the
+communication and synchronization required for parallel and
+distributed execution. By explicitly modeling the \emph{task graph}
+(with tasks as vertices and dependencies as edges), we make it
+possible to explore a wide variety of patterns relevant to
+parallel and distributed computing: trivial parallelism, halo exchanges (as
+in structured and unstructured mesh codes), sweeps (as
+in the discrete ordinates method of simulating radiation), FFTs, trees
+(for divide and conquer), DNNs, graph analytics, etc. Tasks execute kernels with a
 variety of computational properties, including compute- and
-memory-bound operations of varying durations. Dependencies between
-tasks can be configured to carry communication payloads of varying size, permitting
-the design of communication-bound cases. Finally, multiple
+memory-bound loops of varying duration. Dependencies can be configured to carry communication payloads of varying size. Finally, multiple
 (potentially heterogeneous) task graphs can be executed concurrently
 to introduce task parallelism into the workload. Together, these
-design elements enable the exploration of a large space of application
-behaviors.
+elements enable the exploration of a large space of application
+behaviors---and make it easy to explore cases limited by runtime
+overhead as well as ones where computation or communication is
+dominant.
 
@@
 convenient representation of the computation, and the underlying
-system need not provide any native support for tasks. Tasks are simply
-coarse units of work; dependencies induce communication. We provide
+system need not provide any native support for tasks. We provide
 Task Bench implementations in systems as diverse as MPI and
@@
 
-This approach has allowed us to incorporate 15 very different parallel
+This approach has allowed us to benchmark 15 very different parallel
 and distributed programming systems (listed in
@@
 How does one predict whether performance will be good enough for a given application?
-Surprisingly, the most commonly reported metrics,
+The most commonly reported metrics,
 \emph{weak} and \emph{strong} scaling, do not directly characterize

### 2019-11-22 2_task_bench.tex
Edits.
Apply Alex's comments.

@@
 
-Task graphs are a combination of an iteration space (with a task for
-each point in the space) with a dependence relation.
+Task graphs are a combination of an \emph{iteration space} (with a task for
+each point in the space) with a \emph{dependence relation}.
 For simplicity, but without loss of generality, the iteration space in
@@
 graphs that can be implemented with Task Bench. Note that layout is
-significant---in particular, column $i$ represents all tasks
-with index $i$ in the iteration space over all the time
-steps. Generally speaking each column will be
+significant: generally speaking each column will be
 assigned to execute on a different processor core.
@@
 dependence relation enumerates the set of tasks from the
-previous time step each task depends on. This permits a wide variety
-of dependence patterns to be implemented that are relevant to real
-applications in high-performance scientific computing and data analysis: stencils,
-sweeps, FFTs, trees, etc. Dependence patterns may also be
+previous time step each task depends on, permitting a wide variety
+of patterns to be implemented that are relevant to real
+applications: stencils,
+sweeps, FFTs, trees, etc. Dependence relations may be
 parameterized, such as picking the $K$ nearest neighbors, or $K$
-distant neighbors. Table~\ref{tab:equations} shows equations for the
+distant neighbors.
+They may also vary over time, such as in the FFT pattern.
+The set of dependence relations is extensible, making it easy to add patterns to represent new classes applications.
+Table~\ref{tab:equations} shows equations for the
 dependence relations of the patterns in Figure~\ref{fig:task-graphs},
@@
 
-In addition specifying the shape and structure of the task graph, the
+%% (task(t,i) refers to the task at column i of timestep t)
+
+\input{t3_api}
+\input{l0_compute_kernel}
+%% \input{l1_code_sample}
+\input{l2_mpi_sample}
+
+In addition to specifying the shape and structure of the task graph, the
 Task Bench core API also provides implementations of the kernels
@@
 source of performance disparity that can be a pitfall for mini-apps.
-
-%% (task(t,i) refers to the task at column i of timestep t)
-
-\input{t3_api}
-\input{l0_compute_kernel}
-%% \input{l1_code_sample}
-\input{l2_mpi_sample}
 

### 2019-11-22 4_metg.tex
Apply Alex's comments.

@@
 performance and efficiency of the systems under study. As noted
-earlier, the overheads of the systems we consider varies by more than
-five orders of magnitude. This makes it challenging to extract useful
+earlier, the overheads of the systems we consider vary by more than
+five orders of magnitude, making it challenging to extract useful
 information from weak and strong scaling runs.
@@
 The choice of 50\% is also a parameter and is not fundamental to
-METG. We use 50\% in our studies in order to avoid certain pathologies
+METG. We use 50\% in our studies to avoid pathologies
 associated with lower efficiency thresholds (see
 Section~\ref{subsec:peak-performance-and-efficiency}), and also
-because it aligns with what we observe in practice. For example, in a
-CSCS application for new projects, users are instructed to "select the
+because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to "select the
 most parallel efficient job size (ratio of benchmark speed-up
@@
 shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, leading
-to a vertical asymptote as overhead dominates useful work.
+to a vertical asymptote as overhead comes to dominate useful work.
 

### 2019-11-22 5_evaluation.tex
Fiddle with figure placement.
Move load imbalance back.
Update reference to supplement.
Edit section title.
Apply Alex's comments.

@@
 runtime usage; these systems take a minor hit in peak FLOP/s compared
-to systems which share all cores between the application and runtime. Some of the
+to systems that share all cores between the application and runtime. Some of the
 higher overhead systems struggle to achieve peak FLOP/s, though in most cases
@@
 ideal scaling. Most notably, Spark is primarily
-intended industrial data center applications with task
-granularities measured in tens of seconds. Spark uses a centralized
+intended for industrial data center applications with task
+granularities measured in seconds. Spark uses a centralized
 controller, which limits throughput, and this is visible in the figure
@@
 
-\input{e5_imbalance}
-
-\input{e10_cuda_efficiency}
-
-\subsection{Communication Hiding}
+\subsection{Overlapping Communication and Computation}
 
@@
 \subsection{Load Imbalance}
+
+\input{e5_imbalance}
 
@@
 
-In order to determine the cost of scheduling tasks on the GPU, Figure~\ref{fig:cuda-efficiency} compares MPI
+\input{e10_cuda_efficiency}
+
+To determine the cost of scheduling tasks on GPUs, Figure~\ref{fig:cuda-efficiency} compares MPI
 and MPI+CUDA on the Piz Daint supercomputer. The CUDA

### 2019-11-22 6_related.tex
Edits from Sean.
Fiddle with figure placement.
Apply Alex's comments.

@@
 % paper.
-\let\oldbrokenpenalty\brokenpenalty
+%% \let\oldbrokenpenalty\brokenpenalty
 \brokenpenalty=0
@@
 Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
-consist mostly of small kernels for dense array computations and has
+consists mostly of small kernels for dense array computations and has
 implementations in OpenMP~\cite{NASOpenMP99}, MPI and

### 2019-11-22 2_task_bench.tex
Uncapitalize after colon.

@@
 a Task Bench implementation is in the core library, implementations of
-Task Bench are small: Our 15 Task Bench implementations range from 88
+Task Bench are small: our 15 Task Bench implementations range from 88
 to 1500 lines, with several hundred lines being

### 2019-11-22 5_evaluation.tex
Uncapitalize after colon.

@@
 One question when considering different programming
-systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
+systems is: how much overhead does the system add? This question is tricky to answer directly because some systems introduce
 overhead \emph{inline} (i.e., by running system internal processes on
@@
 different), the conclusion here is similar to
-Section~\ref{subsec:scalability}: The cost of sending data and
+Section~\ref{subsec:scalability}: the cost of sending data and
 tasks to the GPU imposes a higher task granularity to achieve

### 2019-11-22 7_conclusion.tex
Edit conclusion.

--- 2019-07-30 7_conclusion.tex
@@
 
-\item Newer task-based systems have performance sufficient for
-  weak scaling many HPC workloads, but more work is needed to strong
-  scale the most demanding codes.
+\item Task Bench has proven effective in finding performance issues
+  and has lead to substantial improvements in several of the systems
+  we study.
 

### 2020-02-03 pldi20_reviews/rebuttal.md
Initial rebuttal.

We thank the reviewers for their insightful feedback.

Reviewers A and D note that the bugs described in lines 197-199
currently do not have a larger analysis. We include a draft of this
analysis below and will incorporate it into the final paper.

Regarding novelty, reviewer C is correct that a major contribution of
our paper is a thorough empirical evaluation, but that this is made
possible by substantial contributions in other areas. Namely, the
design of task bench which permits $\mathcal{O}(N+M)$ rather than
$\mathcal{O}(NM)$ implementation effort, and the definition of METG,
are both critical. Without the former, a comparison of 15 systems
would be intractable without substantially simplifying the
benchmark. Without the latter, the results would be prone to bias, and
even more difficult to interpret.

Contrary to the suggestion of reviewer A, we found that the METGs we
found in practice did not always match our intuitions. In fact, some
were off to such a large degree (multiple orders of magnitude) that
they prompted the development teams to conduct major new development
to address our results. These cases are included in our analysis
below.

It is true that METG is a configuration-specific metric. That is, it
can vary depending on the network, CPU, etc. However it is our
experience that METG for a given programming system tends to span a
range of typical values, and that these can guide both application and
system development. Also, more precise measurements can identify
specific bugs as noted above.

Reviewer B asks about the impact of the network. It is our experience
that, except at the absolute lowest METGs, most programming system
overhead is on the CPU. Thus our study can highlight ways in which
these systems can be optimized to reduce unnecessary
overheads. However, we agree that a comparison across network
architectures would be a good direction for future work.

TPS is reported in citations \[6, 28, 32].

Fitting a Task Bench configuration to a specific application is a
direction for future work, but the patterns in Figure 1 are all
derived from common scientific computing workloads (and this is an
extensible set).

Irregular parallelism is enabled by allowing dependencies to vary over
time. For example, the full Task Bench implementation includes a
random dependence pattern.

The OpenMP implementation uses a switch to execute `#pragma omp task`
statements with varying numbers of `depends` clauses. An excerpt is
included below. Note all our implementations are open source and a
link will be included in the final paper.

Responding to reviewer C: the task graph is streamed whenever
possible. For example, both Realm and Dask operate on explicit task
graphs, but Realm is streamed (the API permits executing one chunk of
the graph while the next is being constructed), whereas Dask requires
the entire graph to be constructed. But see also our note on Dask
below.

Problem size corresponds to the `iterations` parameter to Listing
1. Note that since these are idealized kernels, what matters is the
total number of FLOPs executed, which directly corresponds to
`iterations`.

We use weak scaling. Height is fixed at 1000 (so the duration of the
run remains constant) while width varies with the number of
processors.

### 2020-02-03 pldi20_reviews/rebuttal.md
Additional citations.

@@
 
-TPS is reported in citations \[6, 28, 32].
+TPS is reported in citations \[6, 28, 32]. See also \[A-C] below.
 
@@
 processors.
+
+## Additional Citations
+
+\[A]: Mashayekhi, Omid, Hang Qu, Chinmayee Shah, and Philip Levis.
+"Execution Templates: Caching Control Plane Decisions for Strong Scaling of Data Analytics." In 2017 USENIX Annual Technical Conference (USENIX ATC 17), pp.
+513-526.
+2017.
+
+\[B]: Moritz, Philipp, Robert Nishihara, Stephanie Wang, Alexey Tumanov, Richard Liaw, Eric Liang, Melih Elibol et al.
+"Ray: A Distributed Framework for Emerging AI Applications." In 13th USENIX Symposium on Operating Systems Design and Implementation (OSDI 18), pp.
+561-577.
+2018.
+
+\[C]: Sadooghi, Iman, Geet Kumar, Ke Wang, Dongfang Zhao, Tonglin Li, and Ioan Raicu.
+"Albatross: An Efficient Cloud-Enabled Task Scheduling and Execution Framework Using Distributed Message Queues." In 2016 IEEE 12th International Conference on e-Science (e-Science), pp.
+11-20.
+IEEE, 2016.

### 2020-02-03 pldi20_reviews/rebuttal.md
Push OpenMP code.

@@
 
+MPI p2p is the implementation shown in Listing 2.
+
+## OpenMP Code Excerpt
+
+```c++
+void issue_task(Graph g, long t, long p,
+                  char *output,
+                  char **inputs, long num_inputs,
+                  char *scratch) {
+  switch (num_inputs) {
+  case 1:
+    {
+      #pragma omp task depend(inout: output) depend(in: inputs[0]) depend(inout: scratch)
+      task_body1(g, t, p, output, inputs[0], scratch);
+    }
+    break;
+  case 2:
+    {
+      #pragma omp task depend(inout: output) depend(in: inputs[0]) depend(in: inputs[1]) depend(inout: scratch)
+      task_body2(g, t, p, output, inputs[0], inputs[1], scratch);
+    }
+    break;
+  // ...
+  }
+}
+
+void execute_task_graph(Graph g) {
+  char **outputs = (char **)malloc(g.width * sizeof(char *));
+  char **scratch = (char **)malloc(g.width * sizeof(char *));
+  char **inputs = (char **)malloc(g.width * sizeof(char *));
+  char **input_ptrs = (char **)malloc(/* ... */);
+  // initialize data structures...
+
+  for (long t = 0; t < g.height; ++t) {
+    for (long p = 0; p < g.width; ++p) {
+      if (g.contains_point(t, p)) {
+        long idx = 0;
+        for (long dep : g.deps(t, p)) {
+          input_ptrs[idx] = inputs[dep];
+          idx++;
+        }
+
+        issue_task(g, t, p, outputs[p], input_ptrs, idx, scratch[p]);
+      }
+    }
+    std::swap(inputs, outputs);
+  }
+}
+```
+
 ## Additional Citations

### 2020-02-03 pldi20_reviews/rebuttal.md
Fiddle with OpenMP implementation.

--- 2020-02-03 pldi20_reviews/rebuttal.md
@@
 
+MPI p2p is the implementation shown in Listing 2.
+
+## OpenMP Code Excerpt
+
+```c++
+void issue_task(Graph g, long t, long p,
+                char *output,
+                char **inputs, long num_inputs,
+                char *scratch) {
+  switch (num_inputs) {
+  case 1:
+    {
+      #pragma omp task depend(inout: output) depend(in: inputs[0]) depend(inout: scratch)
+      g.execute_point(t, p, output, inputs, scratch);
+    }
+    break;
+  case 2:
+    {
+      #pragma omp task depend(inout: output) depend(in: inputs[0]) depend(in: inputs[1]) depend(inout: scratch)
+      g.execute_point(t, p, output, inputs, scratch);
+    }
+    break;
+  // ...
+  }
+}
+
+void execute_task_graph(Graph g) {
+  char **outputs = (char **)malloc(g.width * sizeof(char *));
+  char **scratch = (char **)malloc(g.width * sizeof(char *));
+  char **inputs = (char **)malloc(g.width * sizeof(char *));
+  char ***input_ptrs = (char ***)malloc(/* ... */);
+  // initialize data structures...
+
+  for (long t = 0; t < g.height; ++t) {
+    for (long p = 0; p < g.width; ++p) {
+      if (g.contains_point(t, p)) {
+        long idx = 0;
+        for (long dep : g.deps(t, p)) {
+          input_ptrs[p][idx] = inputs[dep];
+          idx++;
+        }
+
+        issue_task(g, t, p, outputs[p], input_ptrs[p], idx, scratch[p]);
+      }
+    }
+    std::swap(inputs, outputs);
+  }
+}
+```
+
 ## Additional Citations

### 2020-02-03 pldi20_reviews/rebuttal.md
More work on rebuttal.

@@
 
+## Analysis of Performance Bugs Found
+
+Realm and Chapel use DMA subsystems that are optimized for large
+copies. Early Task Bench experiments revealed high METG values that
+were diagnosed by the Realm/Chapel developers as overhead due to the
+cost of scheduling small copies. Subsequent improvements in Realm and
+Chapel improved small copy overheads (and thus METG) by over an order
+of magnitude in the case of Realm, and by 2x in the case of
+Chapel. These improvements affect any application where fine-grained
+tasks are used.
+
+PaRSEC uses a task pruning algorithm to improve scalability at large
+node counts. Initial Task Bench results achieved less than the
+expected scalability: METG was rising too quickly with node
+count. This was diagnosed by the PaRSEC developers as a bug in task
+pruning, and subsequently fixed. We also implemented a version of the
+PaRSEC code, `shard`, which uses manual pruning to demonstrate that
+additional gains may be possible.
+
+Early Task Bench results for Dask revealed that the cost of scheduling
+a task was $\mathcal{O}(N)$ where $N$ is the number of tasks in a task
+graph, causing overall cost for a task graph with $N$ nodes to be
+$\mathcal{O}(N^2)$. This issue was reported to and confirmed by the
+Dask developers. As a work around, our result in the paper use a
+lower-level interface which does not suffer from this asymptotic
+slowdown.
+
+Initial experiments with TensorFlow revealed that the TensorFlow Task
+Bench implementation was not able to use all cores on a node (thus
+making it impossible to measure METG since the code did not pass the
+efficiency threshold). This was diagnosed by the TensorFlow developers
+as an issue in constant folding. The entire Task Bench task graph was
+being detected as constant, and the constant-folding pass runs on one
+core. As a workaround, the developers suggested marking the task graph
+inputs as non-constant so that TensorFlow's normal task scheduler
+could be used.
+
 ## OpenMP Code Excerpt

### 2020-02-04 pldi20_reviews/rebuttal.md
Wrap.
Edit rebuttal.

@@
-We thank the reviewers for their insightful feedback.
+We thank the reviewers for their feedback.
+
+Responding to reviewer A on novelty, we provide a quantification (via
+METG) of typical programming system overheads. Contrary to the
+suggestion of reviewer A, we found that the METGs we found in practice
+did not always match our intuitions. In fact, some were off to such a
+large degree (multiple orders of magnitude) that they prompted the
+development teams to conduct major new development to address our
+results (more on this below).
+
+METG itself is also a contribution. Without METG, it is difficult to
+quantify the programming system's contribution to application
+overheads. Papers that look at system overheads typically report
+measurements in tasks per second, but either use empty tasks or do not
+report the level of efficiency achieved. (See below for citations.) In
+either case it is impossible to determine how much useful work is
+being performed. METG constrains efficiency so that a fixed amount of
+work is being performed relative to overhead.
+
+METG is parameterized on the efficiency metric. This is necessary, as
+it would not make sense to measure FLOPS on a memory bound benchmark,
+or vice versa. However, in spite of this, we find that the METG values
+tend to fall into a certain characteristic range which is consistent
+across efficiency measures. For example, compare Figures 6 and 8. The
+systems in each graph occupy roughly the same parts of each graph, and
+in most cases the relative ordering of systems is the same (with a
+small number of close ties flipped between one and the other). This
+holds when plotting Figure 8 on axes of efficiency vs task granularity
+(which we will include in the final paper).
+
+METG is also specific to the machine configuration: it can vary with
+network, CPU, etc. However it is again our experience that METG for a
+given programming system tends to span a range of typical values, and
+that these can guide both application and system development. Also,
+more precise measurements can identify specific bugs as noted below.
+
+Beyond this, it would simply be impossible to conduct any quantitative
+evaluation of a large number of systems without Task Bench. This is a
+major contribution. Task Bench permits $\mathcal{O}(N+M)$ rather than
+$\mathcal{O}(NM)$ implementation effort. In looking at the related
+work, we find that the only benchmarks with a comparable number of
+implementations are substantially more limited than Task Bench.
 
@@
 
-Regarding novelty, reviewer C is correct that a major contribution of
-our paper is a thorough empirical evaluation, but that this is made
-possible by substantial contributions in other areas. Namely, the
-design of task bench which permits $\mathcal{O}(N+M)$ rather than
-$\mathcal{O}(NM)$ implementation effort, and the definition of METG,
-are both critical. Without the former, a comparison of 15 systems
-would be intractable without substantially simplifying the
-benchmark. Without the latter, the results would be prone to bias, and
-even more difficult to interpret.
-
-Contrary to the suggestion of reviewer A, we found that the METGs we
-found in practice did not always match our intuitions. In fact, some
-were off to such a large degree (multiple orders of magnitude) that
-they prompted the development teams to conduct major new development
-to address our results. These cases are included in our analysis
-below.
-
-It is true that METG is a configuration-specific metric. That is, it
-can vary depending on the network, CPU, etc. However it is our
-experience that METG for a given programming system tends to span a
-range of typical values, and that these can guide both application and
-system development. Also, more precise measurements can identify
-specific bugs as noted above.
-
 Reviewer B asks about the impact of the network. It is our experience
@@
 these systems can be optimized to reduce unnecessary
-overheads. However, we agree that a comparison across network
-architectures would be a good direction for future work.
+overheads. However, we agree that comparing networks is a good
+direction for future work.
 
@@
 
-Fitting a Task Bench configuration to a specific application is a
-direction for future work, but the patterns in Figure 1 are all
-derived from common scientific computing workloads (and this is an
-extensible set).
+Fitting Task Bench configurations to specific applications is a
+direction for future work, but the patterns in Figure 1 are derived
+from common scientific workloads (and this is an extensible set).
 
 Irregular parallelism is enabled by allowing dependencies to vary over
-time. For example, the full Task Bench implementation includes a
-random dependence pattern.
+time. For example, Task Bench includes a randomized dependence
+pattern.
 
-The OpenMP implementation uses a switch to execute `#pragma omp task`
-statements with varying numbers of `depends` clauses. An excerpt is
-included below. Note all our implementations are open source and a
-link will be included in the final paper.
+An excerpt from the OpenMP implementation is included below. Note all
+our implementations are open source and a link will be included in the
+final paper.
 
@@
 
-Problem size corresponds to the `iterations` parameter to Listing
-1. Note that since these are idealized kernels, what matters is the
-total number of FLOPs executed, which directly corresponds to
-`iterations`.
+Problem size is the `iterations` parameter in Listing 1. Note that
+since these are idealized kernels, what matters is the total number of
+FLOPs executed, which directly corresponds to `iterations`.
 
@@
 
-MPI p2p is the implementation shown in Listing 2.
+MPI p2p is shown in Listing 2.
 
@@
 tasks are used.
+
+Further analysis of Realm efficiency indicated many overheads due to
+dynamic graph construction. The Realm developers implemented a new
+"subgraph" API in response to this feedback to amortize the cost of
+repeatedly constructing isomorphic task graphs. This API is used in
+Task Bench to achieve further speedups, up to nearly an order of
+magnitude.
+
+Analysis of the schedulers in Chapel and X10 revealed that certain
+poor scheduling decisions could result in poor utilization of the CPUs
+on a node. This resulted in poor peak performance, even at large task
+granularities, which in some cases made it impossible to measure METG
+(because peak efficiency did not exceed 50%). We reported these issues
+to the systems' developers and used fixes or workarounds in our
+experiments.
 
@@
 Bench implementation was not able to use all cores on a node (thus
-making it impossible to measure METG since the code did not pass the
-efficiency threshold). This was diagnosed by the TensorFlow developers
-as an issue in constant folding. The entire Task Bench task graph was
-being detected as constant, and the constant-folding pass runs on one
-core. As a workaround, the developers suggested marking the task graph
-inputs as non-constant so that TensorFlow's normal task scheduler
-could be used.
+making it impossible to measure METG since the efficiency was below
+50%). This was diagnosed by the TensorFlow developers as an issue in
+constant folding. The entire Task Bench task graph was being detected
+as constant, and the constant-folding pass runs on one core. As a
+workaround, the developers suggested marking the task graph inputs as
+non-constant so that TensorFlow's normal task scheduler could be used.
 
@@
 
-\[A]: Mashayekhi, Omid, Hang Qu, Chinmayee Shah, and Philip Levis.
-"Execution Templates: Caching Control Plane Decisions for Strong Scaling of Data Analytics." In 2017 USENIX Annual Technical Conference (USENIX ATC 17), pp.
-513-526.
-2017.
+\[A]: Mashayekhi, Omid, Hang Qu, Chinmayee Shah, and Philip
+Levis. "Execution Templates: Caching Control Plane Decisions for
+Strong Scaling of Data Analytics." In 2017 USENIX Annual Technical
+Conference (USENIX ATC 17), pp. 513-526. 2017.
 
-\[B]: Moritz, Philipp, Robert Nishihara, Stephanie Wang, Alexey Tumanov, Richard Liaw, Eric Liang, Melih Elibol et al.
-"Ray: A Distributed Framework for Emerging AI Applications." In 13th USENIX Symposium on Operating Systems Design and Implementation (OSDI 18), pp.
-561-577.
-2018.
+\[B]: Moritz, Philipp, Robert Nishihara, Stephanie Wang, Alexey
+Tumanov, Richard Liaw, Eric Liang, Melih Elibol et al. "Ray: A
+Distributed Framework for Emerging AI Applications." In 13th USENIX
+Symposium on Operating Systems Design and Implementation (OSDI 18),
+pp. 561-577. 2018.
 
-\[C]: Sadooghi, Iman, Geet Kumar, Ke Wang, Dongfang Zhao, Tonglin Li, and Ioan Raicu.
-"Albatross: An Efficient Cloud-Enabled Task Scheduling and Execution Framework Using Distributed Message Queues." In 2016 IEEE 12th International Conference on e-Science (e-Science), pp.
-11-20.
-IEEE, 2016.
+\[C]: Sadooghi, Iman, Geet Kumar, Ke Wang, Dongfang Zhao, Tonglin Li,
+and Ioan Raicu. "Albatross: An Efficient Cloud-Enabled Task Scheduling
+and Execution Framework Using Distributed Message Queues." In 2016
+IEEE 12th International Conference on e-Science (e-Science),
+pp. 11-20. IEEE, 2016.

### 2020-02-04 pldi20_reviews/rebuttal.md
Apply feedback from Pat.

@@
 report the level of efficiency achieved. (See below for citations.) In
-either case it is impossible to determine how much useful work is
+either case, it is impossible to determine how much useful work is
 being performed. METG constrains efficiency so that a fixed amount of
@@
 it would not make sense to measure FLOPS on a memory bound benchmark,
-or vice versa. However, in spite of this, we find that the METG values
-tend to fall into a certain characteristic range which is consistent
-across efficiency measures. For example, compare Figures 6 and 8. The
-systems in each graph occupy roughly the same parts of each graph, and
-in most cases the relative ordering of systems is the same (with a
-small number of close ties flipped between one and the other). This
-holds when plotting Figure 8 on axes of efficiency vs task granularity
-(which we will include in the final paper).
+or vice versa. However, in spite of this, we find that the METGs for a
+given system tend to fall into a certain characteristic range, which
+is consistent across efficiency measures. For example, compare Figures
+6 and 8. The systems in each graph occupy roughly the same regions of
+each graph, and in most cases the relative ordering of systems is the
+same (with a small number of close ties flipped between one and the
+other). This holds when plotting Figure 8 on axes of efficiency vs
+task granularity (which we will include in the final paper).
 
 METG is also specific to the machine configuration: it can vary with
-network, CPU, etc. However it is again our experience that METG for a
+network, CPU, etc. However, it is again our experience that METG for a
 given programming system tends to span a range of typical values, and
@@
 An excerpt from the OpenMP implementation is included below. Note all
-our implementations are open source and a link will be included in the
+our implementations are open source, and a link will be included in the
 final paper.
@@
 node counts. Initial Task Bench results achieved less than the
-expected scalability: METG was rising too quickly with node
-count. This was diagnosed by the PaRSEC developers as a bug in task
-pruning, and subsequently fixed. We also implemented a version of the
-PaRSEC code, `shard`, which uses manual pruning to demonstrate that
+expected scalability: METG was rising too quickly with node count.
+The PaRSEC developers diagnosed this as a bug in task pruning, and it
+was subsequently fixed. We also implemented a version of the PaRSEC
+code, `shard`, which uses manual pruning to demonstrate that
 additional gains may be possible.
@@
 $\mathcal{O}(N^2)$. This issue was reported to and confirmed by the
-Dask developers. As a work around, our result in the paper use a
+Dask developers. As a work around, our results in the paper use a
 lower-level interface which does not suffer from this asymptotic
@@
 making it impossible to measure METG since the efficiency was below
-50%). This was diagnosed by the TensorFlow developers as an issue in
+50%). The TensorFlow developers diagnosed this as an issue in
 constant folding. The entire Task Bench task graph was being detected
@@
 workaround, the developers suggested marking the task graph inputs as
-non-constant so that TensorFlow's normal task scheduler could be used.
+non-constant so that TensorFlow's standard task scheduler could be used.
 

### 2020-02-04 pldi20_reviews/rebuttal.md
Apply Sean's edits.

@@
 METG) of typical programming system overheads. Contrary to the
-suggestion of reviewer A, we found that the METGs we found in practice
-did not always match our intuitions. In fact, some were off to such a
-large degree (multiple orders of magnitude) that they prompted the
-development teams to conduct major new development to address our
-results (more on this below).
+suggestion of reviewer A, we found that METGs we found in practice did
+not always match intuition. In fact, the actual overheads of some
+systems differed from what the developers of those systems expected to
+such a large degree (multiple orders of magnitude) that it prompted
+substantial new development to address our results (more on this
+below).
 
@@
 each graph, and in most cases the relative ordering of systems is the
-same (with a small number of close ties flipped between one and the
-other). This holds when plotting Figure 8 on axes of efficiency vs
-task granularity (which we will include in the final paper).
+same (with a small number of close ties flipped between the two). This
+holds when plotting Figure 8 on axes of efficiency vs task granularity
+(which we will include in the final paper).
 

### 2020-02-05 pldi20_reviews/rebuttal.md
Apply Alex's edits.

--- 2020-02-04 pldi20_reviews/rebuttal.md
@@
 
-Responding to reviewer A on novelty, we provide a quantification (via
-METG) of typical programming system overheads. Contrary to the
-suggestion of reviewer A, we found that the METGs we found in practice
-did not always match our intuitions. In fact, some were off to such a
-large degree (multiple orders of magnitude) that they prompted the
-development teams to conduct major new development to address our
-results (more on this below).
+Responding to reviewer A's concerns about novelty, we provide a way to
+directly quantify (through METG) runtime overhead. The METGs we found
+in practice did not always match intuition and were sometimes
+surprising. Some system overheads differed so much from what the
+developers of those systems expected (multiple orders of magnitude)
+that it prompted substantial new development to address our results
+(more on this below).
 
-METG itself is also a contribution. Without METG, it is difficult to
-quantify the programming system's contribution to application
-overheads. Papers that look at system overheads typically report
-measurements in tasks per second, but either use empty tasks or do not
-report the level of efficiency achieved. (See below for citations.) In
-either case, it is impossible to determine how much useful work is
-being performed. METG constrains efficiency so that a fixed amount of
-work is being performed relative to overhead.
+Without METG, it is difficult to quantify the programming system's
+contribution to application overheads. Papers that look at system
+overheads typically report measurements in tasks per second, but
+either use empty tasks or do not report the level of efficiency
+achieved. (See below for citations.) In either case, it is impossible
+to determine how much useful work is being performed. METG constrains
+efficiency so that a fixed amount of work is performed relative to
+overhead.
 
-METG is parameterized on the efficiency metric. This is necessary, as
+It is necessary that METG is parameterized on the efficiency metric;
 it would not make sense to measure FLOPS on a memory bound benchmark,
-or vice versa. However, in spite of this, we find that the METGs for a
-given system tend to fall into a certain characteristic range, which
-is consistent across efficiency measures. For example, compare Figures
-6 and 8. The systems in each graph occupy roughly the same regions of
-each graph, and in most cases the relative ordering of systems is the
-same (with a small number of close ties flipped between one and the
-other). This holds when plotting Figure 8 on axes of efficiency vs
-task granularity (which we will include in the final paper).
+or vice versa. However, we also find that the METGs for a given system
+tend to fall in a characteristic range, which is consistent across
+efficiency measures. For example, compare Figures 6 and 8. The systems
+in each graph occupy roughly the same regions of each graph, and in
+most cases the relative ordering of systems is the same (with a small
+number of near ties reversed). This holds when plotting Figure 8 on
+axes of efficiency vs task granularity (which we will include in the
+final paper).
 
@@
 
-Beyond this, it would simply be impossible to conduct any quantitative
-evaluation of a large number of systems without Task Bench. This is a
-major contribution. Task Bench permits $\mathcal{O}(N+M)$ rather than
-$\mathcal{O}(NM)$ implementation effort. In looking at the related
-work, we find that the only benchmarks with a comparable number of
-implementations are substantially more limited than Task Bench.
+Separately, it would simply be impossible to conduct any quantitative
+evaluation of a large number of systems without Task Bench.  Task
+Bench permits $\mathcal{O}(N+M)$ rather than $\mathcal{O}(NM)$
+implementation effort; so far as we know, we are the first to
+recognize how to factor a benchmark framework to achieve such a
+productivity gain.  Compared with related work, the only benchmarks
+with a similar number of implementations are substantially more
+limited than Task Bench.
 
-Reviewers A and D note that the bugs described in lines 197-199
-currently do not have a larger analysis. We include a draft of this
-analysis below and will incorporate it into the final paper.
+Reviewers A and D asked about a depper analysis of the bugs described
+in lines 197-199.  We include a draft of this analysis below and will
+incorporate it into the final paper.
 
@@
 An excerpt from the OpenMP implementation is included below. Note all
-our implementations are open source, and a link will be included in the
-final paper.
+our implementations are open source, and a link will be included in
+the final paper.
 

### 2020-02-05 pldi20_reviews/rebuttal.md
Fix a typo.
Another small tweak.
Update analysis text.

@@
 
-Reviewers A and D asked about a depper analysis of the bugs described
-in lines 197-199.  We include a draft of this analysis below and will
+Reviewers A and D ask about a deeper analysis of the bugs described in
+lines 197-199. We include a draft of this analysis below and will
 incorporate it into the final paper.
@@
 
-Analysis of the schedulers in Chapel and X10 revealed that certain
-poor scheduling decisions could result in poor utilization of the CPUs
-on a node. This resulted in poor peak performance, even at large task
-granularities, which in some cases made it impossible to measure METG
-(because peak efficiency did not exceed 50%). We reported these issues
-to the systems' developers and used fixes or workarounds in our
-experiments.
+Analysis of the schedulers in Chapel and X10 revealed poor CPU
+utilization. Chapel uses a naive round-robin scheduler by default,
+which can lead to unexpected load balance issues because certain
+statements in the language (such as remote array assignment)
+implicitly generate tasks. In X10, `async at(p)` and `at(p) async` are
+not equivalent; the former generates a local task before place
+switching while the latter generates the task remotely. The presence
+of additional local tasks could interfere with scheduling and result
+in idle CPUs. In both cases this resulted in poor peak performance,
+even at large task granularities, which in some cases made it
+impossible to measure METG (because peak efficiency did not exceed
+50%). We reported these issues to the systems' developers and used
+fixes or workarounds in our experiments.
 

### 2020-03-23 5_evaluation.tex
Updates for new TensorFlow results.

@@
 
-Spark, Swift/T and TensorFlow
+Spark and Swift/T
 are omitted from comparisons with more complicated dependencies, as the overheads of these systems require

### 2020-03-23 3_implementation.tex
Rewrite implementation to save space.

@@
 programming systems listed in Table~\ref{tab:systems}. These include
-traditional HPC programming models: MPI~\cite{MPI} and MPI+X (OpenMP
-and CUDA); PGAS and actor models: Chapel~\cite{Chapel15},
-Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and task-based
-systems: OmpSs~\cite{OmpSs11}, OpenMP 4.0~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13, PARSEC_DTD}, Realm~\cite{Realm14},
-Regent~\cite{Regent15}, and StarPU~\cite{StarPU11}. Also, we consider
-several systems used in large scale data analytics and workflows:
-Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
-and TensorFlow~\cite{TensorFlow15}. We describe the systems and implementations below.
+traditional HPC programming models (MPI and MPI+X), PGAS and actor
+models (Chapel, Charm++ and X10), task-based systems (OmpSs, OpenMP
+4.0, PaRSEC, Realm, Regent, and StarPU) and systems for large scale
+data analytics, machine learning and workflows (Dask, Spark, Swift/T,
+and TensorFlow). We briefly describe the systems and implementations
+below.
 
@@
 
-\subsection{Chapel}
+\subsection{Traditional HPC Programming Models}
 
-Chapel~\cite{Chapel15} is a parallel programming language
-with a \emph{multi-res\-o\-lu\-tion} approach, supporting parallelism at a variety of
-levels. Chapel's core features are a partitioned global address space
-(PGAS), data distributions, tasks,
-synchronization primitives, and array promotion. For Task Bench, we target a low level of
-Chapel, using explicit task instantiation (via
-\lstinline[language=Chapel]{coforall}), bulk access to distributed
-arrays for efficient data movement, and atomic integers for
-synchronization.
-
-\subsection{Charm++}
-
-Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors
-(\emph{chares}) do not share memory and communicate via messages. Our Task Bench implementation uses a chare
-array for the task graph, with one chare for each column, and uses messages to implement dependencies.
-
-%% Collectives synchronize the
-%% completion of the task graph execution.
-
-% FIXME: Discuss SMP vs non-SMP versions
-
-\subsection{Dask}
-
-Dask~\cite{Dask15} is an implicitly parallel task-based programming system for large
-scale data analytics with Python. Dask provides abstractions over
-distributed NumPy arrays and Pandas dataframes, and also a lower-level
-interface for launching tasks. We implement Task Bench in the lower-level interface.
-
-\subsection{MPI and MPI+X}
-
-We provide two MPI~\cite{MPI} implementations, written in the style of
-communicating sequential processes (CSP) and bulk synchronous
-programming (BSP). An excerpt from the CSP
-implementation is shown in Listing~\ref{lst:code-sample}. The BSP
-implementation uses \lstinline[language=C++]{MPI_Barrier} to enforce
-boundaries between phases.
+MPI~\cite{MPI} is a message-passing API for HPC. We provide an MPI
+implementation written in the style of communicating sequential
+processes (CSP). An excerpt of this implementation is shown in
+Listing~\ref{lst:code-sample}.
 
@@
 
-\subsection{OpenMP and OmpSs}
+\subsection{PGAS and Actor Models}
 
-OpenMP~\cite{OpenMPSpec40} and OmpSs~\cite{OmpSs11} are programming
-models for loop- and task-based parallelism. (OmpSs is
-source-compatible with OpenMP.) Our implementations use OpenMP 4.0
-tasking.  Because tasks have a fixed number of dependencies,
-we use a switch statement to implement dynamic dependencies for Task
-Bench. For OpenMP, we tested GNU GOMP and Intel KMP and found KMP to be better
-performing.
+PGAS and actor models, such as Chapel~\cite{Chapel15},
+Charm++~\cite{Charmpp93} and X10~\cite{X1005}, provide support for
+asynchronous tasks that make them amenable to a straightforward
+implementation of task bench. Synchronization is explicit and may be
+provided by messages (in actor models) or other primitives, such as
+locks or atomics (in PGAS models). PGAS models such as Chapel and X10
+provide global references to data anywhere in the machine, but vary in
+whether data can be accessed remotely or not (Chapel allows this, X10
+does not).
 
-% FIXME: More to say about OmpSs? Is there anything different vs the OpenMP implementation?
+Chapel also provides support for implicit parallelism which we do not
+evaluate in this paper.
 
-%\subsection{OpenMP}
+\subsection{Task-Based Programming Models}
 
-%OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
-%parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
-%for OpenMP is similar to OmpSs and uses
-%task dependencies. We tested with GNU GOMP and Intel KMP
-%and found KMP to be better performing.
+Task-based systems include OmpSs~\cite{OmpSs11}, OpenMP
+4.0~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13, PARSEC_DTD},
+Realm~\cite{Realm14}, Regent~\cite{Regent15}, and
+StarPU~\cite{StarPU11}. Though the details vary, these systems
+typically provide implicit parallelism, where tasks are enumerated
+sequentially (in program order) and the dependencies between tasks are
+analyzed automatically to construct a \emph{dependence graph} that
+guides the execution of tasks. PaRSEC provides two modes,
+\emph{dynamic task discovery} (DTD)~\cite{PARSEC_DTD}, which operates
+as above, and \emph{parameterized task graphs} (PTG)~\cite{PARSEC13},
+where the dependence graph is constructed automatically from an
+analytical representation of the task graph. Realm (unlike the others
+above) is explicitly parallel and requires tasks to be connected
+explicitly via \emph{events}.
 
-\subsection{PaRSEC}
+Among the distributed task-based systems, one question is how to
+resolve the inherent scalability bottleneck posed by enumerating tasks
+sequentially. PaRSEC and StarPU resolve this by permitting the user to
+omit tasks that are not directly connected in the dependence
+graph. The program is replicated (for efficiency of the analysis), but
+each task is executed only once. Because the dynamic checks to see if
+a task should be executed are not free of cost, we also provide a
+PaRSEC implementation (under the label \lstinline{shard}) which has
+been hand-written to avoid all such checks. Regent performs an
+equivalent optimization at compile time~\cite{ControlReplication17}
+that does not require user intervention.
 
-PaRSEC supports two distinct
-programming models: \emph{parameterized task graphs}
-(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
-(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which users
-write a concise, algebraic description of the tasks and dataflows in
-the program. This compressed representation is expanded into a full
-task graph by a source-to-source compiler.
+\subsection{Data Analytics, Machine Learning and Workflows}
 
-In DTD, tasks are enumerated in program order and dependencies are
-identified automatically based on task arguments. To improve
-scalability, the program is executed in SPMD fashion on all nodes, and
-the user is responsible for eliding tasks that are not direct
-predecessors or successors to those executed on the current node.
+Dask~\cite{Dask15}, Spark~\cite{Spark10} and
+TensorFlow~\cite{TensorFlow15} are programming models for large scale
+data analytics and machine learning. Dask and TensorFlow provide
+domain-specific abstractions built on top of task-based runtimes. Our
+implementations directly create tasks and are similar to other
+task-based systems above.
 
-We implement Task Bench with both PTG and DTD, as well as a third
-option based on DTD that has additional manual optimizations applied
-to further reduce the cost of dynamic checks for task elision.
+Spark provides support for functional operators that implicitly map to
+tasks. We use \lstinline[language=Scala]{flatMap} and
+\lstinline[language=Scala]{groupByKey} to generate dependencies, and
+then \lstinline[language=Scala]{mapPartitions} to execute a series of
+tasks. An explicit hash partitioner ensures the correct task
+granularity.
 
-%% Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
-%% distributed across nodes in pre-defined patterns,
-%% and can be passed to tasks on other nodes, but
-%% can only be written on the node where the data is originally
-%% allocated. In our Task Bench implementation, each task writes to a
-%% tile as output, and these tiles are passed as inputs to the
-%% subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
-%% the task launch code uses a switch statement as the PaRSEC API assumes
-%% tasks have a fixed number of dependencies.
-
-\subsection{Realm}
-
-Realm~\cite{Realm14} is an explicitly-parallel task-based programming
-model used internally by Regent~\cite{Regent15}. Realm is thus a limit
-study of what can be achieved with Regent.
-
-Tasks in Realm are asynchronous, with dependencies
-defined by \emph{events} passed from one task
-to another. Realm's data model supports collections that live in a
-specific memory. Data must be explicitly copied, and the copies
-synchronized with tasks via events.
-Realm also supports a \emph{subgraph} API that enables additional
-optimizations. We use subgraphs in the Task
-Bench implementation.
-
-\subsection{Regent}
-
-Regent~\cite{Regent15} is an implicitly-parallel task-based language
-which implements the Legion programming model~\cite{Legion12}. We use
-Regent rather than Legion since the Regent compiler
-provides a critical optimization to
-improve scalability~\cite{ControlReplication17}.
-
-\subsection{Spark}
-
-Spark~\cite{Spark10} is an implicitly-parallel programming model for
-data analytics.
-The core abstractions in Spark are functional operators such as map,
-reduce, and join. Data in Spark are stored in
-\emph{resilient
-  distributed datasets} (RDDs): globally-visible,
-dynamically single-assignment data structures. Spark caches RDDs in memory to avoid unnecessary
-disk traffic.
-
-Spark has tasks internally but they are not exposed to the user,
-so a Task Bench implementation must map the task graph to operators that result in the
-desired execution. We use
-\lstinline[language=Scala]{flatMap} and
-\lstinline[language=Scala]{groupByKey} to generate
-dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
-series of tasks. An explicit hash partitioner ensures that
-Spark does not attempt to group multiple Task Bench tasks into a
-single Spark task, as tasks in Task Bench already represent coarse-grained units of work.
-
-We performed extensive experiments to verify that no
-extraneous factors interfered with our Spark measurements. We
-disabled logging, ensured no RDDs are written
-to disk, confirmed there is no measurable overhead due to JNI
-calls from Java to C or due to serializing data, and cross-checked results with known cases
-with optimal scheduler throughput.
-
-\subsection{StarPU}
-
-StarPU~\cite{StarPU11} is a task-based system utilizing a \emph{sequential task flow}
-programming model similar to PaRSEC's DTD. Our Task Bench implementation in
-StarPU is very similar to PaRSEC.
-%% We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
-%% project~\cite{Chameleon}.
-
-\subsection{Swift/T}
-
-Swift/T~\cite{Wozniak13} is a parallel scripting language intended
-primarily for the composition of HPC workflows. Swift/T
-programs follow dataflow semantics, where every statement may
-potentially execute in parallel as soon as its dependencies are
-satisfied; statements only execute sequentially when explicitly
-requested. The Swift/T compiler performs a number of optimizations to
-improve performance of highly parallel programs~\cite{Armstrong14}. 
-Our Task Bench in Swift/T uses its dataflow
-semantics to capture dependencies on other tasks.
-
-\subsection{TensorFlow}
-
-TensorFlow~\cite{TensorFlow15} is designed for deep learning
-workloads. Although TensorFlow's API exposes machine learning
-concepts, internally TensorFlow is a task graph execution engine,
-making it a good fit for Task Bench. TensorFlow programs operate via
-explicit graph construction. Task graphs are composed in Python and
-run by a C++ execution engine.
-
-\subsection{X10}
-
-X10~\cite{X1005} is an explicitly parallel programming language for
-place-based programming. Features include \emph{places}
-which represent distributed memories, a PGAS model where references to
-remote objects can be held (but can only be dereferenced on the local
-place), asynchronous tasks, and a place-changing construct to move the
-execution of a task to a remote place. X10 also supports a variety of
-synchronization primitives.
-
-Our Task Bench implementation uses
-\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
-between places, along with place-changing and atomic integers for
-synchronization. We use the native backend of X10, which compiles to
-C++.
+Swift/T~\cite{Wozniak13, Armstrong14} is a parallel scripting language
+with dataflow semantics, used primarily for workflow automation. Our
+implementation is straightforward.

### 2020-03-23 1_introduction.tex
Tweaks for spacing and wording.

@@
 scaling does not separate runtime system overhead from application costs
-(such as increased communication) that scale with the number of nodes when
+(such as communication) that scale with the number of nodes when
 using progressively larger portions of a machine.

### 2020-03-23 3_implementation.tex
Tweaks for spacing and wording.

@@
 \label{sec:implementation}
-
-\input{t1_systems}
 
@@
 implementation written in the style of communicating sequential
-processes (CSP). An excerpt of this implementation is shown in
-Listing~\ref{lst:code-sample}.
+processes (CSP). An excerpt is shown in Listing~\ref{lst:code-sample}.
 
@@
 PGAS and actor models, such as Chapel~\cite{Chapel15},
-Charm++~\cite{Charmpp93} and X10~\cite{X1005}, provide support for
-asynchronous tasks that make them amenable to a straightforward
+Charm++~\cite{Charmpp93} and X10~\cite{X1005} offer
+asynchronous tasks, making them amenable to a straightforward
 implementation of task bench. Synchronization is explicit and may be
@@
 whether data can be accessed remotely or not (Chapel allows this, X10
-does not).
+does not). Chapel also provides support for implicit parallelism which
+we do not evaluate in this paper.
 
-Chapel also provides support for implicit parallelism which we do not
-evaluate in this paper.
+\input{t1_systems}
 
@@
 
-Among the distributed task-based systems, one question is how to
-resolve the inherent scalability bottleneck posed by enumerating tasks
-sequentially. PaRSEC and StarPU resolve this by permitting the user to
-omit tasks that are not directly connected in the dependence
-graph. The program is replicated (for efficiency of the analysis), but
-each task is executed only once. Because the dynamic checks to see if
-a task should be executed are not free of cost, we also provide a
-PaRSEC implementation (under the label \lstinline{shard}) which has
-been hand-written to avoid all such checks. Regent performs an
-equivalent optimization at compile time~\cite{ControlReplication17}
-that does not require user intervention.
+Among distributed task-based systems, there can be a scalability
+bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU
+resolve this by permitting the user to omit tasks that are not
+directly connected in the dependence graph. The program is replicated
+(for efficiency of the analysis), but each task is executed only
+once. Because the dynamic checks to see if a task should be executed
+are not free of cost, we also provide a PaRSEC implementation (under
+the label \lstinline{shard}) which has been hand-written to avoid all
+such checks. Regent performs an equivalent optimization at compile
+time~\cite{ControlReplication17} that does not require user
+intervention.
 
@@
 tasks. We use \lstinline[language=Scala]{flatMap} and
-\lstinline[language=Scala]{groupByKey} to generate dependencies, and
-then \lstinline[language=Scala]{mapPartitions} to execute a series of
-tasks. An explicit hash partitioner ensures the correct task
-granularity.
+\lstinline[language=Scala]{groupByKey} to generate dependencies and
+\lstinline[language=Scala]{mapPartitions} to execute tasks. An
+explicit hash partitioner ensures the correct task granularity.
 
-Swift/T~\cite{Wozniak13, Armstrong14} is a parallel scripting language
+Swift/T~\cite{Wozniak13} is a parallel scripting language
 with dataflow semantics, used primarily for workflow automation. Our

### 2020-03-23 5_evaluation.tex
Tweaks for spacing and wording.

@@
 
-To answer this question, we use the METG metric to determine the
-smallest task granularity that can be executed at a given level of
-efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
+To answer this question, we use METG %% to determine the
+%% smallest task granularity that can be executed at a given level of
+%% efficiency
+as a proxy for overhead. Figure~\ref{fig:metg-compute}
 shows how METG(50\%) varies with node count for a variety of
@@
 remembering when considering this metric that this is a \emph{minimum}
-effective task granularity. Therefore applications with an average
+effective task granularity. Applications with an average
 task granularity of \emph{at least} this value can usually be expected
@@
 
-One advantage of systems with asynchronous execution capabilities is
+One advantage of asynchronous execution is
 the ability to mitigate load imbalance with little or no additional programmer effort, especially in the presence of

### 2020-03-23 l2_mpi_sample.tex
Simplify code sample.
Tweaks for spacing and wording.

@@
-\begin{lstlisting}[language=C++,caption={Excerpt from Task Bench implementation in MPI.},label={lst:code-sample},style=codeblock,float]
+\begin{lstlisting}[language=C++,caption={Excerpt from Task Bench implementation in MPI.},label={lst:code-sample},style=codeblocktop,float]
 void execute_task_graph(Graph g) {
@@
 
-      g.execute_point(t, rank, output, g.output_bytes_per_task,
-        inputs, input_bytes, idx,
-        scratch, g.scratch_bytes_per_task);
+      g.execute_point(t, rank, output, inputs, scratch);
     }

### 2020-03-23 4_metg.tex
Some fiddling with text.

@@
 
-Given that Task Bench permits the rapid exploration of a large space
+Since Task Bench permits rapid exploration of a large space
 of application scenarios, one question is how to characterize the
-performance and efficiency of the systems under study. As noted
-earlier, the overheads of the systems we consider vary by more than
+performance and efficiency of systems under study. As noted
+above, the overheads of the systems we consider vary by more than
 five orders of magnitude, making it challenging to extract useful

### 2020-03-23 e3_metg_compute.tex
Tweak spacing around figure captions.
Update text and fiddle with wording and figures to make everything fit again.
Fiddle with figure placement, caption size, prune communication experiment.

@@
 }
-\subfloat[Nearest pattern, 5 dependencies per task.\label{fig:metg-compute-nearest}]{
+\subfloat[Nearest pattern, 5 deps/task.\label{fig:metg-compute-nearest}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_nearest.pdf}
@@
 
-\subfloat[Spread pattern, 5 dependencies per task.\label{fig:metg-compute-spread}]{
+\subfloat[Spread pattern, 5 deps/task.\label{fig:metg-compute-spread}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_spread.pdf}
 }
-\subfloat[Nearest pattern, 5 dependencies per task, 4 task graphs.\label{fig:metg-compute-4x-nearest}]{
+\subfloat[Nearest pattern, 5 deps/task, 4 graphs.\label{fig:metg-compute-4x-nearest}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_ngraphs_4_nearest.pdf}
@@
 
+\vspace{-0.15cm}
 \caption{METG vs node count for different dependence patterns. Lower is better.\label{fig:metg-compute}}
+%% \vspace{-0.05cm}
 \end{figure*}

### 2020-03-23 e6_communication.tex
Fiddle with figure placement, caption size, prune communication experiment.

@@
 }
-\subfloat[256 bytes per task dependency.\label{fig:efficiency-communication-256}]{
-\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_256.pdf}
-}
-
-\subfloat[4096 bytes per task dependency.\label{fig:efficiency-communication-4096}]{
-\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_4096.pdf}
-}
+%% \subfloat[256 bytes per task dependency.\label{fig:efficiency-communication-256}]{
+%% \includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_256.pdf}
+%% }
+%%
+%% \subfloat[4096 bytes per task dependency.\label{fig:efficiency-communication-4096}]{
+%% \includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_4096.pdf}
+%% }
 \subfloat[65536 bytes per task dependency.\label{fig:efficiency-communication-65536}]{
@@
 
-\vspace{-0.25cm}
-\caption{Efficiency vs task granularity for varying communication (spread pattern, 5 dependencies per task, 4 task graphs, 64 nodes). Higher is better.\label{fig:efficiency-communication}}
+\vspace{-0.15cm}
+\caption{Efficiency vs task granularity for varying communication (spread, 5 deps/task, 4 graphs, 64 nodes). Higher is better.\label{fig:efficiency-communication}}
 \vspace{-0.05cm}

### 2020-03-23 6_related.tex
short the realted work

@@
 
-Parallel and distributed programming systems are most
-commonly evaluated using proxy- or mini-apps, or
-microbenchmarks. Mini-apps are explicitly derived from larger
-applications and therefore have the advantage of bearing some
-relationship to the original. This advantage typically does not hold
+Parallel and distributed programming systems are usually
+evaluated using proxy-/mini-apps, or
+microbenchmarks. Mini-apps are explicitly derived from real
+applications and hence have the advantage of bearing some
+relationships to the original. This advantage typically does not hold
 for microbenchmarks.
@@
 \item
-The initial paper on PENNANT~\cite{PENNANT} includes an
-implementation that supports MPI/OpenMP/MPI+OpenMP; follow-up papers present an implementation in
-Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.
+%The initial paper on PENNANT~\cite{PENNANT} includes an
+%implementation supporting MPI/OpenMP/MPI+OpenMP; follow-up papers present an implementation in
+%Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.
+PENNANT has an initial implementation supporting MPI/OpenMP/MPI+OpenMP
+published in~\cite{PENNANT}, and another one in Regent published
+in~\cite{Regent15, ControlReplication17, LegionTracing18}.
 
@@
 apples-to-apples manner as the last implementation uses a structured
-rather than an unstructured mesh. Follow-up papers describe an
-additional implementation in Regent~\cite{Regent15,
+rather than an unstructured mesh. Follow-up papers describe another
+implementation in Regent~\cite{Regent15,
   ControlReplication17, LegionTracing18} (comparison against reference
@@
 
-Microbenchmarks can be easier to implement, but risk being
-unrepresentative of real applications. PRK Stencil~\cite{PRK14}
-contains 2 kernels (which combined fit in under 50 lines of C++)
+Microbenchmarks is easier to implement, but risk being
+unrepresentative of real applications. 
+PRK Stencil~\cite{PRK14}
+contains 2 kernels (that combined fit in under 50 lines of C++)
 and a halo exchange on a structured grid; an
-evaluation~\cite{PRKRuntimes16} compares MPI variants, SHMEM, UPC,
+evaluation~\cite{PRKRuntimes16} compares MPI, SHMEM, UPC,
 Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
@@
 MPI+OpenMP~\cite{NASMPIOpenMP00}, and
-Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieve the breadth of
+Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieves the breadth of
 coverage, flexibility, or ease of implementation of Task Bench. We
@@
 microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
-it is even easier to use Task Bench to generate new benchmarks.
+can be adopted to generate new benchmarks easily.
 

### 2020-03-23 8_case_study.tex

\section{Case Study}
\label{sec:case-study}

Task Bench is useful not only as a tool for evaluating programming
system performance, but also for discovering potential areas of
improvements. During the development of Task Bench, we identified a
number of performance issues in the underlying programming
systems. These discoveries were possible because of the flexibility of
Task Bench, and our ability to rapidly run new experiments with a
variety of application scenarios. All issues below have been reported
to and acknowledged by the respective system's developers, and are
either fixed or worked around in our experiments.

Realm and Chapel use DMA subsystems that are optimized for large
copies. Early Task Bench experiments revealed high METG values that
were diagnosed by the Realm/Chapel developers as overhead due to the
cost of scheduling small copies. Subsequent improvements in Realm and
Chapel improved small copy overheads (and thus METG) by over an order
of magnitude in the case of Realm, and by $2\times$ in the case of
Chapel. These improvements affect any application where fine-grained
tasks are used.

Further analysis of Realm efficiency indicated many overheads due to
dynamic graph construction. The Realm developers implemented a new
``subgraph'' API in response to this feedback to amortize the cost of
repeatedly constructing isomorphic task graphs. This API is used in
Task Bench to achieve further speedups in the Realm implementation.

Chapel uses a naive round-robin scheduler by default, which can lead
to unexpected load balance issues because certain language features
(such as remote array assignment) implicitly generate tasks. This
resulted in poor peak performance, even at large task granularities,
which in some cases made it impossible to measure METG (because peak
efficiency did not exceed 50\%). Chapel's other schedulers add
overhead, resulting in higher METGs. We worked around this with
\lstinline{serial}.

PaRSEC uses task pruning to improve scalability at large node
counts. Initial Task Bench results achieved less than the expected
scalability: METG was rising too quickly with node count.  The PaRSEC
developers diagnosed this as a bug in task pruning, and it was
subsequently fixed. We also implemented a version of the PaRSEC code,
\lstinline{shard}, which uses manual pruning to demonstrate that
additional gains may be possible.

Early Task Bench results for Dask revealed that the cost of scheduling
a task was $\mathcal{O}(N)$ where $N$ is the number of tasks in a task
graph, causing overall cost for a task graph with $N$ nodes to be
$\mathcal{O}(N^2)$. This issue was reported to and confirmed by the
Dask developers. As a work around, our results in the paper use a
lower-level interface which does not suffer from this asymptotic
slowdown.

Initial experiments with TensorFlow revealed that the TensorFlow Task
Bench implementation was not able to use all cores on a node (thus
making it impossible to measure METG since the efficiency was below
50\%). The TensorFlow developers diagnosed this as an issue in
constant folding. The entire Task Bench task graph was being detected
as constant, and the constant-folding pass runs on one core. As a
workaround, the developers suggested marking the task graph inputs as
non-constant so that TensorFlow's standard task scheduler could be
used.

### 2020-03-23 6_related.tex
Tweaks to avoid stray words.

--- 2019-11-22 6_related.tex
@@
 
-Parallel and distributed programming systems are most
-commonly evaluated using proxy- or mini-apps, or
+Parallel and distributed programming systems are often
+evaluated using proxy- or mini-apps, or
 microbenchmarks. Mini-apps are explicitly derived from larger
-applications and therefore have the advantage of bearing some
+applications and hence have the advantage of bearing some
 relationship to the original. This advantage typically does not hold
@@
 \item
-The initial paper on PENNANT~\cite{PENNANT} includes an
-implementation that supports MPI/OpenMP/MPI+OpenMP; follow-up papers present an implementation in
-Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.
+The PENNANT reference implementation supports
+MPI/OpenMP/MPI+OpenMP~\cite{PENNANT}. A follow-up paper presents a
+Regent implementation~\cite{Regent15}.
 
@@
 apples-to-apples manner as the last implementation uses a structured
-rather than an unstructured mesh. Follow-up papers describe an
-additional implementation in Regent~\cite{Regent15,
-  ControlReplication17, LegionTracing18} (comparison against reference
+rather than an unstructured mesh. A follow-up paper describes another
+implementation in Regent~\cite{Regent15} (comparison against reference
 only).
@@
 microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
-it is even easier to use Task Bench to generate new benchmarks.
+makes it easy to generate new benchmarks.
 

### 2020-04-07 3_implementation.tex
Add addition text and prune case study section.

@@
 data analytics, machine learning and workflows (Dask, Spark, Swift/T,
-and TensorFlow). We briefly describe the systems and implementations
-below.
+and TensorFlow). Implementing Task Bench in such a wide range of
+systems is possible because the separation between core API (in
+Table~\ref{tab:api}) and system implmentation enables an overall
+effort of $\mathcal{O}(m + n)$ (for $m$ benchmarks on $n$ systems)
+rather than $\mathcal{O}(mn)$ as has been the case for all previous
+benchmarks that we know of. We briefly describe the systems and
+implementations below.
 

### 2020-04-07 7_conclusion.tex
Small tweak.
Add addition text and prune case study section.

@@
 specification of a benchmark from implementations in various
-programming systems, it is possible to explore a broad space of
+programming systems, Task Bench reduces overall developer effort to
+$\mathcal{O}(m + n)$ (for $m$ benchmarks on $n$ systems) rather than
+$\mathcal{O}(mn)$ as has been the case for all previous
+benchmarks that we know of. This has enabled us to explore a broad space of
 application scenarios and to do so with a large number of programming
@@
 applications do not operate at the absolute extreme of runtime-limited
-performance, and thus may choose to trade overhead for better
+performance, and thus may choose to trade overhead for
 usability. Our study helps to quantify the performance side of that

### 2020-04-07 8_case_study.tex
Add addition text and prune case study section.

@@
 Task Bench, and our ability to rapidly run new experiments with a
-variety of application scenarios. All issues below have been reported
-to and acknowledged by the respective system's developers, and are
-either fixed or worked around in our experiments.
+variety of application scenarios. All issues were reported to and
+acknowledged by the respective system's developers, and are either
+fixed or worked around in our experiments. For brevity, we describe
+one salient example below and summarize the others briefly.
 
-Realm and Chapel use DMA subsystems that are optimized for large
+Realm and Chapel both use DMA subsystems that are optimized for large
 copies. Early Task Bench experiments revealed high METG values that
@@
 Chapel. These improvements affect any application where fine-grained
-tasks are used.
+tasks are used, which is particularly relevant in strong scaling
+regimes when running on large numbers of nodes.
 
+We found further areas for improvement in schedulers (Realm, Chapel
+and Dask), task pruning for scalability (PaRSEC), and optimizations on
+the task graph (TensorFlow). The impact of these optimizations ranged
+from small constant factors to even asymptotic improvements in
+performance and scalability (in the case of PaRSEC and Dask). All
+users will benefit from the improvements driven by our discoveries.
+
+\zap{
 Further analysis of Realm efficiency indicated many overheads due to
@@
 used.
+}

### 2020-04-07 0_abstract.tex
Apply Alex's comments.

--- 2019-07-30 0_abstract.tex
@@
 programming systems under a variety of application scenarios. Task
-Bench lowers the barrier to benchmarking multiple programming systems by making
+Bench dramatically lowers the barrier to benchmarking and comparing multiple programming systems by making
 the implementation for a given system orthogonal to the benchmarks

### 2020-04-07 1_introduction.tex
Apply Alex's comments.

@@
 current technologies. Our study includes several asynchronous systems
-which are designed to provide benefits such as overlapped computation
+designed to provide benefits such as overlapped computation
 and communication. While small-scale benchmarks of these systems
@@
 tangible at scale (provided the runtime overhead doesn't increase
-beyond about 100~\textmu{}s).
+beyond about 100~\textmu{}s per task).
 
@@
 systems---benefits which apply well beyond Task Bench to all classes
-of applications.
+of applications. The bugs are described in more detail in Section~\ref{sec:case-study}.
 
@@
 application developers.  Section~\ref{sec:evaluation} provides a
-comprehensive evaluation on Cori. Section~\ref{sec:related-work} relates to
+comprehensive evaluation on Cori. Section~\ref{sec:case-study} describes bugs found with Task Bench. Section~\ref{sec:related-work} relates to
 previous efforts; Section~\ref{sec:conclusion} concludes.

### 2020-04-07 4_metg.tex
Apply Alex's comments.

@@
 Section~\ref{subsec:peak-performance-and-efficiency}), and also
-because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to "select the
+because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to ``select the
 most parallel efficient job size (ratio of benchmark speed-up
-vs. linear speed-up above 50\%)" \cite{CSCSReport}. This corresponds
+vs. linear speed-up above 50\%)'' \cite{CSCSReport}, which corresponds
 to METG(50\%).
@@
 run on a single Haswell node of Cori with a problem size large enough
-that runtime is dominated by the application's inner loops. This confirms
+that runtime is dominated by the application's inner loops. This result confirms
 that the application is properly configured and that the
@@
 4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
-configuration. We use 50\% because that is generally an acceptable level
-of efficiency in practice, and values above 90\% can misrepresent the
-performance of some systems (see
-Section~\ref{subsec:peak-performance-and-efficiency}).
+configuration. %% We use 50\% because that is generally an acceptable level
+%% of efficiency in practice, and values above 90\% can misrepresent the
+%% performance of some systems (see
+%% Section~\ref{subsec:peak-performance-and-efficiency}).
 

### 2020-04-07 5_evaluation.tex
Fiddle with figure placement, caption size, prune communication experiment.
Add note on omitted load imbalance results.
Apply Alex's comments.

@@
 \label{sec:evaluation}
+
+\input{e1_flops}
+\input{e2_efficiency}
+\input{e7_bytes}
+
+\input{e3_metg_compute}
 
@@
 
+\emph{Due to space constraints we omit our load balancing results, which will be included in the revision and final paper.}
+
 \subsection{Compute Kernel Performance}
@@
 
-Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
-kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations). This is the full version of
-Figure~\ref{fig:flops-mpi}.
+Figure~\ref{fig:flops}, which is the full version of
+Figure~\ref{fig:flops-mpi}, shows the FLOP/s achieved with a compute-bound
+kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations).
 Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern.
@@
 the baseline for 100\% efficiency.
-
-\input{e1_flops}
-\input{e2_efficiency}
-\input{e7_bytes}
-
-\input{e3_metg_compute}
 
@@
 One question when considering different programming
-systems is: how much overhead does the system add? This question is tricky to answer directly because some systems introduce
+systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
 overhead \emph{inline} (i.e., by running system internal processes on
@@
 
+\input{e4_radix}
+
 Implicitly parallel systems such as PaRSEC, StarPU and Regent that
@@
 
-\input{e4_radix}
-
 \zap{
@@
 
+\zap{
 \subsection{Load Imbalance}
@@
 We leave analysis of persistent load imbalance to future work.
+} % zap
 

### 2020-04-07 8_case_study.tex
Apply Alex's comments.

@@
 Task Bench is useful not only as a tool for evaluating programming
-system performance, but also for discovering potential areas of
-improvements. During the development of Task Bench, we identified a
+system performance, but also for discovering potential areas for
+improvement. During the development of Task Bench, we identified a
 number of performance issues in the underlying programming
@@
 Chapel. These improvements affect any application where fine-grained
-tasks are used, which is particularly relevant in strong scaling
+data movement is needed, which is particularly relevant in strong scaling
 regimes when running on large numbers of nodes.
@@
 efficiency did not exceed 50\%). Chapel's other schedulers add
-overhead, resulting in higher METGs. We worked around this with
+overhead, resulting in higher METGs. We worked around this issue with
 \lstinline{serial}.

### 2020-04-08 1_introduction.tex
Tweak text to save space.

@@
 
-One approach to comparing the performance of different systems is through \emph{proxy-} or
-\emph{mini-apps}. Because they distill the key computational characteristics of a larger
-application, mini-apps can provide insight
-without the expense of developing a production code. However, despite the name, our experience is that
+One approach to comparing performance is through \emph{proxy-} or
+\emph{mini-apps}. Because they distill key computational characteristics of larger
+applications, mini-apps offer insight
+without the expense of developing production codes. However, despite the name, our experience is that
 mini-apps still require significant investment to develop
@@
 Adding a system to Task Bench involves implementing a set of standard
-services, such as executing a task or data transfer. Note that, though
-the benchmarks are described in terms of task graphs, this is simply a
+services, such as executing a task or data transfer. Though
+benchmarks are described in terms of task graphs, this is simply a
 convenient representation of the computation, and the underlying
@@
 This approach has allowed us to benchmark 15 very different parallel
-and distributed programming systems (listed in
+and distributed programming systems (see
 Table~\ref{tab:systems}).  By running all systems on common benchmarks
@@
 that have never before been measured.
-Most strikingly, the overheads of the systems we examine vary by more than five orders
-of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly
-the slower systems have ``good enough'' performance for some class of applications, while presumably
+Most strikingly, the overheads of systems we examine vary by more than five orders
+of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly,
+slower systems have ``good enough'' performance for some applications, while presumably
 providing advantages in programmer productivity.
@@
 
-METG captures the essence of what is important in a
+METG captures the important essence of a
 weak or strong scaling study, the behavior at the limit of
@@
 efficiency. For strong scaling, METG(50\%) can be used to compute the
-scale at which performance can be expected to dip below 50\%
-efficiency.  We note that METG(50\%) for a given runtime system will
-vary with the particular application and the underlying hardware---i.e., METG(50\%)
-is not a constant for a given system, but we will see that individual systems have
+scale at which efficiency can be expected to dip below 50\%.
+We note that METG(50\%) for a given runtime system will
+vary with the application and the underlying hardware---i.e., METG(50\%)
+is not a constant for a given system, but we find that systems have
 a characteristic range of METG(50\%) values and that there is additional insight

### 2020-04-08 2_task_bench.tex
Tweaks for spacing and wording.
Tweak text to save space.

@@
 To explore as broad a space of application scenarios as possible, Task
-Bench provides a large number of configuration parameters. The most
-important parameters are described in
-Table~\ref{tab:parameters}. These parameters control the size and
-structure of the task graph, the type and duration of the kernels
+Bench provides a large number of configuration parameters. These
+parameters are described in
+Table~\ref{tab:parameters}, and control the size and
+structure of the task graph, the type and duration of kernels
 associated with each task, and the amount of data associated with
@@
 
-Dependencies between tasks are controlled by a configurable dependence
+Dependencies between tasks are determined by a dependence
 relation. The
-dependence relation enumerates the set of tasks from the
+dependence relation identifies the tasks from the
 previous time step each task depends on, permitting a wide variety
@@
 
-\input{t3_api}
-\input{l0_compute_kernel}
-%% \input{l1_code_sample}
-\input{l2_mpi_sample}
-
-In addition to specifying the shape and structure of the task graph, the
-Task Bench core API also provides implementations of the kernels
+In addition to specifying the shape of the task graph, the
+core API also provides implementations of the kernels
 executed by each task as well as other utility routines (to parse
@@
 source of performance disparity that can be a pitfall for mini-apps.
+
+\input{t3_api}
+\input{l0_compute_kernel}
+%% \input{l1_code_sample}
+\input{l2_mpi_sample}
 

### 2020-04-08 3_implementation.tex
Apply Alex's comments.
Tweak text to save space.

@@
 bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU
-resolve this by permitting the user to omit tasks that are not
+allow users to omit tasks that are not
 directly connected in the dependence graph. The program is replicated
-(for efficiency of the analysis), but each task is executed only
+(for efficiency of analysis), but each task is executed only
 once. Because the dynamic checks to see if a task should be executed
-are not free of cost, we also provide a PaRSEC implementation (under
-the label \lstinline{shard}) which has been hand-written to avoid all
+are not free of cost, we also provide a PaRSEC implementation (labeled \lstinline{shard}) hand-written to avoid all
 such checks. Regent performs an equivalent optimization at compile

### 2020-04-08 4_metg.tex
Tweak text to save space.

@@
 
-Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
-calculated. The application, in this case an MPI implementation of the
-Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
-run on a single Haswell node of Cori with a problem size large enough
-that runtime is dominated by the application's inner loops. This result confirms
+Figure~\ref{fig:flops-mpi} shows how METG is
+measured. We run the application (MPI Task Bench) on a Cori Haswell node
+with a problem size large enough that runtime is dominated by kernel
+execution. This result confirms
 that the application is properly configured and that the
@@
 expectation is that as problem size shrinks,
-performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
+performance will begin to drop and eventually approach zero. Systems
 with lower runtime overheads maintain higher performance at smaller
@@
 
-METG(50\%) is the intersection of this curve with 50\% efficiency, as
+METG(50\%) is the intersection of the curve at 50\% efficiency, as
 shown by the red, dashed lines in Figures~\ref{fig:flops-mpi} and
@@
 METG has a well-defined
-relationship with quantities of interest to application developers,
-namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
-\ref{fig:strong-scaling-mpi} show the weak and strong scaling behavior of the MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
+relationship with quantities of interest such as
+weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
+\ref{fig:strong-scaling-mpi} show the weak and strong scaling of MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
 figures, the vertical axis is shown as wall time to emphasize the
@@
 very bottom, the lines compress together as running time becomes
-dominated by overhead. Note that the
-shapes at the bottom of the strong and weak scaling curves are identical,
-and conform to the shape of the METG curve (marked by the red, dashed line).
+dominated by overhead. Note that the countour of the bottom of each
+graph is identical and conforms to the METG curve (marked by the red,
+dashed line).
 

### 2020-04-08 5_evaluation.tex
Tweak text to save space.

@@
 systems reserve a number of cores (usually 1 or 2) for internal
-runtime usage; these systems take a minor hit in peak FLOP/s compared
+use; these systems take a minor hit in peak FLOP/s compared
 to systems that share all cores between the application and runtime. Some of the
-higher overhead systems struggle to achieve peak FLOP/s, though in most cases
+higher-overhead systems struggle to achieve peak FLOP/s, though in most cases
 the curves suggest that performance would continue to improve if we
@@
 Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
-peak FLOP/s) against task granularity. As described in Section~\ref{sec:metg}, this plot is
+peak FLOP/s) vs. task granularity. As described in Section~\ref{sec:metg}, this is
 used to calculate METG(50\%). The
-red, dashed line shows the point where 50\% efficiency is achieved.
-In most cases, task granularity asymptotes prior to the 50\% efficiency point,
-though some systems continue to improve as efficiency drops further. Accounting for this effect is one of the main arguments in
-favor of METG with a reasonable efficiency threshold instead of
-measuring task scheduling throughput of empty tasks
-(effectively METG(0\%)). Measuring performance using empty tasks can
-reward implementation strategies, such as devoting nearly 100\% of
+red, dashed line indicates 50\% efficiency.
+In most cases, task granularity asymptotes prior to this point,
+though some systems continue to improve at lower values. Accounting for this effect is one of the main arguments in
+favor of using reasonable efficiency thresholds for METG instead of
+empty tasks
+(i.e., METG(0\%)). Empty tasks
+reward strategies, such as devoting 100\% of
 system resources to the runtime system, that make no sense for real
@@
 to MPI, as it provides no
-opportunity for task parallelism. For the stencil pattern, the
-dominating factor is the basic overhead of executing a task, which is
-minimal for MPI as the Task Bench implementation simply executes tasks one after another in alternation with
-communication phases. In contrast, the asynchronous execution
+opportunity for task parallelism. The
+dominating factor in this case is the overhead of executing a task, which is
+minimal for MPI as the Task Bench implementation simply executes tasks in alternation with
+communication phases. The asynchronous
 mechanisms of other systems are pure overhead in this scenario.
-The gap between MPI and other systems shrinks as the complexity of the
-communication pattern grows, and even reverses as task parallelism is
+The gap between MPI and other systems shrinks as complexity grows, and even reverses as task parallelism is
 added in the form of multiple task graphs.
@@
 
-Implicitly parallel systems such as PaRSEC, StarPU and Regent that
-rely on runtime analysis to build the DAG can suffer from
+PaRSEC, StarPU and Regent
+rely on runtime analysis that can suffer from
 scalability bottlenecks if every node must consider the tasks
@@
 of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
-issued at all, so MPI has nothing to do aside from executing kernels
-as quickly as possible. Clearly, choosing a representative dependence
+issued at all, so MPI has nothing to do aside from executing kernels. Clearly, choosing a representative dependence
 pattern is important when estimating the performance of a workload or

### 2020-04-08 5_evaluation.tex
More tweaks to reduce page count.

@@
 
-METG is useful in part because it summarizes the overhead of each programming system in a single number. This makes it
-possible to evaluate METG at different node counts (shown in
-Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
-communication topology and latency. A key finding is that
-the systems with the smallest METG on one node have roughly an order
-of magnitude higher METG at 256 nodes---increased communication latencies require significantly larger tasks to
+METG summarizes system overheads in a single number. This makes it possible to
+evaluate how communication topology and latency impact METG at
+different node counts, as shown in Figure~\ref{fig:metg-compute}.
+We find that
+systems with the smallest METG on one node have roughly an order
+of magnitude higher METG at 256 nodes. Increased communication latencies require significantly larger tasks to
 achieve the same level of efficiency, so apparent differences in
-runtime overhead at small node counts can matter much less or not at
+overhead at small node counts can matter much less or not at
 all at larger node counts.
@@
 partially by allowing the user to omit tasks not directly
-dependent on those executed by a given node. However, this DAG
-trimming approach requires dynamic checks that scale with
-the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
+dependent on those executed by a given node. However, this approach still requires checks that scale with
+node count and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
 partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
@@
 ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
-a $12\times$ increase. This is unsurprising as in the case
-of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
-issued at all, so MPI has nothing to do aside from executing kernels. Clearly, choosing a representative dependence
+a $12\times$ increase. This is unsurprising, as with
+0 dependencies no \lstinline[language=C++]{MPI_Isend} calls are
+issued at all. Clearly, choosing a representative dependence
 pattern is important when estimating the performance of a workload or
@@
 To determine the cost of scheduling tasks on GPUs, Figure~\ref{fig:cuda-efficiency} compares MPI
-and MPI+CUDA on the Piz Daint supercomputer. The CUDA
-compute kernel achieves a peak performance of $4.759 \times 10^{12}$
+and MPI+CUDA on Piz Daint~\cite{PizDaint}. The CUDA
+compute kernel achieves $4.759 \times 10^{12}$
 FLOP/s, which is very close to the officially reported number
@@
 
-Our MPI+CUDA implementation uses an offload model in which data is copied to and from the GPU on each step. We test
+Our MPI+CUDA implementation uses an offload model with data copied to and from the GPU on each step. We test
 two configurations: \lstinline{w1} uses 1 task per GPU, whereas
-\lstinline{w4} overdecomposes by $4\times$, using 4 MPI
+\lstinline{w4} overdecomposes, using 4 MPI
 ranks per GPU to push work to the GPU in parallel.
 \lstinline{w4} achieves higher FLOP/s but
-drops more rapidly at smaller problem sizes, due to the overhead of
-scheduling $4\times$ as many CUDA kernel launches. In either case, the
-GPU requires more work to achieve high performance, and the overhead
+drops more rapidly at small problem sizes, due to the overhead of
+scheduling $4\times$ as many CUDA kernel launches. Either way,
+GPUs require more work to achieve high performance, and the overhead
 of copying data dominates at small task
-granularities, where the CPU achieves higher
+granularities, where CPUs achieve higher
 performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
@@
 Section~\ref{subsec:scalability}: the cost of sending data and
-tasks to the GPU imposes a higher task granularity to achieve
-the same efficiency than running on the CPU, reducing the advantage at small task granularities of
+tasks to the GPU imposes a floor on task granularity relative to the CPU, reducing the advantage at small task granularities of
 very lightweight mechanisms such as those in MPI.

### 2020-04-08 8_case_study.tex
More tweaks to reduce page count.

@@
 
-Realm and Chapel both use DMA subsystems that are optimized for large
-copies. Early Task Bench experiments revealed high METG values that
-were diagnosed by the Realm/Chapel developers as overhead due to the
+Realm and Chapel both use DMA subsystems optimized for large
+copies. Early Task Bench experiments revealed high METG values,
+diagnosed by the Realm/Chapel developers as overhead due to the
 cost of scheduling small copies. Subsequent improvements in Realm and

### 2020-04-08 3_implementation.tex
Fix typos.

@@
 systems is possible because the separation between core API (in
-Table~\ref{tab:api}) and system implmentation enables an overall
+Table~\ref{tab:api}) and system implementation enables an overall
 effort of $\mathcal{O}(m + n)$ (for $m$ benchmarks on $n$ systems)

### 2020-04-09 0_abstract.tex
Trim abstract down to the word limit for SC.

@@
+% 150 words
+
 \begin{abstract}
@@
 
-We conduct a comprehensive study with implementations of Task Bench in
+We conduct a comprehensive study with
 15 programming systems on up to 256 Haswell nodes of the Cori
@@
   granularity} to study the baseline runtime overhead of each
-system. We show that when running at scale, 100~\textmu{}s is the
-smallest granularity that even the most efficient systems can reliably
+system. Running at scale, 100~\textmu{}s is the
+smallest granularity that any system can reliably
 support with current technologies. We also study each system's

### 2020-04-21 6_related.tex
Rewrite related work microbenchmark paragraph per Brad Chamberlain's feedback.

@@
 
-Microbenchmarks can be easier to implement, but risk being
-unrepresentative of real applications. PRK Stencil~\cite{PRK14}
-contains 2 kernels (which combined fit in under 50 lines of C++)
-and a halo exchange on a structured grid; an
-evaluation~\cite{PRKRuntimes16} compares MPI variants, SHMEM, UPC,
-Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
-consists mostly of small kernels for dense array computations and has
-implementations in OpenMP~\cite{NASOpenMP99}, MPI and
-MPI+OpenMP~\cite{NASMPIOpenMP00}, and
-Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieve the breadth of
-coverage, flexibility, or ease of implementation of Task Bench. We
-believe the evidence for ease of implementation is clear; Task Bench
-is implemented for more systems than all previous mini-apps and
-microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
-makes it easy to generate new benchmarks.
+Microbenchmarks can be easier to implement, but do not solve the
+asymptotic costs of implementation. PRK Stencil~\cite{PRK14} contains a
+2D stencil and is evaluated on implementations in MPI, SHMEM, UPC,
+Charm++, and Grappa~\cite{PRKRuntimes16}. The NAS benchmark
+suite~\cite{NAS91, NAS95} consists mostly of small kernels for dense
+or sparse matrix computations and has implementations in
+OpenMP~\cite{NASOpenMP99}, MPI and MPI+OpenMP~\cite{NASMPIOpenMP00},
+and Charm++~\cite{NASCharm96}. PRK requires $\mathcal{O}(n)$ effort to
+implement (for $n$ systems) by virtue of being only a single
+computational pattern, while NAS requires $\mathcal{O}(mn)$ overall
+effort (for $m$ patterns and $n$ systems). In contrast, Task Bench
+requires $\mathcal{O}(m+n)$ effort and is easily extended to cover new
+systems or patterns with $\mathcal{O}(1)$ effort for each additional
+system or pattern.
 

### 2020-05-27 sc20_reviews/response.md
Response skeleton.

--- 2019-05-25 reviews/response.md
@@
 
-We have added four new implementations to Task Bench: Dask, bulk synchronous MPI,
-MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
-section, including preliminary results from experiments with GPUs on
-the Piz Daint supercomputer.
+Changes made in response to reviewer comments:
 
-Major changes in the revised paper:
-
-  * New implementations: Dask, bulk synchronous MPI, MPI+OpenMP and MPI+CUDA
-  * New sections:
-      - Implementation: Sections 4.3 (Dask) and 4.5 (MPI+X)
-      - Evaluation: Section 5.8 (GPU experiment)
-  * New figures:
-      - Listings 1 and 2 (code samples)
-      - Figure 13 (GPU experiment)
-  * Updated Figures 3 and 4 (see also lines 300-312)
-  * Updated Section 7 with insights derived from Task Bench
-
-In response to specific reviewer questions:
-
-Reviewer 1:
-
-The memory bound kernel has been written to perform a varying number
-of passes (or partial passes) over a fixed-sized buffer, in order to
-keep working set size constant and avoid interference due to cache
-effects, which in our opinion makes the results difficult to
-interpret.
-
-See the code samples in Listing 1 and 2.
-
-Reviewer 2:
-
-The conclusion has been updated with references to additional systems.
-
-Figures 3 and 4 have been updated and new text added.
-
-Reviewer 3:
-
-Weak/strong scaling and TPS are also application and system dependent
-metrics. METG differs from these metrics in that can simultaneously
-(a) capture the behavior of the system at the limit of its scalability
-(which is not true of weak/strong scaling) and (b) while achieving a
-specified amount of useful work (which is not true of TPS). METG also
-isolates the impact of changes to the communication topology at scale,
-whereas in a strong scaling study this effect would be entangled with
-issues of system overhead. Examples of how to compute METG from
-application wall time have been added to Section 2.
-
-METG captures the average task granularity that can be supported. It
-is acceptable for some tasks to be smaller and others larger, as
-overheads will even out across tasks.
-
-Reviewer 4:
-
-We have added a bulk synchronous implementation in MPI which uses
-MPI_Barrier to enforce the boundary between communication and
-computation phases. Also, the new MPI+OpenMP implementation also
-shares some characteristics with BSP as the OpenMP parallel for
-construct acts as an implicit barrier for the cores within a node.
-
-One of the salient features of Task Bench is that it permits the
-implementation of a benchmark with M configurations and N systems with
-an amount of effort that is O(M+N) rather than O(MN). We have updated
-the introduction to reflect this.
-
-The tasks in Task Bench are idealized compute or memory-bound loops;
-problem size is simulated by controlling the number of iterations in
-these loops. Listing 1 is an example.
-
-Shared memory is permitted and is used in our MPI+OpenMP
-implementation (among others).
-
-The code for determining the dependence pattern and implementations of
-kernels are contained in the core. A code sample from an
-implementation is included in Listing 2. As the reviewer can see,
-implementations can be quite simple.
-
-We have added an initial evaluation of MPI+CUDA. Our implementation
-uses an offload model in which data is copied to and from the GPU on
-each timestep. We find that even when using multiple ranks per GPU to
-achieve concurrency in data movement and task scheduling, GPUs
-increase the overhead of scheduling small tasks quite substantially.
-
-Reviewer 5:
-
-As described in Section 6, citations [27] and [37] are the only papers
-we're aware of which provide an empirical performance evaluation of a
-substantial number of systems. Many mini-app and benchmark papers only
-include a small number of systems, or describe systems but do not
-include an empirical evaluation of performance. When there are
-follow-up papers which describe additional implementations, they
-rarely make comparisons beyond the reference implementation of the
-code.
-
-The output of a task is an array of bytes overwritten with a repeating
-pattern (timestep, point). If the input does not match an assertion is
-thrown.
-
-Any measure of efficiency can be used with METG. For example, a
-mesh-based application could use mesh cells processed per second (as a
-percentage of the highest throughput achieved with any problem
-size). We use peak performance as the measure of efficiency in Task
-Bench because it makes it straightforward to verify that all
-implementations are configured correctly.
+Responding to reviewer questions:

### 2020-05-27 8_case_study.tex
Some more text on case study.

@@
 acknowledged by the respective system's developers, and are either
-fixed or worked around in our experiments. For brevity, we describe
-one salient example below and summarize the others briefly.
+fixed or worked around in our experiments. We describe them below.
 
@@
 
-We found further areas for improvement in schedulers (Realm, Chapel
-and Dask), task pruning for scalability (PaRSEC), and optimizations on
-the task graph (TensorFlow). The impact of these optimizations ranged
-from small constant factors to even asymptotic improvements in
-performance and scalability (in the case of PaRSEC and Dask). All
-users will benefit from the improvements driven by our discoveries.
-
-\zap{
+{\color{blue}
 Further analysis of Realm efficiency indicated many overheads due to
@@
 
-PaRSEC uses task pruning to improve scalability at large node
+PaRSEC uses task pruning to reduce analysis work performed on each
+node, and thus improve scalability at large node
 counts. Initial Task Bench results achieved less than the expected
 scalability: METG was rising too quickly with node count.  The PaRSEC
-developers diagnosed this as a bug in task pruning, and it was
+developers diagnosed this as a bug in task pruning---the optimization was failing to trigger---and it was
 subsequently fixed. We also implemented a version of the PaRSEC code,
 \lstinline{shard}, which uses manual pruning to demonstrate that
-additional gains may be possible.
+additional gains may still be possible.
 
@@
 $\mathcal{O}(N^2)$. This issue was reported to and confirmed by the
-Dask developers. As a work around, our results in the paper use a
+Dask developers; the bug is suspected to be in optimizations performed
+on the task graph by Dask. As a work around, our results in the paper use a
 lower-level interface which does not suffer from this asymptotic
@@
 used.
-}
+
+In all cases, we found bugs that are applicable outside of Task Bench
+and that impact metrics other than METG. Notably, the scalability
+issues and asymptotic complexity have a growing impact with larger
+node counts and will eventually become visible with nearly any
+application. In other cases, we found performance degradation at the
+extremes of application configurations which may be more difficult to
+test with full-size applications, but which are nonetheless relevant
+in strong-scaling regimes, particularly as node counts grow.
+
+} % color
+
+\zap{
+We found further areas for improvement in schedulers (Realm, Chapel
+and Dask), task pruning for scalability (PaRSEC), and optimizations on
+the task graph (TensorFlow). The impact of these optimizations ranged
+from small constant factors to even asymptotic improvements in
+performance and scalability (in the case of PaRSEC and Dask). All
+users will benefit from the improvements driven by our discoveries.
+} % zap

### 2020-05-27 8_case_study.tex
More edits on case study.

@@
 
-Initial experiments with TensorFlow revealed that the TensorFlow Task
-Bench implementation was not able to use all cores on a node (thus
-making it impossible to measure METG since the efficiency was below
+Initial experiments with TensorFlow revealed that the Task
+Bench implementation was not able to use all cores on a node (making it impossible to measure METG since efficiency was below
 50\%). The TensorFlow developers diagnosed this as an issue in

### 2020-05-27 8_case_study.tex
Update case study wording.

@@
 node counts and will eventually become visible with nearly any
-application. In other cases, we found performance degradation at the
-extremes of application configurations which may be more difficult to
-test with full-size applications, but which are nonetheless relevant
-in strong-scaling regimes, particularly as node counts grow.
+application. In other cases, Task Bench and METG made it possible to
+identify performance degradation at the extremes of application
+configurations which might remain hidden with full-size applications,
+particularly when evaluated with weak and strong scaling alone. The
+improvements motivated by our findings are nonetheless relevant in
+strong-scaling regimes in a variety of applications, particularly as
+node and core counts grow.
 

### 2020-05-27 sc20_reviews/response.md
Mark additions.
Update response.
Responses to reviewer 1.

@@
 
+  * Added Section V.G: Load Imbalance.
+  * Expanded Section VI: Case Study.
+
 Responding to reviewer questions:
+
+## Reviewer 1
+
+  * Task graphs can be arbitrary, including irregular graphs. We added an example to Figure 1.
+
+  * METG has already proven useful for exposing behaviors at the limits of system performance. We updated Section VI to clarify.
+
+  * **FIXME: Please justify the use of small task granularity values (lower or equal than 1ms).**
+
+## Reviewer 2
+
+## Reviewer 3
+
+## Reviewer 4
+
+## Reviewer 5

### 2020-05-27 6_related.tex
Update related work.

@@
 implement to a level of quality sufficient for conducting comparative
-studies between programming systems. The largest study we know of,
-for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
-systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI), each of
-which require a separate, tuned implementation (in contrast to
-Task Bench). Other
-studies usually lack a comprehensive evaluation, even if multiple
-implementations are available:
+studies between programming systems. {\color{blue} The largest studies
+we know of consider at most 7 and 6 programming systems,
+respectively~\cite{LULESH13, Deakin19}, and the latter only considers
+on-node programming models.} In both cases, the mini-apps under study
+require a separate, tuned implementation (in contrast to Task
+Bench). Other studies usually lack a comprehensive evaluation, even if
+multiple implementations are available:
+
+%% for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
+%% systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI), each of
+%% which
+
+% Deakin19:
+%% BabelStream: OpenMP, Kokkos, CUDA, OpenACC, OpenCL
+%% TeaLeaf: OpenMP, Kokkos, CUDA, OpenACC
+%% CloverLeaf: OpenMP, Kokkos, CUDA, OpenACC, OpenCL
+%% Neutral: OpenMP, Kokkos, CUDA, OpenACC, OpenCL
+%% MiniFMM: OpenMP, Kokkos, CUDA, OpenCL
 

### 2020-05-27 2_task_bench.tex
More tweaks to reduce page count.
Apply Brad Chamberlain's comments.
Update text on validation.

@@
 \input{t2_parameters}
-
-\brokenpenalty=0
 
@@
 each dependence edge in the graph.
-
-\brokenpenalty=\oldbrokenpenalty
 
@@
 They may also vary over time, such as in the FFT pattern.
-The set of dependence relations is extensible, making it easy to add patterns to represent new classes applications.
+The set of dependence relations is extensible, making it easy to add patterns to represent new classes of applications.
 Table~\ref{tab:equations} shows equations for the
@@
 
-The Task Bench core library is fully
-validating. Because the task graph configuration is explicitly
-represented (though unmaterialized) in Task Bench, this representation
-can be queried to determine exactly what dependencies a task should
-expect. The output of every task in Task Bench is unique,
-and all inputs are verified. An assertion is thrown if validation
-fails. These checks ensure that every execution of Task Bench, if it
-completes successfully, is correct. An evaluation of the performance impact of
+The Task Bench core library is fully validating. {\color{blue} The output of each
+task is a tuple (time step, column) and is unique within a task
+graph. All inputs are verified by checking the expected dependencies
+against those received, and an assertion is thrown if validation
+fails. These checks ensure that every execution of Task Bench is
+correct. Note that the graph representation is concise, making these
+checks very inexpensive.} An evaluation of the performance impact of
 validation showed it to be less than 3\% at the smallest task

### 2020-05-27 4_metg.tex
Fix typos.
Clarify CSCS quote.

@@
 
-The choice of 50\% is also a parameter and is not fundamental to
-METG. We use 50\% in our studies to avoid pathologies
-associated with lower efficiency thresholds (see
-Section~\ref{subsec:peak-performance-and-efficiency}), and also
-because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to ``select the
-most parallel efficient job size (ratio of benchmark speed-up
-vs. linear speed-up above 50\%)'' \cite{CSCSReport}, which corresponds
-to METG(50\%).
+The choice of 50\% is a parameter and not fundamental to METG. We use
+50\% in our studies to avoid pathologies associated with lower
+thresholds (see Section~\ref{subsec:peak-performance-and-efficiency}),
+and also because it aligns with what we observe in practice. {\color{blue} For
+example, one supercomputer center instructs users applying for
+projects to run a strong scaling study and then ``select the most
+parallel efficient job size,'' i.e., the largest number of nodes with
+a ``ratio of benchmark speed-up vs. linear speed-up above 50\%''
+\cite{CSCSReport}, which corresponds to METG(50\%).}
 
@@
 very bottom, the lines compress together as running time becomes
-dominated by overhead. Note that the countour of the bottom of each
+dominated by overhead. Note that the contours of the bottom of each
 graph is identical and conforms to the METG curve (marked by the red,

### 2020-05-28 sc20_reviews/response.md
Update response.
Update text on validation.
Clarify CSCS quote.
Add random pattern.
More response to reviewer 1.

@@
   * Added Section V.G: Load Imbalance.
-  * Expanded Section VI: Case Study.
+  * Expanded Section VI: Performance Issues Discovered.
+  * Minor additional clarifications described below.
 
@@
 
-  * METG has already proven useful for exposing behaviors at the limits of system performance. We updated Section VI to clarify.
+  * METG has already proven useful for exposing behaviors at the limits of system performance. We updated Section VI to provide additional details.
 
-  * **FIXME: Please justify the use of small task granularity values (lower or equal than 1ms).**
+  * Our experiments are chosen to span the entire range of task granularities, from very efficient (low overhead) to inefficient (high overhead).
+METG naturally determines the point at which runtime overheads come to dominate application execution.
+We find empirically that the most efficient systems have METG(50%) values ranging from about 5 us to about 100 us depending on the scale and dependency pattern, when data movement per dependency is small.
+Keep in mind that this is the *limit* of what the system can achieve; not all applications will require tasks so fine-grained.
+Also, as noted in the paper, clearly there are popular and widely used systems on both ends of this spectrum; some classes of applications can tolerate much larger task granularities.
+
+  * For a specific example of an application where small task granularities are useful, see https://ieeexplore.ieee.org/document/8948618.
+Figure 2 shows that with a tile size of 1024 the algorithm achieves about 7 TFLOPS/s.
+The task executed on each tile is a straightforward matrix multiply, making it easy to determine the the FLOPS (and therefore task granularity): (2\*1024\*1024\*1024/10^9)/7000gflops = 307us.
 
@@
 
+  * We added commit IDs to Table 5.
+
+  * We renamed Section VI.
+
+  * We added the Deakin citation to our related work.
+
 ## Reviewer 4
 
+  * Text has been added to Section II to clarify validation.
+
+  * We have clarified the quote about 50% in Section IV.
+
 ## Reviewer 5

### 2020-05-28 5_evaluation.tex
Restore load imbalance section.
Mark additions.
Clarify paragraph on peak performance.

@@
 
-\emph{Due to space constraints we omit our load balancing results, which will be included in the revision and final paper.}
-
 \subsection{Compute Kernel Performance}
@@
 
-In theory, any system should achieve peak performance if
-the kernels are well-tuned and of sufficiently large granularity. In
-practice, many subtle pitfalls of implementation or configuration can easily lead to poor performance. Verifying that peak performance is
-achieved helps to ensure that evaluations of overhead and
-efficiency are well-grounded.
+{\color{blue}
+In order to ensure that all systems are properly configured, we
+analyze the peak performance achieved by each. There should exist some
+task granularity which is sufficient to offset the runtime overheads
+of any system, regardless of how large those are. In practice, it's
+best to verify this because system configuration can be unintuitive
+and can lead to a loss in performance (e.g., due to not using all
+available cores).
+}
 
@@
 
-\zap{
+{\color{blue}
 \subsection{Load Imbalance}
@@
 We leave analysis of persistent load imbalance to future work.
-} % zap
+}
 

### 2020-05-28 sc20_reviews/response.md
Clarify paragraph on peak performance.
Update response to reviewer 5.

@@
 
-  * We have clarified the quote about 50% in Section IV.
+  * We clarified the quote about 50% in Section IV.
+
+  * We clarified the paragraph at the top of Section V.A.
+
+  * We expanded Section VI.
 
 ## Reviewer 5
+
+  * We added Section V.G.
+
+  * Figure 11 is comparable to Figure 3.
+As described in the text, Figure 3 is generated from Figure 2 by computing, for each data point, efficiency and task granularity.
+These are derived metrics, resulting in a parametric graph.
+Intuitively, there is a data point per problem size, but task granularity does not necessarily decrease with problem size.
+This view makes it intuitive how efficiency and task granularity are related.

### 2020-05-28 sc20_reviews/response.md
Add response to reviewer 5.

@@
 This view makes it intuitive how efficiency and task granularity are related.
+
+  * We use MPS in the case of MPI with 4 ranks per GPU.
+Our code uses an offload model where any communicated data is copied to and from the GPU for each task execution; this is an intentional decision to model the most common way in which MPI+CUDA is used.
+Using 4 ranks per node allows the copies to/from the GPU and kernel launches to be performed in parallel, increasing the utilization of the GPU (at the cost of higher overheads at small task granularities).

### 2020-05-28 sc20_reviews/response.md
More responses to reviewer 5.
Edit response.

@@
 
+  * Whether dependencies are constructured ahead of time or on the fly is a feature of a specific implementation of Task Bench, not the core API.
+
+    FIXME: Add note about added Dask/TensorFlow impls, or whether we add more text, or whatever.
+
   * We added Section V.G.
@@
 Using 4 ranks per node allows the copies to/from the GPU and kernel launches to be performed in parallel, increasing the utilization of the GPU (at the cost of higher overheads at small task granularities).
+All of these factors, including the kernel launch latency, are included in our task granularity measurements.

### 2020-05-29 1_introduction.tex
Update related work.
Try out some new text in the introduction.

@@
 
+{\color{blue}
+
+Existing approaches to comparing performance focus on \emph{mini-apps}
+or \emph{microbenchmarks}. These smaller codes distill key
+computational characteristics of larger applications: mini-apps are
+often derived from a larger code, and thus directly inherit its
+behaviors, while benchmarks are typically chosen to reflect a more
+narrow set of behaviors. In either case, while a variety of insight
+can be gained, the overall effort required is proportional to the
+product of the number of systems and behaviors one would like to
+evaluate. As a result, few published studies compare more than a
+handful of systems~\cite{LULESH13, Deakin19}.
+
+}
+
+\zap{
 One approach to comparing performance is through \emph{proxy-} or
@@
 often do not reach the level of maturity required to compare
-systems. Few published studies compare more than a handful of systems~\cite{LULESH13}.
+systems. Few published studies compare more than a handful of systems~\cite{LULESH13, Deakin19}.
+} % zap
 

### 2020-05-29 2_task_bench.tex
Fiddle with text.

@@
 implementations can be scripted uniformly and eliminates a potential
-source of performance disparity that can be a pitfall for mini-apps.
+source of performance disparity that can be a pitfall for other benchmarks.
 

### 2020-05-29 1_introduction.tex
Fiddle with text.
More fiddling with wording.

@@
 
-Existing approaches to comparing performance focus on \emph{mini-apps}
+Existing approaches to comparing performance focus on \emph{proxy-}/\emph{mini-apps}
 or \emph{microbenchmarks}. These smaller codes distill key
@@
 often derived from a larger code, and thus directly inherit its
-behaviors, while benchmarks are typically chosen to reflect a more
+properties, while benchmarks are typically chosen to reflect a more
 narrow set of behaviors. In either case, while a variety of insight

### 2020-05-29 sc20_reviews/response.md
Edits to response.

@@
 
-  * Added Section V.G: Load Imbalance.
+  * Added Section V-G: Load Imbalance.
   * Expanded Section VI: Performance Issues Discovered.
@@
 
-  * Our experiments are chosen to span the entire range of task granularities, from very efficient (low overhead) to inefficient (high overhead).
+  * Our experiments are chosen to span the entire range of task granularities, from very efficient (large, and low overhead) to inefficient (small, and high overhead).
 METG naturally determines the point at which runtime overheads come to dominate application execution.
-We find empirically that the most efficient systems have METG(50%) values ranging from about 5 us to about 100 us depending on the scale and dependency pattern, when data movement per dependency is small.
-Keep in mind that this is the *limit* of what the system can achieve; not all applications will require tasks so fine-grained.
-Also, as noted in the paper, clearly there are popular and widely used systems on both ends of this spectrum; some classes of applications can tolerate much larger task granularities.
+What we report in the paper are empirical results: the limit of what each system can achieve.
+This is a lower bound, and generally, such systems will be efficient as long as tasks are at least that large.
+Also, as noted in the paper, clearly there are popular and widely used systems on both ends of this spectrum; therefore some classes of applications must be able to tolerate much larger task granularities.
 
-  * For a specific example of an application where small task granularities are useful, see https://ieeexplore.ieee.org/document/8948618.
-Figure 2 shows that with a tile size of 1024 the algorithm achieves about 7 TFLOPS/s.
+  * For a specific example of an application where small task granularities are useful, see <https://ieeexplore.ieee.org/document/8948618>.
+Figure 2 shows that with a tile size of 1024 the algorithm achieves about 7 TFLOPS/s per GPU.
 The task executed on each tile is a straightforward matrix multiply, making it easy to determine the the FLOPS (and therefore task granularity): (2\*1024\*1024\*1024/10^9)/7000gflops = 307us.
@@
 
+  * Task Bench provides access to a large space of behaviors (kernel computation, load imbalance, shape of the task graph, number of dependencies and their structure, amount of data communicated, etc.).
+We agree that there are dimensions of this space that are not currently covered by Task Bench, and that are covered by specific mini-apps; we consider this a good diretion for future work.
+In spite of this, the space covered by Task Bench today is already larger than the existing benchmarks, or any single mini-app.
+Moreover, Task Bench makes it dramatically easier to implement for a wide variety of systems, making comparative study feasible.
+We have added some clarifications to the introduction.
+
+  * We updated the terminology in the introduction.
+
   * Text has been added to Section II to clarify validation.
@@
 
-  * We clarified the paragraph at the top of Section V.A.
+  * We clarified the paragraph at the top of Section V-A.
 
@@
 
+  * We agree that cross-validation with mini-apps is a good direction for future work, but that concrete benefits have already been demonstrated in the form of performance bugs found and fixed via Task Bench (see updated Section VI).
+
   * Whether dependencies are constructured ahead of time or on the fly is a feature of a specific implementation of Task Bench, not the core API.
+You can see in Listing 2 that the MPI implementation queries the core API to generate send and receive calls.
+Similar, these could be used to determine the set of tasks to issue (Dask, etc.) or the connectivity of the dataflow graph (TensorFlow, etc.).
+We ensure that all core APIs are very efficient, making these checks low-cost so that there is no penalty to querying the APIs on the fly (if that is how the programming model is designed to be used).
 
-    FIXME: Add note about added Dask/TensorFlow impls, or whether we add more text, or whatever.
-
-  * We added Section V.G.
+  * We added Section V-G.
 
@@
 These are derived metrics, resulting in a parametric graph.
-Intuitively, there is a data point per problem size, but task granularity does not necessarily decrease with problem size.
+Intuitively, there is a data point per problem size, but task granularity does not necessarily decrease with problem size (due to runtime overheads and other factors).
 This view makes it intuitive how efficiency and task granularity are related.
@@
   * We use MPS in the case of MPI with 4 ranks per GPU.
-Our code uses an offload model where any communicated data is copied to and from the GPU for each task execution; this is an intentional decision to model the most common way in which MPI+CUDA is used.
+Our code uses an offload model where any communicated data is copied to and from the GPU for each task execution.
 Using 4 ranks per node allows the copies to/from the GPU and kernel launches to be performed in parallel, increasing the utilization of the GPU (at the cost of higher overheads at small task granularities).

### 2020-06-01 1_introduction.tex
Tweak intro.
Apply Alex's edits.

--- 2020-04-08 1_introduction.tex
@@
 remain difficult to find. Our
-goal in this paper is to develop a meaningful and useful framework for
+goal in this paper is to develop a useful framework for
 comparing the performance of parallel and distributed programming
@@
 
+{\color{blue}
+
+Existing approaches to this problem focus on \emph{proxy-}/\emph{mini-apps}
+or \emph{microbenchmarks}. These smaller codes distill key
+computational characteristics of larger applications: mini-apps are
+often derived from a larger code, and thus inherit some subset of its
+properties, while benchmarks are typically chosen to reflect a more
+narrow set of behaviors. In either case, while a variety of insight
+can be gained, the overall programming effort required is proportional to the
+product of the number of systems and behaviors being
+evaluated. Few published studies compare more than a
+handful of systems~\cite{LULESH13, Deakin19}.
+
+}
+
+\zap{
 One approach to comparing performance is through \emph{proxy-} or
@@
 How does one predict whether performance will be good enough for a given application?
-The most commonly reported metrics,
+The most commonly reported {\color{blue} measures},
 \emph{weak} and \emph{strong} scaling, do not directly characterize
@@
 others---individually or in combination contribute to an order of
-magnitude or greater increase in METG, even in the most efficient
-systems. While some systems can achieve sub-microsecond METG(50\%) in
+magnitude or greater increase in METG, even in systems with the lowest
+overheads. While some systems can achieve sub-microsecond METG(50\%) in
 best-case scenarios, we show that a more realistic
@@
 systems. (All were either fixed or worked around in our experiments.)
-In some cases these correspond to order of magnitude (or even
-asymptotic) improvements in the performance of the underlying
+In some cases these correspond to order of magnitude or even
+asymptotic improvements in the performance of the underlying
 systems---benefits which apply well beyond Task Bench to all classes

### 2020-06-01 2_task_bench.tex
Apply Alex's edits.

@@
 
-The Task Bench core library is fully validating. {\color{blue} The output of each
-task is a tuple (time step, column) and is unique within a task
+The Task Bench core library is fully self-validating: {\color{blue} The output of each
+task is a tuple $\langle \textrm{row, col}\rangle$ and is unique within a task
 graph. All inputs are verified by checking the expected dependencies

### 2020-06-01 4_metg.tex
Apply Alex's edits.

@@
 efficiency starts at 100\%. Initially task granularity
-shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, leading
-to a vertical asymptote as overhead comes to dominate useful work.
+shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, approaching a vertical asymptote as overhead comes to dominate useful work.
 
@@
 very bottom, the lines compress together as running time becomes
-dominated by overhead. Note that the contours of the bottom of each
+dominated by overhead. Note that the contour of the bottom of each
 graph is identical and conforms to the METG curve (marked by the red,

### 2020-06-01 5_evaluation.tex
Apply Alex's edits.

@@
 {\color{blue}
-In order to ensure that all systems are properly configured, we
-analyze the peak performance achieved by each. There should exist some
-task granularity which is sufficient to offset the runtime overheads
-of any system, regardless of how large those are. In practice, it's
-best to verify this because system configuration can be unintuitive
-and can lead to a loss in performance (e.g., due to not using all
-available cores).
+
+We first consider the peak performance achieved by each system. There
+should exist some task granularity which is sufficient to offset the
+runtime overheads of any system, regardless of how large those
+are. Even so, a variety of issues can lead to performance loss
+(e.g. due to not using all available cores). Verifying that peak
+performance is achieved ensures that there are no such flaws in our
+configuration.
+
 }

### 2020-06-01 5_evaluation.tex
Add section on predicting strong scalability.

@@
 
-\zap{
-\input{e8_weak}
-\input{e9_strong}
-\subsection{Comparison with Weak and Strong Scaling}
-
-Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
-conventional strong and weak scaling of the stencil pattern for
-comparison with the METG results in
-Figure~\ref{fig:metg-compute-stencil}. At small node counts, where the
-average task granularity is larger than the METG for most systems, the
-performance is ideal. For several systems METG rises with node count,
-and this becomes visible in the weak and strong scaling graphs as
-non-ideal performance. Thus the shape of the weak and strong scaling
-graphs effectively interpolates between ideal performance and the METG
-curve depending on how close each system is to being limited by
-overhead.
-} % zap
-
 \input{e6_communication}
@@
 very lightweight mechanisms such as those in MPI.
+
+\input{e9_strong}
+
+{\color{blue}
+\subsection{Predicting Strong Scaling with METG}
+
+% Summary:
+
+% Charm++:
+% Limit/Ideal intersects at: 55 nodes, 0.068 s
+% Limit/Actual intersects at: 67 nodes, 0.086 s
+% Limit/Ideal vs Limit/Actual gap in nodes: 1.22x
+% Limit/Ideal vs Limit/Actual gap in time: 1.27x
+
+% Overall (out of 11 systems that perform well enough to have an intersection):
+% Limit/Ideal vs Limit/Actual gap in nodes:
+%   mean: 1.61x
+%   median: 1.60x
+%   min: 1.13x
+%   max: 2.19x
+% Limit/Ideal vs Limit/Actual gap in time:
+%   mean: 1.28x
+%   median: 1.27x
+%   min: 1.01x
+%   max: 1.69x
+
+We can use METG to predict the scalability of a code. Figure
+\ref{fig:strong-scaling} shows a strong scaling execution of Task
+Bench stencil. Lines marked ``actual'' represent strong scaling
+measurements, while ones marked ``limit'' are computed by multiplying
+METG(50\%) by tasks per core (in this case, 1000). Intuitively,
+``limit'' is the smallest time to solution that could be achieved for
+\emph{any} problem size at that node count, while maintaining 50\%
+efficiency. Note that ``actual'' eventually falls below ``limit''
+because ``actual'' is not constrained to 50\% efficiency and
+eventually drops below this level.
+
+There are two points of intest on each ``limit'' curve: the point
+where it intersects ``actual'' and the point where it intersects an
+ideal scaling curve, computed by taking the initial time to solution
+and assuming linear scaling. These points are marked in
+Figure~\ref{fig:strong-scaling} with a black square and a black
+circle, respectively. Notably, the intersection between ideal and
+``limit'' requires only a single node run, along with measurements of
+METG(50\%) at each node count. Using this data it is possible to
+predict the strong scalability of the code: i.e., the smallest time to
+solution, and the number of nodes at which that time to solution is
+achieved, while maintaining at least 50\% efficiency. In
+Figure~\ref{fig:strong-scaling} this can be seen as the distance
+between the black square and circle, which differ by a factor of
+$1.22\times$ in node count, and $1.27\times$ in time to solution.
+
+When we compute this same measure across all of the 11 programming
+systems that scale well enough to perform the experiment, we find that
+the average separation between the ideal and actual intersection
+points is $1.61\times$ in node count (max $2.19\times$, min
+$1.13\times$) and $1.28\times$ in time to solution (max $1.69x$, min
+$1.01\times$).
+
+} % color
+
+\zap{
+\input{e8_weak}
+\input{e9_strong}
+\subsection{Comparison with Weak and Strong Scaling}
+
+Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
+conventional strong and weak scaling of the stencil pattern for
+comparison with the METG results in
+Figure~\ref{fig:metg-compute-stencil}. At small node counts, where the
+average task granularity is larger than the METG for most systems, the
+performance is ideal. For several systems METG rises with node count,
+and this becomes visible in the weak and strong scaling graphs as
+non-ideal performance. Thus the shape of the weak and strong scaling
+graphs effectively interpolates between ideal performance and the METG
+curve depending on how close each system is to being limited by
+overhead.
+} % zap

### 2020-06-02 5_evaluation.tex
Tighten text.

--- 2020-06-01 5_evaluation.tex
@@
 Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
-efficiency curves under load imbalance where each task's duration is multiplied by a uniform random variable between [0,~1). The task durations are generated with a deterministic
+efficiency under load imbalance where each task's duration is multiplied by a uniform random variable in [0,~1). Task durations are generated with a deterministic
 pseudo random number generator with a consistent seed to ensure
-identical task durations for all systems.
+identical durations for all systems.
 
-The MPI implementation of Task Bench, with its distinct computation and communication phases,
+The MPI Task Bench, with its distinct computation and communication phases,
 suffers the most under load imbalance. The biggest
 difference is at large task granularities, where the imbalance
-effectively puts an upper bound on MPI efficiency. At smaller task
+effectively puts an upper bound on efficiency. At smaller task
 granularities the effect shrinks and may even reverse as systems hit
@@
 very lightweight mechanisms such as those in MPI.
+
+\input{e9_strong}
+
+{\color{blue}
+\subsection{Predicting Strong Scaling with METG}
+
+% Summary:
+
+% Charm++:
+% Limit/Ideal intersects at: 55 nodes, 0.068 s
+% Limit/Actual intersects at: 67 nodes, 0.086 s
+% Limit/Ideal vs Limit/Actual gap in nodes: 1.22x
+% Limit/Ideal vs Limit/Actual gap in time: 1.27x
+
+% Overall (out of 11 systems that perform well enough to have an intersection):
+% Limit/Ideal vs Limit/Actual gap in nodes:
+%   mean: 1.61x
+%   median: 1.60x
+%   min: 1.13x
+%   max: 2.19x
+% Limit/Ideal vs Limit/Actual gap in time:
+%   mean: 1.28x
+%   median: 1.27x
+%   min: 1.01x
+%   max: 1.69x
+
+We can use METG to predict the scalability of a code. Figure
+\ref{fig:strong-scaling} shows strong scaling for three Task Bench
+implementations with the stencil pattern. Lines marked ``actual''
+represent strong scaling measurements, while ones marked ``limit'' are
+computed by multiplying METG(50\%) by tasks per core (in this case,
+1000). Intuitively, ``limit'' is the smallest time to solution that
+can be achieved for \emph{any} problem size at that node count, while
+maintaining 50\% efficiency. Note that ``actual'' eventually falls
+below ``limit'' because ``actual'' is not constrained to 50\%
+efficiency and eventually drops below this level.
+
+There are two points of interest on each ``limit'' curve: the point
+where it intersects ``actual'' and the point where it intersects an
+ideal scaling curve, computed by taking the initial time to solution
+and assuming linear scaling. These points are marked in
+Figure~\ref{fig:strong-scaling} with a black square and black circle,
+respectively. Notably, the intersection between ideal and ``limit''
+requires only a single node run, along with measurements of METG(50\%)
+at each node count. Using this data it is possible to predict the
+strong scalability of the code: i.e., the smallest time to solution,
+and the number of nodes at which that time is achieved, while
+maintaining at least 50\% efficiency. In
+Figure~\ref{fig:strong-scaling} this can be seen as the distance
+between the black square and circle; these are separated by a factor of
+$1.22\times$ in node count, and $1.27\times$ in time to solution.
+
+When we consider all 11 programming systems that scale well enough to
+perform this experiment, we find that the average separation between
+the ideal and actual intersection points is $1.61\times$ in node count
+(max $2.19\times$, min $1.13\times$) and $1.28\times$ in time to
+solution (max $1.69\times$, min $1.01\times$).
+
+} % color
+
+\zap{
+\input{e8_weak}
+\input{e9_strong}
+\subsection{Comparison with Weak and Strong Scaling}
+
+Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
+conventional strong and weak scaling of the stencil pattern for
+comparison with the METG results in
+Figure~\ref{fig:metg-compute-stencil}. At small node counts, where the
+average task granularity is larger than the METG for most systems, the
+performance is ideal. For several systems METG rises with node count,
+and this becomes visible in the weak and strong scaling graphs as
+non-ideal performance. Thus the shape of the weak and strong scaling
+graphs effectively interpolates between ideal performance and the METG
+curve depending on how close each system is to being limited by
+overhead.
+} % zap

### 2020-06-02 sc20_reviews/response.md
Update repsonse.

@@
   * Added Section V-G: Load Imbalance.
+  * Added Section V-I: Predicting Strong Scaling with METG.
   * Expanded Section VI: Performance Issues Discovered.
@@
 ## Reviewer 2
+
+  * We added Section V-I.
+This section validates METG against time to solution.
+We show that it is possible to use METG, along with a single-node measurement of time to solution, to predict the strong scalability of a code (i.e., the point where it reaches 50% parallel efficiency, in terms of both time to solution, and node count).
+We find that we can predict time to solution (at 50% efficiency) to within a factor of 1.28X on average across 11 programming systems.
 

### 2020-06-02 sc20_reviews/response.md
Update response.

@@
 
-  * Task graphs can be arbitrary, including irregular graphs. We added an example to Figure 1.
+  * Task graphs can be arbitrary, including irregular graphs. We added an example to Figure 1 showing a random graph.
 
@@
   * Task Bench provides access to a large space of behaviors (kernel computation, load imbalance, shape of the task graph, number of dependencies and their structure, amount of data communicated, etc.).
-We agree that there are dimensions of this space that are not currently covered by Task Bench, and that are covered by specific mini-apps; we consider this a good diretion for future work.
+We agree that there are dimensions of this space that are not currently covered by Task Bench, and that are covered by specific mini-apps; we consider this a good direction for future work.
 In spite of this, the space covered by Task Bench today is already larger than the existing benchmarks, or any single mini-app.
@@
 
-  * Whether dependencies are constructured ahead of time or on the fly is a feature of a specific implementation of Task Bench, not the core API.
-You can see in Listing 2 that the MPI implementation queries the core API to generate send and receive calls.
-Similar, these could be used to determine the set of tasks to issue (Dask, etc.) or the connectivity of the dataflow graph (TensorFlow, etc.).
-We ensure that all core APIs are very efficient, making these checks low-cost so that there is no penalty to querying the APIs on the fly (if that is how the programming model is designed to be used).
+  * Whether dependencies are constructed ahead of time or on the fly is a feature of the programming model being used to implement of Task Bench, not a feature of Task Bench itself.
+E.g., Dask, PaRSEC, Regent and StarPU all construct the task graph on the fly.
+The core APIs shown in Table 3 are sufficient to do this, it is simply a matter of calling the right programming model APIs for each task.
+Listing 2 shows how we do this for MPI.
+All of our implementations are open source and will linked in the final paper.
+
+  * We ensure that all core APIs are very efficient, making these checks low-cost so that there is no penalty to querying the APIs on the fly (if that is how the programming model is designed to be used).
 
@@
 These are derived metrics, resulting in a parametric graph.
-Intuitively, there is a data point per problem size, but task granularity does not necessarily decrease with problem size (due to runtime overheads and other factors).
-This view makes it intuitive how efficiency and task granularity are related.
+Intuitively, there is a data point per problem size, but task granularity does not necessarily decrease with problem size (due to overheads).
+Noise can also introduce variations in the measurements, especially at higher node counts, and particularly with MPI which uses a model of communicating sequential processes.
 

### 2020-06-02 1_introduction.tex
Tweaks for space.

@@
 runtime-limited application throughput. But it is not a tight bound, as the
-cost of non-trivial dependencies can be significant. This issue can be easily fixed by running non-trivial
+cost of nontrivial dependencies can be significant. This issue can be easily fixed by running nontrivial
 configurations of Task Bench.
@@
 % Dask (asymptotic complexity of dask.delayed)
-% PaRSEC (task pruning with non-trivial task graphs)
+% PaRSEC (task pruning with nontrivial task graphs)
 % Realm (DMA system overhead)

### 2020-06-03 1_introduction.tex
Apply Pat's comments.

@@
 properties, while benchmarks are typically chosen to reflect a more
-narrow set of behaviors. In either case, while a variety of insight
+narrow set of behavior(s). In either case, while a variety of insight
 can be gained, the overall programming effort required is proportional to the

### 2020-06-03 2_task_bench.tex
Fiddle with figure size.
Apply Pat's comments.

--- 2020-05-29 2_task_bench.tex
@@
 
-\input{t3_api}
-\input{l0_compute_kernel}
-%% \input{l1_code_sample}
-\input{l2_mpi_sample}
-
 \zap{
@@
 
-The Task Bench core library is fully validating. {\color{blue} The output of each
-task is a tuple (time step, column) and is unique within a task
-graph. All inputs are verified by checking the expected dependencies
+The Task Bench core library is fully self-validating: {\color{blue} The output of each
+task is a tuple $\langle \textrm{row, col}\rangle$ and is unique for a given task
+graph. Inputs are verified by checking the expected dependencies
 against those received, and an assertion is thrown if validation
@@
 effect on overall results.
+
+\input{t3_api}
+\input{l0_compute_kernel}
+%% \input{l1_code_sample}
+\input{l2_mpi_sample}
 

### 2020-06-03 5_evaluation.tex
Tweaks for space.
Apply Pat's comments.

@@
 as the line for Spark immediately rises with node count. Keep in mind
-that Spark is being evaluated here with a non-trivial dependence
+that Spark is being evaluated here with a nontrivial dependence
 pattern that is relatively unrepresentative of Spark's normal use
@@
 task parallelism. To quantify this effect,
-Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
+Figure~\ref{fig:efficiency-imbalance} plots task granularity vs.
 efficiency under load imbalance where each task's duration is multiplied by a uniform random variable in [0,~1). Task durations are generated with a deterministic
@@
 load imbalance between tasks. Systems that provide an
-additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
+additional on-node work stealing capability (such as Chapel with the \texttt{distrib} scheduler) see additional gains in
 efficiency at large task granularities. However, the use of
@@
 granularities.
-For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities.
+For example, Chapel's default (non-work-stealing) scheduler outperforms \texttt{distrib} at very small task granularities.
 We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$).

### 2020-06-03 8_case_study.tex
Rename case study section.
Apply Alex's edits.
Apply Pat's comments.

@@
-\section{Case Study}
+\section{Performance Issues Discovered}
 \label{sec:case-study}
@@
 efficiency did not exceed 50\%). Chapel's other schedulers add
-overhead, resulting in higher METGs. We worked around this issue with
-\lstinline{serial}.
+overhead, resulting in higher METGs. We worked around this with
+a \lstinline{serial} block.
 
@@
 scalability: METG was rising too quickly with node count.  The PaRSEC
-developers diagnosed this as a bug in task pruning---the optimization was failing to trigger---and it was
+developers diagnosed a bug in task pruning---the optimization was failing to trigger---and it was
 subsequently fixed. We also implemented a version of the PaRSEC code,

### 2020-06-03 5_evaluation.tex
Fiddle with wording to save space.

@@
 
-Our MPI+CUDA implementation uses an offload model with data copied to and from the GPU on each step. We test
-two configurations: \lstinline{w1} uses 1 task per GPU, whereas
+Our MPI+CUDA code uses an offload model with data copied to/from the GPU every step. In our tests, \lstinline{w1} uses 1 task per GPU, whereas
 \lstinline{w4} overdecomposes, using 4 MPI
@@
 drops more rapidly at small problem sizes, due to the overhead of
-scheduling $4\times$ as many CUDA kernel launches. Either way,
+running $4\times$ as many CUDA kernels. Either way,
 GPUs require more work to achieve high performance, and the overhead
@@
 Section~\ref{subsec:scalability}: the cost of sending data and
-tasks to the GPU imposes a floor on task granularity relative to the CPU, reducing the advantage at small task granularities of
+tasks to GPUs imposes a floor on task granularity relative to CPUs, reducing the advantage at small task granularities of
 very lightweight mechanisms such as those in MPI.

### 2020-06-03 7_conclusion.tex
edit figures and indent to bring the paper back to 11 pages

@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
 
@@
 
-\item While some systems support task granularities as small as 390~ns, this applies only to trivial dependencies and small CPU-based clusters.
-A number of factors (non-trivial dependencies, accelerators and cluster sizes in the hundreds of nodes) raise the
-  METG that any system can reliably achieve by over an order of magnitude: 100~\textmu{}s is a reasonable bound for most applications running at scale with current technologies.
+\item While some systems support METG(50\%) as small as 390~ns, this applies only to trivial dependencies and small CPU-based clusters.
+A number of factors (nontrivial dependencies, accelerators and cluster sizes in the hundreds of nodes) raise
+  the METGs that can be achieved by over an order of magnitude: 100~\textmu{}s is a reasonable bound for most applications running at scale with current technologies.
 

### 2020-06-03 5_evaluation.tex
Fiddle with figure placement.
Some new text and tweaks for space.

@@
 \label{sec:evaluation}
-
-\input{e1_flops}
-\input{e2_efficiency}
-\input{e7_bytes}
-
-\input{e3_metg_compute}
 
@@
 
-For GPU experiments we use the Piz Daint supercomputer~\cite{PizDaint}, a Cray XC50 with
-one Intel Xeon E5-2690 v3 (12 physical cores) and one NVIDIA Tesla
+\input{e1_flops}
+\input{e2_efficiency}
+\input{e7_bytes}
+
+\input{e3_metg_compute}
+
+For GPU experiments we use Piz Daint~\cite{PizDaint}, a Cray XC50 with
+a Intel Xeon E5-2690 v3 (12 physical cores) and one NVIDIA Tesla
 P100 per node. We use GCC 6.2.0, Cray MPICH 7.7.2, and
@@
 kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations).
-Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern.
+Each data point in the graph is a mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern.
 In the best case, we measure peak FLOP/s of
 $1.26 \times 10^{12}$, which compares favorably with the officially
-reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
-of measuring efficiency, we use our empirically determined number as
-the baseline for 100\% efficiency.
+reported number of $1.2 \times 10^{12}$ \cite{Cori}. We use our empirically determined number as
+the baseline for 100\% efficiency below.
 
@@
 systems reserve a number of cores (usually 1 or 2) for internal
-use; these systems take a minor hit in peak FLOP/s compared
-to systems that share all cores between the application and runtime. Some of the
+use (see below); these systems take a minor hit in peak FLOP/s compared
+to systems that share all cores between application and runtime. Some
 higher-overhead systems struggle to achieve peak FLOP/s, though in most cases
@@
 
-Not all cores are required to saturate memory bandwidth, reducing the
-impact of reserving cores for internal use. Most systems hit 100\% of
-peak, unlike the compute-bound case.
+Not all cores are required to saturate memory bandwidth. {\color{blue}
+  This reduces the impact of reserving cores for system use
+  (e.g. task-based systems that perform dependence analysis).} Nearly
+all systems hit 100\% of peak, unlike the compute-bound case.
 
@@
 as a proxy for overhead. Figure~\ref{fig:metg-compute}
-shows how METG(50\%) varies with node count for a variety of
+shows how METG(50\%) varies with node count for a subset of
 dependence patterns supported by Task Bench. METG(50\%) is calculated
 separately at each node count, to distinguish runtime system behavior from
-changes in communication latencies and topology when using
+changes in communication latency and topology when using
 progressively larger portions of the machine.
@@
 
-We observe that the baseline overheads of different
-systems vary by over 5 orders of magnitude. It is worth
-remembering when considering this metric that this is a \emph{minimum}
-effective task granularity. Applications with an average
+We observe that overheads vary by over 5 orders of
+magnitude. {\color{blue} The most efficient systems are explicitly parallel
+and provide very lightweight mechanisms for parallelism. Task-based systems for HPC tend to be next most efficient, and
+provide additional features such as automatic dependency discovery and
+data movement. Slower systems tend to be designed primarily for
+large-scale data analysis or workflows.} It is worth
+remembering that these are \emph{minimum}
+effective task granularities. Applications with an average
 task granularity of \emph{at least} this value can usually be expected
@@
 dominating factor in this case is the overhead of executing a task, which is
-minimal for MPI as the Task Bench implementation simply executes tasks in alternation with
-communication phases. The asynchronous
-mechanisms of other systems are pure overhead in this scenario.
-The gap between MPI and other systems shrinks as complexity grows, and even reverses as task parallelism is
+minimal for MPI as the code simply executes tasks in alternation with
+communication. The asynchrony
+of other systems is pure overhead in this scenario.
+MPI's advantage shrinks as complexity grows, and even reverses as task parallelism is
 added in the form of multiple task graphs.
 
-Spark and Swift/T
-are omitted from comparisons with more complicated dependencies, as the overheads of these systems require
-excessive problem sizes (beyond what can be completed in 6
+We omit Spark and Swift/T
+with more complicated dependencies, as their higher overheads require
+excessive problem sizes (beyond what completes in 6
 hours) to reach 50\% efficiency.

### 2020-06-03 e3_metg_compute.tex
Clarify that multiple graphs are independent.
edit figures and indent to bring the paper back to 11 pages
Copy some of Wei's changes to save space.

@@
 
+{\captionsetup[subfloat]{farskip=-0.1cm}
 \subfloat[Stencil pattern.\label{fig:metg-compute-stencil}]{
@@
 }
+}
 
@@
 }
-\subfloat[Nearest pattern, 5 deps/task, 4 graphs.\label{fig:metg-compute-4x-nearest}]{
+\subfloat[Nearest pattern, 5 deps/task, 4 independent graphs.\label{fig:metg-compute-4x-nearest}]{
 \includegraphics[width=\columnwidth]{figs/task-bench-results/compute/metg_ngraphs_4_nearest.pdf}
@@
 \caption{METG vs node count for different dependence patterns. Lower is better.\label{fig:metg-compute}}
-%% \vspace{-0.05cm}
+\vspace{-0.25cm}
 \end{figure*}

### 2020-06-04 sc20_reviews/response.md
Update response.
Fix a typo.

@@
   * Expanded Section VI: Performance Issues Discovered.
-  * Minor additional clarifications described below.
+  * Additional changes requested by specific reviewers are described below.
+
+All changes are marked in blue.
 
@@
 We find that we can predict time to solution (at 50% efficiency) to within a factor of 1.28X on average across 11 programming systems.
+
+  * We expanded Section V-B and V-C to describe how the properties of the various systems influence our results.
 
@@
 Listing 2 shows how we do this for MPI.
-All of our implementations are open source and will linked in the final paper.
+All of our implementations are open source and will be linked in the final paper.
 

### 2020-06-05 5_evaluation.tex
Add table and explanation.

@@
 \input{e9_strong}
+\input{e11_metg_predict_strong}
 
@@
 
+Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
+patterns tested in Figure~\ref{fig:metg-compute} for all 11
+programming systems that scale well enough to evaluate the separation
+between the limit-ideal and limit-actual intersections. We see that
+overall, the mean separation is less than $2.15\times$ in node count
+and less than $1.31\times$ in time to solution, making this an
+accurate way to predict strong scaling in 4 patterns we tested in
+Section~\ref{subsec:scalability}.
+
+\zap{
 When we consider all 11 programming systems that scale well enough to
@@
 solution (max $1.69\times$, min $1.01\times$).
+}
 

### 2020-06-05 e11_metg_predict_strong.tex
Add table and explanation.

\begin{table}[t]
\color{blue}
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std. dev. & mean & std. dev. \\
\hline
stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
nearest & $1.77\times$ & 0.253 & $1.17\times$ & 0.341 \\
spread  & $1.39\times$ & 0.322 & $1.28\times$ & 0.207 \\
nearest, 4 graphs & $2.15\times$ & 0.635 & $1.31\times$ & 0.777
\end{tabular}

\vspace{-0.20cm}
\caption{\color{blue} Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Figure~\ref{fig:metg-compute}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}

### 2020-06-05 4_metg.tex
Define METTS.

@@
 relationship with quantities of interest such as
-weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
-\ref{fig:strong-scaling-mpi} show the weak and strong scaling of MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
-figures, the vertical axis is shown as wall time to emphasize the
-relationship to time-to-solution, but it could equivalently be shown
-as task granularity (as the number of tasks per execution is
-fixed). Intuitively, at
+time to solution. Figures~\ref{fig:weak-scaling-mpi} and
+\ref{fig:strong-scaling-mpi} show the weak and strong scaling of MPI
+Task Bench running a stencil pattern at a variety of problem
+sizes. {\color{blue} To show the relationship with METG, the red,
+  dashed line in each graph marks the \emph{minimum effective time to
+    solution} (METTS). This is computed by multiplying METG by the
+  number of tasks per core (in this case 1000). METG is computed
+  separately for each node count, so METTS varies at scale.}
+Intuitively, at
 larger problem sizes MPI is perfectly efficient. This can be seen at
@@
 dominated by overhead. Note that the contour of the bottom of each
-graph is identical and conforms to the METG curve (marked by the red,
-dashed line).
+graph is identical and conforms to the METTS curve.
 

### 2020-06-05 5_evaluation.tex
Define METTS.

@@
 
-We can use METG to predict the scalability of a code. Figure
-\ref{fig:strong-scaling} shows strong scaling for three Task Bench
-implementations with the stencil pattern. Lines marked ``actual''
-represent strong scaling measurements, while ones marked ``limit'' are
-computed by multiplying METG(50\%) by tasks per core (in this case,
-1000). Intuitively, ``limit'' is the smallest time to solution that
+We can use METG to predict scalability. Figure
+\ref{fig:strong-scaling} shows three
+implementations strong scaling the stencil pattern. Lines marked ``actual''
+represent strong scaling measurements.
+The METTS lines are calculated as described in Section~\ref{sec:metg}, by multiplying multiplying METG(50\%) by tasks per core (1000).
+Intuitively, METTS is the smallest time to solution that
 can be achieved for \emph{any} problem size at that node count, while
 maintaining 50\% efficiency. Note that ``actual'' eventually falls
-below ``limit'' because ``actual'' is not constrained to 50\%
+below METTS because ``actual'' is not constrained to 50\%
 efficiency and eventually drops below this level.
 
-There are two points of interest on each ``limit'' curve: the point
+There are two points of interest on each METTS curve: the point
 where it intersects ``actual'' and the point where it intersects an
@@
 Figure~\ref{fig:strong-scaling} with a black square and black circle,
-respectively. Notably, the intersection between ideal and ``limit''
+respectively. Notably, the intersection between ideal and METTS
 requires only a single node run, along with measurements of METG(50\%)
@@
 Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
-patterns tested in Figure~\ref{fig:metg-compute} for all 11
+patterns tested in Section~\ref{subsec:scalability} for all 11
 programming systems that scale well enough to evaluate the separation
-between the limit-ideal and limit-actual intersections. We see that
-overall, the mean separation is less than $2.15\times$ in node count
-and less than $1.31\times$ in time to solution, making this an
-accurate way to predict strong scaling in 4 patterns we tested in
+between the METTS-ideal and METTS-actual intersections. We see that
+overall, the mean separation is at most $2.15\times$ in node count
+and at most $1.31\times$ in time to solution, making this an
+accurate way to predict strong scaling in the 4 patterns we tested in
 Section~\ref{subsec:scalability}.

### 2020-06-05 e11_metg_predict_strong.tex
Minor fix.
Define METTS.

@@
         & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
-Pattern & mean & std. dev. & mean & std. dev. \\
+Pattern & mean & std.~dev. & mean & std.~dev. \\
 \hline
@@
 \vspace{-0.20cm}
-\caption{\color{blue} Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Figure~\ref{fig:metg-compute}.\label{tab:metg-predict-strong}}
+\caption{\color{blue} Average factor of separation between METTS-ideal and METTS-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
 \vspace{-0.5cm}

### 2020-06-05 6_related.tex
edit figures and indent to bring the paper back to 11 pages
Un-itemize the bulleted lists in sections 7 and 8.

@@
 
-Though smaller than full applications, mini-apps can be challenging to
-implement to a level of quality sufficient for conducting comparative
-studies between programming systems. {\color{blue} The largest studies
-we know of consider at most 7 and 6 programming systems,
-respectively~\cite{LULESH13, Deakin19}, and the latter only considers
-on-node programming models.} In both cases, the mini-apps under study
-require a separate, tuned implementation (in contrast to Task
-Bench). Other studies usually lack a comprehensive evaluation, even if
-multiple implementations are available:
-
 %% for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
@@
 
-\begin{itemize}
-
-\item
-The PENNANT reference implementation supports
+Though smaller than full applications, mini-apps can be challenging to
+implement to a level of quality sufficient for conducting comparative
+studies between programming systems. {\color{blue} The largest studies
+we know of consider at most 7 and 6 programming systems,
+respectively~\cite{LULESH13, Deakin19}, and the latter only considers
+on-node programming models.} In both cases, the mini-apps under study
+require a separate, tuned implementation (in contrast to Task
+Bench). Other studies usually lack a comprehensive evaluation, even if
+multiple implementations are available:
+(a) The PENNANT reference implementation supports
 MPI/OpenMP/MPI+OpenMP~\cite{PENNANT}. A follow-up paper presents a
 Regent implementation~\cite{Regent15}.
-
-\item
-One follow-up paper for the mini-app CoMD describes a Chapel
+(b) One follow-up paper for the mini-app CoMD describes a Chapel
 implementation~\cite{CoMDChapel16} (comparison against reference
@@
   CoMDThreadedModels14}.
-
-\item
-A report on the Mantevo project~\cite{Mantevo09} describes a number of
+(c) A report on the Mantevo project~\cite{Mantevo09} describes a number of
 mini-apps, but only includes self-comparisons based on reference
 implementations.
-
-\item
-A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
+And (d) A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
 four implementations of the mini-app, but only includes performance
@@
 
-\end{itemize}
-
-Microbenchmarks can be easier to implement, but do not solve the
+Microbenchmarks can be easier to implement, but do not address the
 asymptotic costs of implementation. PRK Stencil~\cite{PRK14} contains a

### 2020-06-05 7_conclusion.tex
Un-itemize the bulleted lists in sections 7 and 8.

@@
 insights:
-
-\begin{itemize}[leftmargin=*]
-
-\item METG for current distributed programming systems varies by over
+(1) METG for current distributed programming systems varies by over
   5 orders of magnitude.  Clearly understanding the needed task
@@
   system for a new application.
-
-\item While some systems support METG(50\%) as small as 390~ns, this applies only to trivial dependencies and small CPU-based clusters.
+(2) While some systems support METG(50\%) as small as 390~ns, this applies only to trivial dependencies and small CPU-based clusters.
 A number of factors (nontrivial dependencies, accelerators and cluster sizes in the hundreds of nodes) raise
   the METGs that can be achieved by over an order of magnitude: 100~\textmu{}s is a reasonable bound for most applications running at scale with current technologies.
-
-\item Systems that support asynchronous execution show benefits under
+(3) Systems that support asynchronous execution show benefits under
   balanced computation
@@
   high baseline overheads.
-
-\item Systems for large scale data analysis require very large tasks
+(4) Systems for large scale data analysis require very large tasks
   (tens of seconds) to scale beyond small node counts,
@@
   performance in current workloads.
-
-\item Task Bench has proven effective in finding performance issues
+And (5) Task Bench has proven effective in finding performance issues
   and has lead to substantial improvements in several of the systems
   we study.
-
-\end{itemize}
 

### 2020-06-05 5_evaluation.tex
reduce spaces
Revert "Define METTS."
This reverts commit 8f71ba6a8c85b0e2cefe46a815aa459134ae66c9.

--- 2020-06-05 5_evaluation.tex
@@
 Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
-patterns tested in Figure~\ref{fig:metg-compute} for all 11
+patterns tested in Section~\ref{subsec:scalability} for all 11
 programming systems that scale well enough to evaluate the separation
 between the limit-ideal and limit-actual intersections. We see that
-overall, the mean separation is less than $2.15\times$ in node count
-and less than $1.31\times$ in time to solution, making this an
-accurate way to predict strong scaling in 4 patterns we tested in
+overall, the mean separation is at most $2.15\times$ in node count
+and at most $1.31\times$ in time to solution, making this an
+accurate way to predict strong scaling in the 4 patterns we tested in
 Section~\ref{subsec:scalability}.

### 2020-06-05 e11_metg_predict_strong.tex
Revert "Define METTS."
This reverts commit 8f71ba6a8c85b0e2cefe46a815aa459134ae66c9.

@@
 \vspace{-0.20cm}
-\caption{\color{blue} Average factor of separation between METTS-ideal and METTS-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
+\caption{\color{blue} Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
 \vspace{-0.5cm}

### 2020-06-05 5_evaluation.tex
Update text for "limit 50%".

@@
 implementations with the stencil pattern. Lines marked ``actual''
-represent strong scaling measurements, while ones marked ``limit'' are
+represent strong scaling measurements, while ones marked ``limit 50\%'' are
 computed by multiplying METG(50\%) by tasks per core (in this case,
@@
 
-There are two points of interest on each ``limit'' curve: the point
+There are two points of interest on each ``limit 50\%'' curve: the point
 where it intersects ``actual'' and the point where it intersects an
@@
 Figure~\ref{fig:strong-scaling} with a black square and black circle,
-respectively. Notably, the intersection between ideal and ``limit''
+respectively. Notably, the intersection between ideal and ``limit 50\%''
 requires only a single node run, along with measurements of METG(50\%)
@@
 overall, the mean separation is at most $2.15\times$ in node count
-and at most $1.31\times$ in time to solution, making this an
+and at most $1.42\times$ in time to solution, making this an
 accurate way to predict strong scaling in the 4 patterns we tested in

### 2020-06-05 5_evaluation.tex
Minor wording tweaks.

@@
 can be achieved for \emph{any} problem size at that node count, while
-maintaining 50\% efficiency. Note that ``actual'' eventually falls
-below ``limit'' because ``actual'' is not constrained to 50\%
-efficiency and eventually drops below this level.
+maintaining 50\% efficiency. Data points where ``actual'' falls below
+``limit'' mark points where strong scaling parallel efficiency is less
+than 50\%.
 
@@
 ideal scaling curve, computed by taking the initial time to solution
-and assuming linear scaling. These points are marked in
+and assuming linear scaling. These points are marked for Charm++ in
 Figure~\ref{fig:strong-scaling} with a black square and black circle,

### 2020-06-06 5_evaluation.tex
Tweak wording and space.

@@
 Figure~\ref{fig:strong-scaling} with a black square and black circle,
-respectively. Notably, the intersection between ideal and ``limit 50\%''
-requires only a single node run, along with measurements of METG(50\%)
-at each node count. Using this data it is possible to predict the
-strong scalability of the code: i.e., the smallest time to solution,
+respectively.
+Compared with brute-force benchmarking of an application at various
+scales, the METG(50\%) measurements allow a single run of the
+application at small scale to be used to estimate the strong scalability
+of the code:
+%% Notably, computing the ideal scaling line requires only
+%% the running time on a single node, so the intersection between ideal
+%% and ``limit'' can be found with only the single-node time to
+%% solution, along with measurements of METG(50\%) at each node
+%% count. Using this data it is possible to predict the
+%% strong scalability of the code:
+i.e., the smallest time to solution,
 and the number of nodes at which that time is achieved, while

### 2020-06-06 6_related.tex
Revert "Un-itemize the bulleted lists in sections 7 and 8."
This reverts commit b75938db0edc33113608b01a9966f9370966d97a.
Tweak wording and space.

@@
 four implementations of the mini-app, but only includes performance
-results for three, and of those three only two can be compared in an
-apples-to-apples manner as the last implementation uses a structured
-rather than an unstructured mesh. A follow-up paper describes another
-implementation in Regent~\cite{Regent15} (comparison against reference
+results for three, of which only two can be compared in an
+apples-to-apples manner as the last implementation uses structured
+rather than unstructured meshes. A follow-up describes another
+implementation in Regent~\cite{Regent15} (comparison vs.~reference
 only).

### 2020-06-06 7_conclusion.tex
Revert "Un-itemize the bulleted lists in sections 7 and 8."
This reverts commit b75938db0edc33113608b01a9966f9370966d97a.
Tweak wording and space.

@@
   reflecting the very coarse tasks and lack of need for strong scaling
-  performance in current workloads.
+  in current workloads.
 And (5) Task Bench has proven effective in finding performance issues
-  and has lead to substantial improvements in several of the systems
+  and has lead to substantial improvements in several systems
   we study.

### 2020-06-06 sc20_reviews/response.md
Fix numbers in response.

@@
 We show that it is possible to use METG, along with a single-node measurement of time to solution, to predict the strong scalability of a code (i.e., the point where it reaches 50% parallel efficiency, in terms of both time to solution, and node count).
-We find that we can predict time to solution (at 50% efficiency) to within a factor of 1.28X on average across 11 programming systems.
+We find that we can predict time to solution (at 50% efficiency) to within a factor of 1.42X on average across 11 programming systems and 4 Task Bench configurations.
 

### 2020-06-06 5_evaluation.tex
Tweak wording.

@@
 overall, the mean separation is at most $2.15\times$ in node count
-and at most $1.42\times$ in time to solution, making this an
-accurate way to predict strong scaling in the 4 patterns we tested in
+and at most $1.42\times$ in time to solution, making this a
+useful way to predict strong scaling in the 4 patterns we tested in
 Section~\ref{subsec:scalability}.

### 2020-06-06 5_evaluation.tex
Update paper and response.

@@
 {\color{blue}
-\subsection{Predicting Strong Scaling with METG}
+\subsection{Validating METG with Time to Solution}
 
@@
 computed by multiplying METG(50\%) by tasks per core (in this case,
-1000). Intuitively, ``limit'' is the smallest time to solution that
+1000) to obtain wall clock time. Intuitively, ``limit'' is the smallest time to solution that
 can be achieved for \emph{any} problem size at that node count, while
@@
 respectively.
-Compared with brute-force benchmarking of an application at various
-scales, the METG(50\%) measurements allow a single run of the
-application at small scale to be used to estimate the strong scalability
-of the code:
+Notably, the ideal-limit intersection (black circle) requires only a
+run of the application on one node, combined with METG(50\%)
+measurements at each node count, to estimate the strong scalability of
+the code:
 %% Notably, computing the ideal scaling line requires only

### 2020-06-06 5_evaluation.tex
Fiddle with wording.

@@
 run of the application on one node, combined with METG(50\%)
-measurements at each node count, to estimate the strong scalability of
+measurements, to estimate the strong scalability of
 the code:

### 2020-06-30 2_task_bench.tex
Strip out color from revision.

@@
 
-The Task Bench core library is fully self-validating: {\color{blue} The output of each
+The Task Bench core library is fully self-validating: The output of each
 task is a tuple $\langle \textrm{row, col}\rangle$ and is unique for a given task
@@
 correct. Note that the graph representation is concise, making these
-checks very inexpensive.} An evaluation of the performance impact of
+checks very inexpensive. An evaluation of the performance impact of
 validation showed it to be less than 3\% at the smallest task

### 2020-06-30 4_metg.tex
reduce spaces
Strip out color from revision.

--- 2020-06-01 4_metg.tex
@@
 thresholds (see Section~\ref{subsec:peak-performance-and-efficiency}),
-and also because it aligns with what we observe in practice. {\color{blue} For
+and also because it aligns with what we observe in practice. For
 example, one supercomputer center instructs users applying for
@@
 a ``ratio of benchmark speed-up vs. linear speed-up above 50\%''
-\cite{CSCSReport}, which corresponds to METG(50\%).}
+\cite{CSCSReport}, which corresponds to METG(50\%).
 

### 2020-06-30 5_evaluation.tex
Strip out color from revision.

@@
 
-{\color{blue}
-
 We first consider the peak performance achieved by each system. There
@@
 configuration.
-
-}
 
@@
 
-Not all cores are required to saturate memory bandwidth. {\color{blue}
+Not all cores are required to saturate memory bandwidth.
   This reduces the impact of reserving cores for system use
-  (e.g. task-based systems that perform dependence analysis).} Nearly
+  (e.g. task-based systems that perform dependence analysis). Nearly
 all systems hit 100\% of peak, unlike the compute-bound case.
@@
 We observe that overheads vary by over 5 orders of
-magnitude. {\color{blue} The most efficient systems are explicitly parallel
+magnitude. The most efficient systems are explicitly parallel
 and provide very lightweight mechanisms for parallelism. Task-based systems for HPC tend to be next most efficient, and
@@
 data movement. Slower systems tend to be designed primarily for
-large-scale data analysis or workflows.} It is worth
+large-scale data analysis or workflows. It is worth
 remembering that these are \emph{minimum}
@@
 
-{\color{blue}
 \subsection{Load Imbalance}
@@
 We leave analysis of persistent load imbalance to future work.
-}
 
@@
 
-{\color{blue}
 \subsection{Validating METG with Time to Solution}
@@
 
-} % color
-
 \zap{

### 2020-06-30 6_related.tex
Strip out color from revision.

@@
 implement to a level of quality sufficient for conducting comparative
-studies between programming systems. {\color{blue} The largest studies
+studies between programming systems. The largest studies
 we know of consider at most 7 and 6 programming systems,
 respectively~\cite{LULESH13, Deakin19}, and the latter only considers
-on-node programming models.} In both cases, the mini-apps under study
+on-node programming models. In both cases, the mini-apps under study
 require a separate, tuned implementation (in contrast to Task

### 2020-06-30 8_case_study.tex
Tweak wording and space.
Strip out color from revision.

@@
 
-{\color{blue}
 Further analysis of Realm efficiency indicated many overheads due to
@@
 efficiency did not exceed 50\%). Chapel's other schedulers add
-overhead, resulting in higher METGs. We worked around this with
+overhead, resulting in higher METGs. Based on guidance from the Chapel team, we worked around this with
 a \lstinline{serial} block.

### 2020-06-30 e11_metg_predict_strong.tex
Fix numbers.
Strip out color from revision.

@@
 \begin{table}[t]
-\color{blue}
 \centering
@@
 stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
-nearest & $1.77\times$ & 0.253 & $1.17\times$ & 0.341 \\
-spread  & $1.39\times$ & 0.322 & $1.28\times$ & 0.207 \\
-nearest, 4 graphs & $2.15\times$ & 0.635 & $1.31\times$ & 0.777
+nearest & $1.77\times$ & 0.253 & $1.24\times$ & 0.310 \\
+spread  & $1.39\times$ & 0.322 & $1.30\times$ & 0.171 \\
+nearest, 4 graphs & $2.15\times$ & 0.635 & $1.42\times$ & 0.723
 \end{tabular}
@@
 \vspace{-0.20cm}
-\caption{\color{blue} Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
+\caption{Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
 \vspace{-0.5cm}

### 2020-06-30 6_related.tex
Re-itemize lists that had to be compressed for the revision.

@@
 multiple implementations are available:
-(a) The PENNANT reference implementation supports
+
+\begin{itemize}
+\item The PENNANT reference implementation supports
 MPI/OpenMP/MPI+OpenMP~\cite{PENNANT}. A follow-up paper presents a
 Regent implementation~\cite{Regent15}.
-(b) One follow-up paper for the mini-app CoMD describes a Chapel
+
+\item One follow-up paper for the mini-app CoMD describes a Chapel
 implementation~\cite{CoMDChapel16} (comparison against reference
@@
   CoMDThreadedModels14}.
-(c) A report on the Mantevo project~\cite{Mantevo09} describes a number of
+
+\item A report on the Mantevo project~\cite{Mantevo09} describes a number of
 mini-apps, but only includes self-comparisons based on reference
 implementations.
-And (d) A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
+
+\item A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
 four implementations of the mini-app, but only includes performance
@@
 only).
+\end{itemize}
 

### 2020-06-30 7_conclusion.tex
Re-itemize lists that had to be compressed for the revision.

--- 2020-06-03 7_conclusion.tex
@@
 
-\begin{itemize}[leftmargin=*]
-
+\begin{itemize}
 \item METG for current distributed programming systems varies by over
@@
   reflecting the very coarse tasks and lack of need for strong scaling
-  performance in current workloads.
+  in current workloads.
 
 \item Task Bench has proven effective in finding performance issues
-  and has lead to substantial improvements in several of the systems
+  and has lead to substantial improvements in several systems
   we study.
-
 \end{itemize}

### 2020-07-30 7_conclusion.tex
Add acknowledgment.

@@
 systems.
+
+\section*{Acknowledgment}
+
+This material is based upon work supported by the U.S. Department of
+Energy, Office of Science, Office of ASCR, under the contract number
+DE-AC02-76SF00515, by National Science Foundation under Grant
+No. ACI-1450300, and the Exascale Computing Project (17-SC-20-SC), a
+collaborative effort of the U.S. Department of Energy Office of
+Science and the National Nuclear Security Administration, under prime
+contract DE-AC05-00OR22725, and UT Battelle subawards 4000151974 and
+89233218CNA000001. Experiments on the Cori supercomputer were
+supported by the National Energy Research Scientific Computing Center,
+a DOE Office of Science User Facility supported by the Office of
+Science of the U.S. Department of Energy under Contract
+No. DE-AC02-05CH11231, and experiments on Piz Daint were supported by
+the Swiss National Supercomputing Centre (CSCS) under project ID d80.

### 2020-08-10 t1_systems.tex
Tweak text and font sizes.
Shrink space around tables.
Update StarPU version and information.

--- 2019-04-08 t1_systems.tex
@@
 \begin{table}[t]
-\begin{tabular}{l | l | l | l | l}
+\small
+\begin{tabular}{@{} l | l | l | l | l @{}}
 System & Paradigm & Parallelism & Distrib. & Network \\
@@
 Spark & functional & implicit & yes & sockets \\
-StarPU & task-based & implicit & yes & MPI \\
+StarPU & task-based & \emph{expl., impl.} & yes & MPI \\
 Swift/T & dataflow & implicit & yes & MPI \\
 TensorFlow & dataflow & explicit & yes\footnote{Our evaluation only considers TensorFlow on a single node.} & sockets \\
-X10 & place-based & explicit & yes & MPI\footnote{X10 also provides support for PAMI on supported networks.}
+X10 & place-based & explicit & yes & MPI\footnote{X10 also provides a PAMI backend on supported networks.}
 \end{tabular}

### 2020-08-11 3_implementation.tex
Fiddle with figures.
Update StarPU text.

@@
 
-\input{t1_systems}
-
 \subsection{Task-Based Programming Models}
@@
 above) is explicitly parallel and requires tasks to be connected
-explicitly via \emph{events}.
+explicitly via \emph{events}. StarPU, in addition to its usual,
+implicitly parallel mode, also provides a mode where MPI is used for
+synchronization (and is thus explicitly parallel).
 
@@
 bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU
-allow users to omit tasks that are not
-directly connected in the dependence graph. The program is replicated
-(for efficiency of analysis), but each task is executed only
-once. Because the dynamic checks to see if a task should be executed
-are not free of cost, we also provide a PaRSEC implementation (labeled \lstinline{shard}) hand-written to avoid all
-such checks. Regent performs an equivalent optimization at compile
-time~\cite{ControlReplication17} that does not require user
-intervention.
+allow users to omit tasks that are not directly connected in the
+dependence graph. The program is replicated (for efficiency of
+analysis), but each task is executed only once. Because the dynamic
+checks to see if a task should be executed are not free of cost, we
+also provide versions of the PaRSEC and StarPU implementations
+(labeled \lstinline{shard} and \lstinline{expl}, respectively)
+hand-written to minimize such costs. PaRSEC \lstinline{shard} uses the
+DTD mode but manually minimizes dynamic checks. StarPU
+\lstinline{expl} uses the MPI integration described above. We see in
+Section~\ref{subsec:scalability} that such modifications are needed to
+achieve optimal scalability. Regent performs an equivalent
+optimization at compile time~\cite{ControlReplication17} that does not
+require user intervention and preserves the original, implicitly
+parallel programming model of the language.
+
+\input{t1_systems}
 

### 2020-08-13 5_evaluation.tex

@@
 by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
-retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
-uses a compile-time optimization to improve
-scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.
+retains dynamic checks that limit scalability. PaRSEC
+\lstinline{shard} includes additional manual optimizations over DTD,
+completely eliminating these dynamic checks. StarPU \lstinline{expl}
+is written in an explicitly parallel style using MPI for
+communication, and thus avoids this bottleneck. Regent uses a
+compile-time optimization that improves scalability with no user
+intervention~\cite{ControlReplication17}. We see that all of PaRSEC
+\lstinline{shard}, StarPU \lstinline{expl}, and Regent are able to
+achieve optimal scalability.
 

### 2020-08-13 3_implementation.tex
Update text for task pruning.

@@
 bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU
-allow users to omit tasks that are not directly connected in the
-dependence graph. The program is replicated (for efficiency of
-analysis), but each task is executed only once. Because the dynamic
+allow users to manually prune the task graph, skipping tasks not
+mapped for execution onto a given node, plus a ``halo'' consisting of
+tasks connected via dependencies to the set of node-local tasks. Because the dynamic
 checks to see if a task should be executed are not free of cost, we

### 2020-08-13 5_evaluation.tex
Push new paragraph for scalability.

--- 2020-06-30 5_evaluation.tex
@@
 
-PaRSEC, StarPU and Regent
-rely on runtime analysis that can suffer from
-scalability bottlenecks if every node must consider the tasks
-executing on all other nodes. PaRSEC DTD and StarPU mitigate this
-partially by allowing the user to omit tasks not directly
-dependent on those executed by a given node. However, this approach still requires checks that scale with
-node count and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
-partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
-by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
-retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
-uses a compile-time optimization to improve
-scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.
+PaRSEC, StarPU and Regent rely on runtime analysis that can suffer
+from scalability bottlenecks if every node must consider the tasks
+executing on all other nodes. Although all three systems offer ways to
+improve scalability, these methods are not equally effective. PaRSEC
+DTD and StarPU allow users to manually prune tasks to reduce overhead;
+however the checks are not reduced to zero. Similarly, PaRSEC PTG uses
+compile-time optimizations to avoid the need for manual pruning, but
+still targets the PaRSEC DTD runtime and thus incurs some
+overhead. Figure~\ref{fig:metg-compute} shows that these models do not
+provide ideal scalability, as seen by METG values that rise with
+increasing node count. Regent uses a compile-time optimization to
+generate code with a constant overhead per
+node~\cite{ControlReplication17}. Of the three systems, Regent is the
+only one that achieves ideal scalability while preserving its
+original, implicitly parallel programming model. The others can
+achieve ideal scalability but require increasing levels of manual
+intervention. PaRSEC \lstinline{shard} includes additional manual
+optimizations over DTD to completely eliminate dynamic checks. StarPU
+\lstinline{expl} is written in an explicitly parallel style using MPI
+for communication, and thus avoids any analysis bottleneck. These
+results indicate that the underlying systems are capable of scalable
+execution, but that the dynamic checks incurred by the implicitly
+parallel programming models hinder that scalability.
 

### 2020-08-17 7_conclusion.tex
Update acknowledgements to include all developers.

@@
 No. DE-AC02-05CH11231, and experiments on Piz Daint were supported by
-the Swiss National Supercomputing Centre (CSCS) under project ID d80.
+the Swiss National Supercomputing Centre (CSCS) under project ID
+d80. A special thanks to thank Katie Antypas for her support.
+
+The authors would like to thank the developers of the programming
+systems tested in this paper. Their support enabled us to produce
+high-quality implementations of the systems under study. The
+developers consulted include (alphabetically by system): Chapel:
+Bradford L.~Chamberlain and Elliot Ronaghan; Charm++: Laxmikant Kalé
+and Sam White; Dask: Matthew Rocklin; MPI: Samuel K.~Gutiérrez and Wei
+Wu; OmpSs: Víctor López and Vicenç Beltran Querol; OpenMP: Alejandro
+Duran and Jeff R.~Hammond; PaRSEC: George Bosilca and Qinglei Cao;
+Realm: Seema Mirchandaney and Sean Treichler; Regent: Michael Bauer,
+Wonchan Lee, Seema Mirchandaney and Elliott Slaughter; Spark: Matei
+Zaharia; StarPU: Samuel Thibault; Swift/T: Justin M.~Wozniak;
+TensorFlow: Jing Dong, Peter Hawkins and Mingsheng Hong; X10: David
+Grove, Sara S.~Hamouda and Josh Milthorpe.

### 2020-08-20 5_evaluation.tex
Add paragraph for Chapel distrib in communication.

@@
 
+The effectiveness of such overlap can be influenced by the scheduling
+policies of the underlying system. For example, Chapel's default
+scheduler uses a round-robin policy; we see in
+Figure~\ref{fig:efficiency-communication} that this approach fails to
+take full advantage of the available task parallelism. A work-stealing
+scheduler (Chapel \lstinline{distrib}) is able to recover this
+performance.
+
 \subsection{Load Imbalance}

### 2020-08-20 8_case_study.tex
Rewrite TensorFlow case study.

@@
 
+The Task Bench implementation for TensorFlow uses aggressive loop
+unrolling to minimize overhead, and this initially resulted in the
+entire graph being marked constant. TensorFlow performs optimizations
+on task graphs prior to execution, including a constant folding
+pass. These optimizations are performed on a single core. In the case
+of Task Bench, this resulted in sequential execution of the entire
+graph, resulting in experiments timing out. As a workaround,
+additional inputs can be fed to nodes in the graph to make them
+non-constant, forcing them to be executed by TensorFlow's normal
+scheduler. Though this was exacerbated by Task Bench's approach to
+unrolling loops, it affects any task graph with constants that are
+expensive to compute where the users may wish to perform
+constant-folding in parallel.
+
+\zap{
 Initial experiments with TensorFlow revealed that the Task
@@
 used.
+}
 

### 2020-08-20 8_case_study.tex
Fiddle with TensorFlow paragraph.

@@
 
-The Task Bench implementation for TensorFlow uses aggressive loop
-unrolling to minimize overhead, and this initially resulted in the
-entire graph being marked constant. TensorFlow performs optimizations
-on task graphs prior to execution, including a constant folding
-pass. These optimizations are performed on a single core. In the case
-of Task Bench, this resulted in sequential execution of the entire
-graph, resulting in experiments timing out. As a workaround,
-additional inputs can be fed to nodes in the graph to make them
-non-constant, forcing them to be executed by TensorFlow's normal
+TensorFlow performs optimizations on task graphs prior to execution,
+including a constant folding pass. The Task Bench implementation uses
+aggressive loop unrolling to minimize overheads, resulting in the task
+graph being marked as constant. Because TensorFlow optimizations are
+performed on a single core, this resulted in sequential execution of
+the entire graph, and Task Bench experiments timed out. As a
+workaround, additional inputs can be fed to nodes in the graph to make
+them non-constant, forcing them to be executed by TensorFlow's normal
 scheduler. Though this was exacerbated by Task Bench's approach to

### 2020-08-24 0_abstract.tex
Update abstract.

@@
-% 150 words
+% 149 words
 
@@
 We present Task Bench, a \emph{parameterized} benchmark designed to
-explore the performance of parallel and distributed
-programming systems under a variety of application scenarios. Task
-Bench dramatically lowers the barrier to benchmarking and comparing multiple programming systems by making
-the implementation for a given system orthogonal to the benchmarks
-themselves: every benchmark constructed with Task Bench
-runs on every Task Bench implementation. Furthermore,
-Task Bench's parameterization enables a wide variety of benchmark
-scenarios that distill the key characteristics of larger applications.
+explore the performance of distributed programming systems under a
+variety of application scenarios. Task Bench dramatically lowers the
+barrier to benchmarking and comparing multiple programming systems by
+making the implementation for a given system orthogonal to the
+benchmarks themselves: every benchmark constructed with Task Bench
+runs on every Task Bench implementation. Furthermore, Task Bench's
+parameterization enables a wide variety of benchmark scenarios that
+distill the key characteristics of larger applications.
 
-We conduct a comprehensive study with
-15 programming systems on up to 256 Haswell nodes of the Cori
-supercomputer. We introduce a novel metric, \emph{minimum effective task
-  granularity} to study the baseline runtime overhead of each
-system. Running at scale, 100~\textmu{}s is the
-smallest granularity that any system can reliably
-support with current technologies. We also study each system's
-scalability, ability to hide communication and mitigate load
+To assess the effectiveness and overheads of the tested systems, we
+introduce a novel metric, \emph{minimum effective task granularity}
+(METG). We conduct a comprehensive study with 15 programming systems
+on up to 256 Haswell nodes of the Cori supercomputer. Running at
+scale, 100\textmu{}s-long tasks are the finest granularity that any
+system runs efficiently with current technologies. We also study each
+system's scalability, ability to hide communication and mitigate load
 imbalance.

### 2020-08-24 8_case_study.tex

@@
 Dask developers; the bug is suspected to be in optimizations performed
-on the task graph by Dask. As a work around, our results in the paper use a
+on the task graph by Dask. As a workaround, our results in the paper use a
 lower-level interface which does not suffer from this asymptotic
@@
 the entire graph, and Task Bench experiments timed out. As a
-workaround, additional inputs can be fed to nodes in the graph to make
-them non-constant, forcing them to be executed by TensorFlow's normal
+workaround, additional inputs are fed to nodes in the graph to make
+them non-constant, forcing them to be executed by TensorFlow's parallel
 scheduler. Though this was exacerbated by Task Bench's approach to

### 2020-08-24 3_implementation.tex
Apply Alex's comments.

@@
 above) is explicitly parallel and requires tasks to be connected
-explicitly via \emph{events}. StarPU, in addition to its usual,
+explicitly via \emph{events}. Realm is the low-level execution engine for Regent and thus serves as a limit study of what can be achieved with Regent. StarPU, in addition to its usual,
 implicitly parallel mode, also provides a mode where MPI is used for
@@
 
-Among distributed task-based systems, there can be a scalability
+Among implicitly parallel, distributed task-based systems, there can be a scalability
 bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU

### 2020-08-24 5_evaluation.tex
minor edit to say parsec also support both inline and offline configurations
Apply Alex's comments.

@@
 overhead \emph{out-of-line} (i.e., by dedicating one or more cores
-solely to runtime use). Some systems, like Charm++, Realm, and Regent,
+solely to runtime use). Some systems, like Charm++, PaRSEC, Realm, and Regent,
 support both configurations.
@@
 provide additional features such as automatic dependency discovery and
-data movement. Slower systems tend to be designed primarily for
+data movement. Higher overhead systems tend to be designed primarily for
 large-scale data analysis or workflows. It is worth
@@
 maintaining at least 50\% efficiency. In
-Figure~\ref{fig:strong-scaling} this can be seen as the distance
+Figure~\ref{fig:strong-scaling}, the error in the estimate is the distance
 between the black square and circle; these are separated by a factor of

### 2020-08-24 7_conclusion.tex
Apply Alex's comments.

@@
 \begin{itemize}
+\item Evaluations of programming system performance should avoid using
+  TPS or weak scaling to characterize overheads, as these metrics do
+  not constrain the useful work achieved. Instead a metric with
+  constrained efficiency, such as METG(50\%), is needed to ensure that
+  measurements are representative and fair.
+
 \item METG for current distributed programming systems varies by over
@@
   high baseline overheads.
+
+\item Task-based systems that rely on runtime analysis for the
+  discovery of parallelism can suffer from sequential bottlenecks that
+  limit scaling. Existing, dynamic task pruning techniques are not
+  sufficient to fully mitigate this bottleneck, while static,
+  compile-time approaches are able to do so.
 

### 2020-08-24 e11_metg_predict_strong.tex
Apply Alex's comments.

@@
 \vspace{-0.20cm}
-\caption{Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
+\caption{Factor of separation between limit-ideal and limit-actual intersections across all 11 programming systems tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
 \vspace{-0.5cm}

### 2020-08-24 7_conclusion.tex
Tweak text.

@@
 \item Evaluations of programming system performance should avoid using
-  TPS or weak scaling to characterize overheads, as these metrics do
+  TPS, or strong or weak scaling to characterize overheads, as these metrics do
   not constrain the useful work achieved. Instead a metric with

### 2020-08-25 2_task_bench.tex
Clarify that listing 1 is an excerpt.
Edit paragraph about MPI implementation.

@@
 Table~\ref{tab:api}. The MPI implementation follows the style of
-communicating sequential processes, and simply executes a set of send
+communicating sequential processes (CSP), and executes a set of send
 and receive calls (lines 24 and 16, respectively) followed by
-executing the task body (line 32). For simplicity, the excerpt
-assumes the number of MPI ranks is equal to the width of the graph;
-the full implementation is more general and provides additional
-optimizations. Despite its simplicity, this MPI implementation is
-among our best performing, and shows that Task Bench can be mapped to
-idiomatic code even in programming systems with no native concept of
-task.
+executing the task body (line 32). Despite MPI having no notion of
+task, the execution of a task graph maps into the CSP style in a
+straightforward way. The implementation is both simple and efficient,
+but do to the choice of CSP makes no attempt to exploit task
+parallelism, and leaves performance on the table when executing task
+graphs with load imbalance or communication. Note the excerpt is
+simplified for presentation and the full implementation is more
+general and provides additional optimizations.
 
@@
 executed by each task as well as other utility routines (to parse
-inputs and display results). A sample of the core API compute kernel
+inputs and display results). An excerpt from the core API compute kernel
 is shown in Listing~\ref{lst:compute-kernel}. In addition to reducing

### 2020-08-25 2_task_bench.tex
Fix typo.

@@
 straightforward way. The implementation is both simple and efficient,
-but do to the choice of CSP makes no attempt to exploit task
+but due to the choice of CSP makes no attempt to exploit task
 parallelism, and leaves performance on the table when executing task

### 2020-08-26 2_task_bench.tex
Tweak paragraph on MPI.

--- 2020-08-25 2_task_bench.tex
@@
 Table~\ref{tab:api}. The MPI implementation follows the style of
-communicating sequential processes (CSP), and executes a set of send
+communicating sequential processes (CSP)~\cite{HoareCSP1978}, and executes a set of send
 and receive calls (lines 24 and 16, respectively) followed by
@@
 straightforward way. The implementation is both simple and efficient,
-but do to the choice of CSP makes no attempt to exploit task
+but due to the choice of CSP makes no attempt to exploit task
 parallelism, and leaves performance on the table when executing task
-graphs with load imbalance or communication. Note the excerpt is
+graphs with load imbalance or significant communication. Note the excerpt is
 simplified for presentation and the full implementation is more

### 2020-08-27 5_evaluation.tex
Update figure placement and performance numbers.

@@
 For comparison, the
-OpenMP-enabled STREAM benchmarks~\cite{STREAM} report up to 98~GB/s.
+OpenMP-enabled STREAM benchmarks~\cite{STREAM} report up to 98~GB/s on the same hardware.
 
@@
 generally be determined by the application domain being
-considered. Most notably, for large-scale data analytics workloads, the higher METG values observed for Spark and
-TensorFlow are sufficient. In contrast, for high-performance
+considered. Most notably, for large-scale data analytics workloads, the higher METG values observed for Spark are sufficient. In contrast, for high-performance
 scientific simulations, task granularities in the millisecond range
@@
 
+\input{e6_communication}
+
 PaRSEC, StarPU and Regent rely on runtime analysis that can suffer
@@
 
-%% From 1 to 3
-%% dependencies, the ratio varies from $1.0\times$ to $150\times$ (median
-%% $2.3\times$).
-
 The ratio in METG between 0 and 3 dependencies per task ranges from
-$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
-perform runtime work inline. For example, MPI achieves an METG of 390
-ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
-a $12\times$ increase. This is unsurprising, as with
-0 dependencies no \lstinline[language=C++]{MPI_Isend} calls are
-issued at all. Clearly, choosing a representative dependence
-pattern is important when estimating the performance of a workload or
-class of workloads.
-
-\input{e6_communication}
+$0.82\times$ to $250\times$ (mean $21\times$, std. dev. $64$). The
+large standard deviation shows that the sensitivity of system overhead
+to the dependency pattern varies widely. The largest ratios are among
+systems that perform runtime work inline. For example, MPI achieves an
+METG of 390 ns with 0 dependencies, but this rises to 4.6 \textmu{}s
+with 3 dependencies, a $12\times$ increase. This is unsurprising, as
+with 0 dependencies no \lstinline[language=C++]{MPI_Isend} calls are
+issued at all. Clearly, choosing a representative dependence pattern
+is important when estimating the performance of a workload or class of
+workloads.
 
@@
 
-\input{e9_strong}
-\input{e11_metg_predict_strong}
-
 \subsection{Validating METG with Time to Solution}
@@
 % Limit/Ideal vs Limit/Actual gap in time: 1.27x
-
-% Overall (out of 11 systems that perform well enough to have an intersection):
-% Limit/Ideal vs Limit/Actual gap in nodes:
-%   mean: 1.61x
-%   median: 1.60x
-%   min: 1.13x
-%   max: 2.19x
-% Limit/Ideal vs Limit/Actual gap in time:
-%   mean: 1.28x
-%   median: 1.27x
-%   min: 1.01x
-%   max: 1.69x
 
@@
 
+\input{e9_strong}
+\input{e11_metg_predict_strong}
+
 Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
@@
 between the limit-ideal and limit-actual intersections. We see that
-overall, the mean separation is at most $2.15\times$ in node count
-and at most $1.42\times$ in time to solution, making this a
+overall, the mean separation is at most $1.96\times$ in node count
+and at most $1.29\times$ in time to solution, making this a
 useful way to predict strong scaling in the 4 patterns we tested in

### 2020-08-27 5_evaluation.tex
Update number of systems in validation comparison.

@@
 Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
-patterns tested in Section~\ref{subsec:scalability} for all 11
+patterns tested in Section~\ref{subsec:scalability} for all 12
 programming systems that scale well enough to evaluate the separation
@@
 Section~\ref{subsec:scalability}.
-
-\zap{
-When we consider all 11 programming systems that scale well enough to
-perform this experiment, we find that the average separation between
-the ideal and actual intersection points is $1.61\times$ in node count
-(max $2.19\times$, min $1.13\times$) and $1.28\times$ in time to
-solution (max $1.69\times$, min $1.01\times$).
-}
 

### 2020-08-27 e11_metg_predict_strong.tex
Update figure placement and performance numbers.
Update number of systems in validation comparison.

--- 2020-06-30 e11_metg_predict_strong.tex
@@
 \hline
-stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
-nearest & $1.77\times$ & 0.253 & $1.24\times$ & 0.310 \\
-spread  & $1.39\times$ & 0.322 & $1.30\times$ & 0.171 \\
-nearest, 4 graphs & $2.15\times$ & 0.635 & $1.42\times$ & 0.723
+stencil & $1.64\times$ & 0.278 & $1.26\times$ & 0.246 \\
+nearest & $1.75\times$ & 0.268 & $1.15\times$ & 0.331 \\
+spread  & $1.43\times$ & 0.288 & $1.29\times$ & 0.206 \\
+nearest, 4 graphs & $1.96\times$ & 0.680 & $1.26\times$ & 0.702
 \end{tabular}
@@
 \vspace{-0.20cm}
-\caption{Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
+\caption{Factor of separation between limit-ideal and limit-actual intersections across all 12 programming systems tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
 \vspace{-0.5cm}
